### 2013-04-18 man/proposal.tex
Add proposal tex

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
%\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09

\usepackage{amsmath, amsthm, amssymb}
% \usepackage{natbib}
\usepackage{graphicx}
\usepackage{wrapfig}

\pagestyle{empty}

% \usepackage{etoolbox}
% \makeatletter
% \patchcmd{\thebibliography}{%
%   \chapter*{\bibname}\@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}}{%
%   \subsubsection*{References}}{}{}
% \makeatother

\title{Optimal strategies in mental rotation tasks}
\author{Jessica B.~Hamrick\\
  Department of Psychology\\
  University of California, Berkeley\\
  Berkeley, CA 94720\\
  \texttt{jhamrick@berkeley.edu}}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

% Your proposal should be one page long (single-spaced), and should have
% four sections:

% \begin{wrapfigure}{r}{0.4\textwidth}
\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.95\textwidth]{MentalRotationFig2.png}
  \end{center}
  \caption{Classic mental rotation task from \cite{Shepard1971}.}
  \label{fig:mental-rotation}
% \end{wrapfigure}
\end{figure}

\section{Background}
% - Background. Identify the topic your project will explore, and
% briefly provide some of the context for your project, describing
% what previous research has found in this area.

% 1. The big picture
% 2. Our contribution
% 3. How we do it
% 4. What's in the paper

Consider the objects in Figure \ref{fig:mental-rotation}. In each
panel, are the two depicted objects identical (except for a rotation),
or distinct? When presented with this mental rotation task, people
default to a strategy in which they visualize one object rotating until
it is congruent with the other \cite{Shepard1971}. There is strong
evidence for such \textit{mental imagery} or \textit{mental
  simulation}: we can imagine three-dimensional objects in our minds
and manipulate them, to a certain extent, as if they were real
\cite{Kosslyn:2009tj}.

However, the use of mental simulation is predicated on determining
appropriate parameters to give the simulation, and people's cognitive
constraints may furthermore place limitations on the duration or
precision of simulation. One hypothesis for how these issues are
handled argues that people use a ``rational'' solution, meaning that
it is optimal under the given constraints
\cite{Lieder:2012wg,Vul:2009wy,Griffiths2012a}.

\section{Question}
% - Question. State the specific question you are going to examine in
% your final project.

In the case of the classic mental rotation task, we might ask: in what
direction should the object be rotated?  What are the requirements for
``congruency''? When should one stop rotating and accept the
hypothesis that the objects are different? In this project, I will
investigate the optimal computational solution to this task. Thus, the
specific question this project aims to answer is: what is the rational
computational solution to the mental rotation task, and does it
predict the people's behavior on the task?

\section{Method}
% - Method. Briefly describe the method you are going to use to try to
% answer this question. Give some of the details behind your
% experimental procedure, your approach to modeling or analyzing the
% data, your plans for analyzing the model, or the position you will
% take in your review.

We can formalize the mental rotation task as an instance of function
learning. Given two images $I_a$ and $I_b$, we wish to determine the
probability of seeing both images given the hypothesis ($h_1$) that
one is merely a rotation of the other: $\Pr(I_b, I_a|h_1)$. We can
generate some mental image $I_M$ using mental rotation, but this is a
costly operation and each $I_M$ must be computed sequentially (that
is, each $I_t$ must be generated from $I_{t-\delta}$, where $\delta$
is a small angle). Let us define the pixel-space similarity between
$I_b$ and $I_M$ as $S(I_b| I_M)$. Then:

\begin{equation}
\Pr(I_b, I_a | h_1)\approx\int_{I_M}S(I_b| I_M)\Pr(I_M|I_a)\Pr(I_a)\ \mathrm{d}I_M\propto \int_{I_M}S(I_b|I_M)\ \mathrm{d}I_M
\end{equation}

We do not know the form of $S$, and so cannot compute this integral
analytically. We cannot rely on a simple Monte Carlo simulation to
estimate it, either: because mental rotation is costly in terms of
cognitive resources and must be performed in a sequence, we cannot
evaluate $S(I_b|I_M)$ at arbitrary $I_M$. Instead, we must choose a
relatively small, sequential set of $I_M$ to estimate the
integral.

Bayesian quadrature \cite{Diaconis:1988uo} provides an alternate
method for evaluating the integral by placing a prior distribution on
functions and then computing a posterior over values of the
integral. While Bayesian quadrature itself does not address the issue
of choosing points to evaluate the function at, recent work in
statistics and machine learning has examined how to optimally select
these samples \cite{Osborne:2012tm}.

In this project, I will apply these methods to the classic mental
rotation task in both two and three dimensions. From the extensive
literature on mental imagery, I will determine appropriate cognitive
limitations and compute the optimal tradeoff between computation time
(precision) and certainty of answer (accuracy). Finally, I will
compare the results of this model with real behavioral data to test
whether people are, in fact, using a rational solution in the mental
rotation task.

% \renewcommand\bibsection{\subsubsection*{\refname}}
\renewcommand\refname{\normalsize{References}}
\bibliographystyle{ieeetr}
\bibliography{references}
% - References. Give at least three references cited in describing the
% background to your project.

\end{document}

### 2013-05-13 man/final-project.tex
Initial version of final project writeup, from notes and proposal

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
%\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09

\usepackage{amsmath, amsthm, amssymb}
% \usepackage{natbib}
\usepackage{graphicx}
\usepackage{wrapfig}

\pagestyle{empty}

% \usepackage{etoolbox}
% \makeatletter
% \patchcmd{\thebibliography}{%
%   \chapter*{\bibname}\@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}}{%
%   \subsubsection*{References}}{}{}
% \makeatother

\title{Optimal strategies in mental rotation tasks}
\author{Jessica B.~Hamrick\\
  Department of Psychology\\
  University of California, Berkeley\\
  Berkeley, CA 94720\\
  \texttt{jhamrick@berkeley.edu}}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

% \begin{wrapfigure}{r}{0.4\textwidth}
\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.95\textwidth]{../figures/shepard-rotation.png}
  \end{center}
  \caption{Classic mental rotation task from \cite{Shepard1971}.}
  \label{fig:mental-rotation}
% \end{wrapfigure}
\end{figure}

\section{Introduction}

% 1. The big picture
% 2. Our contribution
% 3. How we do it
% 4. What's in the paper

Consider the objects in Figure \ref{fig:mental-rotation}. In each
panel, are the two depicted objects identical (except for a rotation),
or distinct? When presented with this mental rotation task, people
default to a strategy in which they visualize one object rotating
until it is congruent with the other \cite{Shepard1971}. There is
strong evidence for such \textit{mental imagery} or \textit{mental
  simulation}: we can imagine three-dimensional objects in our minds
and manipulate them, to a certain extent, as if they were real
\cite{Kosslyn:2009tj}.

However, the use of mental simulation is predicated on determining
appropriate parameters to give the simulation, and people's cognitive
constraints may furthermore place limitations on the duration or
precision of simulation. One hypothesis for how these issues are
handled argues that people use a ``rational'' solution, meaning that
it is optimal under the given constraints
\cite{Lieder:2012wg,Vul:2009wy,Griffiths2012a}.

qIn the case of the classic mental rotation task, we might ask: in what
direction should the object be rotated?  What are the requirements for
``congruency''? When should one stop rotating and accept the
hypothesis that the objects are different? In this project, I will
investigate the optimal computational solution to this task. Thus, the
specific question this project aims to answer is: what is the rational
computational solution to the mental rotation task, and does it
predict the people's behavior on the task?

\section{Background}

\subsection{Mental rotation}

\subsection{Optimal experiment design}
% We do not know the form of $S$, and so cannot compute this integral
% analytically. We cannot rely on a simple Monte Carlo simulation to
% estimate it, either: because mental rotation is costly in terms of
% cognitive resources and must be performed in a sequence, we cannot
% evaluate $S(I_b|I_M)$ at arbitrary $I_M$. Instead, we must choose a
% relatively small, sequential set of $I_M$ to estimate the
% integral.

% Bayesian quadrature \cite{Diaconis:1988uo} provides an alternate
% method for evaluating the integral by placing a prior distribution on
% functions and then computing a posterior over values of the
% integral. While Bayesian quadrature itself does not address the issue
% of choosing points to evaluate the function at, recent work in
% statistics and machine learning has examined how to optimally select
% these samples \cite{Osborne:2012tm}.

\section{Computational model}

People are presented with two images, $X_a$ and $X_b$, which are the
coordinates of the vertices of two shapes. Participants must determine
whether $X_a$ and $X_b$ were generated from the same (albeit possibly
transformed and permuted) original shape, i.e., whether $\exists
R,M\textrm{ s.t. } X_b=MRX_a$, where $M$ is a permutation matrix and
$R$ is a rotation matrix. \TODO{Figure with some example shapes}

We can formulate the judgment of whether $X_a$ and $X_b$ have the same
origins by deciding about two hypotheses:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  $h_0$: $\forall M,R\ X_b\neq MRX_a$
\item
  $h_1$: $\exists M,R\textrm{ s.t. } X_b=MRX_a$
\end{itemize}

To compare the hypotheses, we need to compute the likelihood of each
hypothesis, i.e. $p(X_a,\ X_b\ \vert \ h_0)$ and $p(X_a,\ X_b\ \vert \
h_1)$.

\subsection{Likelihood when the configurations are distinct}

The likelihood is easy to compute under $h_0$, because we assume that
$X_a$ and $X_b$ are independent. Thus:
\begin{equation}
  p(X_a,\ X_b\ \vert \ h_0)=p(X_a)p(X_b)
\end{equation}

\subsection{Likelihood when the configurations are equivalent}

When the configurations are the same, the likelihood becomes more
complicated:
\begin{equation} 
  p(X_a,\ X_b\ \vert \ h_1)=\int_R\int_M p(X_a) p(X_b\vert X_a,R,M) p(R) p(M)\ \mathrm{d}M\ \mathrm{d}R
\end{equation}

For a small number of vertices, we can compute the integral over $M$
by enumerating every possible mapping between $X_a$ and $X_b$. After
doing so, we obtain:
\begin{equation} 
  p(X_a,\ X_b\ \vert \ h_1)=\int_R p(X_a) p(X_b\vert X_a,R) p(R)\ \mathrm{d}R
\end{equation}

However, we cannot compute $p(X_b\vert X_a, R)$ directly. Instead, we
introduce a new variable $X_R$ denoting a mental image, which
approximates $RX_a$. The $X_R$ are generated sequentially by repeated
application of a function $\tau$:
\begin{align*}
  X_R&=RX_a\\
  &=\tau(X_{R-r}, r)\\
  &=\tau(\tau(X_{R-2r}, r), r)\\
  &\ldots{}\\
  &=\tau^{(\frac{R}{r})}(X_a, r)
\end{align*} 
Where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
applications of $\tau$. Using this sequential function, we get:
\begin{align}
  p(X_a, X_b\ \vert \ h_1)&=\int_R \int_{X} p(X_b\vert X) p(X\vert X_a, R)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
  x&= \int_R \int_X p(X_b\vert X)\delta(\tau^{(\frac{R}{r})}(X_a, r)-X)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
  &= \int_R p(X_b\vert X_R)p(X_a)p(R)\ \mathrm{d}R
\end{align}

\subsection{Likelihood ratio}

Once we have computed both likelihoods, we compute their ratio:
\begin{equation}
  \ell=\frac{p(X_a, X_b\ \vert \ h_1)}{p(X_a, X_b\ \vert \ h_0)}
\end{equation} 
If $\ell<1$, then $h_0$ is the more likely hypothesis. If $\ell>1$,
then $h_1$ is the more likely hypothesis.

\section{Implementation}

\subsection{Likelihood when the configurations are distinct}

We define the prior probabilities over stimuli based on how they are
generated. For shapes with $n$ vertices, each vertex is at a random
angle with a radius random chosen between 0 and 1, in polar
coordinates, and could be chosen in any of $n!$ different ways. Thus,
the prior over shapes is:
\begin{equation}
  p(X)=n!\left(\frac{1}{2\pi}\right)^n
\end{equation} 
And so the joint likelihood is just:
\begin{equation}
  p(X_a, X_b\ \vert \ h_0)=\left[n!\left(\frac{1}{2\pi}\right)^n\right]^2
\end{equation}

\subsection{Likelihood when the configurations are equivalent}

We cannot know the quantity $p(X_b\vert X_R)$ exactly, so we
approximate it with a similarity function $S(X_b, X_R)$, which also
takes into account the different possible mappings of vertices. There
are $n$ mappings $M$ of the $n$ vertices: because some vertices are
connected to form a closed loop, we assume the uncertainty is only in
which is the ``first'' vertex. So, the possible orderings are of the
form $M=\lbrace{}0, 1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0,
\ldots{}, n-1\rbrace{}$, and so on. This gives us the final form of
the similarity function:
\begin{equation}
  S(X_b, X_R)=\frac{1}{n}\sum_{M\in\mathbb{M}}\prod_{i=1}^n\mathcal{N}(X_b[i]\ \vert \ (MX_R)[i], \Sigma)
\end{equation}
Where $i$ denotes the $i^{th}$ vertex. We do not know the closed,
analytic form of this function, but we can evaluate it for any $X_b$
and $X_R$. Given this, the desired quantity for $h_1$ becomes:
\begin{equation}
  Z = p(X_a,X_b\ \vert \ h_1)\approx \int_R S(X_b, X_R)p(X_a)p(R)\ \mathrm{d}R
\end{equation}
Note that for brevity, from here on we will refer to $p(X_a, X_b\
\vert \ h_1)$ as $Z$.

Finally, computing any single $X_R$ is a costly operation in that it
is both a lengthy operation and memory-intensive, so we want to
minimize the number of $X_R$ that are computed. With this in mind, we
now examine several different approaches to estimating $S$ and,
ultimately, $Z$.

\subsubsection{Gold standard}

For comparison, we compute a ``gold standard'' by evaluating $S(X_b,
X_R)$ at 360 values spaced evenly between $0$ and $2\pi$.

\subsubsection{Linear Interpolation}

Na\"ively, we can simply use linear interpolation to estimate $S$ and
the trapezoidal rule to estimate $Z$.

\paragraph{Choosing the sequence of mental images}

In this naïve strategy, we perform a hill-climbing search to find a
local maxima. In other words, we rotate in the direction for which
there is a local increase in similarity.

\paragraph{Stopping criterion}

Because we are using hill-climbing, we choose to stop when we hit a
local maxima.

\subsubsection{Parametric Regression}

Another strategy is to assume a parametric shape for $S$ and fit the
appropriate parameters. Here, we assume the likelihood follows a
\emph{Von Mises} distribution:
\begin{equation}
  S(X_b, X_R) \approx \hat{S}_{VM}=\frac{h}{2\pi X_0(\kappa)}e^{\kappa\cos(R-\hat{\theta})}
\end{equation}

Where $\kappa$ is the concentration parameter, $\hat{\theta}$ is the
preferred direction, and $h$ is a scale parameter. We can fit these
parameters by minimizing the mean squared error between this PDF and
the observed values of $S$ which we are given. Then:
\begin{equation}
E[Z\ \vert \ S]\approx \int_R \hat{S}_{VM}p(R)\ \mathrm{d}R
\end{equation}

\paragraph{Choosing the sequence of mental images}

As above, we use hill-climbing, with the stopping criteria being when
we find a local maxima.

\subsubsection{Bayesian Quadrature}

A more flexible strategy uses what is known as \emph{Bayesian
  quadrature} \cite{Diaconis:1988uo,OHagan:1991tx} to estimate $Z$.
Bayesian quadrature allows us to compute a posterior distrbution over
$Z$ by placing a Gaussian Process (GP) \cite{Rasmussen:2006vz} prior
on the log likelihood function $\log S$ and evaluating $\log S$ at a
particular set of points. For our purposes, we choose these points to
be a sequence of $N$ consecutive mental rotations be
$\mathbf{R}=\{R_1, \ldots{}, R_N\}$. To obtain a single estimate of
$Z$, we want compute the posterior mean:
\begin{equation}
  E[Z\ \vert \ \log S]=\int_{\log S}\left(\int_R \exp(\log{S(X_b,X_R)})p(R)\ \mathrm{d}R\right)\mathcal{N}\left(\log{S}\ \vert \ \mu_{\log S}, \Sigma_{\log S}\right)\ \mathrm{d}\log S
\end{equation}
As in \cite{Osborne:2012tm}, we can approximate this with:
\begin{equation}
  E[Z\ \vert \ \log S] \approx \int_R \mu_{S}(1 + \mu_\Delta) p(R)\ \mathrm{d}R
\end{equation} 
Where $\mu_x$ is the mean of a GP, as follows:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item $\mu_\Delta$ is a regression over $\Delta=\mu_{\log S} - \log
  \mu_S$ given $\mathbf{R}_\Delta$, which consists of $\mathbf{R}$ and
  a set of intermediate \emph{candidate points}, as described in
  \cite{Osborne:2012tm}.
\item $\mu_S$ is the mean of a GP regression over $S$ given
  $\mathbf{R}$.
\item $\mu_{\log S}$ is the mean of a GP regression over $\log S$
  given $\mathbf{R}$.
\end{itemize}

\paragraph{Choosing the sequence of mental images}

Because we can only evaluate $S(X_a, X_R)$ at a small number $k$ of
rotations, we must make an explicit decision about which
$\mathbf{R}=\{R_1, \ldots{}, R_k\}$ to use. Ideally, we should choose
$\mathbf{R}$ such that it will give us the best approximation of
$p(X_a, X_b\vert h_1)$. One option is to choose each $R_i$ in a greedy
manner to minimize the variance of $Z$ (i.e.~give a more accurate
estimate of $S$).

\section{Results}

\section{Discussion}

\section{Method}

% \renewcommand\bibsection{\subsubsection*{\refname}}
\renewcommand\refname{\normalsize{References}}
\bibliographystyle{ieeetr}
\bibliography{references}
% - References. Give at least three references cited in describing the
% background to your project.

\end{document}

### 2013-05-13 man/final-project.tex
Add figures
Remove random comments
Add abstract with TODO
Fix typo
Clean up computational model section

@@
 \usepackage{wrapfig}
+\usepackage{subcaption}
 
 \pagestyle{empty}
-
-% \usepackage{etoolbox}
-% \makeatletter
-% \patchcmd{\thebibliography}{%
-%   \chapter*{\bibname}\@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}}{%
-%   \subsubsection*{References}}{}{}
-% \makeatother
 
@@
 
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
 \newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}
@@
 
-% \begin{wrapfigure}{r}{0.4\textwidth}
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[width=0.95\textwidth]{../figures/shepard-rotation.png}
-  \end{center}
-  \caption{Classic mental rotation task from \cite{Shepard1971}.}
-  \label{fig:mental-rotation}
-% \end{wrapfigure}
-\end{figure}
+\begin{abstract}
+\TODO{}
+\end{abstract}
 
@@
 
-qIn the case of the classic mental rotation task, we might ask: in what
+In the case of the classic mental rotation task, we might ask: in what
 direction should the object be rotated?  What are the requirements for
@@
 
-\subsection{Mental rotation}
-
-\subsection{Optimal experiment design}
 % We do not know the form of $S$, and so cannot compute this integral
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.95\textwidth]{../figures/shepard-rotation.png}
+  \caption{Classic mental rotation task from \cite{Shepard1971}.}
+  \label{fig:mental-rotation}
+\end{figure}
+
 \section{Computational model}
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/stimuli_shapes.pdf}
+    \vspace{0pt}
+    \caption{Example stimuli}
+    \label{fig:stimuli}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/likelihood_function.pdf}
+    \caption{Likelihood function}
+    \label{fig:likelihood}
+  \end{subfigure}
+  \caption{Rotated shapes and corresponding similarities.}
+\end{figure}
 
@@
 
-\subsection{Likelihood when the configurations are distinct}
-
 The likelihood is easy to compute under $h_0$, because we assume that
@@
   p(X_a,\ X_b\ \vert \ h_0)=p(X_a)p(X_b)
+  \label{eq:lh-h0}
 \end{equation}
-
-\subsection{Likelihood when the configurations are equivalent}
 
@@
 application of a function $\tau$:
-\begin{align*}
-  X_R&=RX_a\\
-  &=\tau(X_{R-r}, r)\\
-  &=\tau(\tau(X_{R-2r}, r), r)\\
-  &\ldots{}\\
+\begin{align}
+  X_R&=RX_a\nonumber \\
+  &=\tau(X_{R-r}, r)\nonumber \\
+  &=\tau(\tau(X_{R-2r}, r), r)\nonumber \\
+  &\ldots{}\nonumber \\
   &=\tau^{(\frac{R}{r})}(X_a, r)
-\end{align*} 
+  \label{eq:tau}
+\end{align} 
 Where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
@@
   &= \int_R p(X_b\vert X_R)p(X_a)p(R)\ \mathrm{d}R
+  \label{eq:lh-h1}
 \end{align}
-
-\subsection{Likelihood ratio}
 
@@
 \begin{equation}
-  \ell=\frac{p(X_a, X_b\ \vert \ h_1)}{p(X_a, X_b\ \vert \ h_0)}
+  \ell=\frac{p(X_a, X_b\ \vert \ h_1)}{p(X_a, X_b\ \vert \ h_0)}=\frac{\int_R p(X_b\ \vert\ X_R)p(R)\ \mathrm{d}R}{p(X_b)}
+  \label{eq:lh-ratio}
 \end{equation}
@@
 \section{Implementation}
-
-\subsection{Likelihood when the configurations are distinct}
 
@@
 \begin{equation}
+\subsection{Likelihood when the configurations are distinct}
+
   p(X)=n!\left(\frac{1}{2\pi}\right)^n
@@
 
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
+    \caption{Linear interpolation}
+    \label{fig:li}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/vm_regression.pdf}
+    \caption{Parametric regression}
+    \label{fig:vm}
+  \end{subfigure}
+\end{figure}
+
 Na\"ively, we can simply use linear interpolation to estimate $S$ and
@@
 \subsubsection{Bayesian Quadrature}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/bq_regression.pdf}
+  \caption{Bayesian Quadrature}
+  \label{fig:bq}
+\end{figure}
 
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
+  \caption{Model rotations}
+  \label{fig:rotations}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/Z_accuracy.pdf}
+  \caption{Accuracy in estimating $Z$.}
+  \label{fig:accuracy}
+\end{figure}
+
 \section{Discussion}
-
-\section{Method}
 
@@
 \bibliography{references}
-% - References. Give at least three references cited in describing the
-% background to your project.
 

### 2013-05-13 man/final-project.tex
Clean up implementation section, up to gold standard model

@@
 coordinates, and could be chosen in any of $n!$ different ways. Thus,
-the prior over shapes is:
+the prior over a shape $X$ is:
 \begin{equation}
-\subsection{Likelihood when the configurations are distinct}
+  p(X)=n!\left(\frac{1}{2\pi}\right)^n
+  \label{eq:prior}
+\end{equation} 
+which gives us the denominator in Equation \ref{eq:lh-ratio}.
 
-  p(X)=n!\left(\frac{1}{2\pi}\right)^n
-\end{equation} 
-And so the joint likelihood is just:
+Computing the numerator of Equation \ref{eq:lh-ratio} is more
+difficult, as we do know $p(X_b\vert X_R)$. We approximate it with a
+similarity function $S(X_b, X_R)$, which also takes into account the
+different possible mappings of vertices:
 \begin{equation}
-  p(X_a, X_b\ \vert \ h_0)=\left[n!\left(\frac{1}{2\pi}\right)^n\right]^2
+  Z=\int_R S(X_b, X_R)p(R)\ \mathrm{d}R\ \approx\int_Rp(X_b\ \vert\ X_R)p(R)\ \mathrm{d}R
+  \label{eq:Z}
 \end{equation}
 
-\subsection{Likelihood when the configurations are equivalent}
-
-We cannot know the quantity $p(X_b\vert X_R)$ exactly, so we
-approximate it with a similarity function $S(X_b, X_R)$, which also
-takes into account the different possible mappings of vertices. There
-are $n$ mappings $M$ of the $n$ vertices: because some vertices are
-connected to form a closed loop, we assume the uncertainty is only in
-which is the ``first'' vertex. So, the possible orderings are of the
-form $M=\lbrace{}0, 1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0,
-\ldots{}, n-1\rbrace{}$, and so on. This gives us the final form of
-the similarity function:
+Because the vertices are connected in a way which forms a closed loop,
+we need only consider $n$ mappings, $M$, of the $n$ vertices (we
+assume the uncertainty is only in which is the ``first''
+vertex). Thus, the possible orderings are of the form $M=\lbrace{}0,
+1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0, \ldots{}, n-1\rbrace{}$,
+and so on. This gives us an explicit form for the similarity function:
 \begin{equation}
   S(X_b, X_R)=\frac{1}{n}\sum_{M\in\mathbb{M}}\prod_{i=1}^n\mathcal{N}(X_b[i]\ \vert \ (MX_R)[i], \Sigma)
+  \label{eq:similarity}
 \end{equation}
-Where $i$ denotes the $i^{th}$ vertex. We do not know the closed,
-analytic form of this function, but we can evaluate it for any $X_b$
-and $X_R$. Given this, the desired quantity for $h_1$ becomes:
-\begin{equation}
-  Z = p(X_a,X_b\ \vert \ h_1)\approx \int_R S(X_b, X_R)p(X_a)p(R)\ \mathrm{d}R
-\end{equation}
-Note that for brevity, from here on we will refer to $p(X_a, X_b\
-\vert \ h_1)$ as $Z$.
+where $i$ denotes the $i^{th}$ vertex. An example of $S$ for the
+stimuli shown in Figure \ref{fig:stimuli} is illustrated in Figure
+\ref{fig:likelihood}.
 
-Finally, computing any single $X_R$ is a costly operation in that it
-is both a lengthy operation and memory-intensive, so we want to
-minimize the number of $X_R$ that are computed. With this in mind, we
-now examine several different approaches to estimating $S$ and,
-ultimately, $Z$.
+This process of mental rotation (i.e., generating a single $X_R$ (as
+in Equation \ref{eq:tau}) and then computing $S(X_b, X_R)$) is a
+computationally demanding cognitive process. Thus, our goal is to
+minimize the number of rotations while still obtaining an estimate of
+$Z$ that is accurate enough to choose the correct hypothesis. With
+this in mind, we now examine several approaches to estimating $Z$.
 

### 2013-05-13 man/final-project.tex
Clean up gold standard model section

@@
 
-\subsubsection{Gold standard}
+\subsection{Gold standard model}
 
-For comparison, we compute a ``gold standard'' by evaluating $S(X_b,
-X_R)$ at 360 values spaced evenly between $0$ and $2\pi$.
+To compare the accuracy of other models' estimates of $Z$, we computed
+a ``gold standard''\footnote{This only gives an accurate estimate of
+  $Z$, which is itself an approximation, and is thus not necessarily
+  the true value of the numerator in Equation \ref{eq:lh-ratio}.} by
+evaluating $S(X_b, X_R)$ at 360 values of $R$ spaced evenly between
+$0$ and $2\pi$.
 

### 2013-05-13 man/final-project.tex
Clean up naive model section

@@
 
-\subsubsection{Linear Interpolation}
+\subsection{Na\"ive model}
 
@@
 
-Na\"ively, we can simply use linear interpolation to estimate $S$ and
-the trapezoidal rule to estimate $Z$.
+As a baseline, we defined a \textit{na\"ive} model which performs a
+hill-climbing search over the similarity function until it reaches a
+(possibly local) maximum. Once a maximum as been found, the model
+computes an estimate of $Z$ by linearly interpolating between sampled
+rotations.
 
-\paragraph{Choosing the sequence of mental images}
+When $h_1$ is correct and as the true angle of rotation approaches 0,
+the \textit{na\"ive} model should increase in both response and
+rotation accuracy, because local maxima will be less likely (and
+therefore, it is more likely that the global maxima will be
+discovered). For larger angles of rotation, this model should have a
+tendency to under-rotate, because it will find local maxima and end
+prematurely. Its estimates of $Z$ will usually be overestimates (if it
+finds the global maxima, and then terminates) or underestimates (the
+rest of the time). This model should almost always be accurate when
+$h_0$ is correct.
 
-In this naïve strategy, we perform a hill-climbing search to find a
-local maxima. In other words, we rotate in the direction for which
-there is a local increase in similarity.
-
-\paragraph{Stopping criterion}
-
-Because we are using hill-climbing, we choose to stop when we hit a
-local maxima.
-
-\subsubsection{Parametric Regression}
+\subsection{Parametric (Von Mises) model}
 

### 2013-05-13 man/final-project.tex
Clean up von mises model section

@@
 Another strategy is to assume a parametric shape for $S$ and fit the
-appropriate parameters. Here, we assume the likelihood follows a
-\emph{Von Mises} distribution:
+appropriate parameters. When $h_1$ is correct, it is likely that the
+function will be approximately unimodal (shapes with rotational
+symmetry would be multimodal). A reasonable assumption, then, is that
+the likelihood follows a ``wrapped Gaussian'', or Von Mises,
+distribution:
 \begin{equation}
-  S(X_b, X_R) \approx \hat{S}_{VM}=\frac{h}{2\pi X_0(\kappa)}e^{\kappa\cos(R-\hat{\theta})}
+  S(X_b, X_R) \approx h\cdot{}p(R\ \vert\ \hat{\theta}, \kappa)=\frac{h}{2\pi X_0(\kappa)}e^{\kappa\cos(R-\hat{\theta})}
 \end{equation}
+where $\kappa$ is the concentration parameter, $\hat{\theta}$ is the
+preferred direction, and $h$ is a scale parameter. We fit these
+parameters by minimizing the mean squared error between this PDF and
+the computed values of $S$. To choose the sequence of rotations, we
+use the same hill-climbing strategy as in the \textit{na\"ive}
+model.\TODO{a better strategy would have been to compute the variance
+  in parameter estimates, or something...} Figure \ref{fig:vm} shows
+an example of this parametric model, prior to estimating $Z$.
 
-Where $\kappa$ is the concentration parameter, $\hat{\theta}$ is the
-preferred direction, and $h$ is a scale parameter. We can fit these
-parameters by minimizing the mean squared error between this PDF and
-the observed values of $S$ which we are given. Then:
-\begin{equation}
-E[Z\ \vert \ S]\approx \int_R \hat{S}_{VM}p(R)\ \mathrm{d}R
-\end{equation}
-
-\paragraph{Choosing the sequence of mental images}
-
-As above, we use hill-climbing, with the stopping criteria being when
-we find a local maxima.
-
-\subsubsection{Bayesian Quadrature}
+\subsection{Nonparametric (Bayesian Quadrature) model}
 
@@
   \includegraphics[width=\textwidth]{../figures/bq_regression.pdf}
-  \caption{Bayesian Quadrature}
+  \caption{Nonparametric model}
   \label{fig:bq}

### 2013-05-13 man/final-project.tex
Update naive/von mises figure caption
Clean up bayesian quadrature section

@@
 
+In each of these models, we denote the computed rotations to be a set
+$\mathbf{R}=\{R_1, R_2, \ldots{}\}$.
+
 \subsection{Gold standard model}
@@
     \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
-    \caption{Linear interpolation}
+    \caption{Na\"ive model}
     \label{fig:li}
@@
     \includegraphics[width=\textwidth]{../figures/vm_regression.pdf}
-    \caption{Parametric regression}
+    \caption{Parametric model}
     \label{fig:vm}
@@
 computes an estimate of $Z$ by linearly interpolating between sampled
-rotations.
+rotations. Figure \ref{fig:li} shows an example of the
+\textit{na\"ive} model prior to estimating $Z$.
 
@@
 $Z$ by placing a Gaussian Process (GP) \cite{Rasmussen:2006vz} prior
-on the log likelihood function $\log S$ and evaluating $\log S$ at a
-particular set of points. For our purposes, we choose these points to
-be a sequence of $N$ consecutive mental rotations be
-$\mathbf{R}=\{R_1, \ldots{}, R_N\}$. To obtain a single estimate of
-$Z$, we want compute the posterior mean:
+on the function $S$ and evaluating $S$ at a particular set of
+points. However, this strategy has its difficulties: while in our
+case, $S$ is a non-negative likelihood function, GP regression
+enforces no such constraint.
+
+\cite{Osborne:2012tm} give a method which more often (but not
+entirely) avoids this problem. Instead of placing a prior over $S$, a
+prior is instead placed over $\log S$, thus ensuring that $S=e^{\log
+  S}$ will be positive\footnote{In practice, as in
+  \cite{Osborne:2012tm}, we use the transform of $\log(S+1)$.}:
+\begin{equation*}
+  E[Z\ \vert \ \log S]=\int_{\log S}\left(\int_R \exp(\log{S(X_b,X_R)})p(R)\ \mathrm{d}R\right)\mathcal{N}\left(\log{S}\ \vert \ \mu_{\log S}, \Sigma_{\log S}\right)\ \mathrm{d}\log S
+\end{equation*}
+where $\mu_{\log S}$ and $\Sigma_{\log S}$ are the mean and
+covariance, respectively, of the GP regression over $\log S$ given
+$\mathbf{R}$. We approximate this according to the method given in
+\cite{Osborne:2012tm}:
 \begin{equation}
-  E[Z\ \vert \ \log S]=\int_{\log S}\left(\int_R \exp(\log{S(X_b,X_R)})p(R)\ \mathrm{d}R\right)\mathcal{N}\left(\log{S}\ \vert \ \mu_{\log S}, \Sigma_{\log S}\right)\ \mathrm{d}\log S
+  \mu_Z=E[Z\ \vert \ \log S] \approx \int_R \mu_{S}(1 + \mu_\Delta) p(R)\ \mathrm{d}R 
+  \label{eq:bq-Z-mean}
 \end{equation}
-As in \cite{Osborne:2012tm}, we can approximate this with:
+where $\mu_S$ is the mean of a GP regression over $S$ given
+$\mathbf{R}$; and $\mu_{\Delta_c}$ is a regression over
+$\Delta_c=\mu_{\log S} - \log \mu_S$ given $\mathbf{R}_c$, which
+consists of $\mathbf{R}$ and a set of intermediate \emph{candidate
+  points}\footnote{These candidate points do not require evaluating
+  the true $S$, only the GP estimates of $S$ and $\log S$.} $c$ as
+described in \cite{Osborne:2012tm}. The variance of the estimate of
+$Z$ is given by:
 \begin{equation}
-  E[Z\ \vert \ \log S] \approx \int_R \mu_{S}(1 + \mu_\Delta) p(R)\ \mathrm{d}R
+  \sigma_Z=\mathrm{Var}(Z\ \vert \ S, \log S) = \int_R\int_{R^\prime} \mu_S(R)\mu_S(R^\prime) \Sigma_{\log S}(R, R^\prime)p(R)p(R^\prime)\ \mathrm{d}R\ \mathrm{d}R^\prime
+  \label{eq:bq-Z-var}
 \end{equation}
-Where $\mu_x$ is the mean of a GP, as follows:
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item $\mu_\Delta$ is a regression over $\Delta=\mu_{\log S} - \log
-  \mu_S$ given $\mathbf{R}_\Delta$, which consists of $\mathbf{R}$ and
-  a set of intermediate \emph{candidate points}, as described in
-  \cite{Osborne:2012tm}.
-\item $\mu_S$ is the mean of a GP regression over $S$ given
-  $\mathbf{R}$.
-\item $\mu_{\log S}$ is the mean of a GP regression over $\log S$
-  given $\mathbf{R}$.
-\end{itemize}
-
-\paragraph{Choosing the sequence of mental images}
-
-Because we can only evaluate $S(X_a, X_R)$ at a small number $k$ of
-rotations, we must make an explicit decision about which
-$\mathbf{R}=\{R_1, \ldots{}, R_k\}$ to use. Ideally, we should choose
-$\mathbf{R}$ such that it will give us the best approximation of
-$p(X_a, X_b\vert h_1)$. One option is to choose each $R_i$ in a greedy
-manner to minimize the variance of $Z$ (i.e.~give a more accurate
-estimate of $S$).
+To choose $\mathbf{R}$, we pick an initial direction of rotation which
+results in the higher value of $S$, and then continue rotating in that
+direction until the variance in Equation \ref{eq:bq-Z-var} is low
+enough that we are confident about the likelihood ratio $\ell$:
+\begin{equation*}
+p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\ \vert\ \mu_Z, \sigma_z)
+\end{equation*}
+Specifically, we choose $h_0$ when $p(\ell < 1)\geq 0.95$, and we
+choose $h_1$ when $p(\ell > 1)\geq 0.95$. Until one of these
+conditions are met (or the shape has been fully rotated), the model
+will continue to compute rotations and update its estimate of $Z$.
 

### 2013-05-13 man/final-project.tex
Add figure TODO
Write some captions and add caption TODOs

@@
   \includegraphics[width=0.95\textwidth]{../figures/shepard-rotation.png}
-  \caption{Classic mental rotation task from \cite{Shepard1971}.}
+  \caption{\textbf{Classic mental rotation task}. Participants in
+    \cite{Shepard1971} saw stimuli such as these, and judged whether
+    each pair of shapes was the same shape in two different
+    orientations, or two different shapes. \textbf{A} shows a ``plane
+    rotation'', \textbf{B} shows a ``depth rotation'', and \textbf{C}
+    shows two distinct objects. \TODO{include Shepard results, too?}}
   \label{fig:mental-rotation}
@@
   \end{subfigure}
-  \caption{Rotated shapes and corresponding similarities.}
+  \caption{\textbf{Rotated shapes and their similarity.}  \textbf{(a)}
+    An example stimulus in which the shapes differ only by a
+    rotation. All stimuli consist of five vertices centered around the
+    origin, and four edges which create a closed loop from the
+    vertices. \textbf{(b)} The approximate likelihood (similarity) of
+    $X_b$ given $X_R$, where $X_R$ is a rotation of $X_a$ by the angle
+    $R$. The true angle of rotation between $X_a$ and $X_b$ is
+    slightly greater than $\pi$, corresponding to the global maximum
+    in the likelihood function.}
 \end{figure}
@@
   \end{subfigure}
+  \caption{\TODO{}}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{../figures/bq_regression.pdf}
-  \caption{Nonparametric model}
+  \caption{\textbf{Nonparametric model.} \TODO{}}
   \label{fig:bq}
@@
   \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
-  \caption{Model rotations}
+  \caption{\textbf{Model rotations.} \TODO{}}
   \label{fig:rotations}
@@
   \includegraphics[width=\textwidth]{../figures/Z_accuracy.pdf}
-  \caption{Accuracy in estimating $Z$.}
+  \caption{\textbf{Accuracy in estimating $Z$.} \TODO{} \TODO{use
+      different markers for $h_0$ and $h_1$ points}}
   \label{fig:accuracy}

### 2013-05-13 man/final-project.tex
Computational model section edits

@@
 People are presented with two images, $X_a$ and $X_b$, which are the
-coordinates of the vertices of two shapes. Participants must determine
-whether $X_a$ and $X_b$ were generated from the same (albeit possibly
-transformed and permuted) original shape, i.e., whether $\exists
-R,M\textrm{ s.t. } X_b=MRX_a$, where $M$ is a permutation matrix and
-$R$ is a rotation matrix. \TODO{Figure with some example shapes}
+coordinates of the vertices of two shapes (e.g., see Figure
+\ref{fig:stimuli}). Participants must determine whether $X_a$ and
+$X_b$ were generated from the same (albeit possibly transformed and
+permuted) original shape, i.e., whether $\exists R,M\textrm{ s.t. }
+X_b=MRX_a$, where $M$ is a permutation matrix and $R$ is a rotation
+matrix.
 
@@
 
-\subsection{Gold standard model}
+\subsection{Gold standard}
 
@@
 
-As a baseline, we defined a \textit{na\"ive} model which performs a
+As a baseline, we defined a na\"ive model which performs a
 hill-climbing search over the similarity function until it reaches a
@@
 computes an estimate of $Z$ by linearly interpolating between sampled
-rotations. Figure \ref{fig:li} shows an example of the
-\textit{na\"ive} model prior to estimating $Z$.
-
-When $h_1$ is correct and as the true angle of rotation approaches 0,
-the \textit{na\"ive} model should increase in both response and
-rotation accuracy, because local maxima will be less likely (and
-therefore, it is more likely that the global maxima will be
-discovered). For larger angles of rotation, this model should have a
-tendency to under-rotate, because it will find local maxima and end
-prematurely. Its estimates of $Z$ will usually be overestimates (if it
-finds the global maxima, and then terminates) or underestimates (the
-rest of the time). This model should almost always be accurate when
-$h_0$ is correct.
+rotations. Figure \ref{fig:li} shows an example of the na\"ive model
+prior to estimating $Z$.
 
@@
 the computed values of $S$. To choose the sequence of rotations, we
-use the same hill-climbing strategy as in the \textit{na\"ive}
-model.\TODO{a better strategy would have been to compute the variance
-  in parameter estimates, or something...} Figure \ref{fig:vm} shows
-an example of this parametric model, prior to estimating $Z$.
+use the same hill-climbing strategy as in the na\"ive model.\TODO{a
+  better strategy would have been to compute the variance in parameter
+  estimates, or something...} Figure \ref{fig:vm} shows an example of
+this parametric model, prior to estimating $Z$.
 
@@
 \begin{equation}
-  \mu_Z=E[Z\ \vert \ \log S] \approx \int_R \mu_{S}(1 + \mu_\Delta) p(R)\ \mathrm{d}R 
+  \mu_Z=E[Z\ \vert \ S, \log S, \Delta_c] \approx \int_R \mu_{S}(1 + \mu_\Delta) p(R)\ \mathrm{d}R 
   \label{eq:bq-Z-mean}
@@
 \begin{equation}
-  \sigma_Z=\mathrm{Var}(Z\ \vert \ S, \log S) = \int_R\int_{R^\prime} \mu_S(R)\mu_S(R^\prime) \Sigma_{\log S}(R, R^\prime)p(R)p(R^\prime)\ \mathrm{d}R\ \mathrm{d}R^\prime
+  \sigma_Z=\mathrm{Var}(Z\ \vert \ S, \log S, \Delta_c) = \int_R\int_{R^\prime} \mu_S(R)\mu_S(R^\prime) \Sigma_{\log S}(R, R^\prime)p(R)p(R^\prime)\ \mathrm{d}R\ \mathrm{d}R^\prime
   \label{eq:bq-Z-var}

### 2013-05-13 man/final-project.tex
Results writeup for everything except bayesian quadrature model

@@
 
+We evaluated each model's performance on 50 randomly generated shapes
+(such as $X_a$ in Figure \ref{fig:stimuli}). For each shape, we picked
+a random angle $R$ and generated two stimuli: a $h_0$ stimulus, by
+reflecting $X_a$ across the $y$-axis and rotating it by $R$; and a
+$h_1$ stimulus, by simply rotating $X_a$ by $R$.
+
+We considered three metrics of performance in particular:
+\begin{itemize}
+\item \textit{Rotations}: for stimuli where $h_1$ is correct, how
+  correlated are the true angles of rotation with the rotations
+  performed by the model? This is defined to be the Pearson's
+  correlation coefficient $\rho$ between $R$ and $\vert
+  \mathbf{R}\vert$. Figure \ref{fig:rotations} shows individual points
+  corresponding to true rotation vs. rotation by the model, for each
+  model.
+\item \textit{Responses}: how accurate is the model at choosing the
+  correct hypothesis? This is defined to be the mean error
+  ($\mathrm{ME}$), or fraction of times the model picks the incorrect
+  hypothesis.
+\item \textit{Estimates of $Z$}: how accurate is the model's estimate
+  of $Z$? This is defined to be the mean squared error
+  ($\mathrm{MSE}$) between the model's estimate of $Z$ and the ``gold
+  standard'' value, where the error has scale such that
+  $\mathrm{MSE}=0$ indicates no error, and $\mathrm{MSE}=1$ indicates
+  maximum error. Figure \ref{fig:accuracy} shows plots of the true
+  (gold standard) value of $Z$ vs. the model's estimate, for each
+  model.
+\end{itemize}
+
+\paragraph{Gold standard}
+
+The gold standard always evaluates $S$ at all angles, so the
+correlation between it and the true angle of rotation is
+undefined. The mean error of responses was $\mathrm{ME}=0.01$,
+indicating that $Z$ is a sufficient approximation\footnote{The one
+  error comes from a $h_0$ stimulus which is highly
+  symmetric. Although it is distinct enough for the author to
+  correctly judge when both images are viewed side by side, it remains
+  to be seen whether human participants, viewing the images
+  sequentially would also make the correct judgment.}  of $\int_R
+p(X_b\ \vert\ X_R)p(R)\ \mathrm{d}R$.
+
+\paragraph{Na\"ive model}
+
+The correlation between the number of rotations computed by the
+na\"ive model and the true angle of rotation was $\rho=-0.11$. The
+shape of the data is more interesting (see Figure \ref{fig:rotations},
+right panel): the na\"ive model actually corresponds quite well to the
+true angle of rotation for $R<\frac{\pi}{2}$ ($\rho=0.63$). This is
+unsurprising, because the closer the true angle is to zero, the less
+the model has to rotate, and the less likely it will get stuck on
+local maxima. Thus, it is more likely to locat the global maximum,
+which generally corresponds to the true angle of rotation. For
+$R>\frac{\pi}{2}$, we see an increasing tendency to under-rotate
+($\rho=-0.42$); this is likely because it finds a local maxima and
+ends prematurely.
+
+The na\"ive model's response error rate was better than chance, but
+still quite high, with $\mathrm{ME}=0.23$. Closer inspection reveals
+that the bulk of this comes from $h_1$ stimuli ($\mathrm{ME}=0.36$
+vs. $\mathrm{ME}=0.1$ for $h_0$ stimuli). This is, as above, probably
+related to finding only local maxima: if the model finds a local
+maxima which is low enough, the area under the estimated curve will
+not be large enough to accept $h_1$.
+
+This is reflected in the accuracy in estimating $Z$ as well
+($\mathrm{MSE}=0.14$). Figure \ref{fig:accuracy} (leftmost panel)
+shows individual estimates for each stimulus, color-coded by true
+hypothesis. The model underestimates $Z$ for significant portion of
+the $h_1$ stimuli. On the other $h_1$ stimuli, the na\"ive model
+actually tends to overestimate because linear interpolation does not
+account for symmetry of global maxima (\TODO{figure?}).
+
+\paragraph{Parametric (Von Mises) model}
+
+Like the na\"ive model, the parametric model uses hill-climbing as a
+sampling strategy, so they have identical correlations with the true
+angle of rotation. However, given the same $\mathbf{R}$, the
+parametric model estimates $Z$ differently, thus giving rise to
+different response and estimated $Z$ error rates. The response error
+for the parametric model was higher than the na\"ive model, with
+$\mathrm{ME}=0.26$. Breaking this down, we see that the error for
+$h_1$ stimuli is $\mathrm{ME}=0.5$: in other words, the parametric
+model is at chance when determining whether two identical shapes are
+the same. This is because it never overestimates (see Figure
+\ref{fig:accuracy}, center panel). So, it accurately estimates $Z$ for
+stimuli which the na\"ive model would overestimate
+($\mathrm{MSE}=0.07$), but still has poor performance when $Z$ is
+underestimated. Indeed, if we exclude stimuli for which the nai\"ive
+model overestimates ($Z\geq 3$), the error for the na\"ive model
+lowers to match that of the parametric model.
+
+\paragraph{Nonparametric (Bayesian Quadrature) model}
+
 \begin{figure}[t]

### 2013-05-13 man/final-project.tex
Small caption edits

@@
     $R$. The true angle of rotation between $X_a$ and $X_b$ is
-    slightly greater than $\pi$, corresponding to the global maximum
+    approximately $\frac{\pi}{4}$, corresponding to the global maximum
     in the likelihood function.}
@@
   \includegraphics[width=\textwidth]{../figures/Z_accuracy.pdf}
-  \caption{\textbf{Accuracy in estimating $Z$.} \TODO{} \TODO{use
-      different markers for $h_0$ and $h_1$ points}}
+  \caption{\textbf{Accuracy in estimating $Z$.} \TODO{}}
   \label{fig:accuracy}

### 2013-05-13 man/final-project.tex
Move stimuli/likelihood function figure
Fix Bessel function in Von Mises PDF

@@
 
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/stimuli_shapes.pdf}
-    \vspace{0pt}
-    \caption{Example stimuli}
-    \label{fig:stimuli}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/likelihood_function.pdf}
-    \caption{Likelihood function}
-    \label{fig:likelihood}
-  \end{subfigure}
-  \caption{\textbf{Rotated shapes and their similarity.}  \textbf{(a)}
-    An example stimulus in which the shapes differ only by a
-    rotation. All stimuli consist of five vertices centered around the
-    origin, and four edges which create a closed loop from the
-    vertices. \textbf{(b)} The approximate likelihood (similarity) of
-    $X_b$ given $X_R$, where $X_R$ is a rotation of $X_a$ by the angle
-    $R$. The true angle of rotation between $X_a$ and $X_b$ is
-    approximately $\frac{\pi}{4}$, corresponding to the global maximum
-    in the likelihood function.}
-\end{figure}
-
 People are presented with two images, $X_a$ and $X_b$, which are the
@@
 \section{Implementation}
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/stimuli_shapes.pdf}
+    \vspace{0pt}
+    \caption{Example stimuli}
+    \label{fig:stimuli}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/likelihood_function.pdf}
+    \caption{Likelihood function}
+    \label{fig:likelihood}
+  \end{subfigure}
+  \caption{\textbf{Rotated shapes and their similarity.}  \textbf{(a)}
+    An example stimulus in which the shapes differ only by a
+    rotation. All stimuli consist of five vertices centered around the
+    origin, and four edges which create a closed loop from the
+    vertices. \textbf{(b)} The approximate likelihood (similarity) of
+    $X_b$ given $X_R$, where $X_R$ is a rotation of $X_a$ by the angle
+    $R$. The true angle of rotation between $X_a$ and $X_b$ is
+    approximately $\frac{\pi}{4}$, corresponding to the global maximum
+    in the likelihood function.}
+\end{figure}
 
@@
 \begin{equation}
-  S(X_b, X_R) \approx h\cdot{}p(R\ \vert\ \hat{\theta}, \kappa)=\frac{h}{2\pi X_0(\kappa)}e^{\kappa\cos(R-\hat{\theta})}
+  S(X_b, X_R) \approx h\cdot{}p(R\ \vert\ \hat{\theta}, \kappa)=\frac{h}{2\pi I_0(\kappa)}e^{\kappa\cos(R-\hat{\theta})}
 \end{equation}
 where $\kappa$ is the concentration parameter, $\hat{\theta}$ is the
-preferred direction, and $h$ is a scale parameter. We fit these
-parameters by minimizing the mean squared error between this PDF and
-the computed values of $S$. To choose the sequence of rotations, we
-use the same hill-climbing strategy as in the na\"ive model.\TODO{a
-  better strategy would have been to compute the variance in parameter
-  estimates, or something...} Figure \ref{fig:vm} shows an example of
-this parametric model, prior to estimating $Z$.
+preferred direction, $h$ is a scale parameter, and $I_0$ is the
+modified Bessel function of order zero. We fit these parameters by
+minimizing the mean squared error between this PDF and the computed
+values of $S$. To choose the sequence of rotations, we use the same
+hill-climbing strategy as in the na\"ive model.\TODO{a better strategy
+  would have been to compute the variance in parameter estimates, or
+  something...} Figure \ref{fig:vm} shows an example of this
+parametric model, prior to estimating $Z$.
 

### 2013-05-13 man/final-project.tex
Add equation for kernel function

@@
 A more flexible strategy uses what is known as \emph{Bayesian
-  quadrature} \cite{Diaconis:1988uo,OHagan:1991tx} to estimate $Z$.
-Bayesian quadrature allows us to compute a posterior distrbution over
-$Z$ by placing a Gaussian Process (GP) \cite{Rasmussen:2006vz} prior
-on the function $S$ and evaluating $S$ at a particular set of
-points. However, this strategy has its difficulties: while in our
-case, $S$ is a non-negative likelihood function, GP regression
-enforces no such constraint.
+  Quadrature} \cite{Diaconis:1988uo,OHagan:1991tx} to estimate $Z$.
+Bayesian Quadrature allows us to compute a posterior distrbution over
+$Z$ by placing a Gaussian Process (GP) prior on the function $S$ and
+evaluating $S$ at a particular set of points. Because our data is
+circular, we can use a periodic kernel \cite{Rasmussen:2006vz}:
+\begin{equation}
+k(R, R^\prime)=h^2\exp\left(-\frac{2\sin^2\left(\frac{1}{2}(R-R^\prime)\right)}{w^2}\right)
+\end{equation}
 
-\cite{Osborne:2012tm} give a method which more often (but not
-entirely) avoids this problem. Instead of placing a prior over $S$, a
-prior is instead placed over $\log S$, thus ensuring that $S=e^{\log
-  S}$ will be positive\footnote{In practice, as in
-  \cite{Osborne:2012tm}, we use the transform of $\log(S+1)$.}:
+Bayesian Quadrature has its difficulties, however. While in our case
+$S$ is a non-negative likelihood function, GP regression enforces no
+such constraint. In an effort to avoid this problem,
+\cite{Osborne:2012tm} give a method which involves instead placing a
+prior over the log liklihood\footnote{In practice, as in
+  \cite{Osborne:2012tm}, we use the transform of $\log(S+1)$.}, thus
+ensuring that $S=e^{\log S}$ will be positive\footnote{We are not
+  guaranteed positivity, however, the approximation to the integral
+  over $\log S$ (Equation \ref{eq:bq-Z-mean}) requires computing
+  $\mu_S$, which may have nonpositive segments.}:
 \begin{equation*}

### 2013-05-13 man/final-project.tex
Add a bit about BQ results (not done)

@@
 
+The rotations computed by the nonparametric model are the most
+strongly correlated with the true rotations, albeit negatively so
+($\rho=-0.60$). In general, the nonparametric model rotates much more
+than should be necessary; this is because the Bayesian quadrature
+estimate of $Z$ is too large to accept either hypothesis until a
+significant portion of the rotations have been sampled.
+
 \begin{figure}[t]

### 2013-05-13 code/plots.py
BQ plot formatting

@@ -90,7 +90,7 @@ def regression(x, y, xi, yi, xo, yo_mean, yo_var, **kwargs):
     sg.outward_ticks()
     sg.clear_right()
     sg.clear_top()
-    sg.set_scientific(-3, 4, axis='y')
+    sg.set_scientific(-2, 3, axis='y')
 
     # title and axis labels
     if opt['title']:
@@ -156,7 +156,10 @@ def bq_regression(model):
     sg.align_ylabels(-0.12, ax_S, ax_logS)
     sg.set_ylabel_coords(-0.16, ax=ax_Dc)
     # sync y-axis limits
-    sg.sync_ylims(ax_S, ax_final)
+    lim = (-0.2, 1.1)
+    ax_S.set_ylim(*lim)
+    ax_logS.set_ylim(*lim)
+    ax_final.set_ylim(*lim)
 
     # overall figure settings
     sg.set_figsize(9, 5)

### 2013-05-13 code/model_bq.py
Change input scale parameter for BQ

@@ -48,15 +48,16 @@ class BayesianQuadratureModel(Model):
         self._ilast = None
 
         # marginal log likelihood objects
+        h = self.opt['scale'] / 8.
         self._mll_S = KernelMLL(
             kernel=self.opt['kernel'],
-            h=self.opt['scale'],
+            h=h,
             w=None,
             s=self.opt['s']
         )
         self._mll_logS = KernelMLL(
             kernel=self.opt['kernel'],
-            h=np.log(self.opt['scale']+1),
+            h=np.log(h + 1),
             w=None,
             s=self.opt['s']
         )

### 2013-05-13 code/model_bq.py
Go back to original way of computing S_mean

@@ -205,15 +205,7 @@ class BayesianQuadratureModel(Model):
             self.theta_Dc = None
 
         # the final regression for S
-        #
-        # According to the Osborne paper, it should be this:
-        #   self.S_mean = ((self.mu_S + 1) * (1 + self.mu_Dc)) - 1
-        #
-        # But then if self.mu_S < 0, self.S_mean will also have
-        # negative parts. To get around this, I am replacing
-        # (self.mu_S + 1) with exp(self.mu_logS - self.mu_Dc)
-        mu_S_plus_one = np.exp(self.mu_logS - self.mu_Dc)
-        self.S_mean = (mu_S_plus_one * (1 + self.mu_Dc)) - 1
+        self.S_mean = ((self.mu_S + 1) * (1 + self.mu_Dc)) - 1
         self.S_var = np.diag(self.cov_logS)
 
     def integrate(self):

### 2013-05-13 man/final-project.tex
Caption for simple models figure

@@
   \end{subfigure}
-  \caption{\TODO{}}
+  \caption{\textbf{Simple methods of estimating $S$.} In both cases,
+    hill-climbing search is used until a maxima is found (in this
+    case, at approximately $\frac{\pi}{6}$). The sampled points
+    $\mathbf{R}$ (red circles) are then used to estimate $S$ (black
+    lines are the true $S$, red lines are the estimate). \textbf{(a)}
+    Linear interpolation. The overall estimate of $Z$ here will be too
+    low, but certain angles will have a disproportionate contribution
+    (e.g., between $\frac{\pi}{4}$ and $\frac{3\pi}{4}$. \textbf{(b)}
+    Best fit of scaled Von Mises PDF parameters. As in (a), $Z$ will
+    be underestimated. However, the fit around the maximum near
+    $\frac{\pi}{6}$ is much more accurate.}
+  \label{fig:simple-models}
 \end{figure}

### 2013-05-13 man/final-project.tex
Caption for BQ regression figure

@@
   \includegraphics[width=\textwidth]{../figures/bq_regression.pdf}
-  \caption{\textbf{Nonparametric model.} \TODO{}}
+  \caption{\textbf{Nonparametric model.} Each panel shows one step of
+    the Bayesian Quadrature regression. Upper left: the original
+    Gaussian Process (GP) regression for $S$. Lower left: the GP
+    regression for $\log(S+1)$. Lower right: the GP regression for
+    $\Delta_c=\mu_{\log S} - \log \mu_S$. Upper right: the adjusted
+    regression for $S$, where the mean is equal to
+    $\mu_S(1+\mu_{\Delta_c})$. The model uses this final estimate to
+    compute $Z$ and will continue rotating until the variance of $Z$
+    is low enough that a hypothesis may be accepted. This method
+    allows the model to avoid local maxima such as the one near
+    $\frac{\pi}{6}$, which causes trouble for the na\"ive and
+    parametric models in Figure \ref{fig:simple-models}.}
   \label{fig:bq}

### 2013-05-13 man/final-project.tex
Resolve some TODOs, add some more

@@
     rotation'', \textbf{B} shows a ``depth rotation'', and \textbf{C}
-    shows two distinct objects. \TODO{include Shepard results, too?}}
+    shows two distinct objects.}
   \label{fig:mental-rotation}
@@
 values of $S$. To choose the sequence of rotations, we use the same
-hill-climbing strategy as in the na\"ive model.\TODO{a better strategy
-  would have been to compute the variance in parameter estimates, or
-  something...} Figure \ref{fig:vm} shows an example of this
-parametric model, prior to estimating $Z$.
+hill-climbing strategy as in the na\"ive model. Figure \ref{fig:vm}
+shows an example of this parametric model, prior to estimating $Z$.
 
@@
 actually tends to overestimate because linear interpolation does not
-account for symmetry of global maxima (\TODO{figure?}).
+account for symmetry of global maxima (e.g., Figure \ref{fig:li}).
 
@@
 
+\TODO{for von mises, a better strategy would have been to compute the
+  variance in parameter estimates, or something...}
+
+\TODO{talk about marginalising hyperparameters}
+
+\TODO{generalize to 3d case, more interesting -- can actually do
+  active sampling, whereas it's sort limited in the 2d case}
+
 % \renewcommand\bibsection{\subsubsection*{\refname}}

### 2013-05-13 man/final-project.tex
Fix some typos

@@
 \begin{equation}
-  \mu_Z=E[Z\ \vert \ S, \log S, \Delta_c] \approx \int_R \mu_{S}(1 + \mu_\Delta) p(R)\ \mathrm{d}R 
+  \mu_Z=E[Z\ \vert \ S, \log S, \Delta_c] \approx \int_R \mu_{S}(1 + \mu_{\Delta_c}) p(R)\ \mathrm{d}R 
   \label{eq:bq-Z-mean}
@@
 ($\mathrm{MSE}=0.07$), but still has poor performance when $Z$ is
-underestimated. Indeed, if we exclude stimuli for which the nai\"ive
+underestimated. Indeed, if we exclude stimuli for which the na\"ive
 model overestimates ($Z\geq 3$), the error for the na\"ive model

### 2013-05-13 man/final-project.tex
Fix Figure 2 caption
Update result numbers

@@
     $R$. The true angle of rotation between $X_a$ and $X_b$ is
-    approximately $\frac{\pi}{4}$, corresponding to the global maximum
-    in the likelihood function.}
+    slightly less than $\pi$, corresponding to the global maximum in
+    the likelihood function.}
 \end{figure}
@@
 right panel): the na\"ive model actually corresponds quite well to the
-true angle of rotation for $R<\frac{\pi}{2}$ ($\rho=0.63$). This is
+true angle of rotation for $R<\frac{\pi}{2}$ ($\rho=0.64$). This is
 unsurprising, because the closer the true angle is to zero, the less
@@
 
-The rotations computed by the nonparametric model are the most
-strongly correlated with the true rotations, albeit negatively so
-($\rho=-0.60$). In general, the nonparametric model rotates much more
-than should be necessary; this is because the Bayesian quadrature
-estimate of $Z$ is too large to accept either hypothesis until a
-significant portion of the rotations have been sampled.
+The rotations computed by the nonparametric model are strongly
+correlated with the true rotations ($\rho=0.73$). In general, the
+nonparametric model rotates much more than should be necessary; this
+is because the Bayesian quadrature estimate of $Z$ is too large to
+accept either hypothesis until a significant portion of the rotations
+have been sampled.
 

### 2013-05-13 man/final-project.tex
Define macros for ME, MSE, naive, and Naive

@@
 \newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}
+\newcommand{\MSE}[0]{\mathrm{MSE}}
+\newcommand{\ME}[0]{\mathrm{ME}}
+\newcommand{\naive}[0]{na\"ive}
+\newcommand{\Naive}[0]{Na\"ive}
 
@@
 
-\subsection{Na\"ive model}
+\subsection{\Naive{} model}
 
@@
     \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
-    \caption{Na\"ive model}
+    \caption{\Naive{} model}
     \label{fig:li}
@@
 
-As a baseline, we defined a na\"ive model which performs a
+As a baseline, we defined a \naive{} model which performs a
 hill-climbing search over the similarity function until it reaches a
@@
 computes an estimate of $Z$ by linearly interpolating between sampled
-rotations. Figure \ref{fig:li} shows an example of the na\"ive model
+rotations. Figure \ref{fig:li} shows an example of the \naive{} model
 prior to estimating $Z$.
@@
 values of $S$. To choose the sequence of rotations, we use the same
-hill-climbing strategy as in the na\"ive model. Figure \ref{fig:vm}
+hill-climbing strategy as in the \naive{} model. Figure \ref{fig:vm}
 shows an example of this parametric model, prior to estimating $Z$.
@@
     allows the model to avoid local maxima such as the one near
-    $\frac{\pi}{6}$, which causes trouble for the na\"ive and
+    $\frac{\pi}{6}$, which causes trouble for the \naive{} and
     parametric models in Figure \ref{fig:simple-models}.}
@@
 \cite{Osborne:2012tm} give a method which involves instead placing a
-prior over the log liklihood\footnote{In practice, as in
+prior over the log likelihood\footnote{In practice, as in
   \cite{Osborne:2012tm}, we use the transform of $\log(S+1)$.}, thus
@@
 \item \textit{Responses}: how accurate is the model at choosing the
-  correct hypothesis? This is defined to be the mean error
-  ($\mathrm{ME}$), or fraction of times the model picks the incorrect
-  hypothesis.
+  correct hypothesis? This is defined to be the mean error ($\ME{}$),
+  or fraction of times the model picks the incorrect hypothesis.
 \item \textit{Estimates of $Z$}: how accurate is the model's estimate
-  of $Z$? This is defined to be the mean squared error
-  ($\mathrm{MSE}$) between the model's estimate of $Z$ and the ``gold
-  standard'' value, where the error has scale such that
-  $\mathrm{MSE}=0$ indicates no error, and $\mathrm{MSE}=1$ indicates
-  maximum error. Figure \ref{fig:accuracy} shows plots of the true
-  (gold standard) value of $Z$ vs. the model's estimate, for each
-  model.
+  of $Z$? This is defined to be the mean squared error ($\MSE{}$)
+  between the model's estimate of $Z$ and the ``gold standard'' value,
+  where the error has scale such that $\MSE{}=0$ indicates no error,
+  and $\MSE{}=1$ indicates maximum error. Figure \ref{fig:accuracy}
+  shows plots of the true (gold standard) value of $Z$ vs. the model's
+  estimate, for each model.
 \end{itemize}
@@
 correlation between it and the true angle of rotation is
-undefined. The mean error of responses was $\mathrm{ME}=0.01$,
-indicating that $Z$ is a sufficient approximation\footnote{The one
-  error comes from a $h_0$ stimulus which is highly
-  symmetric. Although it is distinct enough for the author to
-  correctly judge when both images are viewed side by side, it remains
-  to be seen whether human participants, viewing the images
-  sequentially would also make the correct judgment.}  of $\int_R
-p(X_b\ \vert\ X_R)p(R)\ \mathrm{d}R$.
+undefined. The mean error of responses was $\ME{}=0.01$, indicating
+that $Z$ is a sufficient approximation\footnote{The one error comes
+  from a $h_0$ stimulus which is highly symmetric. Although it is
+  distinct enough for the author to correctly judge when both images
+  are viewed side by side, it remains to be seen whether human
+  participants, viewing the images sequentially would also make the
+  correct judgment.}  of $\int_R p(X_b\ \vert\ X_R)p(R)\ \mathrm{d}R$.
 
-\paragraph{Na\"ive model}
+\paragraph{\Naive{} model}
 
 The correlation between the number of rotations computed by the
-na\"ive model and the true angle of rotation was $\rho=-0.11$. The
+\naive{} model and the true angle of rotation was $\rho=-0.11$. The
 shape of the data is more interesting (see Figure \ref{fig:rotations},
-right panel): the na\"ive model actually corresponds quite well to the
-true angle of rotation for $R<\frac{\pi}{2}$ ($\rho=0.64$). This is
-unsurprising, because the closer the true angle is to zero, the less
-the model has to rotate, and the less likely it will get stuck on
+right panel): the \naive{} model actually corresponds quite well to
+the true angle of rotation for $R<\frac{\pi}{2}$ ($\rho=0.64$). This
+is unsurprising, because the closer the true angle is to zero, the
+less the model has to rotate, and the less likely it will get stuck on
 local maxima. Thus, it is more likely to locat the global maximum,
@@
 
-The na\"ive model's response error rate was better than chance, but
-still quite high, with $\mathrm{ME}=0.23$. Closer inspection reveals
-that the bulk of this comes from $h_1$ stimuli ($\mathrm{ME}=0.36$
-vs. $\mathrm{ME}=0.1$ for $h_0$ stimuli). This is, as above, probably
+The \naive{} model's response error rate was better than chance, but
+still quite high, with $\ME{}=0.23$. Closer inspection reveals that
+the bulk of this comes from $h_1$ stimuli ($\ME{}=0.36$
+vs. $\ME{}=0.1$ for $h_0$ stimuli). This is, as above, probably
 related to finding only local maxima: if the model finds a local
@@
 This is reflected in the accuracy in estimating $Z$ as well
-($\mathrm{MSE}=0.14$). Figure \ref{fig:accuracy} (leftmost panel)
-shows individual estimates for each stimulus, color-coded by true
+($\MSE{}=0.14$). Figure \ref{fig:accuracy} (leftmost panel) shows
+individual estimates for each stimulus, color-coded by true
 hypothesis. The model underestimates $Z$ for significant portion of
-the $h_1$ stimuli. On the other $h_1$ stimuli, the na\"ive model
+the $h_1$ stimuli. On the other $h_1$ stimuli, the \naive{} model
 actually tends to overestimate because linear interpolation does not
@@
 
-Like the na\"ive model, the parametric model uses hill-climbing as a
+Like the \naive{} model, the parametric model uses hill-climbing as a
 sampling strategy, so they have identical correlations with the true
@@
 different response and estimated $Z$ error rates. The response error
-for the parametric model was higher than the na\"ive model, with
-$\mathrm{ME}=0.26$. Breaking this down, we see that the error for
-$h_1$ stimuli is $\mathrm{ME}=0.5$: in other words, the parametric
-model is at chance when determining whether two identical shapes are
-the same. This is because it never overestimates (see Figure
+for the parametric model was higher than the \naive{} model, with
+$\ME{}=0.26$. Breaking this down, we see that the error for $h_1$
+stimuli is $\ME{}=0.5$: in other words, the parametric model is at
+chance when determining whether two identical shapes are the
+same. This is because it never overestimates (see Figure
 \ref{fig:accuracy}, center panel). So, it accurately estimates $Z$ for
-stimuli which the na\"ive model would overestimate
-($\mathrm{MSE}=0.07$), but still has poor performance when $Z$ is
-underestimated. Indeed, if we exclude stimuli for which the na\"ive
-model overestimates ($Z\geq 3$), the error for the na\"ive model
-lowers to match that of the parametric model.
+stimuli which the \naive{} model would overestimate ($\MSE{}=0.07$),
+but still has poor performance when $Z$ is underestimated. Indeed, if
+we exclude stimuli for which the \naive{} model overestimates ($Z\geq
+3$), the error for the \naive{} model lowers to match that of the
+parametric model.
 

### 2013-05-13 man/final-project.tex
Move rotations figure, and add results for BQ model

@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
+  \caption{\textbf{Model rotations.} \TODO{}}
+  \label{fig:rotations}
+\end{figure}
+
 We evaluated each model's performance on 50 randomly generated shapes
@@
 
-The rotations computed by the nonparametric model are strongly
-correlated with the true rotations ($\rho=0.73$). In general, the
-nonparametric model rotates much more than should be necessary; this
-is because the Bayesian quadrature estimate of $Z$ is too large to
-accept either hypothesis until a significant portion of the rotations
-have been sampled.
+The rotations computed by the nonparametric model were strongly
+correlated with the true rotations ($\rho=0.73$, see Figure
+\ref{fig:rotations}, right panel). This is because the nonparametric
+model does not get stuck as easily on local optima: it will continue
+rotating until it is confident that its estimate of $Z$ is accurate.
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
-  \caption{\textbf{Model rotations.} \TODO{}}
-  \label{fig:rotations}
-\end{figure}
+Correspondingly, the nonparametric model is much more accurate in
+choosing the correct hypothesis ($\ME{}=0.06$). As with the \naive{}
+and parametric models, it mostly errs on $h_1$ stimuli
+($\ME{}=0.1$). This appears to mostly be the result of choosing the
+incorrect initial direction of rotation.
+
+Similarly, because the nonparametric model actively attemps to obtain
+an accurate estimate of $Z$, its estimate of $Z$ is indeed fairly
+accurate ($\MSE{}=0.10$). This is slightly higher than the
+nonparametric model, but is due to a few outliers. If we again exclude
+stimuli for which $Z$ is greatly overestimated ($Z\geq 0.3$), the
+nonparametric model's error decreases significantly ($\MSE{}=0.03$).
 

### 2013-05-13 man/final-project.tex
Add footnote about fitting parameters

@@
 prior over the log likelihood\footnote{In practice, as in
-  \cite{Osborne:2012tm}, we use the transform of $\log(S+1)$.}, thus
-ensuring that $S=e^{\log S}$ will be positive\footnote{We are not
+  \cite{Osborne:2012tm}, we use the transform of
+  $\log(S+1)$. Additionally, while \cite{Osborne:2012tm} use a
+  combination of MLII and marginalization to fit the kernel
+  parameters, we set the output scale $h_S=0.125$ and $h_{\log
+    S}=\log(h_S + 1)$, and use MLII to fit all other parameters.},
+thus ensuring that $S=e^{\log S}$ will be positive\footnote{We are not
   guaranteed positivity, however, the approximation to the integral

### 2013-05-13 man/final-project.tex
Minor miscellaneous changes

@@
 shape of the data is more interesting (see Figure \ref{fig:rotations},
-right panel): the \naive{} model actually corresponds quite well to
+left panel): the \naive{} model actually corresponds quite well to
 the true angle of rotation for $R<\frac{\pi}{2}$ ($\rho=0.64$). This
@@
 This is reflected in the accuracy in estimating $Z$ as well
-($\MSE{}=0.14$). Figure \ref{fig:accuracy} (leftmost panel) shows
+($\MSE{}=0.14$). Figure \ref{fig:accuracy} (left panel) shows
 individual estimates for each stimulus, color-coded by true
@@
 we exclude stimuli for which the \naive{} model overestimates ($Z\geq
-3$), the error for the \naive{} model lowers to match that of the
+0.3$), the error for the \naive{} model lowers to match that of the
 parametric model.

### 2013-05-13 man/final-project.tex
Begin work on discussion

@@
 
-\TODO{for von mises, a better strategy would have been to compute the
-  variance in parameter estimates, or something...}
+We investigated strategies for performing the mental rotation task
+\cite{Shepard1971} in two dimensions, and found that a nonparametric
+model that enforces a positivity constraint on the likelihood function
+performs best, as it is able to actively monitor the confidence of its
+estimate. The simpler models performed much worse: they did not
+maintain the linear relationship with the true angle of rotation due
+to lack of robustness against local maxima, and they were inaccurate
+at identifying whether the two shapes were the same or not.
 
-\TODO{talk about marginalising hyperparameters}
+One option for improving the performance of the simple models would be
+to fit a threshold value, below which local maxima would be
+ignored. However, this strategy would be rather brittle, for if the
+distribution of shapes changed, the threshold would have to be
+re-learned. It is possible that people exhibit this behavior, but we
+cannot make any assumptions one way or the other without empirical
+data.
 
-\TODO{generalize to 3d case, more interesting -- can actually do
-  active sampling, whereas it's sort limited in the 2d case}
+A further strength of the nonparametric model over the other models,
+however, is that it is likely to generalize well to three
+dimensions. One aspect of \cite{Osborne:2012tm} which we have not yet
+explored is their main contribution of \textit{active sampling}, in
+which samples are iteratively selected to maximally decrease the
+expected variance of $Z$. Due to the sequential constraint of mental
+rotation, this strategy is not particularly useful in two dimensions,
+as there are only every two directions in which to rotate. In three
+dimensions, however, there are an infinite number of directions that
+could be chosen after every step.
 

### 2013-05-13 man/final-project.tex
Add empty acknowledgments section
Remove footnote

@@
 undefined. The mean error of responses was $\ME{}=0.01$, indicating
-that $Z$ is a sufficient approximation\footnote{The one error comes
-  from a $h_0$ stimulus which is highly symmetric. Although it is
-  distinct enough for the author to correctly judge when both images
-  are viewed side by side, it remains to be seen whether human
-  participants, viewing the images sequentially would also make the
-  correct judgment.}  of $\int_R p(X_b\ \vert\ X_R)p(R)\ \mathrm{d}R$.
+that $Z$ is a sufficient approximation of $\int_R p(X_b\ \vert\
+X_R)p(R)\ \mathrm{d}R$.
 
@@
 
+\textbf{Acknowledgements}
+
+\TODO{}
+
 % \renewcommand\bibsection{\subsubsection*{\refname}}

### 2013-05-13 man/final-project.tex
Add final paragraph to discussion
Write results figures' captions

@@
   \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
-  \caption{\textbf{Model rotations.} \TODO{}}
+  \caption{\textbf{Model rotations.} Each subplot shows the
+    correspondence between the true angle of rotation ($R$) for $h_1$
+    stimuli and the amount of rotation performed by the model. Dotted
+    lines indicate a perfect 1:1 relationship. The \naive{} and
+    parametric models (left and center panels, respectively) are
+    identical; they show modest agreement until rougly
+    $\frac{\pi}{2}$, after which agreement decreases as they tend to
+    encounter local maxima. The nonparametric model (right panel)
+    exhibits a significant correlation between the true angle and the
+    amount the model rotates.}
   \label{fig:rotations}
@@
   \includegraphics[width=\textwidth]{../figures/Z_accuracy.pdf}
-  \caption{\textbf{Accuracy in estimating $Z$.} \TODO{}}
+  \caption{\textbf{Accuracy in estimating $Z$.} Each subplot shows the
+    true (``gold standard'') value of $Z$ vs. the value estimated by
+    the model. Black dotted lines indicate a perfect 1:1
+    correspondence between the true and estimated values. All models
+    perform well for $h_0$ stimuli, but exhibit different behavior for
+    $h_1$ stimuli. The \naive{} model (left panel) tends to either
+    greatly overestimate, or underestimate. The parametric model
+    (center panel) is very accurate for about half the $h_1$ stimuli,
+    and severly underestimates the rest. The nonparametric model
+    (right panel) maintains a decent correspondence with the true $Z$,
+    with the exception of a handful of outliers.}
   \label{fig:accuracy}
@@
 
+We conclude that, from this initial survey, models of mental rotation
+which take an ``active'', directed approach seem well-suited to
+explaining human behavior in these tasks. Future work will collect
+empirical data from participants to perform a more quantitative
+analysis of these modeling tools.
+
 \textbf{Acknowledgements}

### 2013-05-13 man/final-project.tex
Move shepard rotation figure
Comment out abstract and acknowlegements
Spellcheck

@@
 
-\begin{abstract}
-\TODO{}
-\end{abstract}
+% \begin{abstract}
+% \TODO{}
+% \end{abstract}
 
@@
 direction should the object be rotated?  What are the requirements for
-``congruency''? When should one stop rotating and accept the
+``congruence''? When should one stop rotating and accept the
 hypothesis that the objects are different? In this project, I will
@@
 % these samples \cite{Osborne:2012tm}.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.95\textwidth]{../figures/shepard-rotation.png}
-  \caption{\textbf{Classic mental rotation task}. Participants in
-    \cite{Shepard1971} saw stimuli such as these, and judged whether
-    each pair of shapes was the same shape in two different
-    orientations, or two different shapes. \textbf{A} shows a ``plane
-    rotation'', \textbf{B} shows a ``depth rotation'', and \textbf{C}
-    shows two distinct objects.}
-  \label{fig:mental-rotation}
-\end{figure}
 
@@
 \end{equation}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.95\textwidth]{../figures/shepard-rotation.png}
+  \caption{\textbf{Classic mental rotation task}. Participants in
+    \cite{Shepard1971} saw stimuli such as these, and judged whether
+    each pair of shapes was the same shape in two different
+    orientations, or two different shapes. \textbf{A} shows a ``plane
+    rotation'', \textbf{B} shows a ``depth rotation'', and \textbf{C}
+    shows two distinct objects.}
+  \label{fig:mental-rotation}
+\end{figure}
 
@@
   Quadrature} \cite{Diaconis:1988uo,OHagan:1991tx} to estimate $Z$.
-Bayesian Quadrature allows us to compute a posterior distrbution over
+Bayesian Quadrature allows us to compute a posterior distribution over
 $Z$ by placing a Gaussian Process (GP) prior on the function $S$ and
@@
   over $\log S$ (Equation \ref{eq:bq-Z-mean}) requires computing
-  $\mu_S$, which may have nonpositive segments.}:
+  $\mu_S$, which may have non-positive segments.}:
 \begin{equation*}
@@
     parametric models (left and center panels, respectively) are
-    identical; they show modest agreement until rougly
+    identical; they show modest agreement until roughly
     $\frac{\pi}{2}$, after which agreement decreases as they tend to
@@
 less the model has to rotate, and the less likely it will get stuck on
-local maxima. Thus, it is more likely to locat the global maximum,
+local maxima. Thus, it is more likely to locate the global maximum,
 which generally corresponds to the true angle of rotation. For
@@
 
-Similarly, because the nonparametric model actively attemps to obtain
+Similarly, because the nonparametric model actively attempts to obtain
 an accurate estimate of $Z$, its estimate of $Z$ is indeed fairly
@@
     (center panel) is very accurate for about half the $h_1$ stimuli,
-    and severly underestimates the rest. The nonparametric model
+    and severely underestimates the rest. The nonparametric model
     (right panel) maintains a decent correspondence with the true $Z$,
@@
 
-\textbf{Acknowledgements}
-
-\TODO{}
+% \textbf{Acknowledgments}
+% \TODO{}
 

### 2013-05-13 man/final-project.tex
Flesh out introduction

@@
 and manipulate them, to a certain extent, as if they were real
-\cite{Kosslyn:2009tj}.
-
-However, the use of mental simulation is predicated on determining
-appropriate parameters to give the simulation, and people's cognitive
-constraints may furthermore place limitations on the duration or
-precision of simulation. One hypothesis for how these issues are
-handled argues that people use a ``rational'' solution, meaning that
-it is optimal under the given constraints
-\cite{Lieder:2012wg,Vul:2009wy,Griffiths2012a}.
+\cite{Kosslyn:2009tj}.  However, the use of mental simulation is
+predicated on determining appropriate parameters to give the
+simulation, and people's cognitive constraints may furthermore place
+limitations on the duration or precision of simulation. One hypothesis
+for how these issues are handled argues that people use a rational
+solution, meaning that it is optimal under existing constraints
+\cite{Lieder:2012wg,Vul:2009wy,Griffiths2012a}, such as processing
+speed, memory, or perceptual acuity.
 
@@
 ``congruence''? When should one stop rotating and accept the
-hypothesis that the objects are different? In this project, I will
-investigate the optimal computational solution to this task. Thus, the
-specific question this project aims to answer is: what is the rational
-computational solution to the mental rotation task, and does it
-predict the people's behavior on the task?
+hypothesis that the objects are different? We investigate several
+computational solutions to this task and qualitatively compare them to
+general mental rotation results. Of particular interest is a model
+which actively reasons about the amount of evidence it has, and
+performs mental rotations until it reaches a certain level of
+confidence. We contrast this model with simpler approaches which are
+less computationally demanding, but which employ less flexible
+strategies.
+
+The plan of the paper is as follows. First, we give a brief overview
+of the work on mental rotation and active learning. Next, we present
+the problem domain and a computational-level analysis of how to solve
+it. We describe three different models which approximate this
+solution, evaluate their accuracy at approximating the computational
+solution, and qualitatively compare their behavior to that of the
+classic mental rotation results. We end with a discussion of the
+strengths and weaknesses of each model, as well as directions for
+future work.
 

### 2013-05-14 man/final-project.tex
Write background section

@@
 
-% We do not know the form of $S$, and so cannot compute this integral
-% analytically. We cannot rely on a simple Monte Carlo simulation to
-% estimate it, either: because mental rotation is costly in terms of
-% cognitive resources and must be performed in a sequence, we cannot
-% evaluate $S(I_b|I_M)$ at arbitrary $I_M$. Instead, we must choose a
-% relatively small, sequential set of $I_M$ to estimate the
-% integral.
+There is a long and rich history of work on mental rotation beginning
+with \cite{Shepard1971}, who presented participants with pairs of
+images such as those shown in Figure \ref{fig:mental-rotation}. These
+pairs three-dimensional objects could either be: the same object
+rotated in plane (Figure \ref{fig:mental-rotation}A), the same object
+rotated in depth (Figure \ref{fig:mental-rotation}B), or two different
+objects (Figure \ref{fig:mental-rotation}C; in this case, the
+``different'' objects had reflective symmetry, but no rotational
+symmetry). Participants had to determine whether the two images were
+of the same object or not, and \cite{Shepard1971} famously found that
+response times for plane and depth rotations had a strong linear
+relationship with the minimum angle of rotation between the two
+objects. The conclusion was that participants were visually
+``rotating'' the objects in their minds.
 
-% Bayesian quadrature \cite{Diaconis:1988uo} provides an alternate
-% method for evaluating the integral by placing a prior distribution on
-% functions and then computing a posterior over values of the
-% integral. While Bayesian quadrature itself does not address the issue
-% of choosing points to evaluate the function at, recent work in
-% statistics and machine learning has examined how to optimally select
-% these samples \cite{Osborne:2012tm}.
+For many years, this idea was contested, with some researchers arguing
+the underlying cognitive processes were not visual in nature
+(e.g. \cite{Pylyshyn1981}). In particular, \cite{Anderson1978} proved
+that the mental imagery debate could not be resolved on the basis of
+response time data alone. In recent years, however, a significant
+amount of work has investigated the neural underpinnings of the mental
+rotation phenomena, arguing that the results of brain imaging studies
+support the existence of visual mental rotation \cite{Kosslyn1988,
+  Kosslyn:2009tj}.
 
-\section{Computational model}
+We assume that mental imagery is indeed a visual process, and turn to
+the question of how, and when, it is used. Previous work has examined
+how people might use imagery \cite{Hegarty2004, Schwartz1999} to solve
+reasoning problems, and when they might use imagery as opposed to a
+symbolic rule \cite{Schwartz:1996uy}. People seem to use imagery
+methodically, which begs the question: what is the method?
+
+Others studying people's pattern of behavior when engaged in
+self-directed learning have found that they tend to use a strategy of
+\textit{active learning} \cite{Gureckis:2012gu, Markant:2012uu}. When
+given a choice as to the information they can obtain, people will not
+choose the information randomly (as is the case with a \naive{} Monte
+Carlo sampler), but according to some utility function such as
+information gain \cite{Nelson2007, Markant:2012uu}. We propose that
+such a strategy can be combined with the notion of mental imagery as a
+tool: an optimal model of mental rotation may, perhaps, rely on an
+active strategy. In the next sections, we investigate this idea with
+several different approaches.
+
+\section{Computational-level model}
 
 People are presented with two images, $X_a$ and $X_b$, which are the
-coordinates of the vertices of two shapes (e.g., see Figure
+coordinates of the vertices of two-dimensional shapes similar to those
+used by \cite{Cooper:1975wp} (e.g., see Figure
 \ref{fig:stimuli}). Participants must determine whether $X_a$ and

### 2013-05-14 man/final-project.tex
Move shepard rotation figure again
Reword the beginning of the computation model section

@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.95\textwidth]{../figures/shepard-rotation.png}
+  \caption{\textbf{Classic mental rotation task}. Participants in
+    \cite{Shepard1971} saw stimuli such as these, and judged whether
+    each pair of shapes was the same shape in two different
+    orientations, or two different shapes. \textbf{A} shows a ``plane
+    rotation'', \textbf{B} shows a ``depth rotation'', and \textbf{C}
+    shows two distinct objects.}
+  \label{fig:mental-rotation}
+\end{figure}
+
 \section{Computational-level model}
@@
 
-To compare the hypotheses, we need to compute the likelihood of each
-hypothesis, i.e. $p(X_a,\ X_b\ \vert \ h_0)$ and $p(X_a,\ X_b\ \vert \
-h_1)$.
-
-The likelihood is easy to compute under $h_0$, because we assume that
-$X_a$ and $X_b$ are independent. Thus:
+To compare the hypotheses, we need to compute the posterior:
 \begin{equation}
-  p(X_a,\ X_b\ \vert \ h_0)=p(X_a)p(X_b)
-  \label{eq:lh-h0}
+p(h_i\ \vert\ X_a, X_b)\propto p(X_a, X_b\ \vert\ h_i)p(h_i)
 \end{equation}
-
-When the configurations are the same, the likelihood becomes more
-complicated:
-\begin{equation} 
-  p(X_a,\ X_b\ \vert \ h_1)=\int_R\int_M p(X_a) p(X_b\vert X_a,R,M) p(R) p(M)\ \mathrm{d}M\ \mathrm{d}R
-\end{equation}
-
-For a small number of vertices, we can compute the integral over $M$
-by enumerating every possible mapping between $X_a$ and $X_b$. After
-doing so, we obtain:
+Assuming the hypotheses are equally likely \textit{a priori}, the
+prior term $p(h_i)$ will cancel out when comparing $h_0$ and $h_1$,
+thus allowing us to focus on the likelihoods: 
+\begin{align}
+  p(X_a,\ X_b\ \vert \ h_0)&=p(X_a)p(X_b) \label{eq:lh-h0}\\
+  p(X_a,\ X_b\ \vert \ h_1)&=\int_R\int_M p(X_a) p(X_b\vert X_a,R,M) p(R) p(M)\ \mathrm{d}M\ \mathrm{d}R \label{eq:lh-h1}
+\end{align}
+Under $h_0$, the likelihood is easy to compute because we assume that
+$X_a$ and $X_b$ are independent (Equation \ref{eq:lh-h0}). However,
+when the configurations are the same, the likelihood becomes more
+complicated (Equation \ref{eq:lh-h1}). For a small number of vertices,
+we can compute the integral over $M$ by enumerating every possible
+mapping between $X_a$ and $X_b$. After doing so, we obtain:
 \begin{equation} 
@@
 \end{equation}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.95\textwidth]{../figures/shepard-rotation.png}
-  \caption{\textbf{Classic mental rotation task}. Participants in
-    \cite{Shepard1971} saw stimuli such as these, and judged whether
-    each pair of shapes was the same shape in two different
-    orientations, or two different shapes. \textbf{A} shows a ``plane
-    rotation'', \textbf{B} shows a ``depth rotation'', and \textbf{C}
-    shows two distinct objects.}
-  \label{fig:mental-rotation}
-\end{figure}
 
@@
   &= \int_R p(X_b\vert X_R)p(X_a)p(R)\ \mathrm{d}R
-  \label{eq:lh-h1}
 \end{align}

### 2013-05-29 man/nips-2013.tex
Create initial version from final project writeup
Add Tom's comments/changes

--- 2013-05-14 man/final-project.tex
@@
 
-\title{Optimal strategies in mental rotation tasks}
-\author{Jessica B.~Hamrick\\
+\title{Mental simulation as Bayesian Quadrature}
+\author{Jessica B.~Hamrick and Thomas L.~Griffiths\\
   Department of Psychology\\
@@
   Berkeley, CA 94720\\
-  \texttt{jhamrick@berkeley.edu}}
+  \texttt{\{jhamrick,tom\_griffiths\}@berkeley.edu}}
 
@@
 
+% I think the main thing would be doing something more sophisticated
+% in the search, rather than using the gradient. A less sophisticated
+% adaptive sampling method seems like it would be best, and using von
+% Mises makes sense.
+
 \section{Introduction}
@@
 % 4. What's in the paper
+
+% The introduction could be made more general -- start with the
+% question of how a rational agent should use her computational
+% resources, and then get onto mental simulation/rotation as a special
+% case of that, rather than starting with the phenomena. I would make
+% the big picture goals clear at the start, and then segue into the
+% mystery of mental rotation -- how do we know which way to rotate?
 
@@
 images such as those shown in Figure \ref{fig:mental-rotation}. These
-pairs three-dimensional objects could either be: the same object
+pairs of three-dimensional objects could either be: the same object
 rotated in plane (Figure \ref{fig:mental-rotation}A), the same object
@@
 symbolic rule \cite{Schwartz:1996uy}. People seem to use imagery
-methodically, which begs the question: what is the method?
+methodically, which raises the question: what is the method?
 
@@
     rotation'', \textbf{B} shows a ``depth rotation'', and \textbf{C}
-    shows two distinct objects.}
+    shows two distinct objects. \TODO{update this caption}}
   \label{fig:mental-rotation}
@@
 \end{align} 
+% I wasn't clear whether the step that happens in Equation 6 should be
+% in the model, or in the algorithm. I would probably be inclined to
+% define the model in full generality, and then bring up the
+% increments as part of the algorithm.
 Where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
@@
 This process of mental rotation (i.e., generating a single $X_R$ (as
-in Equation \ref{eq:tau}) and then computing $S(X_b, X_R)$) is a
-computationally demanding cognitive process. Thus, our goal is to
-minimize the number of rotations while still obtaining an estimate of
-$Z$ that is accurate enough to choose the correct hypothesis. With
-this in mind, we now examine several approaches to estimating $Z$.
+in Equation \ref{eq:tau}) and then computing $S(X_b, X_R)$) is
+computationally demanding. Thus, our goal is to minimize the number of
+rotations while still obtaining an estimate of $Z$ that is accurate
+enough to choose the correct hypothesis. With this in mind, we now
+examine several approaches to estimating $Z$.
 
@@
 \section{Results}
+
+% I think it would be compelling to add a second row to Equation 5
+% where you show the averages and s.e. at each rotation for many noisy
+% runs of the algorithm (possibly on different stimuli). If you can
+% show that the average is a linear function, in a way that looks like
+% the classic behavioral data, that will be compelling.
 

### 2013-05-29 man/nips-2013.tex
Include analysis tex file
Use analysis macros

@@
 \newcommand{\Naive}[0]{Na\"ive}
+
+\include{nips-2013-analyses}
 
@@
 correlation between it and the true angle of rotation is
-undefined. The mean error of responses was $\ME{}=0.01$, indicating
-that $Z$ is a sufficient approximation of $\int_R p(X_b\ \vert\
-X_R)p(R)\ \mathrm{d}R$.
+undefined. The mean error of responses was $\GSME{}$, indicating that
+$Z$ is a sufficient approximation of $\int_R p(X_b\ \vert\ X_R)p(R)\
+\mathrm{d}R$.
 
@@
 The correlation between the number of rotations computed by the
-\naive{} model and the true angle of rotation was $\rho=-0.11$. The
+\naive{} model and the true angle of rotation was $\Naivecorr{}$. The
 shape of the data is more interesting (see Figure \ref{fig:rotations},
-left panel): the \naive{} model actually corresponds quite well to
-the true angle of rotation for $R<\frac{\pi}{2}$ ($\rho=0.64$). This
+left panel): the \naive{} model actually corresponds quite well to the
+true angle of rotation for $R<\frac{\pi}{2}$ ($\Naivecorrsm{}$). This
 is unsurprising, because the closer the true angle is to zero, the
@@
 $R>\frac{\pi}{2}$, we see an increasing tendency to under-rotate
-($\rho=-0.42$); this is likely because it finds a local maxima and
+($\Naivecorrbg{}$); this is likely because it finds a local maxima and
 ends prematurely.
@@
 The \naive{} model's response error rate was better than chance, but
-still quite high, with $\ME{}=0.23$. Closer inspection reveals that
-the bulk of this comes from $h_1$ stimuli ($\ME{}=0.36$
-vs. $\ME{}=0.1$ for $h_0$ stimuli). This is, as above, probably
+still quite high, with $\NaiveME{}$. Closer inspection reveals that
+the bulk of this comes from $h_1$ stimuli ($\NaiveMEsame{}$
+vs. $\NaiveMEdiff{}$ for $h_0$ stimuli). This is, as above, probably
 related to finding only local maxima: if the model finds a local
@@
 This is reflected in the accuracy in estimating $Z$ as well
-($\MSE{}=0.14$). Figure \ref{fig:accuracy} (left panel) shows
+($\NaiveMSE{}$). Figure \ref{fig:accuracy} (left panel) shows
 individual estimates for each stimulus, color-coded by true
@@
 for the parametric model was higher than the \naive{} model, with
-$\ME{}=0.26$. Breaking this down, we see that the error for $h_1$
-stimuli is $\ME{}=0.5$: in other words, the parametric model is at
-chance when determining whether two identical shapes are the
-same. This is because it never overestimates (see Figure
-\ref{fig:accuracy}, center panel). So, it accurately estimates $Z$ for
-stimuli which the \naive{} model would overestimate ($\MSE{}=0.07$),
-but still has poor performance when $Z$ is underestimated. Indeed, if
-we exclude stimuli for which the \naive{} model overestimates ($Z\geq
-0.3$), the error for the \naive{} model lowers to match that of the
-parametric model.
+$\VMME{}$. Breaking this down, we see that the error for $h_1$ stimuli
+is $\VMMEsame{}$: in other words, the parametric model is at chance
+when determining whether two identical shapes are the same. This is
+because it never overestimates (see Figure \ref{fig:accuracy}, center
+panel). So, it accurately estimates $Z$ for stimuli which the \naive{}
+model would overestimate ($\VMMSE{}$), but still has poor performance
+when $Z$ is underestimated. \TODO{Indeed, if we exclude stimuli for
+  which the \naive{} model overestimates ($Z\geq 0.3$), the error for
+  the \naive{} model lowers to match that of the parametric model.}
 
@@
 The rotations computed by the nonparametric model were strongly
-correlated with the true rotations ($\rho=0.73$, see Figure
+correlated with the true rotations ($\BQcorr$, see Figure
 \ref{fig:rotations}, right panel). This is because the nonparametric
@@
 Correspondingly, the nonparametric model is much more accurate in
-choosing the correct hypothesis ($\ME{}=0.06$). As with the \naive{}
-and parametric models, it mostly errs on $h_1$ stimuli
-($\ME{}=0.1$). This appears to mostly be the result of choosing the
-incorrect initial direction of rotation.
+choosing the correct hypothesis ($\BQME{}$). As with the \naive{} and
+parametric models, it mostly errs on $h_1$ stimuli ($\BQMEsame$). This
+appears to mostly be the result of choosing the incorrect initial
+direction of rotation.
 
@@
 an accurate estimate of $Z$, its estimate of $Z$ is indeed fairly
-accurate ($\MSE{}=0.10$). This is slightly higher than the
-nonparametric model, but is due to a few outliers. If we again exclude
-stimuli for which $Z$ is greatly overestimated ($Z\geq 0.3$), the
-nonparametric model's error decreases significantly ($\MSE{}=0.03$).
+accurate ($\BQMSE{}$). This is slightly higher than the nonparametric
+model, but is due to a few outliers. \TODO{If we again exclude stimuli
+  for which $Z$ is greatly overestimated ($Z\geq 0.3$), the
+  nonparametric model's error decreases significantly
+  ($\MSE{}=0.03$).}
 

### 2013-05-31 code/util.py
Handle upper/lower Z values in run_model

@@ -200,7 +200,7 @@ def run_model(stim, model, opt):
     # how many points were sampled
     samps = np.zeros(nsamp)
     # the estimate of Z
-    Z = np.empty((nsamp, 2))
+    Z = np.empty((nsamp, 3))
     # the likelihood ratio
     ratio = np.empty((nsamp, 3))
     # which hypothesis was accepted
@@ -216,7 +216,8 @@ def run_model(stim, model, opt):
 
         # fill in the data arrays
         samps[i] = len(m.ix) / float(m._rotations.size)
-        Z[i] = (m.Z_mean, m.Z_var)
+        Z[i, 0] = m.Z_mean
+        Z[i, 1:] = m.Z_var
         ratio[i] = m.likelihood_ratio()
         hyp[i] = m.ratio_test(level=10)
 

### 2013-05-31 man/nips-2013.tex
Add TODO
Some paper cleanup

@@
 
+\TODO{Want to talk about previous work on strategies people might use,
+  e.g. eye-tracking results from Just and Carpenter (1975?)}
+
 % 1. The big picture
@@
 
-Consider the objects in Figure \ref{fig:mental-rotation}. In each
-panel, are the two depicted objects identical (except for a rotation),
-or distinct? When presented with this mental rotation task, people
-default to a strategy in which they visualize one object rotating
-until it is congruent with the other \cite{Shepard1971}. There is
-strong evidence for such \textit{mental imagery} or \textit{mental
-  simulation}: we can imagine three-dimensional objects in our minds
-and manipulate them, to a certain extent, as if they were real
-\cite{Kosslyn:2009tj}.  However, the use of mental simulation is
-predicated on determining appropriate parameters to give the
-simulation, and people's cognitive constraints may furthermore place
-limitations on the duration or precision of simulation. One hypothesis
-for how these issues are handled argues that people use a rational
-solution, meaning that it is optimal under existing constraints
-\cite{Lieder:2012wg,Vul:2009wy,Griffiths2012a}, such as processing
-speed, memory, or perceptual acuity.
+Consider the images on the left in Figure
+\ref{fig:mental-rotation}. In each panel, are the two depicted objects
+identical (except for a rotation), or distinct? When presented with
+this mental rotation task, people default to a strategy in which they
+visualize one object rotating until it is congruent with the other
+\cite{Shepard1971}. There is strong evidence for such \textit{mental
+  imagery} or \textit{mental simulation}: we can imagine
+three-dimensional objects in our minds and manipulate them, to a
+certain extent, as if they were real \cite{Kosslyn:2009tj}.  However,
+the use of mental simulation is predicated on determining appropriate
+parameters to give the simulation, and people's cognitive constraints
+may furthermore place limitations on the duration or precision of
+simulation. One hypothesis for how these issues are handled argues
+that people use a rational solution, meaning that it is optimal under
+existing constraints \cite{Lieder:2012wg,Vul:2009wy,Griffiths2012a},
+such as processing speed, memory, or perceptual acuity.
 
@@
 There is a long and rich history of work on mental rotation beginning
-with \cite{Shepard1971}, who presented participants with pairs of
-images such as those shown in Figure \ref{fig:mental-rotation}. These
-pairs of three-dimensional objects could either be: the same object
-rotated in plane (Figure \ref{fig:mental-rotation}A), the same object
-rotated in depth (Figure \ref{fig:mental-rotation}B), or two different
-objects (Figure \ref{fig:mental-rotation}C; in this case, the
-``different'' objects had reflective symmetry, but no rotational
+with Shepard and Metzler \cite{Shepard1971}, who presented
+participants with pairs of images such as those shown in Figure
+\ref{fig:mental-rotation} (left panel). These pairs of
+three-dimensional objects could either be: the same object rotated in
+plane (Figure \ref{fig:mental-rotation}A), the same object rotated in
+depth (Figure \ref{fig:mental-rotation}B), or two different objects
+(Figure \ref{fig:mental-rotation}C; in this case, the ``different''
+objects had reflective symmetry, but no rotational
 symmetry). Participants had to determine whether the two images were
@@
 relationship with the minimum angle of rotation between the two
-objects. The conclusion was that participants were visually
-``rotating'' the objects in their minds.
+objects (Figure \ref{fig:mental-rotation}, right panel). The
+conclusion was that participants were visually ``rotating'' the
+objects in their minds.
 
@@
 Because the vertices are connected in a way which forms a closed loop,
-we need only consider $n$ mappings, $M$, of the $n$ vertices (we
-assume the uncertainty is only in which is the ``first''
-vertex). Thus, the possible orderings are of the form $M=\lbrace{}0,
-1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0, \ldots{}, n-1\rbrace{}$,
-and so on. This gives us an explicit form for the similarity function:
+we need only consider $2n$ mappings, $M$, of the $n$ vertices (we
+assume uncertainty for which is the ``first'' vertex, and then which
+of its two neighbors is the ``second''). Thus, the possible orderings
+are of the form $M=\lbrace{}0, 1, \ldots{}, n\rbrace{}$,
+$M=\lbrace{}n, 0, \ldots{}, n-1\rbrace{}$, and so on. This gives us an
+explicit form for the similarity function:
 \begin{equation}
-  S(X_b, X_R)=\frac{1}{n}\sum_{M\in\mathbb{M}}\prod_{i=1}^n\mathcal{N}(X_b[i]\ \vert \ (MX_R)[i], \Sigma)
+  S(X_b, X_R)=\frac{1}{2n}\sum_{M\in\mathbb{M}}\prod_{i=1}^n\mathcal{N}(X_b[i]\ \vert \ (MX_R)[i], \Sigma)
   \label{eq:similarity}
@@
 
-\subsection{Gold standard}
+\paragraph{Gold standard}
 
@@
 
-\subsection{\Naive{} model}
+\paragraph{\Naive{} model}
 
@@
 
-\subsection{Parametric (Von Mises) model}
+\paragraph{Parametric (Von Mises) model}
 
@@
 
-\subsection{Nonparametric (Bayesian Quadrature) model}
+\paragraph{Nonparametric (Bayesian Quadrature) model}
 
@@
   combination of MLII and marginalization to fit the kernel
-  parameters, we set the output scale $h_S=0.125$ and $h_{\log
+  parameters, we set the output scale $h_S=\sqrt{0.1}$ and $h_{\log
     S}=\log(h_S + 1)$, and use MLII to fit all other parameters.},
@@
 
-To choose $\mathbf{R}$, we pick an initial direction of rotation which
-results in the higher value of $S$, and then continue rotating in that
-direction until the variance in Equation \ref{eq:bq-Z-var} is low
-enough that we are confident about the likelihood ratio $\ell$:
+\TODO{To choose $\mathbf{R}$}, we pick an initial direction of
+rotation which results in the higher value of $S$, and then continue
+rotating in that direction until the variance in Equation
+\ref{eq:bq-Z-var} is low enough that we are confident about the
+likelihood ratio $\ell$:
 \begin{equation*}
@@
 
-We evaluated each model's performance on 50 randomly generated shapes
-(such as $X_a$ in Figure \ref{fig:stimuli}). For each shape, we picked
-a random angle $R$ and generated two stimuli: a $h_0$ stimulus, by
-reflecting $X_a$ across the $y$-axis and rotating it by $R$; and a
-$h_1$ stimulus, by simply rotating $X_a$ by $R$.
+We evaluated each model's performance on 20 randomly generated shapes
+which had between 3 and 6 vertices, inclusive, (e.g. $X_a$ in Figure
+\ref{fig:stimuli}). For each shape, we computed 18 ``different'' and
+18 ``same'' stimuli pairs, with $R$ spaced at $20^\prime$ increments
+between 0 and 360. ``Different'' pairs were created by reflecting
+$X_a$ across the $y$-axis and rotating it by $R$; for ``same'' pairs,
+$X_a$ was rotated by $R$.
 
@@
 \begin{itemize}
-\item \textit{Rotations}: for stimuli where $h_1$ is correct, how
-  correlated are the true angles of rotation with the rotations
-  performed by the model? This is defined to be the Pearson's
-  correlation coefficient $\rho$ between $R$ and $\vert
-  \mathbf{R}\vert$. Figure \ref{fig:rotations} shows individual points
-  corresponding to true rotation vs. rotation by the model, for each
-  model.
+\item \textit{Rotations}: for ``same'' pairs, how correlated are the
+  true angles of rotation with the rotations performed by the model?
+  This is defined to be the Pearson's correlation coefficient $\rho$
+  between $R$ and $\vert \mathbf{R}\vert$. Figure \ref{fig:rotations}
+  shows individual points corresponding to true rotation vs. rotation
+  by the model, for each model.
 \item \textit{Responses}: how accurate is the model at choosing the

### 2013-05-31 code/analysis_tools.py
Update rotations plot

@@ -115,13 +115,13 @@ def calc_Rot(models, data, sdata, only_correct=False):
         dd = samps[ih0[:, None] & (err == 0)] * 100
         diff_raw = dd
         diff_mean = np.mean(dd)
-        diff_sem = scipy.stats.sem(dd, ddof=1)
-        Rot_diff[model] = (diff_raw, diff_mean, diff_sem)
+        diff_std = np.std(dd, ddof=1)
+        Rot_diff[model] = (diff_raw, diff_mean, diff_std)
 
         # calculate for same pairs
         same_raw = []
         same_mean = np.empty(unique_ang.size)
-        same_sem = np.empty(unique_ang.size)
+        same_std = np.empty(unique_ang.size)
         for aidx, ang in enumerate(unique_ang):
             idx = (np.abs(min_ang - ang) < 1e-8).ravel()
             if only_correct:
@@ -131,8 +131,8 @@ def calc_Rot(models, data, sdata, only_correct=False):
             ss = samps[(idx & ih1)[:, None] & correct] * 100
             same_raw.append(ss.copy())
             same_mean[aidx] = np.mean(ss)
-            same_sem[aidx] = scipy.stats.sem(ss, ddof=1)
-        Rot_same[model] = (same_raw, same_mean, same_sem)
+            same_std[aidx] = np.std(ss, ddof=1)
+        Rot_same[model] = (same_raw, same_mean, same_std)
 
     return Rot_diff, Rot_same
 
@@ -142,7 +142,7 @@ def calc_Corr(models, sdata, Rot_same):
     sm = unique_ang < (np.pi / 2.)
     Corr = {}
     for model in models:
-        raw, mean, sem = Rot_same[model]
+        raw, mean, std = Rot_same[model]
         corr = float(xcorr(unique_ang, mean))
         corr_sm = xcorr(unique_ang[sm], mean[sm])
         corr_bg = xcorr(unique_ang[~sm], mean[~sm])

### 2013-05-31 man/nips-2013.tex
Show all rotations, not means
First pass through results

@@
 
-\paragraph{\Naive{} model}
+\paragraph{\Naive{}}
 
@@
 
-\paragraph{Parametric (Von Mises) model}
+\paragraph{Parametric (Von Mises)}
 
@@
 
-\paragraph{Nonparametric (Bayesian Quadrature) model}
+\paragraph{Nonparametric (Bayesian Quadrature)}
 
@@
 $X_a$ across the $y$-axis and rotating it by $R$; for ``same'' pairs,
-$X_a$ was rotated by $R$.
+$X_a$ was rotated by $R$. \TODO{need to mention that shapes were
+  chosen to be discriminable by the gold standard}
 
@@
 \begin{itemize}
-\item \textit{Rotations}: for ``same'' pairs, how correlated are the
-  true angles of rotation with the rotations performed by the model?
-  This is defined to be the Pearson's correlation coefficient $\rho$
-  between $R$ and $\vert \mathbf{R}\vert$. Figure \ref{fig:rotations}
-  shows individual points corresponding to true rotation vs. rotation
-  by the model, for each model.
 \item \textit{Responses}: how accurate is the model at choosing the
@@
   or fraction of times the model picks the incorrect hypothesis.
+\item \textit{Rotations}: for those ``same'' pairs which the model
+  gets correct, how correlated are the true angles of rotation with
+  the rotations performed by the model?  This is defined to be the
+  Pearson's correlation coefficient $\rho$ between $R$ and $\vert
+  \mathbf{R}\vert$. Figure \ref{fig:rotations} shows individual points
+  corresponding to true rotation vs. rotation by the model, for each
+  model.
 \item \textit{Estimates of $Z$}: how accurate is the model's estimate
@@
 
-The gold standard always evaluates $S$ at all angles, so the
-correlation between it and the true angle of rotation is
-undefined. The mean error of responses was $\GSME{}$, indicating that
-$Z$ is a sufficient approximation of $\int_R p(X_b\ \vert\ X_R)p(R)\
-\mathrm{d}R$.
+The mean error of the gold standard model was $\GSME{}$, indicating
+that $Z$ is a sufficient approximation of $\int_R p(X_b\ \vert\
+X_R)p(R)\ \mathrm{d}R$ for these particular stimuli.
 
-\paragraph{\Naive{} model}
+\paragraph{\Naive{}}
 
-The correlation between the number of rotations computed by the
-\naive{} model and the true angle of rotation was $\Naivecorr{}$. The
-shape of the data is more interesting (see Figure \ref{fig:rotations},
-left panel): the \naive{} model actually corresponds quite well to the
-true angle of rotation for $R<\frac{\pi}{2}$ ($\Naivecorrsm{}$). This
-is unsurprising, because the closer the true angle is to zero, the
-less the model has to rotate, and the less likely it will get stuck on
-local maxima. Thus, it is more likely to locate the global maximum,
-which generally corresponds to the true angle of rotation. For
-$R>\frac{\pi}{2}$, we see an increasing tendency to under-rotate
-($\Naivecorrbg{}$); this is likely because it finds a local maxima and
-ends prematurely.
+The \naive{} model's response error rate was better than chance with
+$\NaiveME{}$. Closer inspection reveals that much of this comes from
+$h_0$ stimuli ($\NaiveMEdiff{}$), where the asymmetric linear
+interpolation gives an overestimate of $Z$ (e.g., in Figure
+\ref{fig:li}). There is some error from $h_1$ stimuli as well
+($\NaiveMEsame{}$), which happens when the hill climbing search finds
+a local maxima and terminates prematurely, thus resulting in a low
+estimate of $Z$.
 
-The \naive{} model's response error rate was better than chance, but
-still quite high, with $\NaiveME{}$. Closer inspection reveals that
-the bulk of this comes from $h_1$ stimuli ($\NaiveMEsame{}$
-vs. $\NaiveMEdiff{}$ for $h_0$ stimuli). This is, as above, probably
-related to finding only local maxima: if the model finds a local
-maxima which is low enough, the area under the estimated curve will
-not be large enough to accept $h_1$.
+The correlation between the average number of rotations computed by
+the \naive{} model and the true angle of rotation was
+$\Naivecorr{}$. The shape of the data is more interesting (see Figure
+\ref{fig:rotations}, top left panel): the \naive{} model actually
+corresponds extremeley well to the true angle of rotation for
+$R<\frac{\pi}{2}$ ($\Naivecorrsm{}$). This is unsurprising, because
+the closer the true angle is to zero, the less the model has to
+rotate, and the less likely it will get stuck on local maxima. Thus,
+it is more likely to locate the global maximum, which corresponds to
+the true angle of rotation. For $R>\frac{\pi}{2}$, we see an
+increasing tendency to under-rotate due to getting stuck on local
+maxima, as well as a tendency to over-rotate if the wrong rotation
+direction was chosen.
 
-This is reflected in the accuracy in estimating $Z$ as well
-($\NaiveMSE{}$). Figure \ref{fig:accuracy} (left panel) shows
-individual estimates for each stimulus, color-coded by true
-hypothesis. The model underestimates $Z$ for significant portion of
-the $h_1$ stimuli. On the other $h_1$ stimuli, the \naive{} model
-actually tends to overestimate because linear interpolation does not
-account for symmetry of global maxima (e.g., Figure \ref{fig:li}).
+The \naive{} model's error on $h_0$ stimuli is reflected in its
+accuracy in estimating $Z$ ($\NaiveMSE{}$). Figure \ref{fig:accuracy}
+(left panel) shows individual estimates for each stimulus, color-coded
+by true hypothesis. The model overestimates $Z$ for most of the $h_0$
+and $h_1$ stimuli, and underestimates several of the $h_1$ stimuli as
+well.
 
-\paragraph{Parametric (Von Mises) model}
+\paragraph{Parametric (Von Mises)}
 
-Like the \naive{} model, the parametric model uses hill-climbing as a
-sampling strategy, so they have identical correlations with the true
-angle of rotation. However, given the same $\mathbf{R}$, the
-parametric model estimates $Z$ differently, thus giving rise to
-different response and estimated $Z$ error rates. The response error
-for the parametric model was higher than the \naive{} model, with
-$\VMME{}$. Breaking this down, we see that the error for $h_1$ stimuli
-is $\VMMEsame{}$: in other words, the parametric model is at chance
-when determining whether two identical shapes are the same. This is
-because it never overestimates (see Figure \ref{fig:accuracy}, center
-panel). So, it accurately estimates $Z$ for stimuli which the \naive{}
-model would overestimate ($\VMMSE{}$), but still has poor performance
-when $Z$ is underestimated. \TODO{Indeed, if we exclude stimuli for
-  which the \naive{} model overestimates ($Z\geq 0.3$), the error for
-  the \naive{} model lowers to match that of the parametric model.}
+The parametric model is also fairly inaccurate ($\VMME{}$), making
+more errors on $h_1$ stimuli ($\VMMEsame{}$) than $h_0$ stimuli
+($\VMMEdiff{}$). This is because the Von Mises distribution only has a
+single peak and thus will underestimate $Z$ for multimodal similarity
+functions. However, it does more accurately estimate $Z$ for stimuli
+that the \naive{} model would overestimate ($\VMMSE{}$). This
+underestimation is severe enough that in most cases where it gets
+stuck on a local maxima, it produces the incorrect judgment. So, the
+correlation for \textit{correct} responses is actually improved
+($\VMcorr{}$) and would be near perfect if the correct diretion of
+rotation was initially chosen (see Figure \ref{fig:rotations}, top
+center).
 
-\paragraph{Nonparametric (Bayesian Quadrature) model}
+\paragraph{Nonparametric (Bayesian Quadrature)}
 
-The rotations computed by the nonparametric model were strongly
-correlated with the true rotations ($\BQcorr$, see Figure
-\ref{fig:rotations}, right panel). This is because the nonparametric
+The nonparametric model is much more accurate in choosing the correct
+hypothesis ($\BQME{}$), erring approximately equally for both types of
+stimuli. We note that this is very close to the 3.2\% error rate
+reported by \cite{Shepard1971}, though further experimentation is
+necessary to determine whether the types of errors people make align
+with those of the model.
+
+The average rotations computed by the nonparametric model were
+strongly correlated with the true rotations ($\BQcorr$, see Figure
+\ref{fig:rotations}, bottom right). This is because the nonparametric
 model does not get stuck as easily on local optima: it will continue
-rotating until it is confident that its estimate of $Z$ is accurate.
-
-Correspondingly, the nonparametric model is much more accurate in
-choosing the correct hypothesis ($\BQME{}$). As with the \naive{} and
-parametric models, it mostly errs on $h_1$ stimuli ($\BQMEsame$). This
-appears to mostly be the result of choosing the incorrect initial
-direction of rotation.
-
-Similarly, because the nonparametric model actively attempts to obtain
-an accurate estimate of $Z$, its estimate of $Z$ is indeed fairly
-accurate ($\BQMSE{}$). This is slightly higher than the nonparametric
-model, but is due to a few outliers. \TODO{If we again exclude stimuli
-  for which $Z$ is greatly overestimated ($Z\geq 0.3$), the
-  nonparametric model's error decreases significantly
-  ($\MSE{}=0.03$).}
+rotating until it is confident that its estimate of $Z$ is
+accurate. Indeed, its active attempts to obtain an accurate estimate
+of $Z$ are largely succeessful ($\BQMSE{}$). Moreover, it has the
+capacity to ``restart'' and can recover from initially rotating in the
+incorrect direction, as illustrated by the fewer number of extreme
+over-rotations in Figure \ref{fig:rotations} (top right).
 

### 2013-05-31 man/nips-2013.tex
Add Tom intro edits

@@
 
+One of the challenges of solving any computational problem is
+determining how best to use the available computing resources. For
+example, a computer can render complex graphics faster by recognizing
+that this kind of computation should be carried out by a specialized
+graphics processor. The same challenge arises in designing an
+intelligent agent: how should the agent make best use of its computing
+resources? Recent research on rational models of human cognition has
+provided insight into the nature of the computational problems that
+human beings need to solve (e.g., \cite{something}), but leaves open
+the question of how people allocate their resources in solving those
+problems. In this paper, we take a step towards addressing this
+question, applying rational analysis (in the spirit of
+\cite{marr,anderson, shepard}) to one aspect of human metacognition:
+the use of mental simulation.
+
 Consider the images on the left in Figure
@@
 visualize one object rotating until it is congruent with the other
-\cite{Shepard1971}. There is strong evidence for such \textit{mental
-  imagery} or \textit{mental simulation}: we can imagine
-three-dimensional objects in our minds and manipulate them, to a
-certain extent, as if they were real \cite{Kosslyn:2009tj}.  However,
-the use of mental simulation is predicated on determining appropriate
-parameters to give the simulation, and people's cognitive constraints
-may furthermore place limitations on the duration or precision of
-simulation. One hypothesis for how these issues are handled argues
-that people use a rational solution, meaning that it is optimal under
-existing constraints \cite{Lieder:2012wg,Vul:2009wy,Griffiths2012a},
-such as processing speed, memory, or perceptual acuity.
+\cite{Shepard1971}. There is strong evidence for such mental
+simulation: we can imagine three-dimensional objects in our minds and
+manipulate them, to a certain extent, as if they were real
+\cite{Kosslyn:2009tj}.  However, the use of mental simulation is
+predicated on determining appropriate parameters to give the
+simulation, analogous to determining exactly what computation should
+be passed to a graphics processor.  In the case of the classic mental
+rotation task, we might ask: How do people know which way to rotate
+the object?  When should one stop rotating and accept the hypothesis
+that the objects are different?
 
-In the case of the classic mental rotation task, we might ask: in what
-direction should the object be rotated?  What are the requirements for
-``congruence''? When should one stop rotating and accept the
-hypothesis that the objects are different? We investigate several
-computational solutions to this task and qualitatively compare them to
-general mental rotation results. Of particular interest is a model
-which actively reasons about the amount of evidence it has, and
-performs mental rotations until it reaches a certain level of
-confidence. We contrast this model with simpler approaches which are
-less computationally demanding, but which employ less flexible
-strategies.
+Recent work in cognitive science has shown how problems of allocating
+cognitive resources to solving computational problems can be analyzed
+using the methods of statistical decision theory
+\cite{Lieder:2012wg,Vul:2009wy}. We apply this ``rational
+metacognition'' approach to the problem of mental rotation. We
+investigate several computational solutions to this task and
+qualitatively compare them to human mental rotation performance. In
+particular, we argue that performing mental rotation can be framed as
+an integration problem, with the direction of rotation becoming a
+statistical design problem (or in machine learning parlance, a problem
+of active learning). We show that recent work on methods for Bayesian
+quadrature \cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}
+provides a way to solve this problem, outperforming several simpler
+heuristics for determining both the direction and extent of rotation.
 

### 2013-05-31 man/nips-2013.tex
Polished results section

@@
 
-% I think it would be compelling to add a second row to Equation 5
-% where you show the averages and s.e. at each rotation for many noisy
-% runs of the algorithm (possibly on different stimuli). If you can
-% show that the average is a linear function, in a way that looks like
-% the classic behavioral data, that will be compelling.
-
 \begin{figure}[t]
@@
   \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
-  \caption{\textbf{Model rotations.} Each subplot shows the
-    correspondence between the true angle of rotation ($R$) for $h_1$
-    stimuli and the amount of rotation performed by the model. Dotted
-    lines indicate a perfect 1:1 relationship. The \naive{} and
-    parametric models (left and center panels, respectively) are
-    identical; they show modest agreement until roughly
-    $\frac{\pi}{2}$, after which agreement decreases as they tend to
-    encounter local maxima. The nonparametric model (right panel)
-    exhibits a significant correlation between the true angle and the
-    amount the model rotates.}
+  \caption{\textbf{Model rotations.} Top: each subplot shows the
+    correspondence between the true angle of rotation ($R$) for
+    ``same'' pairs and the amount of rotation performed by the
+    model. Bottom: each subplot shows the models' mean rotations over
+    stimuli pairs that were judged correctly. Black dots correspond to
+    ``same'' pairs, and blue lines correspond to ``different'' pairs
+    (for which the mean rotation is constant across true rotations, as
+    the true rotation is undefined). Error bars/shaded regions
+    indicate one standard deviation, and the dotted lines indicate the
+    least-squares fit to the ``same'' pairs.}
   \label{fig:rotations}
@@
 
-We evaluated each model's performance on 20 randomly generated shapes
-which had between 3 and 6 vertices, inclusive, (e.g. $X_a$ in Figure
-\ref{fig:stimuli}). For each shape, we computed 18 ``different'' and
-18 ``same'' stimuli pairs, with $R$ spaced at $20^\prime$ increments
-between 0 and 360. ``Different'' pairs were created by reflecting
-$X_a$ across the $y$-axis and rotating it by $R$; for ``same'' pairs,
-$X_a$ was rotated by $R$. \TODO{need to mention that shapes were
-  chosen to be discriminable by the gold standard}
+We evaluated each model's performance on 20 shapes which had between 3
+and 6 vertices, inclusive, (e.g. $X_a$ in Figure
+\ref{fig:stimuli}). For each shape, we computed 18 ``same'' and 18
+``different'' stimuli pairs, with $R$ spaced at $20^\prime$ increments
+between 0 and 360, as in \cite{Shepard1971}. ``Same'' pairs were
+created by rotating $X_a$ by $R$; the same was true for ``different'',
+except that $X_a$ was also reflected across the $y$-axis. Each shape
+was generated randomly, but under the constraint that the
+``same''/``different'' pairs created from the shape could both be
+judged according to the ``gold standard''.
 
@@
 \begin{itemize}
-\item \textit{Responses}: how accurate is the model at choosing the
-  correct hypothesis? This is defined to be the mean error ($\ME{}$),
+\item \textit{Responses}: How accurate was the model at choosing the
+  correct hypothesis? This was defined to be the mean error ($\ME{}$),
   or fraction of times the model picks the incorrect hypothesis.
-\item \textit{Rotations}: for those ``same'' pairs which the model
-  gets correct, how correlated are the true angles of rotation with
-  the rotations performed by the model?  This is defined to be the
-  Pearson's correlation coefficient $\rho$ between $R$ and $\vert
-  \mathbf{R}\vert$. Figure \ref{fig:rotations} shows individual points
-  corresponding to true rotation vs. rotation by the model, for each
-  model.
-\item \textit{Estimates of $Z$}: how accurate is the model's estimate
-  of $Z$? This is defined to be the mean squared error ($\MSE{}$)
-  between the model's estimate of $Z$ and the ``gold standard'' value,
-  where the error has scale such that $\MSE{}=0$ indicates no error,
-  and $\MSE{}=1$ indicates maximum error. Figure \ref{fig:accuracy}
-  shows plots of the true (gold standard) value of $Z$ vs. the model's
-  estimate, for each model.
+\item \textit{Rotations}: For those ``same'' pairs which the model
+  judged correct, how correlated were the model's rotations with the
+  true angles of rotation?  We quantified this using the Pearson's
+  correlation coefficient $\rho$ for $R$ vs. $\vert
+  \mathbf{R}\vert$. The top row in Figure \ref{fig:rotations} shows
+  individual points corresponding to true rotations vs. rotations by
+  the models for all stimuli, including those judged incorrectly. The
+  bottom row shows mean model rotations, across only those stimuli
+  which were judged correctly. We structured the analysis in this way
+  to better compare to the results of \cite{Shepard1971}, which also
+  excluded incorrectly judged stimuli.
+\item \textit{Estimates of $Z$}: How accurate was the model's estimate
+  of $Z$? We defined this quantity as the mean squared error
+  ($\MSE{}$) between the model's estimate of $Z$ and the ``gold
+  standard'' value, where the error has scale such that $\MSE{}=0$
+  indicates no error and $\MSE{}=1$ indicates maximum error. Figure
+  \ref{fig:accuracy} shows plots of the true (``gold standard'') value
+  of $Z$ vs. the model's estimate, for each model.
 \end{itemize}
-
-\paragraph{Gold standard}
-
-The mean error of the gold standard model was $\GSME{}$, indicating
-that $Z$ is a sufficient approximation of $\int_R p(X_b\ \vert\
-X_R)p(R)\ \mathrm{d}R$ for these particular stimuli.
 
@@
 
-The \naive{} model's response error rate was better than chance with
+The \naive{} model's response error rate was better than chance, with
 $\NaiveME{}$. Closer inspection reveals that much of this comes from
-$h_0$ stimuli ($\NaiveMEdiff{}$), where the asymmetric linear
-interpolation gives an overestimate of $Z$ (e.g., in Figure
-\ref{fig:li}). There is some error from $h_1$ stimuli as well
-($\NaiveMEsame{}$), which happens when the hill climbing search finds
-a local maxima and terminates prematurely, thus resulting in a low
-estimate of $Z$.
+``different'' pairs ($\NaiveMEdiff{}$), where the asymmetric linear
+interpolation may give an overestimate of $Z$ (e.g. Figure
+\ref{fig:li}).  This intuition is confirmed by the model's accuracy
+($\NaiveMSE{}$) in estimating $Z$: as shown in the left panel of
+Figure \ref{fig:accuracy}, it overestimated for nearly all stimuli
+pairs.
 
-The correlation between the average number of rotations computed by
-the \naive{} model and the true angle of rotation was
-$\Naivecorr{}$. The shape of the data is more interesting (see Figure
-\ref{fig:rotations}, top left panel): the \naive{} model actually
-corresponds extremeley well to the true angle of rotation for
-$R<\frac{\pi}{2}$ ($\Naivecorrsm{}$). This is unsurprising, because
-the closer the true angle is to zero, the less the model has to
-rotate, and the less likely it will get stuck on local maxima. Thus,
-it is more likely to locate the global maximum, which corresponds to
-the true angle of rotation. For $R>\frac{\pi}{2}$, we see an
-increasing tendency to under-rotate due to getting stuck on local
-maxima, as well as a tendency to over-rotate if the wrong rotation
-direction was chosen.
-
-The \naive{} model's error on $h_0$ stimuli is reflected in its
-accuracy in estimating $Z$ ($\NaiveMSE{}$). Figure \ref{fig:accuracy}
-(left panel) shows individual estimates for each stimulus, color-coded
-by true hypothesis. The model overestimates $Z$ for most of the $h_0$
-and $h_1$ stimuli, and underestimates several of the $h_1$ stimuli as
-well.
+The correlation between the \naive{} model's average rotation and the
+true angle of rotation was $\Naivecorr{}$ (Figure \ref{fig:rotations},
+bottom left). More complex patterns are revealed by examining the
+shape of the raw data in Figure \ref{fig:rotations}, top left: the
+\naive{} model corresponds extremeley well to the true angle of
+rotation for approximately $R<\frac{\pi}{2}$. This is unsurprising,
+because the closer the true angle is to zero, the less the model has
+to rotate, and the less likely it will get stuck on local
+maxima. Thus, it is more likely to locate the global maximum, which
+corresponds to the true angle of rotation. For $R>\frac{\pi}{2}$, we
+see an increasing tendency to under-rotate due to getting stuck on
+local maxima, as well as a tendency to over-rotate if the wrong
+direction was initially chosen.
 
@@
 
-The parametric model is also fairly inaccurate ($\VMME{}$), making
-more errors on $h_1$ stimuli ($\VMMEsame{}$) than $h_0$ stimuli
-($\VMMEdiff{}$). This is because the Von Mises distribution only has a
-single peak and thus will underestimate $Z$ for multimodal similarity
-functions. However, it does more accurately estimate $Z$ for stimuli
-that the \naive{} model would overestimate ($\VMMSE{}$). This
-underestimation is severe enough that in most cases where it gets
-stuck on a local maxima, it produces the incorrect judgment. So, the
-correlation for \textit{correct} responses is actually improved
-($\VMcorr{}$) and would be near perfect if the correct diretion of
-rotation was initially chosen (see Figure \ref{fig:rotations}, top
-center).
+The parametric model's error in determining whether the objects were
+identical was $\VMME{}$, making more errors on ``same'' pairs
+($\VMMEsame{}$) than ``different'' pairs ($\VMMEdiff{}$). Rather than
+overestimating as with the \naive{} model, however, the parametric
+model largely underestimated $Z$. This is unsurprising: because the
+Von Mises distribution only has a single peak, it necessarily
+underestimates $Z$ for multimodal similarity functions. Overall, the
+parametric model had an error rate of $\VMMSE{}$ in estimating $Z$,
+which was significantly more accurate than that of the \naive{} model.
 
-\paragraph{Nonparametric (Bayesian Quadrature)}
-
-The nonparametric model is much more accurate in choosing the correct
-hypothesis ($\BQME{}$), erring approximately equally for both types of
-stimuli. We note that this is very close to the 3.2\% error rate
-reported by \cite{Shepard1971}, though further experimentation is
-necessary to determine whether the types of errors people make align
-with those of the model.
-
-The average rotations computed by the nonparametric model were
-strongly correlated with the true rotations ($\BQcorr$, see Figure
-\ref{fig:rotations}, bottom right). This is because the nonparametric
-model does not get stuck as easily on local optima: it will continue
-rotating until it is confident that its estimate of $Z$ is
-accurate. Indeed, its active attempts to obtain an accurate estimate
-of $Z$ are largely succeessful ($\BQMSE{}$). Moreover, it has the
-capacity to ``restart'' and can recover from initially rotating in the
-incorrect direction, as illustrated by the fewer number of extreme
-over-rotations in Figure \ref{fig:rotations} (top right).
+The correlation between the parametric model's average rotation and
+the true angle of rotation was also higher ($\VMcorr{}$), though this
+is because more of the underestimated stimuli pairs were excluded from
+the analysis. The individual rotations performed by the parametric and
+\naive{} models were actually identical (see Figure
+\ref{fig:rotations}, upper left and center) because both models use
+the same hill-climbing stopping strategy.
 
@@
     the model. Black dotted lines indicate a perfect 1:1
-    correspondence between the true and estimated values. All models
-    perform well for $h_0$ stimuli, but exhibit different behavior for
-    $h_1$ stimuli. The \naive{} model (left panel) tends to either
-    greatly overestimate, or underestimate. The parametric model
-    (center panel) is very accurate for about half the $h_1$ stimuli,
+    correspondence between the true and estimated values. The \naive{}
+    model (left panel) tends to overestimate. The parametric model
+    (center panel) is very accurate for about half the ``same'' pairs,
     and severely underestimates the rest. The nonparametric model
-    (right panel) maintains a decent correspondence with the true $Z$,
-    with the exception of a handful of outliers.}
+    (right panel) maintains a decent correspondence with the true
+    $Z$.}
   \label{fig:accuracy}
 \end{figure}
+
+\paragraph{Nonparametric (Bayesian Quadrature)}
+
+The nonparametric model was much more accurate in choosing the correct
+hypothesis than the other two models ($\BQME{}$). We note that this is
+very close to the 3.2\% error rate reported by \cite{Shepard1971},
+though further experimentation is necessary to determine whether the
+types of errors people make align with those of the model.
+
+The average rotations computed by the nonparametric model were
+strongly correlated with the true rotations ($\BQcorr$, see Figure
+\ref{fig:rotations}, bottom right). Because the nonparametric model
+has the capacity to ``restart'', it could recover from rotating in the
+incorrect direction (e.g. Figure \ref{fig:bq}) and thus did not
+over-rotate as frequently.  It also under-rotated less frequently:
+because the model continues to rotate until it is confident that its
+estimate of $Z$ is accurate, it does not get stuck as easily on local
+optima.  Indeed, its active attempts to obtain an accurate estimate of
+$Z$ were largely succeessful, with $\BQMSE{}$.
 

### 2013-05-31 man/nips-2013.tex
Update model sections

@@
 provided insight into the nature of the computational problems that
-human beings need to solve (e.g., \cite{something}), but leaves open
+human beings need to solve (e.g. \cite{something}), but leaves open
 the question of how people allocate their resources in solving those
@@
 coordinates of the vertices of two-dimensional shapes similar to those
-used by \cite{Cooper:1975wp} (e.g., see Figure
+used by \cite{Cooper:1975wp} (e.g. Figure
 \ref{fig:stimuli}). Participants must determine whether $X_a$ and
@@
 We can formulate the judgment of whether $X_a$ and $X_b$ have the same
-origins by deciding about two hypotheses:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  $h_0$: $\forall M,R\ X_b\neq MRX_a$
-\item
-  $h_1$: $\exists M,R\textrm{ s.t. } X_b=MRX_a$
-\end{itemize}
-
-To compare the hypotheses, we need to compute the posterior:
+origins by deciding about two hypotheses, $h_0$: $\forall M,R\ X_b\neq
+MRX_a$ and $h_1$: $\exists M,R\textrm{ s.t. } X_b=MRX_a$.  To compare
+the hypotheses, we need to compute the posterior for each:
 \begin{equation}
-p(h_i\ \vert\ X_a, X_b)\propto p(X_a, X_b\ \vert\ h_i)p(h_i)
+p(h\ \vert\ X_a, X_b)\propto p(X_a, X_b\ \vert\ h)p(h)
 \end{equation}
 Assuming the hypotheses are equally likely \textit{a priori}, the
-prior term $p(h_i)$ will cancel out when comparing $h_0$ and $h_1$,
-thus allowing us to focus on the likelihoods: 
+prior term $p(h)$ will cancel out when comparing $h_0$ and $h_1$, thus
+allowing us to focus on the likelihoods:
 \begin{align}
@@
 Under $h_0$, the likelihood is easy to compute because we assume that
-$X_a$ and $X_b$ are independent (Equation \ref{eq:lh-h0}). However,
-when the configurations are the same, the likelihood becomes more
-complicated (Equation \ref{eq:lh-h1}). For a small number of vertices,
-we can compute the integral over $M$ by enumerating every possible
-mapping between $X_a$ and $X_b$. After doing so, we obtain:
+$X_a$ and $X_b$ are independent (Equation \ref{eq:lh-h0}). When the
+configurations are the same, the likelihood becomes more complicated
+(Equation \ref{eq:lh-h1}). For a small number of vertices, we can
+compute the integral over $M$ by enumerating every possible mapping
+between $X_a$ and $X_b$. After doing so, we obtain:
 \begin{equation} 
@@
 
-However, we cannot compute $p(X_b\vert X_a, R)$ directly. Instead, we
-introduce a new variable $X_R$ denoting a mental image, which
-approximates $RX_a$. The $X_R$ are generated sequentially by repeated
-application of a function $\tau$:
-\begin{align}
-  X_R&=RX_a\nonumber \\
-  &=\tau(X_{R-r}, r)\nonumber \\
-  &=\tau(\tau(X_{R-2r}, r), r)\nonumber \\
-  &\ldots{}\nonumber \\
-  &=\tau^{(\frac{R}{r})}(X_a, r)
-  \label{eq:tau}
-\end{align} 
-% I wasn't clear whether the step that happens in Equation 6 should be
-% in the model, or in the algorithm. I would probably be inclined to
-% define the model in full generality, and then bring up the
-% increments as part of the algorithm.
-Where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
-applications of $\tau$. Using this sequential function, we get:
-\begin{align}
-  p(X_a, X_b\ \vert \ h_1)&=\int_R \int_{X} p(X_b\vert X) p(X\vert X_a, R)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
-  x&= \int_R \int_X p(X_b\vert X)\delta(\tau^{(\frac{R}{r})}(X_a, r)-X)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
-  &= \int_R p(X_b\vert X_R)p(X_a)p(R)\ \mathrm{d}R
-\end{align}
-
 Once we have computed both likelihoods, we compute their ratio:
 \begin{equation}
-  \ell=\frac{p(X_a, X_b\ \vert \ h_1)}{p(X_a, X_b\ \vert \ h_0)}=\frac{\int_R p(X_b\ \vert\ X_R)p(R)\ \mathrm{d}R}{p(X_b)}
+  \ell=\frac{p(X_a, X_b\ \vert \ h_1)}{p(X_a, X_b\ \vert \ h_0)}=\frac{\int_R p(X_b\ \vert\ X_a, R)p(R)\ \mathrm{d}R}{p(X_b)}
   \label{eq:lh-ratio}
@@
     An example stimulus in which the shapes differ only by a
-    rotation. All stimuli consist of five vertices centered around the
-    origin, and four edges which create a closed loop from the
+    rotation. All stimuli consist of three to six vertices centered
+    around the origin, and edges which create a closed loop from the
     vertices. \textbf{(b)} The approximate likelihood (similarity) of
     $X_b$ given $X_R$, where $X_R$ is a rotation of $X_a$ by the angle
-    $R$. The true angle of rotation between $X_a$ and $X_b$ is
-    slightly less than $\pi$, corresponding to the global maximum in
-    the likelihood function.}
+    $R$. The true angle of rotation between $X_a$ and $X_b$ is at
+    $\frac{2\pi}{3}$, corresponding to the global maximum in the
+    likelihood function.}
+  \label{fig:shapes}
 \end{figure}
 
-We define the prior probabilities over stimuli based on how they are
-generated. For shapes with $n$ vertices, each vertex is at a random
-angle with a radius random chosen between 0 and 1, in polar
-coordinates, and could be chosen in any of $n!$ different ways. Thus,
-the prior over a shape $X$ is:
+We define the prior probabilities over stimuli according to a
+generative procedure. A set of $n$ vertices could be chosen in any of
+$n!$ different ways, and each vertex is located at a random angle
+(between 0 and $2\pi$) and radius (between 0 and 1). Thus, the prior
+over a shape $X$ is:
 \begin{equation}
@@
 which gives us the denominator in Equation \ref{eq:lh-ratio}.
-
-Computing the numerator of Equation \ref{eq:lh-ratio} is more
-difficult, as we do know $p(X_b\vert X_R)$. We approximate it with a
-similarity function $S(X_b, X_R)$, which also takes into account the
-different possible mappings of vertices:
-\begin{equation}
-  Z=\int_R S(X_b, X_R)p(R)\ \mathrm{d}R\ \approx\int_Rp(X_b\ \vert\ X_R)p(R)\ \mathrm{d}R
-  \label{eq:Z}
-\end{equation}
-
-Because the vertices are connected in a way which forms a closed loop,
-we need only consider $2n$ mappings, $M$, of the $n$ vertices (we
-assume uncertainty for which is the ``first'' vertex, and then which
-of its two neighbors is the ``second''). Thus, the possible orderings
-are of the form $M=\lbrace{}0, 1, \ldots{}, n\rbrace{}$,
-$M=\lbrace{}n, 0, \ldots{}, n-1\rbrace{}$, and so on. This gives us an
-explicit form for the similarity function:
+Computing the numerator is more difficult, as we we cannot compute
+$p(X_b\vert X_a, R)$ directly. Instead, we introduce a new variable
+$X_R$ denoting a mental image, which approximates $RX_a$. The $X_R$
+are generated sequentially by repeated application of a function
+$\tau$:
+\begin{align}
+  X_R=RX_a&=\tau(X_{R-r}, r)=\tau(\tau(X_{R-2r}, r), r)=\ldots{}\nonumber \\
+  &=\tau^{(\frac{R}{r})}(X_a, r)
+  \label{eq:tau}
+\end{align} 
+% I wasn't clear whether the step that happens in Equation 6 should be
+% in the model, or in the algorithm. I would probably be inclined to
+% define the model in full generality, and then bring up the
+% increments as part of the algorithm.
+Where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
+applications of $\tau$. Using this sequential function, we get:
+\begin{align}
+  p(X_a, X_b\ \vert \ h_1)&=\int_R \int_{X} p(X_b\vert X) p(X\vert X_a, R)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
+  &= \int_R \int_X p(X_b\vert X)\delta(\tau^{(\frac{R}{r})}(X_a, r)-X)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
+  &= \int_R p(X_b\vert X_R)p(X_a)p(R)\ \mathrm{d}R
+\end{align}
+However, the exact form of $p(X_b\vert X_R)$ is still unknown. We
+approximate it with a similarity function $S(X_b, X_R)$ which takes
+into account the different possible mappings of vertices. Because the
+vertices are connected in a way which forms a closed loop, we need
+only consider $2n$ mappings of the $n$ vertices (we assume uncertainty
+for which is the ``first'' vertex, and then which of its two neighbors
+is the ``second''). So, the possible orderings are of the form
+$M=\lbrace{}0, 1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0, \ldots{},
+n-1\rbrace{}$, and so on. Combining this with a Gaussian similarity
+metric, we obtain:
 \begin{equation}
@@
 \end{equation}
-where $i$ denotes the $i^{th}$ vertex. An example of $S$ for the
-stimuli shown in Figure \ref{fig:stimuli} is illustrated in Figure
-\ref{fig:likelihood}.
+where $i$ denotes the $i^{th}$ vertex. An example stimulus and
+corresponding $S$ is shown in Figure \ref{fig:shapes}.
 
-This process of mental rotation (i.e., generating a single $X_R$ (as
-in Equation \ref{eq:tau}) and then computing $S(X_b, X_R)$) is
-computationally demanding. Thus, our goal is to minimize the number of
-rotations while still obtaining an estimate of $Z$ that is accurate
-enough to choose the correct hypothesis. With this in mind, we now
-examine several approaches to estimating $Z$.
-
-In each of these models, we denote the computed rotations to be a set
-$\mathbf{R}=\{R_1, R_2, \ldots{}\}$.
+To summarize, the process of generating a mental image consists of
+computing a single $X_R$ (as in Equation \ref{eq:tau}) and then
+computing $S(X_b, X_R)$. We denote the sequence of rotations computed
+by this procedure as $\mathbf{R}=\{R_1, R_2, \ldots{}\}$. However,
+this sequence cannot be arbitrary, as mental rotation is
+computationally demanding. Our goal is to minimize the number of
+rotations $\vert\mathbf{R}\vert$ while still obtaining an estimate of
+$Z$ that is accurate enough to choose the correct hypothesis. With
+this in mind, we now examine several approaches to solve this problem.
 
@@
 To compare the accuracy of other models' estimates of $Z$, we computed
-a ``gold standard''\footnote{This only gives an accurate estimate of
-  $Z$, which is itself an approximation, and is thus not necessarily
-  the true value of the numerator in Equation \ref{eq:lh-ratio}.} by
-evaluating $S(X_b, X_R)$ at 360 values of $R$ spaced evenly between
-$0$ and $2\pi$.
+a ``gold standard'' by evaluating $S(X_b, X_R)$ at 360 values of $R$
+spaced evenly between $0$ and $2\pi$ and estimating the integral using
+the trapezoidal rule.
 
@@
     low, but certain angles will have a disproportionate contribution
-    (e.g., between $\frac{\pi}{4}$ and $\frac{3\pi}{4}$. \textbf{(b)}
+    (e.g. between $\frac{\pi}{4}$ and $\frac{3\pi}{4}$. \textbf{(b)}
     Best fit of scaled Von Mises PDF parameters. As in (a), $Z$ will
@@
 computes an estimate of $Z$ by linearly interpolating between sampled
-rotations. Figure \ref{fig:li} shows an example of the \naive{} model
-prior to estimating $Z$.
+rotations. Figure \ref{fig:li} shows an example of the \naive{}
+model's estimate of $S$.
 
@@
 Another strategy is to assume a parametric shape for $S$ and fit the
-appropriate parameters. When $h_1$ is correct, it is likely that the
-function will be approximately unimodal (shapes with rotational
-symmetry would be multimodal). A reasonable assumption, then, is that
-the likelihood follows a ``wrapped Gaussian'', or Von Mises,
-distribution:
+appropriate parameters. If we assume that $S$ is approximately
+unimodal when $h_1$ is true\footnote{This is not always a reasonable
+  assumption, see e.g. Figure \ref{fig:vm}}., then a reasonable form
+for the similarity function is that of a ``circular Gaussian'' or Von
+Mises distribution:
 \begin{equation}
@@
 hill-climbing strategy as in the \naive{} model. Figure \ref{fig:vm}
-shows an example of this parametric model, prior to estimating $Z$.
+shows an example of the parametric model's estimate of $S$.
 
@@
 evaluating $S$ at a particular set of points. Because our data is
-circular, we can use a periodic kernel \cite{Rasmussen:2006vz}:
+circular, we use a periodic kernel \cite{Rasmussen:2006vz}:
 \begin{equation}
@@
 such constraint. In an effort to avoid this problem,
-\cite{Osborne:2012tm} give a method which involves instead placing a
-prior over the log likelihood\footnote{In practice, as in
-  \cite{Osborne:2012tm}, we use the transform of
-  $\log(S+1)$. Additionally, while \cite{Osborne:2012tm} use a
-  combination of MLII and marginalization to fit the kernel
-  parameters, we set the output scale $h_S=\sqrt{0.1}$ and $h_{\log
-    S}=\log(h_S + 1)$, and use MLII to fit all other parameters.},
-thus ensuring that $S=e^{\log S}$ will be positive\footnote{We are not
+\cite{Osborne:2012tm} give a method to place a prior over the log
+likelihood\footnote{In practice, as in \cite{Osborne:2012tm}, we use
+  the transform of $\log(S+1)$. Additionally, we follow
+  \cite{Osborne:2012tm} and use a combination of MLII and
+  marginalization to fit the $w$ kernel parameters. The output scales
+  are fixed at $h_S=\sqrt{0.1}$ and $h_{\log S}=\log(h_S + 1)$.}, thus
+ensuring that $S=e^{\log S}$ will be positive\footnote{We are not
   guaranteed positivity, however, the approximation to the integral
@@
 consists of $\mathbf{R}$ and a set of intermediate \emph{candidate
-  points}\footnote{These candidate points do not require evaluating
-  the true $S$, only the GP estimates of $S$ and $\log S$.} $c$ as
-described in \cite{Osborne:2012tm}. The variance of the estimate of
-$Z$ is given by:
-\begin{equation}
-  \sigma_Z=\mathrm{Var}(Z\ \vert \ S, \log S, \Delta_c) = \int_R\int_{R^\prime} \mu_S(R)\mu_S(R^\prime) \Sigma_{\log S}(R, R^\prime)p(R)p(R^\prime)\ \mathrm{d}R\ \mathrm{d}R^\prime
-  \label{eq:bq-Z-var}
-\end{equation}
+  points} $c$ as described in \cite{Osborne:2012tm}. The variance is
+$\tilde{V}(Z\vert S, \log S, \Delta_c)$ as defined in Equation 12 of
+\cite{Osborne:2012tm}.
 
-\TODO{To choose $\mathbf{R}$}, we pick an initial direction of
-rotation which results in the higher value of $S$, and then continue
-rotating in that direction until the variance in Equation
-\ref{eq:bq-Z-var} is low enough that we are confident about the
-likelihood ratio $\ell$:
+To start, we pick an initial direction of rotation which results in
+the higher value of $S$. At each step, we compute $\mu_Z$ and
+$\tilde{V}$ in order to estimate a distribution over the likelihood
+ratio $\ell$:
 \begin{equation*}
@@
 \end{equation*}
-Specifically, we choose $h_0$ when $p(\ell < 1)\geq 0.95$, and we
-choose $h_1$ when $p(\ell > 1)\geq 0.95$. Until one of these
-conditions are met (or the shape has been fully rotated), the model
-will continue to compute rotations and update its estimate of $Z$.
+We choose $h_0$ when $p(\ell < 1)\geq 0.95$, and $h_1$ when $p(\ell >
+1)\geq 0.95$. Until one of these conditions are met (or the shape has
+been fully rotated), the model will continues to compute rotations and
+update its estimate of $Z$.
+
+We additionally allow the model to change direction or ``reset'' based
+on an estimate of the posterior variance of $Z$ given some new sample
+$a$. This is similar to the procedure given in \cite{Osborne:2012tm},
+however we do not compute the full posterior variance. Instead, we
+compute the variance given only the current mean estimate of $a$. If
+this estimated variance is lowered more by ``resetting'', then the
+model will change directions. Thus, it is able to have an active say
+in which mental rotations are computed, unlike the hill-climbing
+procedure.
+
+% \begin{equation}
+%   \sigma_Z=\mathrm{Var}(Z\ \vert \ S, \log S, \Delta_c) = \int_R\int_{R^\prime} \mu_S(R)\mu_S(R^\prime) \Sigma_{\log S}(R, R^\prime)p(R)p(R^\prime)\ \mathrm{d}R\ \mathrm{d}R^\prime
+%   \label{eq:bq-Z-var}
+% \end{equation}
 

### 2013-05-31 man/nips-2013.tex
Add some todos, include Tom's feedback

@@
 \begin{itemize}
-\item \textit{Responses}: How accurate was the model at choosing the
-  correct hypothesis? This was defined to be the mean error ($\ME{}$),
-  or fraction of times the model picks the incorrect hypothesis.
+\item \textit{Response error rates}: How accurate was the model at
+  choosing the correct hypothesis? This was defined to be the mean
+  error ($\ME{}$), or fraction of times the model picks the incorrect
+  hypothesis.
 \item \textit{Rotations}: For those ``same'' pairs which the model
@@
   true angles of rotation?  We quantified this using the Pearson's
-  correlation coefficient $\rho$ for $R$ vs. $\vert
-  \mathbf{R}\vert$. The top row in Figure \ref{fig:rotations} shows
-  individual points corresponding to true rotations vs. rotations by
-  the models for all stimuli, including those judged incorrectly. The
-  bottom row shows mean model rotations, across only those stimuli
-  which were judged correctly. We structured the analysis in this way
-  to better compare to the results of \cite{Shepard1971}, which also
-  excluded incorrectly judged stimuli.
+  correlation coefficient $\rho$ for the true rotation ($R$) versus
+  the extent of the model's rotations ($\vert \mathbf{R}\vert$). The
+  top row in Figure \ref{fig:rotations} shows individual points
+  corresponding to true rotations vs. rotations by the models for all
+  stimuli, including those judged incorrectly. The bottom row shows
+  mean model rotations, across only those stimuli which were judged
+  correctly. We structured the analysis in this way to better compare
+  to the results of \cite{Shepard1971}, which also excluded
+  incorrectly judged stimuli.
 \item \textit{Estimates of $Z$}: How accurate was the model's estimate
@@
 
-The \naive{} model's response error rate was better than chance, with
-$\NaiveME{}$. Closer inspection reveals that much of this comes from
-``different'' pairs ($\NaiveMEdiff{}$), where the asymmetric linear
-interpolation may give an overestimate of $Z$ (e.g. Figure
-\ref{fig:li}).  This intuition is confirmed by the model's accuracy
-($\NaiveMSE{}$) in estimating $Z$: as shown in the left panel of
-Figure \ref{fig:accuracy}, it overestimated for nearly all stimuli
-pairs.
+The \naive{} model's response error rate was $\NaiveME{}$, which is
+better than chance (equivalent to guessing randomly,
+i.e. $\ME{}=0.5$). \TODO{binomial test?} Closer inspection reveals
+that much of this comes from ``different'' pairs ($\NaiveMEdiff{}$),
+where the asymmetric linear interpolation may give an overestimate of
+$Z$ (e.g. Figure \ref{fig:li}).  This intuition is confirmed by the
+model's accuracy in estimating $Z$, which was $\NaiveMSE{}$. As shown
+in the left panel of Figure \ref{fig:accuracy}, it overestimated $Z$
+for nearly all stimuli pairs. \TODO{how many?}
 
@@
 identical was $\VMME{}$, making more errors on ``same'' pairs
-($\VMMEsame{}$) than ``different'' pairs ($\VMMEdiff{}$). Rather than
-overestimating as with the \naive{} model, however, the parametric
-model largely underestimated $Z$. This is unsurprising: because the
-Von Mises distribution only has a single peak, it necessarily
-underestimates $Z$ for multimodal similarity functions. Overall, the
-parametric model had an error rate of $\VMMSE{}$ in estimating $Z$,
-which was significantly more accurate than that of the \naive{} model.
+($\VMMEsame{}$) than ``different'' pairs ($\VMMEdiff{}$). \TODO{is
+  this significant?} Rather than overestimating as with the \naive{}
+model, however, the parametric model largely underestimated $Z$. This
+is unsurprising: because the Von Mises distribution only has a single
+peak, it necessarily underestimates $Z$ for multimodal similarity
+functions. Overall, the parametric model had an error rate of
+$\VMMSE{}$ in estimating $Z$, which was significantly more accurate
+than that of the \naive{} model \TODO{test}.
 
@@
 The nonparametric model was much more accurate in choosing the correct
-hypothesis than the other two models ($\BQME{}$). We note that this is
-very close to the 3.2\% error rate reported by \cite{Shepard1971},
-though further experimentation is necessary to determine whether the
-types of errors people make align with those of the model.
+hypothesis than the other two models ($\BQME{}$) \TODO{test}. We note
+that this is very close to the 3.2\% error rate reported by
+\cite{Shepard1971}, though further experimentation is necessary to
+determine whether the types of errors people make align with those of
+the model.
 
@@
 \ref{fig:rotations}, bottom right). Because the nonparametric model
-has the capacity to ``restart'', it could recover from rotating in the
+has the capacity to ``reset'', it could recover from rotating in the
 incorrect direction (e.g. Figure \ref{fig:bq}) and thus did not

### 2013-05-31 man/nips-2013.tex
Uncomment abstract, so I don't forget!
Make anonymous
Capitalize title
Equation formatting

@@
 
-\title{Mental simulation as Bayesian Quadrature}
+\title{Mental Simulation as Bayesian Quadrature}
 \author{Jessica B.~Hamrick and Thomas L.~Griffiths\\
@@
 
-\nipsfinalcopy % Uncomment for camera-ready version
+%\nipsfinalcopy % Uncomment for camera-ready version
 
@@
 
-% \begin{abstract}
-% \TODO{}
-% \end{abstract}
+\begin{abstract}
+\TODO{}
+\end{abstract}
 
@@
 \begin{equation}
-p(h\ \vert\ X_a, X_b)\propto p(X_a, X_b\ \vert\ h)p(h)
+p(h\ \vert\ X_a, X_b)\propto p(X_a, X_b\ \vert\ h)p(h).
 \end{equation}
@@
   p(X_a,\ X_b\ \vert \ h_0)&=p(X_a)p(X_b) \label{eq:lh-h0}\\
-  p(X_a,\ X_b\ \vert \ h_1)&=\int_R\int_M p(X_a) p(X_b\vert X_a,R,M) p(R) p(M)\ \mathrm{d}M\ \mathrm{d}R \label{eq:lh-h1}
+  p(X_a,\ X_b\ \vert \ h_1)&=\int_R\int_M p(X_a) p(X_b\vert X_a,R,M) p(R) p(M)\ \mathrm{d}M\ \mathrm{d}R. \label{eq:lh-h1}
 \end{align}
@@
 \begin{equation} 
-  p(X_a,\ X_b\ \vert \ h_1)=\int_R p(X_a) p(X_b\vert X_a,R) p(R)\ \mathrm{d}R
+  p(X_a,\ X_b\ \vert \ h_1)=\int_R p(X_a) p(X_b\vert X_a,R) p(R)\ \mathrm{d}R.
 \end{equation}
@@
 \begin{equation}
-  \ell=\frac{p(X_a, X_b\ \vert \ h_1)}{p(X_a, X_b\ \vert \ h_0)}=\frac{\int_R p(X_b\ \vert\ X_a, R)p(R)\ \mathrm{d}R}{p(X_b)}
+  \ell=\frac{p(X_a, X_b\ \vert \ h_1)}{p(X_a, X_b\ \vert \ h_0)}=\frac{\int_R p(X_b\ \vert\ X_a, R)p(R)\ \mathrm{d}R}{p(X_b)}.
   \label{eq:lh-ratio}
@@
 (between 0 and $2\pi$) and radius (between 0 and 1). Thus, the prior
-over a shape $X$ is:
-\begin{equation}
-  p(X)=n!\left(\frac{1}{2\pi}\right)^n
-  \label{eq:prior}
-\end{equation} 
-which gives us the denominator in Equation \ref{eq:lh-ratio}.
-Computing the numerator is more difficult, as we we cannot compute
-$p(X_b\vert X_a, R)$ directly. Instead, we introduce a new variable
-$X_R$ denoting a mental image, which approximates $RX_a$. The $X_R$
-are generated sequentially by repeated application of a function
-$\tau$:
+over a shape $X$ is $p(X)=n!\left(\frac{1}{2\pi}\right)^n$, which
+gives us the denominator in Equation \ref{eq:lh-ratio}. Computing the
+numerator is more difficult, as we we cannot compute $p(X_b\vert X_a,
+R)$ directly. Instead, we introduce a new variable $X_R$ denoting a
+mental image, which approximates $RX_a$. The $X_R$ are generated
+sequentially by repeated application of a function $\tau$:
 \begin{align}
   X_R=RX_a&=\tau(X_{R-r}, r)=\tau(\tau(X_{R-2r}, r), r)=\ldots{}\nonumber \\
-  &=\tau^{(\frac{R}{r})}(X_a, r)
+  &=\tau^{(\frac{R}{r})}(X_a, r).
   \label{eq:tau}
@@
 unimodal when $h_1$ is true\footnote{This is not always a reasonable
-  assumption, see e.g. Figure \ref{fig:vm}}., then a reasonable form
+  assumption, see e.g. Figure \ref{fig:vm}.}, then a reasonable form
 for the similarity function is that of a ``circular Gaussian'' or Von
@@
 \begin{equation}
-k(R, R^\prime)=h^2\exp\left(-\frac{2\sin^2\left(\frac{1}{2}(R-R^\prime)\right)}{w^2}\right)
+k(R, R^\prime)=h^2\exp\left(-\frac{2\sin^2\left(\frac{1}{2}(R-R^\prime)\right)}{w^2}\right).
 \end{equation}
@@
 \begin{equation*}
-p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\ \vert\ \mu_Z, \sigma_z)
+p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\ \vert\ \mu_Z, \sigma_z).
 \end{equation*}

### 2013-05-31 man/nips-2013.tex
Rename implementation section
Define Z

@@
 
-\section{Implementation}
+\section{Metacognitive algorithms}
 
@@
 R)$ directly. Instead, we introduce a new variable $X_R$ denoting a
-mental image, which approximates $RX_a$. The $X_R$ are generated
-sequentially by repeated application of a function $\tau$:
+mental image, which approximates $RX_a$. The $X_R$ are generated by
+repeated application of a function $\tau$:
 \begin{align}
@@
 \end{align} 
-% I wasn't clear whether the step that happens in Equation 6 should be
-% in the model, or in the algorithm. I would probably be inclined to
-% define the model in full generality, and then bring up the
-% increments as part of the algorithm.
-Where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
+where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
 applications of $\tau$. Using this sequential function, we get:
@@
 However, the exact form of $p(X_b\vert X_R)$ is still unknown. We
-approximate it with a similarity function $S(X_b, X_R)$ which takes
-into account the different possible mappings of vertices. Because the
-vertices are connected in a way which forms a closed loop, we need
-only consider $2n$ mappings of the $n$ vertices (we assume uncertainty
-for which is the ``first'' vertex, and then which of its two neighbors
-is the ``second''). So, the possible orderings are of the form
-$M=\lbrace{}0, 1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0, \ldots{},
-n-1\rbrace{}$, and so on. Combining this with a Gaussian similarity
-metric, we obtain:
+approximate it with a similarity function $S(X_b, X_R)$, and denote
+the resulting integral as $Z$:
+\begin{equation}
+Z=\int_R S(X_b, X_R)p(R)\ \mathrm{d}R\approx \int_R p(X_b\vert X_R)p(R)\ \mathrm{d}R.
+\label{eq:Z}
+\end{equation}
+We define the similarity function $S$, which incorporates the
+different mappings as vertices, as follows. Because the vertices are
+connected in a way which forms a closed loop, we need only consider
+$2n$ mappings of the $n$ vertices (we assume uncertainty for which is
+the ``first'' vertex, and then which of its two neighbors is the
+``second''). So, the possible orderings are of the form $M=\lbrace{}0,
+1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0, \ldots{}, n-1\rbrace{}$,
+and so on. Combining this with a Gaussian similarity metric, we
+obtain:
 \begin{equation}

### 2013-05-31 man/nips-2013.tex
Define computational level

@@
 
-People are presented with two images, $X_a$ and $X_b$, which are the
+We begin by first analyzing mental rotation at Marr's
+\textit{computational} level \cite{Marr:1983to}: what is the problem
+to be solved, and what is the optimal way to do so?  Formally, people
+are presented with two images, $X_a$ and $X_b$, which are the
 coordinates of the vertices of two-dimensional shapes similar to those

### 2013-05-31 man/nips-2013.tex
Add intro sentence to metacognitive algorithms section
Citations

@@
 provided insight into the nature of the computational problems that
-human beings need to solve (e.g. \cite{something}), but leaves open
-the question of how people allocate their resources in solving those
-problems. In this paper, we take a step towards addressing this
-question, applying rational analysis (in the spirit of
-\cite{marr,anderson, shepard}) to one aspect of human metacognition:
-the use of mental simulation.
+human beings need to solve (e.g. \cite{Chater:1999wp,Griffiths2012a}),
+but leaves open the question of how people allocate their resources in
+solving those problems. In this paper, we take a step towards
+addressing this question, applying rational analysis (in the spirit of
+\cite{Marr:1983to,anderson90} \TODO{shepard?}) to one aspect of human
+metacognition: the use of mental simulation.
 
@@
 
-We define the prior probabilities over stimuli according to a
-generative procedure. A set of $n$ vertices could be chosen in any of
-$n!$ different ways, and each vertex is located at a random angle
-(between 0 and $2\pi$) and radius (between 0 and 1). Thus, the prior
-over a shape $X$ is $p(X)=n!\left(\frac{1}{2\pi}\right)^n$, which
-gives us the denominator in Equation \ref{eq:lh-ratio}. Computing the
-numerator is more difficult, as we we cannot compute $p(X_b\vert X_a,
-R)$ directly. Instead, we introduce a new variable $X_R$ denoting a
-mental image, which approximates $RX_a$. The $X_R$ are generated by
-repeated application of a function $\tau$:
+Given the computational model defined above, we now explore the ways
+in which it can be expressed algorithmically. First, we define the
+prior probabilities over stimuli according to a generative
+procedure. A set of $n$ vertices could be chosen in any of $n!$
+different ways, and each vertex is located at a random angle (between
+0 and $2\pi$) and radius (between 0 and 1). Thus, the prior over a
+shape $X$ is $p(X)=n!\left(\frac{1}{2\pi}\right)^n$, which gives us
+the denominator in Equation \ref{eq:lh-ratio}. Computing the numerator
+is more difficult, as we we cannot compute $p(X_b\vert X_a, R)$
+directly. Instead, we introduce a new variable $X_R$ denoting a mental
+image, which approximates $RX_a$. The $X_R$ are generated by repeated
+application of a function $\tau$:
 \begin{align}

### 2013-05-31 man/nips-2013.tex
Add citations, background, conclusion

@@
 
-% I think the main thing would be doing something more sophisticated
-% in the search, rather than using the gradient. A less sophisticated
-% adaptive sampling method seems like it would be best, and using von
-% Mises makes sense.
-
 \section{Introduction}
-
-\TODO{Want to talk about previous work on strategies people might use,
-  e.g. eye-tracking results from Just and Carpenter (1975?)}
-
-% 1. The big picture
-% 2. Our contribution
-% 3. How we do it
-% 4. What's in the paper
-
-% The introduction could be made more general -- start with the
-% question of how a rational agent should use her computational
-% resources, and then get onto mental simulation/rotation as a special
-% case of that, rather than starting with the phenomena. I would make
-% the big picture goals clear at the start, and then segue into the
-% mystery of mental rotation -- how do we know which way to rotate?
 
@@
 reasoning problems, and when they might use imagery as opposed to a
-symbolic rule \cite{Schwartz:1996uy}. People seem to use imagery
-methodically, which raises the question: what is the method?
+symbolic rule \cite{Schwartz:1996uy}. Eye-tracking studies of people
+performing mental rotations suggest strategies that people might be
+using, such as aligning smaller features of the images
+\cite{Just1976}. People seem to use imagery methodically, which raises
+the question: what is the method?
 
@@
 
-We investigated strategies for performing the mental rotation task
-\cite{Shepard1971} in two dimensions, and found that a nonparametric
-model that enforces a positivity constraint on the likelihood function
-performs best, as it is able to actively monitor the confidence of its
-estimate. The simpler models performed much worse: they did not
-maintain the linear relationship with the true angle of rotation due
-to lack of robustness against local maxima, and they were inaccurate
-at identifying whether the two shapes were the same or not.
+In this paper, we asked: how do people use mental simulation?  We
+investigated the specific case of mental rotation, taking a rational
+analysis approach to characterizing optimal strategies for performing
+mental rotation. We found that an adaptive, nonparametric model
+performs best at the classic rotation task \cite{Shepard1971}, as it
+is able to actively monitor the confidence of its estimate and
+intelligently choose the direction of rotation. Two simpler models
+based on heuristics performed much worse: they did not maintain the
+linear relationship with the true angle of rotation due to lack of
+robustness against local maxima or incorrect rotation direction. The
+nonparametric model qualitatively approximates human behavioral
+judgments, both in linearity \cite{Shepard1971} and variability
+\cite{Just1976}. Future work will collect empirical data from
+participants to perform a more thorough, quantitative analysis.  We
+conclude that from this initial survey, models of mental rotation
+which take an ``active'', directed approach seem well-suited to
+explaining human behavior in such metacognitive domains as mental
+simulation.
 
-One option for improving the performance of the simple models would be
-to fit a threshold value, below which local maxima would be
-ignored. However, this strategy would be rather brittle, for if the
-distribution of shapes changed, the threshold would have to be
-re-learned. It is possible that people exhibit this behavior, but we
-cannot make any assumptions one way or the other without empirical
-data.
+% One option for improving the performance of the simple models would be
+% to fit a threshold value, below which local maxima would be
+% ignored. However, this strategy would be rather brittle, for if the
+% distribution of shapes changed, the threshold would have to be
+% re-learned. It is possible that people exhibit this behavior, but we
+% cannot make any assumptions one way or the other without empirical
+% data.
 
-A further strength of the nonparametric model over the other models,
-however, is that it is likely to generalize well to three
-dimensions. One aspect of \cite{Osborne:2012tm} which we have not yet
-explored is their main contribution of \textit{active sampling}, in
-which samples are iteratively selected to maximally decrease the
-expected variance of $Z$. Due to the sequential constraint of mental
-rotation, this strategy is not particularly useful in two dimensions,
-as there are only every two directions in which to rotate. In three
-dimensions, however, there are an infinite number of directions that
-could be chosen after every step.
-
-We conclude that, from this initial survey, models of mental rotation
-which take an ``active'', directed approach seem well-suited to
-explaining human behavior in these tasks. Future work will collect
-empirical data from participants to perform a more quantitative
-analysis of these modeling tools.
+% A further strength of the nonparametric model over the other models,
+% however, is that it is likely to generalize well to three
+% dimensions. One aspect of \cite{Osborne:2012tm} which we have not yet
+% explored is their main contribution of \textit{active sampling}, in
+% which samples are iteratively selected to maximally decrease the
+% expected variance of $Z$. Due to the sequential constraint of mental
+% rotation, this strategy is not particularly useful in two dimensions,
+% as there are only every two directions in which to rotate. In three
+% dimensions, however, there are an infinite number of directions that
+% could be chosen after every step.
 

### 2013-05-31 man/nips-2013.tex
Update caption

@@
   \centering
-  \includegraphics[width=0.95\textwidth]{../figures/shepard-rotation.png}
+  \includegraphics[width=0.8\textwidth]{../figures/shepard-rotation.png}
   \caption{\textbf{Classic mental rotation task}. Participants in
-    \cite{Shepard1971} saw stimuli such as these, and judged whether
-    each pair of shapes was the same shape in two different
-    orientations, or two different shapes. \textbf{A} shows a ``plane
-    rotation'', \textbf{B} shows a ``depth rotation'', and \textbf{C}
-    shows two distinct objects. \TODO{update this caption}}
+    \cite{Shepard1971} saw stimuli such as those on the left, and
+    judged whether each pair of shapes was the same shape in two
+    different orientations (``same'' pairs), or two different shapes
+    (``different'' pairs). \textbf{A} and \textbf{B} show ``same''
+    pairs, while \textbf{C} shows a ``different'' pair. The plots on
+    the right indicate mean response times on the mental rotation
+    task, exhibiting a strong linear relationship with increasing
+    variance as a function of true rotation.}
   \label{fig:mental-rotation}

### 2013-05-31 man/nips-2013.tex
Abstract and referrences

@@
 
-\title{Mental Simulation as Bayesian Quadrature}
+\title{Mental Rotation as Bayesian Quadrature}
 \author{Jessica B.~Hamrick and Thomas L.~Griffiths\\
@@
 \begin{abstract}
-\TODO{}
+  Given a computationl resource--for example, the ability to visualize
+  an object rotating--how do you best make use of it? We explored how
+  mental simulation should used in the classic psychological task of
+  determining if two images depict the same object in different
+  orientations. We compared three models on this mental rotation task,
+  and found that a model based on an optimal experiment design for
+  Bayesian quadrature is objectively more accurate and qualitatively
+  more consistent with classic behavioral data than the other models
+  we evaluated. We suggest that rational models which adaptively
+  exploit available resources are promising in their ability to
+  characterize metacognitive processes like mental simulation.
 \end{abstract}
@@
 addressing this question, applying rational analysis (in the spirit of
-\cite{Marr:1983to,anderson90} \TODO{shepard?}) to one aspect of human
+\cite{Marr:1983to,anderson90,Shepard:1987tt}) to one aspect of human
 metacognition: the use of mental simulation.
@@
 an integration problem, with the direction of rotation becoming a
-statistical design problem (or in machine learning parlance, a problem
+problem of optimal experiment design (or in machine learning parlance,
 of active learning). We show that recent work on methods for Bayesian
@@
 \section{Background}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.8\textwidth]{../figures/shepard-rotation.png}
+  \caption{\textbf{Classic mental rotation task}. Participants in
+    \cite{Shepard1971} saw stimuli such as those on the left, and
+    judged whether each pair of shapes was the same shape in two
+    different orientations (``same'' pairs), or two different shapes
+    (``different'' pairs). \textbf{A} and \textbf{B} show ``same''
+    pairs, while \textbf{C} shows a ``different'' pair. The plots on
+    the right indicate mean response times on the mental rotation
+    task, exhibiting a strong linear relationship with increasing
+    variance as a function of true rotation.}
+  \label{fig:mental-rotation}
+\end{figure}
 
@@
 that the mental imagery debate could not be resolved on the basis of
-response time data alone. In recent years, however, a significant
-amount of work has investigated the neural underpinnings of the mental
+response time data alone. In recent years, however, significant
+efforts have investigated the neural underpinnings of the mental
 rotation phenomena, arguing that the results of brain imaging studies
-support the existence of visual mental rotation \cite{Kosslyn1988,
+support the existence of visual imagery \cite{Kosslyn1988,
   Kosslyn:2009tj}.
@@
 several different approaches.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.8\textwidth]{../figures/shepard-rotation.png}
-  \caption{\textbf{Classic mental rotation task}. Participants in
-    \cite{Shepard1971} saw stimuli such as those on the left, and
-    judged whether each pair of shapes was the same shape in two
-    different orientations (``same'' pairs), or two different shapes
-    (``different'' pairs). \textbf{A} and \textbf{B} show ``same''
-    pairs, while \textbf{C} shows a ``different'' pair. The plots on
-    the right indicate mean response times on the mental rotation
-    task, exhibiting a strong linear relationship with increasing
-    variance as a function of true rotation.}
-  \label{fig:mental-rotation}
-\end{figure}
 
@@
   \centering
-  \begin{subfigure}[b]{0.45\textwidth}
+  \begin{subfigure}[b]{0.4\textwidth}
     \centering
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
+  \begin{subfigure}[b]{0.4\textwidth}
     \centering
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/bq_regression.pdf}
+  \includegraphics[width=0.9\textwidth]{../figures/bq_regression.pdf}
   \caption{\textbf{Nonparametric model.} Each panel shows one step of
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
+  \includegraphics[width=0.9\textwidth]{../figures/model_rotations.pdf}
   \caption{\textbf{Model rotations.} Top: each subplot shows the
@@
 better than chance (equivalent to guessing randomly,
-i.e. $\ME{}=0.5$). \TODO{binomial test?} Closer inspection reveals
-that much of this comes from ``different'' pairs ($\NaiveMEdiff{}$),
-where the asymmetric linear interpolation may give an overestimate of
-$Z$ (e.g. Figure \ref{fig:li}).  This intuition is confirmed by the
-model's accuracy in estimating $Z$, which was $\NaiveMSE{}$. As shown
-in the left panel of Figure \ref{fig:accuracy}, it overestimated $Z$
-for nearly all stimuli pairs. \TODO{how many?}
+i.e. $\ME{}=0.5$). Closer inspection reveals that much of this comes
+from ``different'' pairs ($\NaiveMEdiff{}$), where the asymmetric
+linear interpolation may give an overestimate of $Z$ (e.g. Figure
+\ref{fig:li}).  This intuition is confirmed by the model's accuracy in
+estimating $Z$, which was $\NaiveMSE{}$. As shown in the left panel of
+Figure \ref{fig:accuracy}, it overestimated $Z$ for nearly all stimuli
+pairs.
 
@@
 identical was $\VMME{}$, making more errors on ``same'' pairs
-($\VMMEsame{}$) than ``different'' pairs ($\VMMEdiff{}$). \TODO{is
-  this significant?} Rather than overestimating as with the \naive{}
-model, however, the parametric model largely underestimated $Z$. This
-is unsurprising: because the Von Mises distribution only has a single
-peak, it necessarily underestimates $Z$ for multimodal similarity
-functions. Overall, the parametric model had an error rate of
-$\VMMSE{}$ in estimating $Z$, which was significantly more accurate
-than that of the \naive{} model \TODO{test}.
+($\VMMEsame{}$) than ``different'' pairs ($\VMMEdiff{}$). Rather than
+overestimating as with the \naive{} model, however, the parametric
+model largely underestimated $Z$. This is unsurprising: because the
+Von Mises distribution only has a single peak, it necessarily
+underestimates $Z$ for multimodal similarity functions. Overall, the
+parametric model had an error rate of $\VMMSE{}$ in estimating $Z$,
+which was significantly more accurate than that of the \naive{} model.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/Z_accuracy.pdf}
+  \includegraphics[width=0.9\textwidth]{../figures/Z_accuracy.pdf}
   \caption{\textbf{Accuracy in estimating $Z$.} Each subplot shows the
@@
 The nonparametric model was much more accurate in choosing the correct
-hypothesis than the other two models ($\BQME{}$) \TODO{test}. We note
-that this is very close to the 3.2\% error rate reported by
-\cite{Shepard1971}, though further experimentation is necessary to
-determine whether the types of errors people make align with those of
-the model.
+hypothesis than the other two models ($\BQME{}$). We note that this is
+very close to the 3.2\% error rate reported by \cite{Shepard1971},
+though further experimentation is necessary to determine whether the
+types of errors people make align with those of the model.
 
@@
 
-% One option for improving the performance of the simple models would be
-% to fit a threshold value, below which local maxima would be
-% ignored. However, this strategy would be rather brittle, for if the
-% distribution of shapes changed, the threshold would have to be
-% re-learned. It is possible that people exhibit this behavior, but we
-% cannot make any assumptions one way or the other without empirical
-% data.
-
-% A further strength of the nonparametric model over the other models,
-% however, is that it is likely to generalize well to three
-% dimensions. One aspect of \cite{Osborne:2012tm} which we have not yet
-% explored is their main contribution of \textit{active sampling}, in
-% which samples are iteratively selected to maximally decrease the
-% expected variance of $Z$. Due to the sequential constraint of mental
-% rotation, this strategy is not particularly useful in two dimensions,
-% as there are only every two directions in which to rotate. In three
-% dimensions, however, there are an infinite number of directions that
-% could be chosen after every step.
+% two questions: which way to rotate, when to stop, giving formal answer
+% to these questions which produces answers consistent with human
+% behavior. instance of showing people are appropriately using
+% computationl resources
 

### 2013-05-31 man/nips-2013.tex
First submission version

@@
 \begin{abstract}
-  Given a computationl resource--for example, the ability to visualize
-  an object rotating--how do you best make use of it? We explored how
-  mental simulation should used in the classic psychological task of
-  determining if two images depict the same object in different
-  orientations. We compared three models on this mental rotation task,
-  and found that a model based on an optimal experiment design for
-  Bayesian quadrature is objectively more accurate and qualitatively
-  more consistent with classic behavioral data than the other models
-  we evaluated. We suggest that rational models which adaptively
-  exploit available resources are promising in their ability to
-  characterize metacognitive processes like mental simulation.
+  Given a computational resource--for example, the ability to
+  visualize an object rotating--how do you best make use of it? We
+  explored how mental simulation should used in the classic
+  psychological task of determining if two images depict the same
+  object in different orientations. We compared three models on this
+  mental rotation task, and found that a model based on an optimal
+  experiment design for Bayesian quadrature is objectively more
+  accurate and qualitatively more consistent with classic behavioral
+  data than the other models we evaluated. We suggest that rational
+  models which adaptively exploit available resources are promising in
+  their ability to characterize metacognitive processes like mental
+  simulation.
 \end{abstract}
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{../figures/shepard-rotation.png}
+  \includegraphics[width=0.7\textwidth]{../figures/shepard-rotation.png}
   \caption{\textbf{Classic mental rotation task}. Participants in
@@
 \item \textit{Response error rates}: How accurate was the model at
-  choosing the correct hypothesis? This was defined to be the mean
-  error ($\ME{}$), or fraction of times the model picks the incorrect
-  hypothesis.
+  choosing the correct hypothesis? This was defined as the mean error
+  ($\ME{}$), or fraction of times the model chose incorrectly.
 \item \textit{Rotations}: For those ``same'' pairs which the model
@@
 estimating $Z$, which was $\NaiveMSE{}$. As shown in the left panel of
-Figure \ref{fig:accuracy}, it overestimated $Z$ for nearly all stimuli
-pairs.
+Figure \ref{fig:accuracy}, it overestimated $Z$ for nearly all
+stimuli.
 
@@
 shape of the raw data in Figure \ref{fig:rotations}, top left: the
-\naive{} model corresponds extremeley well to the true angle of
+\naive{} model corresponds extremely well to the true angle of
 rotation for approximately $R<\frac{\pi}{2}$. This is unsurprising,
@@
 estimate of $Z$ is accurate, it does not get stuck as easily on local
-optima.  Indeed, its active attempts to obtain an accurate estimate of
-$Z$ were largely succeessful, with $\BQMSE{}$.
+optima.  Indeed, it was largely successful, with $\BQMSE{}$.
 
@@
 participants to perform a more thorough, quantitative analysis.  We
-conclude that from this initial survey, models of mental rotation
-which take an ``active'', directed approach seem well-suited to
-explaining human behavior in such metacognitive domains as mental
+conclude that from this initial survey, models which take an
+``active'', directed approach seem well-suited to explaining mental
 simulation.
@@
 \bibliographystyle{ieeetr}
-\bibliography{references}
+{\small \bibliography{references}}
 

### 2013-05-31 man/nips-2013.tex
Edits from Tom

@@
 provided insight into the nature of the computational problems that
-human beings need to solve (e.g. \cite{Chater:1999wp,Griffiths2012a}),
+human beings need to solve (e.g., \cite{Chater:1999wp,tenenbaumkgg11}),
 but leaves open the question of how people allocate their resources in
@@
 particular, we argue that performing mental rotation can be framed as
-an integration problem, with the direction of rotation becoming a
-problem of optimal experiment design (or in machine learning parlance,
-of active learning). We show that recent work on methods for Bayesian
-quadrature \cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}
-provides a way to solve this problem, outperforming several simpler
-heuristics for determining both the direction and extent of rotation.
+integration over a probability distribution, with the direction of
+rotation becoming an optimal experiment design problem (or in machine
+learning parlance, an active learning problem). We show that recent
+work on methods for Bayesian quadrature
+\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm} provides a way to
+solve this problem, outperforming several simpler heuristics for
+determining both the direction and extent of rotation.
 
@@
 solution, and qualitatively compare their behavior to that of the
-classic mental rotation results. We end with a discussion of the
-strengths and weaknesses of each model, as well as directions for
-future work.
+classic mental rotation results. We end with a brief summary and
+conclusion of our approach.
 
@@
 three-dimensional objects could either be: the same object rotated in
-plane (Figure \ref{fig:mental-rotation}A), the same object rotated in
-depth (Figure \ref{fig:mental-rotation}B), or two different objects
+the plane (Figure \ref{fig:mental-rotation}A), the same object rotated
+in depth (Figure \ref{fig:mental-rotation}B), or two different objects
 (Figure \ref{fig:mental-rotation}C; in this case, the ``different''
@@
 symmetry). Participants had to determine whether the two images were
-of the same object or not, and \cite{Shepard1971} famously found that
+of the same object or not, and Shepard and Metzler famously found that
 response times for plane and depth rotations had a strong linear
@@
 the underlying cognitive processes were not visual in nature
-(e.g. \cite{Pylyshyn1981}). In particular, \cite{Anderson1978} proved
-that the mental imagery debate could not be resolved on the basis of
-response time data alone. In recent years, however, significant
-efforts have investigated the neural underpinnings of the mental
-rotation phenomena, arguing that the results of brain imaging studies
-support the existence of visual imagery \cite{Kosslyn1988,
-  Kosslyn:2009tj}.
+(e.g., \cite{Pylyshyn1981}). In particular, Anderson
+\cite{Anderson1978} proved that the mental imagery debate could not be
+resolved on the basis of response time data alone. In recent years,
+however, significant efforts have investigated the neural
+underpinnings of the mental rotation phenomena, arguing that the
+results of brain imaging studies support the existence of visual
+imagery \cite{Kosslyn:2009tj,Kosslyn1988}.
 
@@
 self-directed learning have found that they tend to use a strategy of
-\textit{active learning} \cite{Gureckis:2012gu, Markant:2012uu}. When
+\textit{active learning} \cite{Gureckis:2012gu,Markant:2012uu}. When
 given a choice as to the information they can obtain, people will not
@@
 Carlo sampler), but according to some utility function such as
-information gain \cite{Nelson2007, Markant:2012uu}. We propose that
+information gain \cite{Markant:2012uu,Nelson2007}. We propose that
 such a strategy can be combined with the notion of mental imagery as a
 tool: an optimal model of mental rotation may, perhaps, rely on an
-active strategy. In the next sections, we investigate this idea with
-several different approaches.
+active learning strategy. In the next sections, we investigate this
+idea with several different approaches.
 
@@
 coordinates of the vertices of two-dimensional shapes similar to those
-used by \cite{Cooper:1975wp} (e.g. Figure
+used by \cite{Cooper:1975wp} (e.g., Figure
 \ref{fig:stimuli}). Participants must determine whether $X_a$ and
@@
 \begin{align}
-  p(X_a,\ X_b\ \vert \ h_0)&=p(X_a)p(X_b) \label{eq:lh-h0}\\
+  p(X_a,\ X_b\ \vert \ h_0)&=p(X_a)p(X_b), \label{eq:lh-h0}\\
   p(X_a,\ X_b\ \vert \ h_1)&=\int_R\int_M p(X_a) p(X_b\vert X_a,R,M) p(R) p(M)\ \mathrm{d}M\ \mathrm{d}R. \label{eq:lh-h1}
@@
     low, but certain angles will have a disproportionate contribution
-    (e.g. between $\frac{\pi}{4}$ and $\frac{3\pi}{4}$. \textbf{(b)}
+    (e.g., between $\frac{\pi}{4}$ and $\frac{3\pi}{4}$. \textbf{(b)}
     Best fit of scaled Von Mises PDF parameters. As in (a), $Z$ will
@@
 unimodal when $h_1$ is true\footnote{This is not always a reasonable
-  assumption, see e.g. Figure \ref{fig:vm}.}, then a reasonable form
+  assumption, see e.g., Figure \ref{fig:vm}.}, then a reasonable form
 for the similarity function is that of a ``circular Gaussian'' or Von
@@
 We evaluated each model's performance on 20 shapes which had between 3
-and 6 vertices, inclusive, (e.g. $X_a$ in Figure
+and 6 vertices, inclusive, (e.g., $X_a$ in Figure
 \ref{fig:stimuli}). For each shape, we computed 18 ``same'' and 18
@@
 from ``different'' pairs ($\NaiveMEdiff{}$), where the asymmetric
-linear interpolation may give an overestimate of $Z$ (e.g. Figure
+linear interpolation may give an overestimate of $Z$ (e.g., Figure
 \ref{fig:li}).  This intuition is confirmed by the model's accuracy in
@@
 has the capacity to ``reset'', it could recover from rotating in the
-incorrect direction (e.g. Figure \ref{fig:bq}) and thus did not
+incorrect direction (e.g., Figure \ref{fig:bq}) and thus did not
 over-rotate as frequently.  It also under-rotated less frequently:
@@
 
-\section{Discussion}
+\section{Conclusion}
 
 In this paper, we asked: how do people use mental simulation?  We
-investigated the specific case of mental rotation, taking a rational
-analysis approach to characterizing optimal strategies for performing
-mental rotation. We found that an adaptive, nonparametric model
-performs best at the classic rotation task \cite{Shepard1971}, as it
-is able to actively monitor the confidence of its estimate and
-intelligently choose the direction of rotation. Two simpler models
-based on heuristics performed much worse: they did not maintain the
-linear relationship with the true angle of rotation due to lack of
-robustness against local maxima or incorrect rotation direction. The
+investigated the specific case of mental rotation, using rational
+analysis to characterize optimal strategies for performing mental
+rotation. We found that an adaptive, nonparametric model performs best
+at the classic rotation task \cite{Shepard1971}, as it is able to
+actively monitor the confidence of its estimate and intelligently
+choose the direction of rotation. Two simpler models based on
+heuristics performed much worse: they did not maintain the linear
+relationship with the true angle of rotation due to lack of robustness
+against local maxima or incorrect rotation direction. The
 nonparametric model qualitatively approximates human behavioral

### 2013-08-04 man/nips-2013-rebuttal.md
First draft of rebuttal

# NIPS 2013 Rebuttal

**Paper ID**: 1352  
**Title**: Mental Rotation as Bayesian Quadrature

## Reviewer 3

Reviewer 3 was primarily concerned with the comparison between models
and human data. While we did not collect human data ourselves, we
compared our model to the human response times reported by Shepard and
Metzler [6] (the key comparison is between Figure 1, depicting the
classic results, and Figure 5, illustrating our modeling results).
The mental rotation results are some of the most famous findings in
cognitive science ([6] alone has at least 2,061 citations according to
Web of Knowledge; 3,640 according to Google Scholar) and have been
replicated many times with many types of stimuli (Thomas 2013). As a
proof-of-concept, we demonstrate that our model is qualitatively
similar to the standard mental rotation result.  No other
computational models have been proposed which can explain how people
choose the correct direction and duration of rotation.

The reviewer was also concerned about how the prior over stimuli
related to the process by which the shapes were created. To clarify,
the shapes were generated in a manner consistent with the prior:
vertex angles were chosen uniformly at random between 0 and $2\pi$,
and vertex radii were chosen uniformly at random between 0 and 1.
Edges were drawn between vertices which were adjacent when sorted by
angle.

## Reviewer 4

Reviewer 4 was concerned with the explanatory power of our model,
suggesting that people do not need to integrate over the full set of
rotations. However, it is not obvious how the strategy "just rotate
until they find a match that is easy to recognize or exhaust the set
of potential rotations" should be implemented. Similarly, stopping
when a local maximum is found is not equivalent to quitting "even if
it's obviously not a match at that rotation". What does it mean for a
match to be easy to recognize, or for something be obviously not a
match? A perfect match cannot truly be identified due to perceptual
uncertainty in people's visual system. Moreover, one cannot just
rotate until a threshold of similarity has been reached, because the
similarity function is not normalized and therefore the value of its
global maximum is not known *a priori*.

If the value of the global maximum of the similarity function is not
known, then it is impossible to know if any particular maximum is the
global maximum or not. To reliably find the global maximum, then,
would require an exhaustive search (a full rotation of 360 degrees),
which is inconsistent with the linear nature of human response times
from [6]. Thus, it is necessary to either be satisfied with a local
maximum (as found by hill-climbing), or by a more sophisticated method.

Even if it was possible to perfectly recognize a match after rotating
the appropriate amount, there is still a separate issue: how do people
know which direction to rotate? Simply choosing one direction at
random and rotating to the true angle produces a pattern of data which
is inconsistent with the behavioral data from [6]. For any true
minimum angle $r$, we would have expected response time proportional
to $E[r] = 0.5r + 0.5(2\pi - r) = \pi$, which is constant rather than
monotonic.

The "model rotations" described around line 372 and shown in Figure 5
do refer to the number of steps the models takes. A more precise label
would be "total distance rotated, in degrees" which is directly
proportional to the number of steps the model takes.

To address the reviewer's question about whether the behavioral
results hold for our stimuli, we emphasize (as above in response to
Reviewer 3) that the Shepard and Metzler results [6] are very robust:
they have been replicated many times and with many other types of
stimuli. In particular, our stimuli are similar to those used by [23].

The reviewer correctly notes that the 95% confidence threshold used by
the BQ model's hypothesis test is likely the source of the 4% error
rate. The reviewer also points out that the MSE on line 382 is
inconsistent with our definition of MSE. This is indeed a mistake: the
MSE cannot be normalized between 0 and 1 because it has no upper
bound. The correct, non-normalized MSEs are 0.03 for the naive model,
0.002 for the parametric model, and 0.001 for the nonparametric model.

## Reviewer 5

We thank Reviewer 5 for their encouraging comments. We do intend to
collect human data in the future and to perform a more careful,
quantitative comparison between the model and participants. We further
hope that the model will be useful in uncovering previously unexamined
nuances in people's behavior.

Regarding people's mental representations, we do not think that people
are necessarily performing full Bayesian quadrature in their
heads. Ours is a computational-level model [3], which means that it is
a formal way of describing how the problem of determining if two
objects are the same could be solved. Although we incorporate the
constraint of sequential rotations, this is not a true
algorithmic-level constraint, as we have not given specific limits on
working memory, processing power, knowledge representation, etc. The
algorithmic-level form of this solution would be an approximation to
our computational-level model and would make the resource limitations
explicit. Currently, however, we have not explored possible
algorithmic approximations. Doing so is a direction for future work.

# References

* Thomas, Nigel J.T., "Supplement: Mental Rotation", *The Stanford
  Encyclopedia of Philosophy*, Edward N. Zalta (ed.). Retrieved on
  08/05/2013 from
  http://plato.stanford.edu/archives/spr2013/entries/mental-imagery/mental-rotation.html

* [6] R. N. Shepard and J. Metzler, "Mental Rotation of
  Three-Dimensional Objects," *Science*, vol. 171, no. 3972,
  pp. 701–703, 1971.

* [23] L. A. Cooper, "Mental Rotation of Random Two-Dimensional
  Shapes," *Cognitive Psychology*, vol. 7, pp. 20-43, 1975.

### 2013-08-04 code/analysis_tools.py
Update analysis
Fix issue with scaled MSE -- it's not correct to try to scale it,
because the error doesn't have an upper bound.

@@ -180,8 +180,7 @@ def calc_MSE(models, data):
     gs_Z_mean, gs_Z_std = process_Z(data["GoldStandardModel"])
     for model in models:
         Z_mean, Z_std = process_Z(data[model])
-        normed_err = (gs_Z_mean - Z_mean) / np.ptp(gs_Z_mean)
-        MSE[model] = np.mean(normed_err ** 2)
+        MSE[model] = np.mean((gs_Z_mean - Z_mean) ** 2)
     return MSE
 
 
@@ -190,7 +189,7 @@ def MSE_latex(models, MSE, shortnames):
     for model in models:
         shortname = shortnames[model]
         mse = MSE[model]
-        latex.append(new_cmd("%sMSE" % shortname, r"\MSE{}=%.2f" % mse))
+        latex.append(new_cmd("%sMSE" % shortname, r"\MSE{}=%.3f" % mse))
     return "".join(latex)
 
 

### 2013-08-05 man/nips-2013-rebuttal.md
Second draft of rebuttal

@@
 classic results, and Figure 5, illustrating our modeling results).
-The mental rotation results are some of the most famous findings in
-cognitive science ([6] alone has at least 2,061 citations according to
-Web of Knowledge; 3,640 according to Google Scholar) and have been
-replicated many times with many types of stimuli (Thomas 2013). As a
-proof-of-concept, we demonstrate that our model is qualitatively
-similar to the standard mental rotation result.  No other
-computational models have been proposed which can explain how people
-choose the correct direction and duration of rotation.
+The mental rotation result is one of the most famous findings in
+cognitive science ([6] alone has 2,061 citations according to Web of
+Knowledge; 3,640 according to Google Scholar) and has been replicated
+many times with many types of stimuli (Thomas, 2013). As a
+proof-of-concept, we showed a high correlation between our model and a
+linear function (which is the same form as the classic result).  No
+other computational models have been proposed which can explain how
+people choose the correct direction and duration of rotation.
 
@@
 the shapes were generated in a manner consistent with the prior:
-vertex angles were chosen uniformly at random between 0 and $2\pi$,
-and vertex radii were chosen uniformly at random between 0 and 1.
-Edges were drawn between vertices which were adjacent when sorted by
-angle.
+vertex angles were chosen uniformly at random between 0 and 2*pi, and
+vertex radii were chosen uniformly at random between 0 and 1.  Edges
+were drawn between vertices which were adjacent when sorted by angle.
 
@@
 Reviewer 4 was concerned with the explanatory power of our model,
-suggesting that people do not need to integrate over the full set of
-rotations. However, it is not obvious how the strategy "just rotate
-until they find a match that is easy to recognize or exhaust the set
-of potential rotations" should be implemented. Similarly, stopping
-when a local maximum is found is not equivalent to quitting "even if
-it's obviously not a match at that rotation". What does it mean for a
-match to be easy to recognize, or for something be obviously not a
-match? A perfect match cannot truly be identified due to perceptual
-uncertainty in people's visual system. Moreover, one cannot just
-rotate until a threshold of similarity has been reached, because the
-similarity function is not normalized and therefore the value of its
-global maximum is not known *a priori*.
+suggesting that people could "just rotate until they find a match that
+is easy to recognize or exhaust the set of potential
+rotations". However, it is not obvious how this strategy should be
+implemented. What does it mean for a match to be easy to recognize?
+Similarly, stopping when a local maximum is found is not equivalent to
+quitting "even if it's obviously not a match at that rotation".  A
+perfect match cannot truly be identified due to perceptual uncertainty
+in people's visual system, and one cannot just rotate until a
+threshold of similarity has been reached, because the similarity
+function is not normalized and therefore the value of its global
+maximum is not known *a priori*.
 
 If the value of the global maximum of the similarity function is not
-known, then it is impossible to know if any particular maximum is the
+known, then it is impossible to know if any local maximum is the
 global maximum or not. To reliably find the global maximum, then,
@@
 from [6]. Thus, it is necessary to either be satisfied with a local
-maximum (as found by hill-climbing), or by a more sophisticated method.
+maximum (as found by hill-climbing), or to use a more sophisticated
+method.
 
-Even if it was possible to perfectly recognize a match after rotating
-the appropriate amount, there is still a separate issue: how do people
-know which direction to rotate? Simply choosing one direction at
-random and rotating to the true angle produces a pattern of data which
-is inconsistent with the behavioral data from [6]. For any true
-minimum angle $r$, we would have expected response time proportional
-to $E[r] = 0.5r + 0.5(2\pi - r) = \pi$, which is constant rather than
-monotonic.
+Even if it is possible to perfectly recognize a match after rotating
+the appropriate amount, there is a separate issue: how do people know
+which direction to rotate? Choosing one direction at random and
+rotating to the true angle produces a pattern of data which is
+inconsistent with the behavioral data from [6]. For any true minimum
+angle r, we would have expected response times proportional to E[r] =
+0.5r + 0.5(2*pi - r) = pi, which is constant rather than monotonic.
 
@@
 results hold for our stimuli, we emphasize (as above in response to
-Reviewer 3) that the Shepard and Metzler results [6] are very robust:
-they have been replicated many times and with many other types of
-stimuli. In particular, our stimuli are similar to those used by [23].
+Reviewer 3) that the Shepard and Metzler results [6] are very robust
+and have been replicated with many other types of stimuli. In
+particular, our stimuli are similar to those used by [23].
 
@@
 We thank Reviewer 5 for their encouraging comments. We do intend to
-collect human data in the future and to perform a more careful,
-quantitative comparison between the model and participants. We further
-hope that the model will be useful in uncovering previously unexamined
-nuances in people's behavior.
+collect human data in the future and to perform a more detailed,
+item-specific comparison between the model and participants. We
+further hope that the model will be useful in uncovering previously
+unexamined nuances in people's behavior.
 
@@
 heads. Ours is a computational-level model [3], which means that it is
-a formal way of describing how the problem of determining if two
-objects are the same could be solved. Although we incorporate the
-constraint of sequential rotations, this is not a true
-algorithmic-level constraint, as we have not given specific limits on
-working memory, processing power, knowledge representation, etc. The
-algorithmic-level form of this solution would be an approximation to
-our computational-level model and would make the resource limitations
+a formal way of describing how to solve the problem of determining if
+two objects are the same. Although we incorporate the constraint of
+sequential rotations, this is not an algorithmic-level constraint
+because we have not given specific limits on working memory,
+processing power, knowledge representation, etc. The algorithmic-level
+form of this solution would be an approximation to our
+computational-level model and would make the resource limitations
 explicit. Currently, however, we have not explored possible
@@
 
+* [3] D. Marr, *Vision: A Computational Investigation into the Human
+  Representation and Processing of Visual Information*. Henry Holt and
+  Company, 1983.
+
 * [6] R. N. Shepard and J. Metzler, "Mental Rotation of

### 2013-10-09 code/plots.py
Revise figures

@@ -125,6 +125,21 @@ def regression(x, y, xi, yi, xo, yo_mean, yo_var, **kwargs):
         plt.legend(loc=0, fontsize=12, frameon=False)
 
 
+def bq_regression_final(model):
+    plt.figure()
+    R = model.R
+    S = model.S
+    Ri = model.Ri
+    Si = model.Si
+    S_mean = model.S_mean
+    S_var = model.S_var
+    regression(
+        R, S, Ri, Si,
+        R, S_mean, S_var,
+        title="BQ regression for $S$")
+    plt.ylim(0, 1.05)
+
+
 def bq_regression(model):
 
     # plot the regression for S
@@ -340,63 +355,52 @@ def likelihood_all(models):
 
 
 def model_rotations(models):
-    fig, axes = plt.subplots(2, len(models), sharex=True, sharey=True)
-    if len(models) == 1:
-        ax0 = axes[1]
-    else:
-        ax0 = axes[1][0]
+    nm = len(models)
+    fig, axes = plt.subplots(1, 2*nm, sharex=True, sharey=True)
+    ax0 = axes[0]
 
     ax0.set_xticks([0, np.pi / 4., np.pi / 2., 3 * np.pi / 4., np.pi])
     ax0.set_xticklabels([0, 45, 90, 135, 180])
     ax0.set_xlim(-np.pi/16, np.pi+(np.pi/16))
 
     ax0.set_yticks([0, 25, 50, 75, 100])
-    ax0.set_yticklabels([0, 90, 180, 270, 360])
+    # ax0.set_yticklabels([0, 90, 180, 270, 360])
     ax0.set_ylim(-5, 105)
 
     for i in xrange(len(models)):
-        if len(models) == 1:
-            tax = axes[0]
-            bax = axes[1]
-        else:
-            tax = axes[0][i]
-            bax = axes[1][i]
+        tax = axes[i]
+        bax = axes[nm + i]
 
         model = models[i]
         tax.set_title(model, fontsize=12)
+        bax.set_title(model, fontsize=12)
+        tax.set_xlabel(r"True rotation ($R$)")
         bax.set_xlabel(r"True rotation ($R$)")
+
         if i == 0:
-            tax.set_ylabel(r"All rotations")
-            bax.set_ylabel(r"Mean correct rotations")
+            tax.set_ylabel(r"% rotated (out of 360$^\circ$)")
         for ax in (tax, bax):
             sg.outward_ticks(ax=ax)
             sg.clear_right(ax=ax)
[…]

### 2013-10-09 man/nips-2013-bayesian-optimization.tex
Begin work on bayesian optimization submission

--- 2013-05-31 man/nips-2013.tex
@@
 \newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}
-\newcommand{\MSE}[0]{\mathrm{MSE}}
 \newcommand{\ME}[0]{\mathrm{ME}}
@@
 
-%\nipsfinalcopy % Uncomment for camera-ready version
+\nipsfinalcopy % Uncomment for camera-ready version
 
@@
 \cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm} provides a way to
-solve this problem, outperforming several simpler heuristics for
-determining both the direction and extent of rotation.
+solve this problem, outperforming simpler heuristics for determining
+the direction and extent of rotation.
 
-The plan of the paper is as follows. First, we give a brief overview
-of the work on mental rotation and active learning. Next, we present
-the problem domain and a computational-level analysis of how to solve
-it. We describe three different models which approximate this
-solution, evaluate their accuracy at approximating the computational
-solution, and qualitatively compare their behavior to that of the
-classic mental rotation results. We end with a brief summary and
-conclusion of our approach.
-
-\section{Background}
+The plan of the paper is as follows. First, we present the problem
+domain and a computational-level analysis of how to solve it. We
+describe two different models which approximate this solution,
+evaluate their accuracy at approximating the computational solution,
+and qualitatively compare their behavior to that of the classic mental
+rotation results. We end with a brief summary and conclusion of our
+approach.
 
@@
 
-There is a long and rich history of work on mental rotation beginning
-with Shepard and Metzler \cite{Shepard1971}, who presented
-participants with pairs of images such as those shown in Figure
-\ref{fig:mental-rotation} (left panel). These pairs of
-three-dimensional objects could either be: the same object rotated in
-the plane (Figure \ref{fig:mental-rotation}A), the same object rotated
-in depth (Figure \ref{fig:mental-rotation}B), or two different objects
-(Figure \ref{fig:mental-rotation}C; in this case, the ``different''
-objects had reflective symmetry, but no rotational
-symmetry). Participants had to determine whether the two images were
-of the same object or not, and Shepard and Metzler famously found that
-response times for plane and depth rotations had a strong linear
-relationship with the minimum angle of rotation between the two
-objects (Figure \ref{fig:mental-rotation}, right panel). The
-conclusion was that participants were visually ``rotating'' the
-objects in their minds.
+% \section{Background}
 
-For many years, this idea was contested, with some researchers arguing
-the underlying cognitive processes were not visual in nature
-(e.g., \cite{Pylyshyn1981}). In particular, Anderson
-\cite{Anderson1978} proved that the mental imagery debate could not be
-resolved on the basis of response time data alone. In recent years,
-however, significant efforts have investigated the neural
-underpinnings of the mental rotation phenomena, arguing that the
-results of brain imaging studies support the existence of visual
-imagery \cite{Kosslyn:2009tj,Kosslyn1988}.
+% There is a long and rich history of work on mental rotation beginning
+% with Shepard and Metzler \cite{Shepard1971}, who presented
+% participants with pairs of images such as those shown in Figure
+% \ref{fig:mental-rotation} (left panel). These pairs of
+% three-dimensional objects could either be: the same object rotated in
+% the plane (Figure \ref{fig:mental-rotation}A), the same object rotated
+% in depth (Figure \ref{fig:mental-rotation}B), or two different objects
+% (Figure \ref{fig:mental-rotation}C; in this case, the ``different''
+% objects had reflective symmetry, but no rotational
+% symmetry). Participants had to determine whether the two images were
+% of the same object or not, and Shepard and Metzler famously found that
+% response times for plane and depth rotations had a strong linear
+% relationship with the minimum angle of rotation between the two
+% objects (Figure \ref{fig:mental-rotation}, right panel). The
+% conclusion was that participants were visually ``rotating'' the
+% objects in their minds.
 
-We assume that mental imagery is indeed a visual process, and turn to
-the question of how, and when, it is used. Previous work has examined
-how people might use imagery \cite{Hegarty2004, Schwartz1999} to solve
-reasoning problems, and when they might use imagery as opposed to a
-symbolic rule \cite{Schwartz:1996uy}. Eye-tracking studies of people
-performing mental rotations suggest strategies that people might be
-using, such as aligning smaller features of the images
-\cite{Just1976}. People seem to use imagery methodically, which raises
-the question: what is the method?
+% For many years, this idea was contested, with some researchers arguing
+% the underlying cognitive processes were not visual in nature (e.g.,
+% \cite{Pylyshyn1981}). In particular, Anderson \cite{Anderson1978}
+% proved that the mental imagery debate could not be resolved on the
+% basis of response time data alone. In recent years, however,
+% significant efforts have investigated the neural underpinnings of the
+% mental rotation phenomena, arguing that the results of brain imaging
+% studies support the existence of visual imagery
+% \cite{Kosslyn:2009tj,Kosslyn1988}.  We assume that mental imagery is
+% indeed a visual process, and turn to the question of how, and when, it
+% is used. Previous work has examined how people might use imagery
+% \cite{Hegarty2004, Schwartz1999} to solve reasoning problems, and when
+% they might use imagery as opposed to a symbolic rule
+% \cite{Schwartz:1996uy}. Eye-tracking studies of people performing
+% mental rotations suggest strategies that people might be using, such
+% as aligning smaller features of the images \cite{Just1976}. People
+% seem to use imagery methodically, which raises the question: what is
+% the method?
 
-Others studying people's pattern of behavior when engaged in
-self-directed learning have found that they tend to use a strategy of
-\textit{active learning} \cite{Gureckis:2012gu,Markant:2012uu}. When
-given a choice as to the information they can obtain, people will not
-choose the information randomly (as is the case with a \naive{} Monte
-Carlo sampler), but according to some utility function such as
-information gain \cite{Markant:2012uu,Nelson2007}. We propose that
-such a strategy can be combined with the notion of mental imagery as a
-tool: an optimal model of mental rotation may, perhaps, rely on an
-active learning strategy. In the next sections, we investigate this
-idea with several different approaches.
+% Others studying people's pattern of behavior when engaged in
+% self-directed learning have found that they tend to use a strategy of
+% \textit{active learning} \cite{Gureckis:2012gu,Markant:2012uu}. When
+% given a choice as to the information they can obtain, people will not
+% choose the information randomly (as is the case with a \naive{} Monte
+% Carlo sampler), but according to some utility function such as
+% information gain \cite{Markant:2012uu,Nelson2007}. We propose that
+% such a strategy can be combined with the notion of mental imagery as a
+% tool: an optimal model of mental rotation may, perhaps, rely on an
+% active learning strategy. In the next sections, we investigate this
+% idea with several different approaches.
 
@@
 MRX_a$ and $h_1$: $\exists M,R\textrm{ s.t. } X_b=MRX_a$.  To compare
-the hypotheses, we need to compute the posterior for each:
-\begin{equation}
-p(h\ \vert\ X_a, X_b)\propto p(X_a, X_b\ \vert\ h)p(h).
-\end{equation}
+the hypotheses, we need to compute the posterior for each: $p(h\
+\vert\ X_a, X_b)\propto p(X_a, X_b\ \vert\ h)p(h)$.
+
 Assuming the hypotheses are equally likely \textit{a priori}, the
@@
 
-Once we have computed both likelihoods, we compute their ratio:
-\begin{equation}
-  \ell=\frac{p(X_a, X_b\ \vert \ h_1)}{p(X_a, X_b\ \vert \ h_0)}=\frac{\int_R p(X_b\ \vert\ X_a, R)p(R)\ \mathrm{d}R}{p(X_b)}.
-  \label{eq:lh-ratio}
-\end{equation}
-If $\ell<1$, then $h_0$ is the more likely hypothesis. If $\ell>1$,
-then $h_1$ is the more likely hypothesis.
-
-\section{Metacognitive algorithms}
+Once we have computed both likelihoods, we compute their ratio $\ell$
+which is given by $\ell=\left(\int_R p(X_b\ \vert\ X_a, R)p(R)\
+  \mathrm{d}R\right) /\ p(X_b)$. If $\ell<1$, then $h_0$ is the more
+likely hypothesis. If $\ell>1$, then $h_1$ is the more likely
+hypothesis.
 
@@
   \centering
-  \begin{subfigure}[b]{0.4\textwidth}
+  \begin{subfigure}[b]{0.32\textwidth}
     \centering
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.4\textwidth}
+  \begin{subfigure}[b]{0.32\textwidth}
     \centering
-    \includegraphics[width=\textwidth]{../figures/likelihood_function.pdf}
-    \caption{Likelihood function}
-    \label{fig:likelihood}
+    \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
+    \caption{\Naive{} model}
+    \label{fig:li}
   \end{subfigure}
-  \caption{\textbf{Rotated shapes and their similarity.}  \textbf{(a)}
-    An example stimulus in which the shapes differ only by a
-    rotation. All stimuli consist of three to six vertices centered
-    around the origin, and edges which create a closed loop from the
-    vertices. \textbf{(b)} The approximate likelihood (similarity) of
-    $X_b$ given $X_R$, where $X_R$ is a rotation of $X_a$ by the angle
-    $R$. The true angle of rotation between $X_a$ and $X_b$ is at
-    $\frac{2\pi}{3}$, corresponding to the global maximum in the
-    likelihood function.}
+  \begin{subfigure}[b]{0.32\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/bq_regression_final.pdf}
+    \caption{Bayesian Quadrature Model}
+    \vspace{0pt}
+    \label{fig:bq}
+  \end{subfigure}
+  \caption{\textbf{(a)} An example stimulus in which the shapes differ
+    only by a rotation. All stimuli consist of three to six vertices
+    centered around the origin, and edges which create a closed loop
+    from the vertices. The true angle of rotation between $X_a$ and
+    $X_b$ is at $\frac{2\pi}{3}$. \textbf{(b-c)} Likelihood function
+    and \naive{} (b) and BQ (c) model estimates. The sampled points
+    $\mathbf{R}$ (red circles) are then used to estimate $S$ (black
+    lines are the true $S$, red lines are the estimate).}
   \label{fig:shapes}
 \end{figure}
+
+\section{Metacognitive algorithms}
 
@@
 shape $X$ is $p(X)=n!\left(\frac{1}{2\pi}\right)^n$, which gives us
-the denominator in Equation \ref{eq:lh-ratio}. Computing the numerator
-is more difficult, as we we cannot compute $p(X_b\vert X_a, R)$
-directly. Instead, we introduce a new variable $X_R$ denoting a mental
-image, which approximates $RX_a$. The $X_R$ are generated by repeated
-application of a function $\tau$:
-\begin{align}
-  X_R=RX_a&=\tau(X_{R-r}, r)=\tau(\tau(X_{R-2r}, r), r)=\ldots{}\nonumber \\
-  &=\tau^{(\frac{R}{r})}(X_a, r).
+the denominator of $\ell$. Computing the numerator is more difficult,
+as we we cannot compute $p(X_b\vert X_a, R)$ directly. Instead, we
+introduce a new variable $X_R$ denoting a mental image, which
+approximates $RX_a$. The $X_R$ are generated by repeated application
+of a function $\tau$:
+\begin{equation}
+  X_R=RX_a=\tau(X_{R-r}, r)=\tau(\tau(X_{R-2r}, r), r)=\ldots{}=\tau^{(\frac{R}{r})}(X_a, r).
   \label{eq:tau}
-\end{align} 
+\end{equation} 
 where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
@@
 
-\paragraph{Gold standard}
-
-To compare the accuracy of other models' estimates of $Z$, we computed
-a ``gold standard'' by evaluating $S(X_b, X_R)$ at 360 values of $R$
-spaced evenly between $0$ and $2\pi$ and estimating the integral using
-the trapezoidal rule.
-
 \paragraph{\Naive{}}
-
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.4\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
-    \caption{\Naive{} model}
-    \label{fig:li}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.4\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/vm_regression.pdf}
-    \caption{Parametric model}
-    \label{fig:vm}
-  \end{subfigure}
-  \caption{\textbf{Simple methods of estimating $S$.} In both cases,
-    hill-climbing search is used until a maxima is found (in this
-    case, at approximately $\frac{\pi}{6}$). The sampled points
-    $\mathbf{R}$ (red circles) are then used to estimate $S$ (black
-    lines are the true $S$, red lines are the estimate). \textbf{(a)}
-    Linear interpolation. The overall estimate of $Z$ here will be too
-    low, but certain angles will have a disproportionate contribution
-    (e.g., between $\frac{\pi}{4}$ and $\frac{3\pi}{4}$. \textbf{(b)}
-    Best fit of scaled Von Mises PDF parameters. As in (a), $Z$ will
-    be underestimated. However, the fit around the maximum near
-    $\frac{\pi}{6}$ is much more accurate.}
-  \label{fig:simple-models}
-\end{figure}
 
@@
 
-\paragraph{Parametric (Von Mises)}
+\paragraph{Bayesian Quadrature}
 
-Another strategy is to assume a parametric shape for $S$ and fit the
-appropriate parameters. If we assume that $S$ is approximately
-unimodal when $h_1$ is true\footnote{This is not always a reasonable
-  assumption, see e.g., Figure \ref{fig:vm}.}, then a reasonable form
-for the similarity function is that of a ``circular Gaussian'' or Von
-Mises distribution:
-\begin{equation}
-  S(X_b, X_R) \approx h\cdot{}p(R\ \vert\ \hat{\theta}, \kappa)=\frac{h}{2\pi I_0(\kappa)}e^{\kappa\cos(R-\hat{\theta})}
-\end{equation}
-where $\kappa$ is the concentration parameter, $\hat{\theta}$ is the
-preferred direction, $h$ is a scale parameter, and $I_0$ is the
-modified Bessel function of order zero. We fit these parameters by
-minimizing the mean squared error between this PDF and the computed
-values of $S$. To choose the sequence of rotations, we use the same
-hill-climbing strategy as in the \naive{} model. Figure \ref{fig:vm}
-shows an example of the parametric model's estimate of $S$.
-
-\paragraph{Nonparametric (Bayesian Quadrature)}
+A more flexible strategy uses what is known as \emph{Bayesian
+  Quadrature} \cite{Diaconis:1988uo,OHagan:1991tx} to estimate $Z$.
+Bayesian Quadrature (BQ) allows us to compute a posterior distribution
+over $Z$ by placing a Gaussian Process (GP) prior on the function $S$
+and evaluating $S$ at a particular set of points. Because our data is
+circular, we use a periodic kernel \cite{Rasmussen:2006vz}.
 
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{../figures/bq_regression.pdf}
-  \caption{\textbf{Nonparametric model.} Each panel shows one step of
-    the Bayesian Quadrature regression. Upper left: the original
-    Gaussian Process (GP) regression for $S$. Lower left: the GP
-    regression for $\log(S+1)$. Lower right: the GP regression for
-    $\Delta_c=\mu_{\log S} - \log \mu_S$. Upper right: the adjusted
-    regression for $S$, where the mean is equal to
-    $\mu_S(1+\mu_{\Delta_c})$. The model uses this final estimate to
-    compute $Z$ and will continue rotating until the variance of $Z$
-    is low enough that a hypothesis may be accepted. This method
-    allows the model to avoid local maxima such as the one near
-    $\frac{\pi}{6}$, which causes trouble for the \naive{} and
-    parametric models in Figure \ref{fig:simple-models}.}
-  \label{fig:bq}
+  \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
+  \caption{\textbf{Model rotations.} Top: each subplot shows the
+    correspondence between the true angle of rotation ($R$) for
+    ``same'' pairs and the amount of rotation performed by the
+    model. Bottom: each subplot shows the models' mean rotations over
+    stimuli pairs that were judged correctly. Black dots correspond to
+    ``same'' pairs, and blue lines correspond to ``different'' pairs
+    (for which the mean rotation is constant across true rotations, as
+    the true rotation is undefined). Error bars/shaded regions
+    indicate one standard deviation, and the dotted lines indicate the
+    least-squares fit to the ``same'' pairs.}
+  \label{fig:rotations}
 \end{figure}
 
-A more flexible strategy uses what is known as \emph{Bayesian
-  Quadrature} \cite{Diaconis:1988uo,OHagan:1991tx} to estimate $Z$.
-Bayesian Quadrature allows us to compute a posterior distribution over
-$Z$ by placing a Gaussian Process (GP) prior on the function $S$ and
-evaluating $S$ at a particular set of points. Because our data is
-circular, we use a periodic kernel \cite{Rasmussen:2006vz}:
-\begin{equation}
-k(R, R^\prime)=h^2\exp\left(-\frac{2\sin^2\left(\frac{1}{2}(R-R^\prime)\right)}{w^2}\right).
-\end{equation}
-
-Bayesian Quadrature has its difficulties, however. While in our case
-$S$ is a non-negative likelihood function, GP regression enforces no
-such constraint. In an effort to avoid this problem,
-\cite{Osborne:2012tm} give a method to place a prior over the log
-likelihood\footnote{In practice, as in \cite{Osborne:2012tm}, we use
-  the transform of $\log(S+1)$. Additionally, we follow
-  \cite{Osborne:2012tm} and use a combination of MLII and
-  marginalization to fit the $w$ kernel parameters. The output scales
-  are fixed at $h_S=\sqrt{0.1}$ and $h_{\log S}=\log(h_S + 1)$.}, thus
-ensuring that $S=e^{\log S}$ will be positive\footnote{We are not
-  guaranteed positivity, however, the approximation to the integral
-  over $\log S$ (Equation \ref{eq:bq-Z-mean}) requires computing
-  $\mu_S$, which may have non-positive segments.}:
+BQ has its difficulties, however. While in our case $S$ is a
+non-negative likelihood function, GP regression enforces no such
+constraint. In an effort to avoid this problem, \cite{Osborne:2012tm}
+give a method to place a prior over the log likelihood\footnote{In
+  practice, as in \cite{Osborne:2012tm}, we use the transform of
+  $\log(S+1)$. Additionally, we follow \cite{Osborne:2012tm} and use a
+  combination of MLII and marginalization to fit the $w$ kernel
+  parameters. The output scales are fixed at $h_S=\sqrt{0.1}$ and
+  $h_{\log S}=\log(h_S + 1)$.}, thus ensuring that $S=e^{\log S}$ will
+be positive:
 \begin{equation*}
@@
 $\tilde{V}$ in order to estimate a distribution over the likelihood
-ratio $\ell$:
-\begin{equation*}
-p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\ \vert\ \mu_Z, \sigma_z).
-\end{equation*}
-We choose $h_0$ when $p(\ell < 1)\geq 0.95$, and $h_1$ when $p(\ell >
-1)\geq 0.95$. Until one of these conditions are met (or the shape has
-been fully rotated), the model will continues to compute rotations and
-update its estimate of $Z$.
+ratio $\ell$, i.e.  $p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\
+\vert\ \mu_Z, \sigma_z)$.  We choose $h_0$ when $p(\ell < 1)\geq
+0.95$, and $h_1$ when $p(\ell > 1)\geq 0.95$. Until one of these
+conditions are met (or the shape has been fully rotated), the model
+will continues to compute rotations and update its estimate of $Z$.
 
@@
 
-% \begin{equation}
-%   \sigma_Z=\mathrm{Var}(Z\ \vert \ S, \log S, \Delta_c) = \int_R\int_{R^\prime} \mu_S(R)\mu_S(R^\prime) \Sigma_{\log S}(R, R^\prime)p(R)p(R^\prime)\ \mathrm{d}R\ \mathrm{d}R^\prime
-%   \label{eq:bq-Z-var}
-% \end{equation}
-
 \section{Results}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.9\textwidth]{../figures/model_rotations.pdf}
-  \caption{\textbf{Model rotations.} Top: each subplot shows the
-    correspondence between the true angle of rotation ($R$) for
-    ``same'' pairs and the amount of rotation performed by the
-    model. Bottom: each subplot shows the models' mean rotations over
-    stimuli pairs that were judged correctly. Black dots correspond to
-    ``same'' pairs, and blue lines correspond to ``different'' pairs
-    (for which the mean rotation is constant across true rotations, as
-    the true rotation is undefined). Error bars/shaded regions
-    indicate one standard deviation, and the dotted lines indicate the
-    least-squares fit to the ``same'' pairs.}
-  \label{fig:rotations}
-\end{figure}
 
@@
 was generated randomly, but under the constraint that the
-``same''/``different'' pairs created from the shape could both be
-judged according to the ``gold standard''.
+``same''/``different'' pairs created from the shape were
+distinguisable.
 
-We considered three metrics of performance in particular:
-\begin{itemize}
-\item \textit{Response error rates}: How accurate was the model at
-  choosing the correct hypothesis? This was defined as the mean error
-  ($\ME{}$), or fraction of times the model chose incorrectly.
-\item \textit{Rotations}: For those ``same'' pairs which the model
-  judged correct, how correlated were the model's rotations with the
-  true angles of rotation?  We quantified this using the Pearson's
-  correlation coefficient $\rho$ for the true rotation ($R$) versus
-  the extent of the model's rotations ($\vert \mathbf{R}\vert$). The
-  top row in Figure \ref{fig:rotations} shows individual points
-  corresponding to true rotations vs. rotations by the models for all
-  stimuli, including those judged incorrectly. The bottom row shows
-  mean model rotations, across only those stimuli which were judged
-  correctly. We structured the analysis in this way to better compare
-  to the results of \cite{Shepard1971}, which also excluded
+We considered two metrics of performance. First, we looked at
+\textit{response error rates}: How accurate was the model at choosing
+the correct hypothesis? This was defined as the mean error ($\ME{}$),
+or fraction of times the model chose incorrectly.
+
+We also looked at \textit{rotations}: For those ``same'' pairs which
+the model judged correct, how correlated were the model's rotations
+with the true angles of rotation?  We quantified this using the
+Pearson's correlation coefficient $\rho$ for the true rotation ($R$)
+versus the extent of the model's rotations ($\vert
+\mathbf{R}\vert$). The top row in Figure \ref{fig:rotations} shows
+individual points corresponding to true rotations vs. rotations by the
+models for all stimuli, including those judged incorrectly. The bottom
+row shows mean model rotations, across only those stimuli which were
+judged correctly. We structured the analysis in this way to better
+compare to the results of \cite{Shepard1971}, which also excluded
 incorrectly judged stimuli.
-\item \textit{Estimates of $Z$}: How accurate was the model's estimate
-  of $Z$? We defined this quantity as the mean squared error
-  ($\MSE{}$) between the model's estimate of $Z$ and the ``gold
-  standard'' value, where the error has scale such that $\MSE{}=0$
-  indicates no error and $\MSE{}=1$ indicates maximum error. Figure
-  \ref{fig:accuracy} shows plots of the true (``gold standard'') value
-  of $Z$ vs. the model's estimate, for each model.
-\end{itemize}
 
@@
 better than chance (equivalent to guessing randomly,
-i.e. $\ME{}=0.5$). Closer inspection reveals that much of this comes
-from ``different'' pairs ($\NaiveMEdiff{}$), where the asymmetric
-linear interpolation may give an overestimate of $Z$ (e.g., Figure
-\ref{fig:li}).  This intuition is confirmed by the model's accuracy in
-estimating $Z$, which was $\NaiveMSE{}$. As shown in the left panel of
-Figure \ref{fig:accuracy}, it overestimated $Z$ for nearly all
-stimuli.
+i.e. $\ME{}=0.5$). The correlation between the \naive{} model's
+average rotation and the true angle of rotation was $\Naivecorr{}$
+(Figure \ref{fig:rotations}, bottom left). More complex patterns are
+revealed by examining the shape of the raw data in Figure
+\ref{fig:rotations}, top left: the \naive{} model corresponds
+extremely well to the true angle of rotation for approximately
+$R<\frac{\pi}{2}$. This is unsurprising, because the closer the true
+angle is to zero, the less the model has to rotate, and the less
+likely it will get stuck on local maxima. Thus, it is more likely to
+locate the global maximum, which corresponds to the true angle of
+rotation. For $R>\frac{\pi}{2}$, we see an increasing tendency to
+under-rotate due to getting stuck on local maxima, as well as a
+tendency to over-rotate if the wrong direction was initially chosen.
 
-The correlation between the \naive{} model's average rotation and the
-true angle of rotation was $\Naivecorr{}$ (Figure \ref{fig:rotations},
-bottom left). More complex patterns are revealed by examining the
-shape of the raw data in Figure \ref{fig:rotations}, top left: the
-\naive{} model corresponds extremely well to the true angle of
-rotation for approximately $R<\frac{\pi}{2}$. This is unsurprising,
-because the closer the true angle is to zero, the less the model has
-to rotate, and the less likely it will get stuck on local
-maxima. Thus, it is more likely to locate the global maximum, which
-corresponds to the true angle of rotation. For $R>\frac{\pi}{2}$, we
-see an increasing tendency to under-rotate due to getting stuck on
-local maxima, as well as a tendency to over-rotate if the wrong
-direction was initially chosen.
+\paragraph{Bayesian Quadrature}
 
-\paragraph{Parametric (Von Mises)}
-
-The parametric model's error in determining whether the objects were
-identical was $\VMME{}$, making more errors on ``same'' pairs
-($\VMMEsame{}$) than ``different'' pairs ($\VMMEdiff{}$). Rather than
-overestimating as with the \naive{} model, however, the parametric
-model largely underestimated $Z$. This is unsurprising: because the
-Von Mises distribution only has a single peak, it necessarily
-underestimates $Z$ for multimodal similarity functions. Overall, the
-parametric model had an error rate of $\VMMSE{}$ in estimating $Z$,
-which was significantly more accurate than that of the \naive{} model.
-
-The correlation between the parametric model's average rotation and
-the true angle of rotation was also higher ($\VMcorr{}$), though this
-is because more of the underestimated stimuli pairs were excluded from
-the analysis. The individual rotations performed by the parametric and
-\naive{} models were actually identical (see Figure
-\ref{fig:rotations}, upper left and center) because both models use
-the same hill-climbing stopping strategy.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.9\textwidth]{../figures/Z_accuracy.pdf}
-  \caption{\textbf{Accuracy in estimating $Z$.} Each subplot shows the
-    true (``gold standard'') value of $Z$ vs. the value estimated by
-    the model. Black dotted lines indicate a perfect 1:1
-    correspondence between the true and estimated values. The \naive{}
-    model (left panel) tends to overestimate. The parametric model
-    (center panel) is very accurate for about half the ``same'' pairs,
-    and severely underestimates the rest. The nonparametric model
-    (right panel) maintains a decent correspondence with the true
-    $Z$.}
-  \label{fig:accuracy}
-\end{figure}
-
-\paragraph{Nonparametric (Bayesian Quadrature)}
-
-The nonparametric model was much more accurate in choosing the correct
-hypothesis than the other two models ($\BQME{}$). We note that this is
-very close to the 3.2\% error rate reported by \cite{Shepard1971},
-though further experimentation is necessary to determine whether the
-types of errors people make align with those of the model.
-
-The average rotations computed by the nonparametric model were
-strongly correlated with the true rotations ($\BQcorr$, see Figure
-\ref{fig:rotations}, bottom right). Because the nonparametric model
-has the capacity to ``reset'', it could recover from rotating in the
-incorrect direction (e.g., Figure \ref{fig:bq}) and thus did not
-over-rotate as frequently.  It also under-rotated less frequently:
+The BQ model was much more accurate in choosing the correct hypothesis
+than the other two models ($\BQME{}$). The average rotations computed
+by the BQ model were strongly correlated with the true rotations
+($\BQcorr$, see Figure \ref{fig:rotations}, bottom right). Because the
+BQ model has the capacity to ``reset'', it could recover from rotating
+in the incorrect direction (e.g., Figure \ref{fig:bq}) and thus did
+not over-rotate as frequently.  It also under-rotated less frequently:
 because the model continues to rotate until it is confident that its
 estimate of $Z$ is accurate, it does not get stuck as easily on local
-optima.  Indeed, it was largely successful, with $\BQMSE{}$.
+optima.
 
@@
 analysis to characterize optimal strategies for performing mental
-rotation. We found that an adaptive, nonparametric model performs best
-at the classic rotation task \cite{Shepard1971}, as it is able to
-actively monitor the confidence of its estimate and intelligently
-choose the direction of rotation. Two simpler models based on
-heuristics performed much worse: they did not maintain the linear
-relationship with the true angle of rotation due to lack of robustness
-against local maxima or incorrect rotation direction. The
-nonparametric model qualitatively approximates human behavioral
-judgments, both in linearity \cite{Shepard1971} and variability
-\cite{Just1976}. Future work will collect empirical data from
-participants to perform a more thorough, quantitative analysis.  We
-conclude that from this initial survey, models which take an
-``active'', directed approach seem well-suited to explaining mental
-simulation.
+rotation. We found that an adaptive, BQ model performs best at the
+classic rotation task \cite{Shepard1971}, as it is able to actively
+monitor the confidence of its estimate and intelligently choose the
+direction of rotation. Two simpler models based on heuristics
+performed much worse: they did not maintain the linear relationship
+with the true angle of rotation due to lack of robustness against
+local maxima or incorrect rotation direction.  More broadly, the BQ
+model provides answers to puzzling questions surrounding the
+incremental nature of mental rotation: which way should the object be
+rotated, and for how long? This model formalizes these answers in a
+way that is qualitatively consistent with human behavior, both in
+response time linearity \cite{Shepard1971} and variability
+\cite{Just1976}. These results further indicate that this may be
+another instance in which people do appropriately use available
+computational resources to solve the task at hand.  Rational, adaptive
+approaches such as this may therefore be successful when applied to
+more complex cognitive processes above and beyond mental rotation.
 

### 2013-10-15 man/nips-2013-bayesian-optimization.tex
Trim down the submission to 4 pages plus references

@@
 \author{Jessica B.~Hamrick and Thomas L.~Griffiths\\
-  Department of Psychology\\
-  University of California, Berkeley\\
+  Department of Psychology, University of California, Berkeley\\
   Berkeley, CA 94720\\
@@
   psychological task of determining if two images depict the same
-  object in different orientations. We compared three models on this
+  object in different orientations. We compared two models on this
   mental rotation task, and found that a model based on an optimal
@@
   accurate and qualitatively more consistent with classic behavioral
-  data than the other models we evaluated. We suggest that rational
-  models which adaptively exploit available resources are promising in
-  their ability to characterize metacognitive processes like mental
+  data than a simpler model. We suggest that rational models which
+  adaptively exploit available resources are promising in their
+  ability to characterize metacognitive processes like mental
   simulation.
@@
 
-Consider the images on the left in Figure
+Consider the images on the left in Fig.
 \ref{fig:mental-rotation}. In each panel, are the two depicted objects
@@
 
-The plan of the paper is as follows. First, we present the problem
-domain and a computational-level analysis of how to solve it. We
-describe two different models which approximate this solution,
-evaluate their accuracy at approximating the computational solution,
-and qualitatively compare their behavior to that of the classic mental
-rotation results. We end with a brief summary and conclusion of our
-approach.
-
 \begin{figure}[t]
   \centering
-  \includegraphics[width=0.7\textwidth]{../figures/shepard-rotation.png}
+  \includegraphics[width=0.9\textwidth]{../figures/shepard-rotation.png}
   \caption{\textbf{Classic mental rotation task}. Participants in
@@
 
-% \section{Background}
-
-% There is a long and rich history of work on mental rotation beginning
-% with Shepard and Metzler \cite{Shepard1971}, who presented
-% participants with pairs of images such as those shown in Figure
-% \ref{fig:mental-rotation} (left panel). These pairs of
-% three-dimensional objects could either be: the same object rotated in
-% the plane (Figure \ref{fig:mental-rotation}A), the same object rotated
-% in depth (Figure \ref{fig:mental-rotation}B), or two different objects
-% (Figure \ref{fig:mental-rotation}C; in this case, the ``different''
-% objects had reflective symmetry, but no rotational
-% symmetry). Participants had to determine whether the two images were
-% of the same object or not, and Shepard and Metzler famously found that
-% response times for plane and depth rotations had a strong linear
-% relationship with the minimum angle of rotation between the two
-% objects (Figure \ref{fig:mental-rotation}, right panel). The
-% conclusion was that participants were visually ``rotating'' the
-% objects in their minds.
-
-% For many years, this idea was contested, with some researchers arguing
-% the underlying cognitive processes were not visual in nature (e.g.,
-% \cite{Pylyshyn1981}). In particular, Anderson \cite{Anderson1978}
-% proved that the mental imagery debate could not be resolved on the
-% basis of response time data alone. In recent years, however,
-% significant efforts have investigated the neural underpinnings of the
-% mental rotation phenomena, arguing that the results of brain imaging
-% studies support the existence of visual imagery
-% \cite{Kosslyn:2009tj,Kosslyn1988}.  We assume that mental imagery is
-% indeed a visual process, and turn to the question of how, and when, it
-% is used. Previous work has examined how people might use imagery
-% \cite{Hegarty2004, Schwartz1999} to solve reasoning problems, and when
-% they might use imagery as opposed to a symbolic rule
-% \cite{Schwartz:1996uy}. Eye-tracking studies of people performing
-% mental rotations suggest strategies that people might be using, such
-% as aligning smaller features of the images \cite{Just1976}. People
-% seem to use imagery methodically, which raises the question: what is
-% the method?
-
-% Others studying people's pattern of behavior when engaged in
-% self-directed learning have found that they tend to use a strategy of
-% \textit{active learning} \cite{Gureckis:2012gu,Markant:2012uu}. When
-% given a choice as to the information they can obtain, people will not
-% choose the information randomly (as is the case with a \naive{} Monte
-% Carlo sampler), but according to some utility function such as
-% information gain \cite{Markant:2012uu,Nelson2007}. We propose that
-% such a strategy can be combined with the notion of mental imagery as a
-% tool: an optimal model of mental rotation may, perhaps, rely on an
-% active learning strategy. In the next sections, we investigate this
-% idea with several different approaches.
-
 \section{Computational-level model}
@@
 \textit{computational} level \cite{Marr:1983to}: what is the problem
-to be solved, and what is the optimal way to do so?  Formally, people
-are presented with two images, $X_a$ and $X_b$, which are the
-coordinates of the vertices of two-dimensional shapes similar to those
-used by \cite{Cooper:1975wp} (e.g., Figure
-\ref{fig:stimuli}). Participants must determine whether $X_a$ and
-$X_b$ were generated from the same (albeit possibly transformed and
-permuted) original shape, i.e., whether $\exists R,M\textrm{ s.t. }
-X_b=MRX_a$, where $M$ is a permutation matrix and $R$ is a rotation
-matrix.
+to be solved?  Formally, people are presented with two images, $X_a$
+and $X_b$, which are the coordinates of the vertices of 2D shapes
+(e.g., Fig.  \ref{fig:stimuli}). Participants must determine whether
+$X_a$ and $X_b$ depict the same shape, i.e., whether $\exists
+R\textrm{ s.t. } X_b=RX_a$, where $R$ is a rotation matrix. We can
+formulate the judgment of whether $X_a$ and $X_b$ have the same
+origins by deciding about two hypotheses, $h_0$: $\forall R\ X_b\neq
+RX_a$ and $h_1$: $\exists R\textrm{ s.t. } X_b=RX_a$.  To compare the
+hypotheses, we need to compute the posterior for each: $p(h\ \vert\
+X_a, X_b)\propto p(X_a, X_b\ \vert\ h)p(h)$. Assuming the hypotheses
+are equally likely \textit{a priori}, the prior term $p(h)$ will
+cancel out when comparing $h_0$ and $h_1$, thus allowing us to focus
+on the likelihoods, which are $p(X_a,\ X_b\ \vert \ h_0)=p(X_a)p(X_b)$
+for $h_0$ and $p(X_a,\ X_b\ \vert \ h_1)=\int_R p(X_a) p(X_b\vert
+X_a,R) p(R)\ \mathrm{d}R$ for $h_1$. From these likelihoods, we
+compute the ratio $\ell$ which is given by $\ell=\left(\int_R p(X_b\
+  \vert\ X_a, R)p(R)\ \mathrm{d}R\right) /\ p(X_b)$. If $\ell<1$, then
+$h_0$ is the more likely hypothesis. If $\ell>1$, then $h_1$ is the
+more likely hypothesis.
 
-We can formulate the judgment of whether $X_a$ and $X_b$ have the same
-origins by deciding about two hypotheses, $h_0$: $\forall M,R\ X_b\neq
-MRX_a$ and $h_1$: $\exists M,R\textrm{ s.t. } X_b=MRX_a$.  To compare
-the hypotheses, we need to compute the posterior for each: $p(h\
-\vert\ X_a, X_b)\propto p(X_a, X_b\ \vert\ h)p(h)$.
+\section{Algorithmic approximation}
 
-Assuming the hypotheses are equally likely \textit{a priori}, the
-prior term $p(h)$ will cancel out when comparing $h_0$ and $h_1$, thus
-allowing us to focus on the likelihoods:
+We define the prior probability of shape $X$ to be
+$p(X)=n!\left(\frac{1}{2\pi}\right)^n$ according to a generative
+procedure\footnote{A set of $n$ vertices could be chosen in any of
+  $n!$ different ways, and each vertex is located at a random angle
+  (between 0 and $2\pi$) and radius (between 0 and 1).}. This gives us
+the denominator of $\ell$. Computing the numerator is more difficult,
+as we we cannot compute $p(X_b\vert X_a, R)$ directly. Instead, we
+introduce a new variable $X_R$ denoting a mental image, which
+approximates $RX_a$. The $X_R$ are generated by repeated application
+of a function $\tau$, i.e. $X_R=RX_a=\tau(X_{R-r},
+r)=\tau(\tau(X_{R-2r}, r), r)=\ldots{}=\tau^{(\frac{R}{r})}(X_a, r)$,
+where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
+applications of $\tau$. Using this sequential function, we get:
 \begin{align}
-  p(X_a,\ X_b\ \vert \ h_0)&=p(X_a)p(X_b), \label{eq:lh-h0}\\
-  p(X_a,\ X_b\ \vert \ h_1)&=\int_R\int_M p(X_a) p(X_b\vert X_a,R,M) p(R) p(M)\ \mathrm{d}M\ \mathrm{d}R. \label{eq:lh-h1}
+  &p(X_a, X_b\ \vert \ h_1)=\int_R \int_{X} p(X_b\vert X) p(X\vert X_a, R)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
+  &= \int_R \int_X p(X_b\vert X)\delta(\tau^{(\frac{R}{r})}(X_a, r)-X)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R = \int_R p(X_b\vert X_R)p(X_a)p(R)\ \mathrm{d}R
 \end{align}
-Under $h_0$, the likelihood is easy to compute because we assume that
-$X_a$ and $X_b$ are independent (Equation \ref{eq:lh-h0}). When the
-configurations are the same, the likelihood becomes more complicated
-(Equation \ref{eq:lh-h1}). For a small number of vertices, we can
-compute the integral over $M$ by enumerating every possible mapping
-between $X_a$ and $X_b$. After doing so, we obtain:
-\begin{equation} 
-  p(X_a,\ X_b\ \vert \ h_1)=\int_R p(X_a) p(X_b\vert X_a,R) p(R)\ \mathrm{d}R.
-\end{equation}
+However, the exact form of $p(X_b\vert X_R)$ is still unknown. We
+approximate it with a similarity function $S(X_b, X_R)$, and denote
+the resulting integral as $Z=\int_R S(X_b, X_R)p(R)\
+\mathrm{d}R\approx \int_R p(X_b\vert X_R)p(R)\ \mathrm{d}R$.  We
+define the similarity based on Gaussian similarity and possible
+mappings $M$ of the vertices\footnote{Because the vertices are
+  connected in a way which forms a closed loop, we need only consider
+  $2n$ mappings of the $n$ vertices (we assume uncertainty for which
+  is the ``first'' vertex, and then which of its two neighbors is the
+  ``second''). So, the possible orderings are of the form
+  $M=\lbrace{}0, 1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0, \ldots{},
+  n-1\rbrace{}$, etc.}, i.e. $S(X_b,
+X_R)=\frac{1}{2n}\sum_{M\in\mathbb{M}}\prod_{i=1}^n\mathcal{N}(X_b[i]\
+\vert \ (MX_R)[i], \Sigma)$ where $i$ denotes the $i^{th}$ vertex. An
+example stimulus and corresponding $S$ is shown in Fig.
+\ref{fig:shapes}.
 
-Once we have computed both likelihoods, we compute their ratio $\ell$
-which is given by $\ell=\left(\int_R p(X_b\ \vert\ X_a, R)p(R)\
-  \mathrm{d}R\right) /\ p(X_b)$. If $\ell<1$, then $h_0$ is the more
-likely hypothesis. If $\ell>1$, then $h_1$ is the more likely
-hypothesis.
+To summarize, the process of generating a mental image consists of
+computing a single $X_R$ and then computing $S(X_b, X_R)$. We denote
+the sequence of rotations computed by this procedure as
+$\mathbf{R}=\{R_1, R_2, \ldots{}\}$. However, this sequence cannot be
+arbitrary, as mental rotation is computationally demanding. Our goal
+is to minimize the number of rotations $\vert\mathbf{R}\vert$ while
+still obtaining an estimate of $Z$ that is accurate enough to choose
+the correct hypothesis.
+
+\paragraph{\Naive{}}
+
+We defined a \naive{} model which performs a hill-climbing search over
+the similarity function until it reaches a (possibly local)
+maximum. Once a maximum as been found, the model computes an estimate
+of $Z$ by linearly interpolating between sampled rotations (e.g.,
+Fig. \ref{fig:li}).
+
+\paragraph{Bayesian Quadrature}
+
+A more flexible strategy uses what is known as \emph{Bayesian
+  Quadrature} (BQ) \cite{Diaconis:1988uo,OHagan:1991tx} to estimate
+$Z$.  BQ computes a posterior distribution over $Z$ by placing a
+Gaussian Process (GP) prior on the function $S$ and evaluating $S$ at
+a particular set of points. However, althought $S$ is a non-negative
+likelihood function, GP regression enforces no such
+constraint. \cite{Osborne:2012tm} give a method to place a prior over
+the log likelihood\footnote{In practice, as in \cite{Osborne:2012tm},
+  we use the transform of $\log(S+1)$. Additionally, we follow
+  \cite{Osborne:2012tm} and use a combination of MLII and
+  marginalization to fit the $w$ kernel parameters. The output scales
+  are fixed at $h_S=\sqrt{0.1}$ and $h_{\log S}=\log(h_S + 1)$.}, thus
+ensuring that $S=e^{\log S}$ will be positive, i.e.  $E[Z\ \vert \
+\log S]=\int_{\log S}\left(\int_R \exp(\log{S(X_b,X_R)})p(R)\
+  \mathrm{d}R\right)\mathcal{N}\left(\log{S}\ \vert \ \mu_{\log S},
+  \Sigma_{\log S}\right)\ \mathrm{d}\log S$, where $\mu_{\log S}$ and
+$\Sigma_{\log S}$ are the mean and covariance, respectively, of the GP
+over $\log S$ given $\mathbf{R}$. We approximate this according to the
+method given in \cite{Osborne:2012tm}, i.e. $\mu_Z=E[Z\ \vert \ S,
+\log S, \Delta_c] \approx \int_R \mu_{S}(1 + \mu_{\Delta_c}) p(R)\
+\mathrm{d}R$, where $\mu_S$ is the mean of a GP over $S$ given
+$\mathbf{R}$; and $\mu_{\Delta_c}$ of a GP over $\Delta_c=\mu_{\log S}
+- \log \mu_S$ given $\mathbf{R}_c$, which consists of $\mathbf{R}$ and
+a set of intermediate \emph{candidate points} $c$ as described in
+\cite{Osborne:2012tm}. The variance is $\tilde{V}(Z\vert S, \log S,
+\Delta_c)$ as defined in Equation 12 of \cite{Osborne:2012tm}.
 
@@
 
-\section{Metacognitive algorithms}
-
-Given the computational model defined above, we now explore the ways
-in which it can be expressed algorithmically. First, we define the
-prior probabilities over stimuli according to a generative
-procedure. A set of $n$ vertices could be chosen in any of $n!$
-different ways, and each vertex is located at a random angle (between
-0 and $2\pi$) and radius (between 0 and 1). Thus, the prior over a
-shape $X$ is $p(X)=n!\left(\frac{1}{2\pi}\right)^n$, which gives us
-the denominator of $\ell$. Computing the numerator is more difficult,
-as we we cannot compute $p(X_b\vert X_a, R)$ directly. Instead, we
-introduce a new variable $X_R$ denoting a mental image, which
-approximates $RX_a$. The $X_R$ are generated by repeated application
-of a function $\tau$:
-\begin{equation}
-  X_R=RX_a=\tau(X_{R-r}, r)=\tau(\tau(X_{R-2r}, r), r)=\ldots{}=\tau^{(\frac{R}{r})}(X_a, r).
-  \label{eq:tau}
-\end{equation} 
-where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
-applications of $\tau$. Using this sequential function, we get:
-\begin{align}
-  p(X_a, X_b\ \vert \ h_1)&=\int_R \int_{X} p(X_b\vert X) p(X\vert X_a, R)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
-  &= \int_R \int_X p(X_b\vert X)\delta(\tau^{(\frac{R}{r})}(X_a, r)-X)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
-  &= \int_R p(X_b\vert X_R)p(X_a)p(R)\ \mathrm{d}R
-\end{align}
-However, the exact form of $p(X_b\vert X_R)$ is still unknown. We
-approximate it with a similarity function $S(X_b, X_R)$, and denote
-the resulting integral as $Z$:
-\begin{equation}
-Z=\int_R S(X_b, X_R)p(R)\ \mathrm{d}R\approx \int_R p(X_b\vert X_R)p(R)\ \mathrm{d}R.
-\label{eq:Z}
-\end{equation}
-We define the similarity function $S$, which incorporates the
-different mappings as vertices, as follows. Because the vertices are
-connected in a way which forms a closed loop, we need only consider
-$2n$ mappings of the $n$ vertices (we assume uncertainty for which is
-the ``first'' vertex, and then which of its two neighbors is the
-``second''). So, the possible orderings are of the form $M=\lbrace{}0,
-1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0, \ldots{}, n-1\rbrace{}$,
-and so on. Combining this with a Gaussian similarity metric, we
-obtain:
-\begin{equation}
-  S(X_b, X_R)=\frac{1}{2n}\sum_{M\in\mathbb{M}}\prod_{i=1}^n\mathcal{N}(X_b[i]\ \vert \ (MX_R)[i], \Sigma)
-  \label{eq:similarity}
-\end{equation}
-where $i$ denotes the $i^{th}$ vertex. An example stimulus and
-corresponding $S$ is shown in Figure \ref{fig:shapes}.
-
-To summarize, the process of generating a mental image consists of
-computing a single $X_R$ (as in Equation \ref{eq:tau}) and then
-computing $S(X_b, X_R)$. We denote the sequence of rotations computed
-by this procedure as $\mathbf{R}=\{R_1, R_2, \ldots{}\}$. However,
-this sequence cannot be arbitrary, as mental rotation is
-computationally demanding. Our goal is to minimize the number of
-rotations $\vert\mathbf{R}\vert$ while still obtaining an estimate of
-$Z$ that is accurate enough to choose the correct hypothesis. With
-this in mind, we now examine several approaches to solve this problem.
-
-\paragraph{\Naive{}}
-
-As a baseline, we defined a \naive{} model which performs a
-hill-climbing search over the similarity function until it reaches a
-(possibly local) maximum. Once a maximum as been found, the model
-computes an estimate of $Z$ by linearly interpolating between sampled
-rotations. Figure \ref{fig:li} shows an example of the \naive{}
-model's estimate of $S$.
-
-\paragraph{Bayesian Quadrature}
-
-A more flexible strategy uses what is known as \emph{Bayesian
-  Quadrature} \cite{Diaconis:1988uo,OHagan:1991tx} to estimate $Z$.
-Bayesian Quadrature (BQ) allows us to compute a posterior distribution
-over $Z$ by placing a Gaussian Process (GP) prior on the function $S$
-and evaluating $S$ at a particular set of points. Because our data is
-circular, we use a periodic kernel \cite{Rasmussen:2006vz}.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
-  \caption{\textbf{Model rotations.} Top: each subplot shows the
-    correspondence between the true angle of rotation ($R$) for
-    ``same'' pairs and the amount of rotation performed by the
-    model. Bottom: each subplot shows the models' mean rotations over
-    stimuli pairs that were judged correctly. Black dots correspond to
-    ``same'' pairs, and blue lines correspond to ``different'' pairs
-    (for which the mean rotation is constant across true rotations, as
-    the true rotation is undefined). Error bars/shaded regions
-    indicate one standard deviation, and the dotted lines indicate the
-    least-squares fit to the ``same'' pairs.}
-  \label{fig:rotations}
-\end{figure}
-
-BQ has its difficulties, however. While in our case $S$ is a
-non-negative likelihood function, GP regression enforces no such
-constraint. In an effort to avoid this problem, \cite{Osborne:2012tm}
-give a method to place a prior over the log likelihood\footnote{In
-  practice, as in \cite{Osborne:2012tm}, we use the transform of
-  $\log(S+1)$. Additionally, we follow \cite{Osborne:2012tm} and use a
-  combination of MLII and marginalization to fit the $w$ kernel
-  parameters. The output scales are fixed at $h_S=\sqrt{0.1}$ and
-  $h_{\log S}=\log(h_S + 1)$.}, thus ensuring that $S=e^{\log S}$ will
-be positive:
-\begin{equation*}
-  E[Z\ \vert \ \log S]=\int_{\log S}\left(\int_R \exp(\log{S(X_b,X_R)})p(R)\ \mathrm{d}R\right)\mathcal{N}\left(\log{S}\ \vert \ \mu_{\log S}, \Sigma_{\log S}\right)\ \mathrm{d}\log S
-\end{equation*}
-where $\mu_{\log S}$ and $\Sigma_{\log S}$ are the mean and
-covariance, respectively, of the GP regression over $\log S$ given
-$\mathbf{R}$. We approximate this according to the method given in
-\cite{Osborne:2012tm}:
-\begin{equation}
-  \mu_Z=E[Z\ \vert \ S, \log S, \Delta_c] \approx \int_R \mu_{S}(1 + \mu_{\Delta_c}) p(R)\ \mathrm{d}R 
-  \label{eq:bq-Z-mean}
-\end{equation}
-where $\mu_S$ is the mean of a GP regression over $S$ given
-$\mathbf{R}$; and $\mu_{\Delta_c}$ is a regression over
-$\Delta_c=\mu_{\log S} - \log \mu_S$ given $\mathbf{R}_c$, which
-consists of $\mathbf{R}$ and a set of intermediate \emph{candidate
-  points} $c$ as described in \cite{Osborne:2012tm}. The variance is
-$\tilde{V}(Z\vert S, \log S, \Delta_c)$ as defined in Equation 12 of
-\cite{Osborne:2012tm}.
-
-To start, we pick an initial direction of rotation which results in
-the higher value of $S$. At each step, we compute $\mu_Z$ and
-$\tilde{V}$ in order to estimate a distribution over the likelihood
-ratio $\ell$, i.e.  $p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\
-\vert\ \mu_Z, \sigma_z)$.  We choose $h_0$ when $p(\ell < 1)\geq
-0.95$, and $h_1$ when $p(\ell > 1)\geq 0.95$. Until one of these
-conditions are met (or the shape has been fully rotated), the model
-will continues to compute rotations and update its estimate of $Z$.
-
-We additionally allow the model to change direction or ``reset'' based
-on an estimate of the posterior variance of $Z$ given some new sample
-$a$. This is similar to the procedure given in \cite{Osborne:2012tm},
-however we do not compute the full posterior variance. Instead, we
-compute the variance given only the current mean estimate of $a$. If
-this estimated variance is lowered more by ``resetting'', then the
-model will change directions. Thus, it is able to have an active say
-in which mental rotations are computed, unlike the hill-climbing
-procedure.
+We pick an initial direction of rotation which results in the higher
+value of $S$. At each step, we compute $\mu_Z$ and $\tilde{V}$ to
+estimate a distribution over the likelihood ratio $\ell$, i.e.
+$p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\ \vert\ \mu_Z,
+\sigma_z)$.  We choose $h_0$ when $p(\ell < 1)\geq 0.95$, and $h_1$
+when $p(\ell > 1)\geq 0.95$. Until one of these conditions are met (or
+the shape has been fully rotated), the model will continue to compute
+rotations and update its estimate of $Z$.  The model may also change
+direction based on the \textit{expected} posterior variance of $Z$
+given some new sample $a$\footnote{This is similar to the full
+  posterior variance calculation given in \cite{Osborne:2012tm},
+  however we compute the variance given only the current mean estimate
+  of $a$.}. The model will change directions if doing so would lower
+this expected variance. Thus, it is able to actively change its
+strategy, unlike the hill-climbing procedure.
 
@@
 
-We evaluated each model's performance on 20 shapes which had between 3
-and 6 vertices, inclusive, (e.g., $X_a$ in Figure
+We evaluated each model's performance on 20 randomly generated shapes
+which had between 3 and 6 vertices, inclusive, (e.g., $X_a$ in Fig.
 \ref{fig:stimuli}). For each shape, we computed 18 ``same'' and 18
@@
 created by rotating $X_a$ by $R$; the same was true for ``different'',
-except that $X_a$ was also reflected across the $y$-axis. Each shape
-was generated randomly, but under the constraint that the
-``same''/``different'' pairs created from the shape were
-distinguisable.
-
-We considered two metrics of performance. First, we looked at
-\textit{response error rates}: How accurate was the model at choosing
-the correct hypothesis? This was defined as the mean error ($\ME{}$),
-or fraction of times the model chose incorrectly.
-
-We also looked at \textit{rotations}: For those ``same'' pairs which
-the model judged correct, how correlated were the model's rotations
-with the true angles of rotation?  We quantified this using the
-Pearson's correlation coefficient $\rho$ for the true rotation ($R$)
-versus the extent of the model's rotations ($\vert
-\mathbf{R}\vert$). The top row in Figure \ref{fig:rotations} shows
-individual points corresponding to true rotations vs. rotations by the
-models for all stimuli, including those judged incorrectly. The bottom
-row shows mean model rotations, across only those stimuli which were
-judged correctly. We structured the analysis in this way to better
-compare to the results of \cite{Shepard1971}, which also excluded
-incorrectly judged stimuli.
+except that $X_a$ was also reflected across the $y$-axis. To gauge
+performance, we looked at \textit{response error rates}: how accurate
+was the model at choosing the correct hypothesis? This was defined as
+the mean error ($\ME{}$), or fraction of times the model chose
+incorrectly.  We additionally looked at \textit{rotations}: for those
+``same'' pairs which the model judged correct, how correlated were the
+model's rotations with the true angles of rotation?  We quantified
+this using the Pearson's correlation coefficient $\rho$ for the true
+rotation ($R$) versus the number of steps/rotations take by the model
+($\vert \mathbf{R}\vert$). Fig. \ref{fig:rotations} shows true
+rotations vs number of steps/rotations by the model for individual
+stimuli (left) and average per true rotation (right). This analysis
+can be qualitatively compared to the results of \cite{Shepard1971}
+(Fig.  \ref{fig:mental-rotation}).
 
@@
 average rotation and the true angle of rotation was $\Naivecorr{}$
-(Figure \ref{fig:rotations}, bottom left). More complex patterns are
-revealed by examining the shape of the raw data in Figure
-\ref{fig:rotations}, top left: the \naive{} model corresponds
-extremely well to the true angle of rotation for approximately
-$R<\frac{\pi}{2}$. This is unsurprising, because the closer the true
-angle is to zero, the less the model has to rotate, and the less
-likely it will get stuck on local maxima. Thus, it is more likely to
-locate the global maximum, which corresponds to the true angle of
-rotation. For $R>\frac{\pi}{2}$, we see an increasing tendency to
-under-rotate due to getting stuck on local maxima, as well as a
-tendency to over-rotate if the wrong direction was initially chosen.
+(Fig. \ref{fig:rotations}, bottom left). For $R<\frac{\pi}{2}$, the
+\naive{} model corresponds extremely well to the true angle of
+rotation; this is because it needs to rotate less and is therefore
+less likely to get stuck on local maxima. For $R>\frac{\pi}{2}$, we
+see an increasing tendency to under-rotate due to getting stuck on
+local maxima, as well as a tendency to over-rotate if the wrong
+direction was initially chosen.
 
@@
 The BQ model was much more accurate in choosing the correct hypothesis
-than the other two models ($\BQME{}$). The average rotations computed
-by the BQ model were strongly correlated with the true rotations
-($\BQcorr$, see Figure \ref{fig:rotations}, bottom right). Because the
-BQ model has the capacity to ``reset'', it could recover from rotating
-in the incorrect direction (e.g., Figure \ref{fig:bq}) and thus did
-not over-rotate as frequently.  It also under-rotated less frequently:
-because the model continues to rotate until it is confident that its
-estimate of $Z$ is accurate, it does not get stuck as easily on local
-optima.
+than the other model ($\BQME{}$). The number of rotations computed by
+the BQ model were strongly correlated with the true rotations
+($\BQcorr$), a result which is qualitatively similar to that exhibited
+by humans (Fig. \ref{fig:mental-rotation}
+vs. Fig. \ref{fig:rotations}). Because the BQ model has the capacity
+to ``reset'', it could recover from rotating in the incorrect
+direction (e.g., Fig. \ref{fig:bq}) and thus did not over-rotate as
+frequently.  It also under-rotated less frequently: because the model
+continues to rotate until it is confident in its estimate of $Z$, it
+does not get stuck as easily on local optima.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
+  \caption{\textbf{Model rotations.} Left: each subplot shows the
+    correspondence between the true angle of rotation ($R$) for
+    ``same'' pairs and the amount of rotation performed by the
+    model. Right: each subplot shows the models' mean rotations over
+    stimuli pairs that were judged correctly. Black dots correspond to
+    ``same'' pairs, and blue lines correspond to ``different''
+    pairs. Error bars/shaded regions indicate one standard deviation,
+    and the dotted lines indicate the least-squares fit to the
+    ``same'' pairs.}
+  \label{fig:rotations}
+\end{figure}
 
@@
 investigated the specific case of mental rotation, using rational
-analysis to characterize optimal strategies for performing mental
-rotation. We found that an adaptive, BQ model performs best at the
-classic rotation task \cite{Shepard1971}, as it is able to actively
-monitor the confidence of its estimate and intelligently choose the
-direction of rotation. Two simpler models based on heuristics
-performed much worse: they did not maintain the linear relationship
-with the true angle of rotation due to lack of robustness against
-local maxima or incorrect rotation direction.  More broadly, the BQ
-model provides answers to puzzling questions surrounding the
-incremental nature of mental rotation: which way should the object be
-rotated, and for how long? This model formalizes these answers in a
-way that is qualitatively consistent with human behavior, both in
-response time linearity \cite{Shepard1971} and variability
-\cite{Just1976}. These results further indicate that this may be
-another instance in which people do appropriately use available
-computational resources to solve the task at hand.  Rational, adaptive
-approaches such as this may therefore be successful when applied to
-more complex cognitive processes above and beyond mental rotation.
-
-% two questions: which way to rotate, when to stop, giving formal answer
-% to these questions which produces answers consistent with human
-% behavior. instance of showing people are appropriately using
-% computationl resources
-
-% \textbf{Acknowledgments}
-% \TODO{}
+analysis to characterize optimal strategies. We found that an
+adaptive, BQ model performs best at the classic rotation task
+\cite{Shepard1971}, as it is able to actively monitor the confidence
+of its estimate and intelligently choose the direction of rotation. A
+simpler model based on heuristics performed much worse: it did not
+maintain the linear relationship with the true angle of rotation due
+to lack of robustness against local maxima or incorrect rotation
+direction.  More broadly, the BQ model provides answers to puzzling
+questions surrounding the incremental nature of mental rotation: which
+way should the object be rotated, and for how long? This model
+formalizes these answers in a way that is qualitatively consistent
+with human behavior, both in response time linearity
+\cite{Shepard1971} and variability \cite{Just1976}. These results
+further indicate that this may be another instance in which people do
+appropriately use available computational resources to solve the task
+at hand.  Rational, adaptive approaches such as this may therefore be
+successful when applied to more complex cognitive processes above and
+beyond mental rotation.
 

### 2013-10-15 code/plots.py
Figure tweaks

@@ -356,20 +356,31 @@ def likelihood_all(models):
 
 def model_rotations(models):
     nm = len(models)
-    fig, axes = plt.subplots(1, 2*nm, sharex=True, sharey=True)
-    ax0 = axes[0]
 
-    ax0.set_xticks([0, np.pi / 4., np.pi / 2., 3 * np.pi / 4., np.pi])
-    ax0.set_xticklabels([0, 45, 90, 135, 180])
-    ax0.set_xlim(-np.pi/16, np.pi+(np.pi/16))
+    figA, axesA = plt.subplots(1, nm, sharex=True, sharey=True)
+    sg.set_figsize(6, 3)
+    plt.subplots_adjust(
+        wspace=0.2, top=0.9, bottom=0.2,
+        hspace=0.4, left=0.12, right=0.95)
 
-    ax0.set_yticks([0, 25, 50, 75, 100])
-    # ax0.set_yticklabels([0, 90, 180, 270, 360])
-    ax0.set_ylim(-5, 105)
+    figB, axesB = plt.subplots(1, nm, sharex=True, sharey=True)
+    sg.set_figsize(6, 3)
+    plt.subplots_adjust(
+        wspace=0.2, top=0.9, bottom=0.2,
+        hspace=0.4, left=0.1, right=0.95)
+
+    for ax0 in [axesA[0], axesB[0]]:
+        ax0.set_xticks([0, np.pi / 4., np.pi / 2., 3 * np.pi / 4., np.pi])
+        ax0.set_xticklabels([0, 45, 90, 135, 180])
+        ax0.set_xlim(-np.pi/16, np.pi+(np.pi/16))
+
+        ax0.set_yticks([0, 25, 50, 75, 100])
+        # ax0.set_yticklabels([0, 90, 180, 270, 360])
+        ax0.set_ylim(-5, 105)
 
     for i in xrange(len(models)):
-        tax = axes[i]
-        bax = axes[nm + i]
+        tax = axesA[i]
+        bax = axesB[i]
 
         model = models[i]
         tax.set_title(model, fontsize=12)
@@ -384,23 +395,18 @@ def model_rotations(models):
             sg.clear_right(ax=ax)
             sg.clear_top(ax=ax)
 
-    sg.set_figsize(12, 3)
-    plt.subplots_adjust(
-        wspace=0.2, top=0.9, bottom=0.2,
-        hspace=0.4, left=0.1, right=0.95)
-
[…]

### 2013-10-15 man/nips-2013-bayesian-optimization.tex
Small wording changes to workshop submission

@@
 $p(X)=n!\left(\frac{1}{2\pi}\right)^n$ according to a generative
-procedure\footnote{A set of $n$ vertices could be chosen in any of
+procedure.\footnote{A set of $n$ vertices could be chosen in any of
   $n!$ different ways, and each vertex is located at a random angle
-  (between 0 and $2\pi$) and radius (between 0 and 1).}. This gives us
+  (between 0 and $2\pi$) and radius (between 0 and 1).} This gives us
 the denominator of $\ell$. Computing the numerator is more difficult,
@@
 define the similarity based on Gaussian similarity and possible
-mappings $M$ of the vertices\footnote{Because the vertices are
+mappings $M$ of the vertices,\footnote{Because the vertices are
   connected in a way which forms a closed loop, we need only consider
@@
   $M=\lbrace{}0, 1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0, \ldots{},
-  n-1\rbrace{}$, etc.}, i.e. $S(X_b,
+  n-1\rbrace{}$, etc.} i.e. $S(X_b,
 X_R)=\frac{1}{2n}\sum_{M\in\mathbb{M}}\prod_{i=1}^n\mathcal{N}(X_b[i]\
@@
 constraint. \cite{Osborne:2012tm} give a method to place a prior over
-the log likelihood\footnote{In practice, as in \cite{Osborne:2012tm},
+the log likelihood,\footnote{In practice, as in \cite{Osborne:2012tm},
   we use the transform of $\log(S+1)$. Additionally, we follow
@@
   marginalization to fit the $w$ kernel parameters. The output scales
-  are fixed at $h_S=\sqrt{0.1}$ and $h_{\log S}=\log(h_S + 1)$.}, thus
+  are fixed at $h_S=\sqrt{0.1}$ and $h_{\log S}=\log(h_S + 1)$.} thus
 ensuring that $S=e^{\log S}$ will be positive, i.e.  $E[Z\ \vert \
@@
   \end{subfigure}
-  \caption{\textbf{(a)} An example stimulus in which the shapes differ
-    only by a rotation. All stimuli consist of three to six vertices
-    centered around the origin, and edges which create a closed loop
-    from the vertices. The true angle of rotation between $X_a$ and
-    $X_b$ is at $\frac{2\pi}{3}$. \textbf{(b-c)} Likelihood function
-    and \naive{} (b) and BQ (c) model estimates. The sampled points
-    $\mathbf{R}$ (red circles) are then used to estimate $S$ (black
-    lines are the true $S$, red lines are the estimate).}
+  \caption{\textbf{Example model behavior.} \textbf{(a)} An example
+    stimulus in which the shapes differ only by a rotation. All
+    stimuli consist of three to six vertices centered around the
+    origin, and edges which create a closed loop from the
+    vertices. The true angle of rotation between $X_a$ and $X_b$ is at
+    $\frac{2\pi}{3}$. \textbf{(b-c)} Likelihood function and \naive{}
+    (b) and BQ (c) model estimates. The sampled points $\mathbf{R}$
+    (red circles) are then used to estimate $S$ (black lines are the
+    true $S$, red lines are the estimate).}
   \label{fig:shapes}
@@
 direction based on the \textit{expected} posterior variance of $Z$
-given some new sample $a$\footnote{This is similar to the full
+given some new sample $a$.\footnote{This is similar to the full
   posterior variance calculation given in \cite{Osborne:2012tm},
   however we compute the variance given only the current mean estimate
-  of $a$.}. The model will change directions if doing so would lower
+  of $a$.} The model will change directions if doing so would lower
 this expected variance. Thus, it is able to actively change its
@@
 We evaluated each model's performance on 20 randomly generated shapes
-which had between 3 and 6 vertices, inclusive, (e.g., $X_a$ in Fig.
-\ref{fig:stimuli}). For each shape, we computed 18 ``same'' and 18
-``different'' stimuli pairs, with $R$ spaced at $20^\prime$ increments
-between 0 and 360, as in \cite{Shepard1971}. ``Same'' pairs were
-created by rotating $X_a$ by $R$; the same was true for ``different'',
-except that $X_a$ was also reflected across the $y$-axis. To gauge
-performance, we looked at \textit{response error rates}: how accurate
-was the model at choosing the correct hypothesis? This was defined as
-the mean error ($\ME{}$), or fraction of times the model chose
-incorrectly.  We additionally looked at \textit{rotations}: for those
-``same'' pairs which the model judged correct, how correlated were the
-model's rotations with the true angles of rotation?  We quantified
-this using the Pearson's correlation coefficient $\rho$ for the true
-rotation ($R$) versus the number of steps/rotations take by the model
-($\vert \mathbf{R}\vert$). Fig. \ref{fig:rotations} shows true
-rotations vs number of steps/rotations by the model for individual
-stimuli (left) and average per true rotation (right). This analysis
-can be qualitatively compared to the results of \cite{Shepard1971}
-(Fig.  \ref{fig:mental-rotation}).
+which had between three and six vertices, inclusive, (e.g., $X_a$ in
+Fig.  \ref{fig:stimuli}). For each shape, we computed 18 ``same'' and
+18 ``different'' stimuli pairs, with $R$ spaced at $20^\prime$
+increments between 0 and 360, as in \cite{Shepard1971}. ``Same'' pairs
+were created by rotating $X_a$ by $R$; the same was true for
+``different'', except that $X_a$ was also reflected across the
+$y$-axis. To gauge performance, we looked at \textit{response error
+  rates}: how accurate was the model at choosing the correct
+hypothesis? This was defined as the mean error ($\ME{}$), or fraction
+of times the model chose incorrectly.  We additionally looked at
+\textit{rotations}: for those ``same'' pairs which the model judged
+correct, how correlated were the model's rotations with the true
+angles of rotation?  We quantified this using the Pearson's
+correlation coefficient $\rho$ for the true rotation ($R$) versus the
+number of steps/rotations take by the model ($\vert
+\mathbf{R}\vert$). Fig. \ref{fig:rotations} shows true rotations vs
+number of steps/rotations by the model for individual stimuli
+(\ref{fig:all-stimuli}) and average per true rotation
+(\ref{fig:correct-stimuli}). This latter analysis can be qualitatively
+compared to the results of \cite{Shepard1971} (Fig.
+\ref{fig:mental-rotation}).
 
@@
 average rotation and the true angle of rotation was $\Naivecorr{}$
-(Fig. \ref{fig:rotations}, bottom left). For $R<\frac{\pi}{2}$, the
+(Fig. \ref{fig:correct-stimuli}, left). For $R<\frac{\pi}{2}$, the
 \naive{} model corresponds extremely well to the true angle of
@@
 by humans (Fig. \ref{fig:mental-rotation}
-vs. Fig. \ref{fig:rotations}). Because the BQ model has the capacity
-to ``reset'', it could recover from rotating in the incorrect
-direction (e.g., Fig. \ref{fig:bq}) and thus did not over-rotate as
-frequently.  It also under-rotated less frequently: because the model
-continues to rotate until it is confident in its estimate of $Z$, it
-does not get stuck as easily on local optima.
+vs. Fig. \ref{fig:correct-stimuli}, right). Because the BQ model has
+the capacity to ``reset'', it could recover from rotating in the
+incorrect direction (e.g., Fig. \ref{fig:bq}) and thus did not
+over-rotate as frequently.  It also under-rotated less frequently:
+because the model continues to rotate until it is confident in its
+estimate of $Z$, it does not get stuck as easily on local optima.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
-  \caption{\textbf{Model rotations.} Left: each subplot shows the
+  \begin{subfigure}[b]{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/model_rotations_A.pdf}
+    % \vspace{0pt}
+    \caption{All stimuli}
+    \label{fig:all-stimuli}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/model_rotations_B.pdf}
+    \caption{Correct stimuli}
+    \label{fig:correct-stimuli}
+  \end{subfigure}
+  \caption{\textbf{Model rotations.} (a) Each subplot shows the
     correspondence between the true angle of rotation ($R$) for
     ``same'' pairs and the amount of rotation performed by the
-    model. Right: each subplot shows the models' mean rotations over
+    model. (b) Each subplot shows the models' mean rotations over
     stimuli pairs that were judged correctly. Black dots correspond to

### 2013-10-15 man/nips-2013-bayesian-optimization.tex
Remove extra space after 'Fig.'

@@
 
-Consider the images on the left in Fig.
-\ref{fig:mental-rotation}. In each panel, are the two depicted objects
-identical (except for a rotation), or distinct? When presented with
-this mental rotation task, people default to a strategy in which they
-visualize one object rotating until it is congruent with the other
+Consider the images on the left in Fig.~\ref{fig:mental-rotation}. In
+each panel, are the two depicted objects identical (except for a
+rotation), or distinct? When presented with this mental rotation task,
+people default to a strategy in which they visualize one object
+rotating until it is congruent with the other
 \cite{Shepard1971}. There is strong evidence for such mental
@@
 and $X_b$, which are the coordinates of the vertices of 2D shapes
-(e.g., Fig.  \ref{fig:stimuli}). Participants must determine whether
+(e.g., Fig.~\ref{fig:stimuli}). Participants must determine whether
 $X_a$ and $X_b$ depict the same shape, i.e., whether $\exists
@@
 \vert \ (MX_R)[i], \Sigma)$ where $i$ denotes the $i^{th}$ vertex. An
-example stimulus and corresponding $S$ is shown in Fig.
-\ref{fig:shapes}.
+example stimulus and corresponding $S$ is shown in
+Fig.~\ref{fig:shapes}.
 
@@
 of $Z$ by linearly interpolating between sampled rotations (e.g.,
-Fig. \ref{fig:li}).
+Fig.~\ref{fig:li}).
 
@@
 which had between three and six vertices, inclusive, (e.g., $X_a$ in
-Fig.  \ref{fig:stimuli}). For each shape, we computed 18 ``same'' and
+Fig.~\ref{fig:stimuli}). For each shape, we computed 18 ``same'' and
 18 ``different'' stimuli pairs, with $R$ spaced at $20^\prime$
@@
 number of steps/rotations take by the model ($\vert
-\mathbf{R}\vert$). Fig. \ref{fig:rotations} shows true rotations vs
+\mathbf{R}\vert$). Fig.~\ref{fig:rotations} shows true rotations vs
 number of steps/rotations by the model for individual stimuli
@@
 (\ref{fig:correct-stimuli}). This latter analysis can be qualitatively
-compared to the results of \cite{Shepard1971} (Fig.
-\ref{fig:mental-rotation}).
+compared to the results of \cite{Shepard1971}
+(Fig.~\ref{fig:mental-rotation}).
 
@@
 average rotation and the true angle of rotation was $\Naivecorr{}$
-(Fig. \ref{fig:correct-stimuli}, left). For $R<\frac{\pi}{2}$, the
+(Fig.~\ref{fig:correct-stimuli}, left). For $R<\frac{\pi}{2}$, the
 \naive{} model corresponds extremely well to the true angle of
@@
 ($\BQcorr$), a result which is qualitatively similar to that exhibited
-by humans (Fig. \ref{fig:mental-rotation}
-vs. Fig. \ref{fig:correct-stimuli}, right). Because the BQ model has
+by humans (Fig.~\ref{fig:mental-rotation}
+vs. Fig.~\ref{fig:correct-stimuli}, right). Because the BQ model has
 the capacity to ``reset'', it could recover from rotating in the
-incorrect direction (e.g., Fig. \ref{fig:bq}) and thus did not
+incorrect direction (e.g., Fig.~\ref{fig:bq}) and thus did not
 over-rotate as frequently.  It also under-rotated less frequently:

### 2013-10-16 man/nips-2013-bayesian-optimization.tex
Small edits, mostly rewording

@@
   visualize an object rotating--how do you best make use of it? We
-  explored how mental simulation should used in the classic
+  explored how mental simulation should be used in the classic
   psychological task of determining if two images depict the same
@@
 
-We begin by first analyzing mental rotation at Marr's
-\textit{computational} level \cite{Marr:1983to}: what is the problem
-to be solved?  Formally, people are presented with two images, $X_a$
-and $X_b$, which are the coordinates of the vertices of 2D shapes
-(e.g., Fig.~\ref{fig:stimuli}). Participants must determine whether
-$X_a$ and $X_b$ depict the same shape, i.e., whether $\exists
-R\textrm{ s.t. } X_b=RX_a$, where $R$ is a rotation matrix. We can
-formulate the judgment of whether $X_a$ and $X_b$ have the same
-origins by deciding about two hypotheses, $h_0$: $\forall R\ X_b\neq
-RX_a$ and $h_1$: $\exists R\textrm{ s.t. } X_b=RX_a$.  To compare the
-hypotheses, we need to compute the posterior for each: $p(h\ \vert\
-X_a, X_b)\propto p(X_a, X_b\ \vert\ h)p(h)$. Assuming the hypotheses
-are equally likely \textit{a priori}, the prior term $p(h)$ will
-cancel out when comparing $h_0$ and $h_1$, thus allowing us to focus
-on the likelihoods, which are $p(X_a,\ X_b\ \vert \ h_0)=p(X_a)p(X_b)$
-for $h_0$ and $p(X_a,\ X_b\ \vert \ h_1)=\int_R p(X_a) p(X_b\vert
-X_a,R) p(R)\ \mathrm{d}R$ for $h_1$. From these likelihoods, we
-compute the ratio $\ell$ which is given by $\ell=\left(\int_R p(X_b\
-  \vert\ X_a, R)p(R)\ \mathrm{d}R\right) /\ p(X_b)$. If $\ell<1$, then
-$h_0$ is the more likely hypothesis. If $\ell>1$, then $h_1$ is the
-more likely hypothesis.
+We begin by analyzing mental rotation at Marr's \textit{computational}
+level \cite{Marr:1983to}: what is the problem to be solved?  Formally,
+people are presented with two images, $X_a$ and $X_b$, which are the
+coordinates of the vertices of 2D shapes (e.g.,
+Fig.~\ref{fig:stimuli}). Participants must determine whether $X_a$ and
+$X_b$ depict the same shape, i.e., whether $\exists R\textrm{ s.t. }
+X_b=RX_a$, where $R$ is a rotation matrix. We can formulate the
+judgment of whether $X_a$ and $X_b$ have the same origins by deciding
+about two hypotheses, $h_0$: $\forall R\ X_b\neq RX_a$ and $h_1$:
+$\exists R\textrm{ s.t. } X_b=RX_a$.  To compare the hypotheses, we
+need to compute the posterior for each: $p(h\ \vert\ X_a, X_b)\propto
+p(X_a, X_b\ \vert\ h)p(h)$. Assuming the hypotheses are equally likely
+\textit{a priori}, the prior term $p(h)$ will cancel out when
+comparing $h_0$ and $h_1$, thus allowing us to focus on the
+likelihoods, which are $p(X_a,\ X_b\ \vert \ h_0)=p(X_a)p(X_b)$ for
+$h_0$ and $p(X_a,\ X_b\ \vert \ h_1)=\int_R p(X_a) p(X_b\vert X_a,R)
+p(R)\ \mathrm{d}R$ for $h_1$. From these likelihoods, we compute the
+ratio $\ell$ which is given by $\ell=\left(\int_R p(X_b\ \vert\ X_a,
+  R)p(R)\ \mathrm{d}R\right) /\ p(X_b)$. If $\ell<1$, then $h_0$ is
+the more likely hypothesis. If $\ell>1$, then $h_1$ is the more likely
+hypothesis.
 
@@
 Gaussian Process (GP) prior on the function $S$ and evaluating $S$ at
-a particular set of points. However, althought $S$ is a non-negative
+a particular set of points. However, while $S$ is a non-negative
 likelihood function, GP regression enforces no such
 constraint. \cite{Osborne:2012tm} give a method to place a prior over
-the log likelihood,\footnote{In practice, as in \cite{Osborne:2012tm},
-  we use the transform of $\log(S+1)$. Additionally, we follow
-  \cite{Osborne:2012tm} and use a combination of MLII and
-  marginalization to fit the $w$ kernel parameters. The output scales
-  are fixed at $h_S=\sqrt{0.1}$ and $h_{\log S}=\log(h_S + 1)$.} thus
-ensuring that $S=e^{\log S}$ will be positive, i.e.  $E[Z\ \vert \
-\log S]=\int_{\log S}\left(\int_R \exp(\log{S(X_b,X_R)})p(R)\
+the \textit{log} likelihood, thus ensuring that $S=e^{\log S}$ will be
+positive, i.e.  $E[Z\ \vert \ \log S]=\int_{\log S}\left(\int_R
+  \exp(\log{S(X_b,X_R)})p(R)\
   \mathrm{d}R\right)\mathcal{N}\left(\log{S}\ \vert \ \mu_{\log S},
@@
 
-We pick an initial direction of rotation which results in the higher
-value of $S$. At each step, we compute $\mu_Z$ and $\tilde{V}$ to
-estimate a distribution over the likelihood ratio $\ell$, i.e.
-$p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\ \vert\ \mu_Z,
-\sigma_z)$.  We choose $h_0$ when $p(\ell < 1)\geq 0.95$, and $h_1$
-when $p(\ell > 1)\geq 0.95$. Until one of these conditions are met (or
-the shape has been fully rotated), the model will continue to compute
-rotations and update its estimate of $Z$.  The model may also change
-direction based on the \textit{expected} posterior variance of $Z$
-given some new sample $a$.\footnote{This is similar to the full
+We pick the initial direction of rotation which results in the higher
+value of $S$. From then on, at each step we compute $\mu_Z$ and
+$\tilde{V}$ to estimate a distribution over the likelihood ratio
+$\ell$, i.e.  $p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\ \vert\
+\mu_Z, \sigma_z)$.  We choose $h_0$ when $p(\ell < 1)\geq 0.95$, and
+$h_1$ when $p(\ell > 1)\geq 0.95$. Until one of these conditions are
+met (or the shape has been fully rotated), the model will continue to
+compute rotations and update its estimate of $Z$.  The model may also
+change direction based on the \textit{expected} posterior variance of
+$Z$ given some new sample $a$.\footnote{This is similar to the full
   posterior variance calculation given in \cite{Osborne:2012tm},
@@
 were created by rotating $X_a$ by $R$; the same was true for
-``different'', except that $X_a$ was also reflected across the
+``different'' pairs, except that $X_a$ was also reflected across the
 $y$-axis. To gauge performance, we looked at \textit{response error
@@
 \textit{rotations}: for those ``same'' pairs which the model judged
-correct, how correlated were the model's rotations with the true
+correctly, how correlated were the model's rotations with the true
 angles of rotation?  We quantified this using the Pearson's
-correlation coefficient $\rho$ for the true rotation ($R$) versus the
-number of steps/rotations take by the model ($\vert
-\mathbf{R}\vert$). Fig.~\ref{fig:rotations} shows true rotations vs
-number of steps/rotations by the model for individual stimuli
-(\ref{fig:all-stimuli}) and average per true rotation
-(\ref{fig:correct-stimuli}). This latter analysis can be qualitatively
-compared to the results of \cite{Shepard1971}
-(Fig.~\ref{fig:mental-rotation}).
+correlation coefficient $\rho$ for the true rotation, $R$, versus the
+number of steps/rotations take by the model, $\vert
+\mathbf{R}\vert$. Fig.~\ref{fig:rotations} shows true rotations vs
+(\ref{fig:all-stimuli}) the number of steps/rotations taken by the
+model for individual stimuli, and (\ref{fig:correct-stimuli}) the
+average number of steps taken for each true rotation. This latter
+analysis can be qualitatively compared to the results of
+\cite{Shepard1971}, as shown in Fig.~\ref{fig:mental-rotation}.
 
@@
 The \naive{} model's response error rate was $\NaiveME{}$, which is
-better than chance (equivalent to guessing randomly,
+better than chance (where chance is equivalent to guessing randomly,
 i.e. $\ME{}=0.5$). The correlation between the \naive{} model's
 average rotation and the true angle of rotation was $\Naivecorr{}$
-(Fig.~\ref{fig:correct-stimuli}, left). For $R<\frac{\pi}{2}$, the
-\naive{} model corresponds extremely well to the true angle of
-rotation; this is because it needs to rotate less and is therefore
-less likely to get stuck on local maxima. For $R>\frac{\pi}{2}$, we
-see an increasing tendency to under-rotate due to getting stuck on
-local maxima, as well as a tendency to over-rotate if the wrong
-direction was initially chosen.
+(Fig.~\ref{fig:correct-stimuli}, left). As shown in
+Fig.~\ref{fig:correct-stimuli} (right), the \naive{} model corresponds
+extremely well to the true angle of rotation when $R<\frac{\pi}{2}$;
+this is because it needs to rotate less and is therefore less likely
+to get stuck on local maxima. For $R>\frac{\pi}{2}$, we see an
+increasing tendency to under-rotate due to getting stuck on local
+maxima, as well as a tendency to over-rotate if the wrong direction
+was initially chosen.
 
@@
 The BQ model was much more accurate in choosing the correct hypothesis
-than the other model ($\BQME{}$). The number of rotations computed by
-the BQ model were strongly correlated with the true rotations
+than the \naive{} model ($\BQME{}$). The number of rotations computed
+by the BQ model were strongly correlated with the true rotations
 ($\BQcorr$), a result which is qualitatively similar to that exhibited
@@
 incorrect direction (e.g., Fig.~\ref{fig:bq}) and thus did not
-over-rotate as frequently.  It also under-rotated less frequently:
-because the model continues to rotate until it is confident in its
-estimate of $Z$, it does not get stuck as easily on local optima.
+over-rotate as frequently as the \naive{} model.  The BQ model also
+under-rotated less frequently because it rotates until it is confident
+in its estimate of $Z$ and thus does not get stuck on local optima.
 

### 2013-10-17 man/nips-2013-bayesian-optimization.tex
Split caption into subcaptions

@@
   \centering
-  \begin{subfigure}[b]{0.49\textwidth}
+  \begin{subfigure}[b]{0.46\textwidth}
     \centering
     \includegraphics[width=\textwidth]{../figures/model_rotations_A.pdf}
-    % \vspace{0pt}
-    \caption{All stimuli}
+    \caption{\textbf{All stimuli}. Each subplot shows the
+      correspondence between the true angle of rotation ($R$) for
+      ``same'' pairs and the amount of rotation performed by the
+      model.}
     \label{fig:all-stimuli}
   \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
+  \hspace{0.05\textwidth}
+  \begin{subfigure}[b]{0.46\textwidth}
     \centering
     \includegraphics[width=\textwidth]{../figures/model_rotations_B.pdf}
-    \caption{Correct stimuli}
+    \caption{\textbf{Correct stimuli}. Each subplot shows the models'
+      mean rotations over stimuli pairs that were judged
+      correctly. Black dots correspond to ``same'' pairs, and blue
+      lines to ``different'' pairs.}
     \label{fig:correct-stimuli}
   \end{subfigure}
-  \caption{\textbf{Model rotations.} (a) Each subplot shows the
-    correspondence between the true angle of rotation ($R$) for
-    ``same'' pairs and the amount of rotation performed by the
-    model. (b) Each subplot shows the models' mean rotations over
-    stimuli pairs that were judged correctly. Black dots correspond to
-    ``same'' pairs, and blue lines correspond to ``different''
-    pairs. Error bars/shaded regions indicate one standard deviation,
-    and the dotted lines indicate the least-squares fit to the
-    ``same'' pairs.}
+  \caption{\textbf{Model rotations.} Error bars/shaded regions
+    indicate one standard deviation, and the dotted lines indicate the
+    least-squares fit to the ``same'' pairs.}
   \label{fig:rotations}

### 2013-11-27 code/kernel.py
Catch floating point error in cholesky decomposition

@@ -12,7 +12,10 @@ import periodic_mll as mll
 
 
 def cholesky(mat):
-    m = np.mean(np.abs(mat))
+    try:
+        m = np.mean(np.abs(mat))
+    except FloatingPointError:
+        raise np.linalg.LinAlgError("floating point error")
     try:
         L = np.linalg.cholesky(mat)
     except np.linalg.LinAlgError:

### 2013-11-27 man/nips-2013.tex
Uncomment \nipsfinalcopy

@@
 \cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm} provides a way to
-solve this problem, outperforming several simpler heuristics for
-determining both the direction and extent of rotation.
+solve this problem, outperforming simpler heuristics for determining
+the direction and extent of rotation.
 
@@
 For many years, this idea was contested, with some researchers arguing
-the underlying cognitive processes were not visual in nature
-(e.g., \cite{Pylyshyn1981}). In particular, Anderson
-\cite{Anderson1978} proved that the mental imagery debate could not be
-resolved on the basis of response time data alone. In recent years,
-however, significant efforts have investigated the neural
-underpinnings of the mental rotation phenomena, arguing that the
-results of brain imaging studies support the existence of visual
-imagery \cite{Kosslyn:2009tj,Kosslyn1988}.
+the underlying cognitive processes were not visual in nature (e.g.,
+\cite{Pylyshyn1981}). In particular, Anderson \cite{Anderson1978}
+proved that the mental imagery debate could not be resolved on the
+basis of response time data alone. In recent years, however,
+significant efforts have investigated the neural underpinnings of the
+mental rotation phenomena, arguing that the results of brain imaging
+studies support the existence of visual imagery
+\cite{Kosslyn:2009tj,Kosslyn1988}.  We assume that mental imagery is
+indeed a visual process, and turn to the question of how, and when, it
+is used. Previous work has examined how people might use imagery
+\cite{Hegarty2004, Schwartz1999} to solve reasoning problems, and when
+they might use imagery as opposed to a symbolic rule
+\cite{Schwartz:1996uy}. Eye-tracking studies of people performing
+mental rotations suggest strategies that people might be using, such
+as aligning smaller features of the images \cite{Just1976}. People
+seem to use imagery methodically, which raises the question: what is
+the method?
 
-We assume that mental imagery is indeed a visual process, and turn to
-the question of how, and when, it is used. Previous work has examined
-how people might use imagery \cite{Hegarty2004, Schwartz1999} to solve
-reasoning problems, and when they might use imagery as opposed to a
-symbolic rule \cite{Schwartz:1996uy}. Eye-tracking studies of people
-performing mental rotations suggest strategies that people might be
-using, such as aligning smaller features of the images
-\cite{Just1976}. People seem to use imagery methodically, which raises
-the question: what is the method?
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/stimuli_shapes.pdf}
+    \vspace{0pt}
+    \caption{Example stimuli}
+    \label{fig:stimuli}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{../figures/likelihood_function.pdf}
+    \caption{Likelihood function}
+    \label{fig:likelihood}
+  \end{subfigure}
+  \caption{\textbf{Rotated shapes and their similarity.}  \textbf{(a)}
+    An example stimulus in which the shapes differ only by a
+    rotation. All stimuli consist of three to six vertices centered
+    around the origin, and edges which create a closed loop from the
+    vertices. \textbf{(b)} The approximate likelihood (similarity) of
+    $X_b$ given $X_R$, where $X_R$ is a rotation of $X_a$ by the angle
+    $R$. The true angle of rotation between $X_a$ and $X_b$ is at
+    $\frac{2\pi}{3}$, corresponding to the global maximum in the
+    likelihood function.}
+  \label{fig:shapes}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/stimuli_shapes.pdf}
-    \vspace{0pt}
-    \caption{Example stimuli}
-    \label{fig:stimuli}
+    \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
+    \caption{\Naive{} model}
+    \label{fig:li}
   \end{subfigure}
@@
     \centering
-    \includegraphics[width=\textwidth]{../figures/likelihood_function.pdf}
-    \caption{Likelihood function}
-    \label{fig:likelihood}
+    \includegraphics[width=\textwidth]{../figures/vm_regression.pdf}
+    \caption{Parametric model}
+    \label{fig:vm}
   \end{subfigure}
-  \caption{\textbf{Rotated shapes and their similarity.}  \textbf{(a)}
-    An example stimulus in which the shapes differ only by a
-    rotation. All stimuli consist of three to six vertices centered
-    around the origin, and edges which create a closed loop from the
-    vertices. \textbf{(b)} The approximate likelihood (similarity) of
-    $X_b$ given $X_R$, where $X_R$ is a rotation of $X_a$ by the angle
-    $R$. The true angle of rotation between $X_a$ and $X_b$ is at
-    $\frac{2\pi}{3}$, corresponding to the global maximum in the
-    likelihood function.}
-  \label{fig:shapes}
+  \caption{\textbf{Simple methods of estimating $S$.} In both cases,
+    hill-climbing search is used until a maxima is found (in this
+    case, at approximately $\frac{\pi}{6}$). The sampled points
+    $\mathbf{R}$ (red circles) are then used to estimate $S$ (black
+    lines are the true $S$, red lines are the estimate). \textbf{(a)}
+    Linear interpolation. The overall estimate of $Z$ here will be too
+    low, but certain angles will have a disproportionate contribution
+    (e.g., between $\frac{\pi}{4}$ and $\frac{3\pi}{4}$. \textbf{(b)}
+    Best fit of scaled Von Mises PDF parameters. As in (a), $Z$ will
+    be underestimated. However, the fit around the maximum near
+    $\frac{\pi}{6}$ is much more accurate.}
+  \label{fig:simple-models}
 \end{figure}
@@
 
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.4\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
-    \caption{\Naive{} model}
-    \label{fig:li}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.4\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{../figures/vm_regression.pdf}
-    \caption{Parametric model}
-    \label{fig:vm}
-  \end{subfigure}
-  \caption{\textbf{Simple methods of estimating $S$.} In both cases,
-    hill-climbing search is used until a maxima is found (in this
-    case, at approximately $\frac{\pi}{6}$). The sampled points
-    $\mathbf{R}$ (red circles) are then used to estimate $S$ (black
-    lines are the true $S$, red lines are the estimate). \textbf{(a)}
-    Linear interpolation. The overall estimate of $Z$ here will be too
-    low, but certain angles will have a disproportionate contribution
-    (e.g., between $\frac{\pi}{4}$ and $\frac{3\pi}{4}$. \textbf{(b)}
-    Best fit of scaled Von Mises PDF parameters. As in (a), $Z$ will
-    be underestimated. However, the fit around the maximum near
-    $\frac{\pi}{6}$ is much more accurate.}
-  \label{fig:simple-models}
-\end{figure}
-
 As a baseline, we defined a \naive{} model which performs a
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.85\textwidth]{../figures/model_rotations.pdf}
+  \caption{\textbf{Model rotations.} Top: each subplot shows the
+    correspondence between the true angle of rotation ($R$) for
+    ``same'' pairs and the amount of rotation performed by the
+    model. Bottom: each subplot shows the models' mean rotations over
+    stimuli pairs that were judged correctly. Black dots correspond to
+    ``same'' pairs, and blue lines correspond to ``different'' pairs
+    (for which the mean rotation is constant across true rotations, as
+    the true rotation is undefined). Error bars/shaded regions
+    indicate one standard deviation, and the dotted lines indicate the
+    least-squares fit to the ``same'' pairs.}
+  \label{fig:rotations}
+\end{figure}
+
 Bayesian Quadrature has its difficulties, however. While in our case
@@
 ensuring that $S=e^{\log S}$ will be positive\footnote{We are not
-  guaranteed positivity, however, the approximation to the integral
+  guaranteed positivity, however, as the approximation to the integral
   over $\log S$ (Equation \ref{eq:bq-Z-mean}) requires computing
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{../figures/model_rotations.pdf}
-  \caption{\textbf{Model rotations.} Top: each subplot shows the
-    correspondence between the true angle of rotation ($R$) for
-    ``same'' pairs and the amount of rotation performed by the
-    model. Bottom: each subplot shows the models' mean rotations over
-    stimuli pairs that were judged correctly. Black dots correspond to
-    ``same'' pairs, and blue lines correspond to ``different'' pairs
-    (for which the mean rotation is constant across true rotations, as
-    the true rotation is undefined). Error bars/shaded regions
-    indicate one standard deviation, and the dotted lines indicate the
-    least-squares fit to the ``same'' pairs.}
-  \label{fig:rotations}
+  \includegraphics[width=0.85\textwidth]{../figures/Z_accuracy.pdf}
+  \caption{\textbf{Accuracy in estimating $Z$.} Each subplot shows the
+    true (``gold standard'') value of $Z$ vs. the value estimated by
+    the model. Black dotted lines indicate a perfect 1:1
+    correspondence between the true and estimated values. The \naive{}
+    model (left panel) tends to overestimate. The parametric model
+    (center panel) is very accurate for about half the ``same'' pairs,
+    and severely underestimates the rest. The nonparametric model
+    (right panel) maintains a decent correspondence with the true
+    $Z$.}
+  \label{fig:accuracy}
 \end{figure}
@@
 We considered three metrics of performance in particular:
-\begin{itemize}
-\item \textit{Response error rates}: How accurate was the model at
-  choosing the correct hypothesis? This was defined as the mean error
-  ($\ME{}$), or fraction of times the model chose incorrectly.
-\item \textit{Rotations}: For those ``same'' pairs which the model
-  judged correct, how correlated were the model's rotations with the
-  true angles of rotation?  We quantified this using the Pearson's
-  correlation coefficient $\rho$ for the true rotation ($R$) versus
-  the extent of the model's rotations ($\vert \mathbf{R}\vert$). The
-  top row in Figure \ref{fig:rotations} shows individual points
-  corresponding to true rotations vs. rotations by the models for all
-  stimuli, including those judged incorrectly. The bottom row shows
-  mean model rotations, across only those stimuli which were judged
-  correctly. We structured the analysis in this way to better compare
-  to the results of \cite{Shepard1971}, which also excluded
-  incorrectly judged stimuli.
-\item \textit{Estimates of $Z$}: How accurate was the model's estimate
-  of $Z$? We defined this quantity as the mean squared error
-  ($\MSE{}$) between the model's estimate of $Z$ and the ``gold
-  standard'' value, where the error has scale such that $\MSE{}=0$
-  indicates no error and $\MSE{}=1$ indicates maximum error. Figure
-  \ref{fig:accuracy} shows plots of the true (``gold standard'') value
-  of $Z$ vs. the model's estimate, for each model.
-\end{itemize}
+
+\textit{Response error rates}: How accurate was the model at choosing
+the correct hypothesis? This was defined as the mean error ($\ME{}$),
+or fraction of times the model chose incorrectly.
+
+\textit{Rotations}: For those ``same'' pairs which the model judged
+correct, how correlated were the model's rotations with the true
+angles of rotation?  We quantified this using the Pearson's
+correlation coefficient $\rho$ for the true rotation ($R$) versus the
+extent of the model's rotations ($\vert \mathbf{R}\vert$). The top row
+in Figure \ref{fig:rotations} shows individual points corresponding to
+true rotations vs. rotations by the models for all stimuli, including
+those judged incorrectly. The bottom row shows mean model rotations,
+across only those stimuli which were judged correctly. We structured
+the analysis in this way to better compare to the results of
+\cite{Shepard1971}, which also excluded incorrectly judged stimuli.
+
+\textit{Estimates of $Z$}: How accurate was the model's estimate of
+$Z$? We defined this quantity as the mean squared error ($\MSE{}$)
+between the model's estimate of $Z$ and the ``gold standard'' value,
+where the error has scale such that $\MSE{}=0$ indicates no error and
+$\MSE{}=1$ indicates maximum error. Figure \ref{fig:accuracy} shows
+plots of the true (``gold standard'') value of $Z$ vs. the model's
+estimate, for each model.
 
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.9\textwidth]{../figures/Z_accuracy.pdf}
-  \caption{\textbf{Accuracy in estimating $Z$.} Each subplot shows the
-    true (``gold standard'') value of $Z$ vs. the value estimated by
-    the model. Black dotted lines indicate a perfect 1:1
-    correspondence between the true and estimated values. The \naive{}
-    model (left panel) tends to overestimate. The parametric model
-    (center panel) is very accurate for about half the ``same'' pairs,
-    and severely underestimates the rest. The nonparametric model
-    (right panel) maintains a decent correspondence with the true
-    $Z$.}
-  \label{fig:accuracy}
-\end{figure}
-
 \paragraph{Nonparametric (Bayesian Quadrature)}
@@
 relationship with the true angle of rotation due to lack of robustness
-against local maxima or incorrect rotation direction. The
-nonparametric model qualitatively approximates human behavioral
-judgments, both in linearity \cite{Shepard1971} and variability
-\cite{Just1976}. Future work will collect empirical data from
-participants to perform a more thorough, quantitative analysis.  We
-conclude that from this initial survey, models which take an
-``active'', directed approach seem well-suited to explaining mental
-simulation.
+against local maxima or incorrect rotation direction.  More broadly,
+the nonparametric model provides answers to puzzling questions
+surrounding the incremental nature of mental rotation: which way
+should the object be rotated, and for how long? This model formalizes
+these answers in a way that is qualitatively consistent with human
+behavior, both in response time linearity \cite{Shepard1971} and
+variability \cite{Just1976}. These results further indicate that this
+may be another instance in which people do appropriately use available
+computational resources to solve the task at hand.  Rational, adaptive
+approaches such as this may therefore be successful when applied to
+more complex cognitive processes above and beyond mental rotation.
 

### 2013-12-02 man/nips-2013-bayesian-optimization.tex
Tone down the introduction and conclusion; other minor edits

@@
   mental rotation task, and found that a model based on an optimal
-  experiment design for Bayesian quadrature is objectively more
-  accurate and qualitatively more consistent with classic behavioral
-  data than a simpler model. We suggest that rational models which
-  adaptively exploit available resources are promising in their
-  ability to characterize metacognitive processes like mental
-  simulation.
+  experiment design for Bayesian quadrature is qualitatively more
+  consistent with classic behavioral data than a simpler model. We
+  suggest that rational models which adaptively exploit available
+  resources are promising in their ability to characterize
+  metacognitive processes like mental simulation.
 \end{abstract}
@@
 rotating until it is congruent with the other
-\cite{Shepard1971}. There is strong evidence for such mental
-simulation: we can imagine three-dimensional objects in our minds and
-manipulate them, to a certain extent, as if they were real
+\cite{Shepard1971}. There is strong evidence for such ``mental
+simulation'': we can imagine three-dimensional objects in our minds
+and manipulate them, to a certain extent, as if they were real
 \cite{Kosslyn:2009tj}.  However, the use of mental simulation is
@@
 
-Recent work in cognitive science has shown how problems of allocating
+Recent work in cognitive science has shown how the allocation of
 cognitive resources to solving computational problems can be analyzed
 using the methods of statistical decision theory
-\cite{Lieder:2012wg,Vul:2009wy}. We apply this ``rational
-metacognition'' approach to the problem of mental rotation. We
-investigate several computational solutions to this task and
-qualitatively compare them to human mental rotation performance. In
-particular, we argue that performing mental rotation can be framed as
-integration over a probability distribution, with the direction of
-rotation becoming an optimal experiment design problem (or in machine
-learning parlance, an active learning problem). We show that recent
-work on methods for Bayesian quadrature
-\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm} provides a way to
-solve this problem, outperforming simpler heuristics for determining
+\cite{Lieder:2012wg,Vul:2009wy}. We suggest that this ``rational
+metacognition'' approach may also be applied to the problem of mental
+rotation. Specifically, we hypothesize that mental rotation can be
+framed as integration over a probability distribution, with the
+direction of rotation becoming an optimal experiment design problem
+(or in machine learning parlance, an active learning problem). In an
+initial investigation into this hypothesis, we find that recent
+methods for Bayesian quadrature
+\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}, in contrast to a
+simpler heuristic model, provide a possible solution to determining
 the direction and extent of rotation.
@@
 
-We defined a \naive{} model which performs a hill-climbing search over
-the similarity function until it reaches a (possibly local)
-maximum. Once a maximum as been found, the model computes an estimate
-of $Z$ by linearly interpolating between sampled rotations (e.g.,
-Fig.~\ref{fig:li}).
+As a lower bound on performance, we defined a \naive{} model which
+performs a hill-climbing search over the similarity function until it
+reaches a (possibly local) maximum. Once a maximum as been found, the
+model computes an estimate of $Z$ by linearly interpolating between
+sampled rotations (e.g., Fig.~\ref{fig:li}).
 
@@
 \mu_Z, \sigma_z)$.  We choose $h_0$ when $p(\ell < 1)\geq 0.95$, and
-$h_1$ when $p(\ell > 1)\geq 0.95$. Until one of these conditions are
-met (or the shape has been fully rotated), the model will continue to
-compute rotations and update its estimate of $Z$.  The model may also
-change direction based on the \textit{expected} posterior variance of
-$Z$ given some new sample $a$.\footnote{This is similar to the full
-  posterior variance calculation given in \cite{Osborne:2012tm},
-  however we compute the variance given only the current mean estimate
-  of $a$.} The model will change directions if doing so would lower
-this expected variance. Thus, it is able to actively change its
-strategy, unlike the hill-climbing procedure.
+$h_1$ when $p(\ell > 1)\geq 0.95$.\footnote{We chose a threshold of
+  0.95 because the standard confidence interval is 95\%. In the
+  future, however, this threshold could be fit to human data.} Until
+one of these conditions are met (or the shape has been fully rotated),
+the model will continue to compute rotations and update its estimate
+of $Z$.  Additionally, the model will change direction if doing so
+would lower the \textit{expected} posterior variance of $Z$ given some
+new sample $a$.\footnote{This is similar to the full posterior
+  variance calculation given in \cite{Osborne:2012tm}, however we
+  compute the variance given only the current mean estimate of $a$.}
+Thus, it is able to actively change its strategy, unlike the
+hill-climbing procedure.
 
@@
 
-In this paper, we asked: how do people use mental simulation?  We
-investigated the specific case of mental rotation, using rational
-analysis to characterize optimal strategies. We found that an
-adaptive, BQ model performs best at the classic rotation task
-\cite{Shepard1971}, as it is able to actively monitor the confidence
-of its estimate and intelligently choose the direction of rotation. A
-simpler model based on heuristics performed much worse: it did not
-maintain the linear relationship with the true angle of rotation due
-to lack of robustness against local maxima or incorrect rotation
-direction.  More broadly, the BQ model provides answers to puzzling
+In this paper, we asked: how do people perform mental simulation?  We
+performed an initial investigation into the specific case of mental
+rotation, using rational analysis to characterize optimal strategies.
+We demonstrated that the classic mental rotation task
+\cite{Shepard1971} presents a non-trivial computational problem and
+cannot be solved with a simple, heuristic-based model. In contrast, an
+adaptive, Bayesian Quadrature model provides answers to puzzling
 questions surrounding the incremental nature of mental rotation: which
@@
 with human behavior, both in response time linearity
-\cite{Shepard1971} and variability \cite{Just1976}. These results
-further indicate that this may be another instance in which people do
-appropriately use available computational resources to solve the task
-at hand.  Rational, adaptive approaches such as this may therefore be
-successful when applied to more complex cognitive processes above and
-beyond mental rotation.
+\cite{Shepard1971} and variability \cite{Just1976}. Although this
+research is still in its first stages, these initial results support
+the idea that mental rotation may be another instance in which people
+do appropriately use available computational resources to solve the
+task at hand.
 

### 2014-01-16 man/cogsci-2014/cogsci-2014.tex
Move old papers into their own folders; add cogsci template

% 
% Annual Cognitive Science Conference
% Sample LaTeX Paper -- Proceedings Format
%

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified: Niels Taatgen (taatgen@cmu.edu) 10/24/2006

%% Change ``a4paper'' in the following line to ``letterpaper'' if you are
%% producing a letter-format document.

\documentclass[10pt,letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}

\title{How to Make a Camera-Ready Proceedings Contribution}

\author{{\large \bf Morton Ann Gernsbacher (MAG@Macc.Wisc.Edu)} \\
  Department of Psychology, 1202 W. Johnson Street \\
  Madison, WI 53706 USA
  \AND {\large \bf Sharon J. Derry (SDJ@Macc.Wisc.Edu)} \\
  Department of Educational Psychology, 1025 W. Johnson Street \\
  Madison, WI 53706 USA}

\begin{document}

\maketitle

\begin{abstract}
The abstract should be one paragraph, indented 1/8~inch on both sides,
in 9~point font with single spacing. The heading {\bf Abstract} should
be 10~point, bold, centered, with one line space below it. This
one-paragraph abstract section is required only for standard spoken
papers and standard posters (i.e., those presentations that will be
represented by six page papers in the Proceedings).

\textbf{Keywords:} 
Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term.
\end{abstract}

\section{General Formatting Instructions}

For standard spoken papers and standard posters, the entire contribution (including figures, references, everything) can be no longer than six pages. For abstract posters, the entire contribution can be no longer than one page. For symposia, the entire contribution can be no longer than two pages.

The text of the paper should be formatted in two columns with an
overall width of 7 inches (17.8 cm) and length of 9.25 inches (23.5
cm), with 0.25 inches between the columns. Leave two line spaces
between the last author listed and the text of the paper. The left
margin should be 0.75 inches and the top margin should be 1 inch.
\textbf{The right and bottom margins will depend on whether you use
  U.S. letter or A4 paper, so you must be sure to measure the width of
  the printed text.} Use 10~point Times Roman with 12~point vertical
spacing, unless otherwise specified.

The title should be in 14~point, bold, and centered. The title should
be formatted with initial caps (the first letter of content words
capitalized and the rest lower case). Each author's name should appear
on a separate line, 11~point bold, and centered, with the author's
email address in parentheses. Under each author's name list the
author's affiliation and postal address in ordinary 10~point type.

Indent the first line of each paragraph by 1/8~inch (except for the
first paragraph of a new section). Do not add extra vertical space
between paragraphs.

\section{First-Level Headings}

First level headings should be in 12~point, initial caps, bold and
centered. Leave one line space above the heading and 1/4~ line space
below the heading.

\subsection{Second-Level Headings}

Second level headings should be 11~point, initial caps, bold, and
flush left. Leave one line space above the heading and 1/4~ line
space below the heading.

\subsubsection{Third-Level Headings}

Third-level headings should be 10~point, initial caps, bold, and flush
left. Leave one line space above the heading, but no space after the
heading.

\section{Formalities, Footnotes, and Floats}

Use standard APA citation format. Citations within the text should
include the author's last name and year. If the authors' names are
included in the sentence, place only the year in parentheses, as in
\citeA{NewellSimon1972a}, but otherwise place the entire reference in
parentheses with the authors and year separated by a comma
\cite{NewellSimon1972a}. List multiple references alphabetically and
separate them by semicolons
\cite{ChalnickBillman1988a,NewellSimon1972a}. Use the
et~al. construction only after listing all the authors to a
publication in an earlier reference and for citations with four or
more authors.

\subsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first
  footnote.} in the text. Place the footnotes in 9~point type at the
bottom of the page on which they appear. Precede the footnote with a
horizontal rule.\footnote{Sample of the second footnote.}

\subsection{Tables}

Number tables consecutively; place the table number and title (in
10~point) above the table with one line space above the caption and
one line space below it, as in Table~\ref{sample-table}. You may float
tables to the top or bottom of a column, set wide tables across both
columns.

\begin{table}[!ht]
\begin{center} 
\caption{Sample table title.} 
\label{sample-table} 
\vskip 0.12in
\begin{tabular}{ll} 
\hline
Error type    &  Example \\
\hline
Take smaller        &   63 - 44 = 21 \\
Always borrow~~~~   &   96 - 42 = 34 \\
0 - N = N           &   70 - 47 = 37 \\
0 - N = 0           &   70 - 47 = 30 \\
\hline
\end{tabular} 
\end{center} 
\end{table}

\subsection{Figures}

All artwork must be very dark for purposes of reproduction and should
not be hand drawn. Number figures sequentially, placing the figure
number and caption, in 10~point, after the figure with one line space
above the caption and one line space below it, as in
Figure~\ref{sample-figure}. If necessary, leave extra white space at
the bottom of the page to avoid splitting the figure and figure
caption. You may float figures to the top or bottom of a column, or
set wide figures across both columns.

\begin{figure}[ht]
\begin{center}
\fbox{CoGNiTiVe ScIeNcE}
\end{center}
\caption{This is a figure.} 
\label{sample-figure}
\end{figure}

\section{Acknowledgments}

Place acknowledgments (including funding information) in a section at
the end of the paper.

\section{References Instructions}

Follow the APA Publication Manual for citation format, both within the
text and in the reference list, with the following exceptions: (a) do
not cite the page numbers of any book, including chapters in edited
volumes; (b) use the same format for unpublished references as for
published ones. Alphabetize references by the surnames of the authors,
with single author entries preceding multiple author entries. Order
references by the same authors by the year of publication, with the
earliest first.

Use a first level section heading for the reference list. Use a
hanging indent style, with the first line of the reference flush
against the left margin and subsequent lines indented by 1/8~inch.
Below are example references for a conference paper, book chapter,
journal article, technical report, dissertation, book, and edited
volume, respectively.

\nocite{ChalnickBillman1988a}
\nocite{Feigenbaum1963a}
\nocite{Hill1983a}
\nocite{OhlssonLangley1985a}
\nocite{Lewis1978a}
\nocite{NewellSimon1972a}
\nocite{ShragerLangley1990a}

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{CogSci_Template}

\end{document}

### 2014-01-19 man/cogsci-2014/cogsci-2014.tex
Work on modeling section of paper

@@
 \usepackage{apacite}
+\usepackage{graphicx}
+\usepackage{color}
+\usepackage{amsmath, amsthm, amssymb}
 
-\title{How to Make a Camera-Ready Proceedings Contribution}
+\newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}
+\newcommand{\Xa}[0]{\mathbf{X}_a}
+\newcommand{\Xb}[0]{\mathbf{X}_b}
+\newcommand{\Xt}[0]{\mathbf{X}_t}
+\newcommand{\R}[0]{\mathbf{R}_\theta}
+\newcommand{\F}[0]{\mathbf{F}}
+\newcommand{\M}[0]{\mathbf{M}}
+\newcommand{\I}[0]{\mathbb{I}}
+\newcommand{\hi}[0]{h=0}
+\newcommand{\hf}[0]{h=1}
 
-\author{{\large \bf Morton Ann Gernsbacher (MAG@Macc.Wisc.Edu)} \\
-  Department of Psychology, 1202 W. Johnson Street \\
-  Madison, WI 53706 USA
-  \AND {\large \bf Sharon J. Derry (SDJ@Macc.Wisc.Edu)} \\
-  Department of Educational Psychology, 1025 W. Johnson Street \\
-  Madison, WI 53706 USA}
+\newenvironment{pitemize}{
+\begin{itemize}
+  \setlength{\itemsep}{1pt}
+  \setlength{\parskip}{0pt}
+  \setlength{\parsep}{0pt}
+}{\end{itemize}}
+
+\title{An active sampling model of mental rotation}
+
+\author{{\large \bf Jessica B. Hamrick (jhamrick@berkeley.edu)} \\
+  {\large \bf Thomas L. Griffiths (tom\_griffiths@berkeley.edu)} \\
+  Department of Psychology, University of California, Berkeley, CA
+  94720 USA}
 
@@
 \begin{abstract}
-The abstract should be one paragraph, indented 1/8~inch on both sides,
-in 9~point font with single spacing. The heading {\bf Abstract} should
-be 10~point, bold, centered, with one line space below it. This
-one-paragraph abstract section is required only for standard spoken
-papers and standard posters (i.e., those presentations that will be
-represented by six page papers in the Proceedings).
+\TODO{}
 
 \textbf{Keywords:} 
-Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term.
+\TODO{}
 \end{abstract}
 
-\section{General Formatting Instructions}
+\section{Introduction}
 
-For standard spoken papers and standard posters, the entire contribution (including figures, references, everything) can be no longer than six pages.
-For abstract posters, the entire contribution can be no longer than one page.
-For symposia, the entire contribution can be no longer than two pages.
+One of our most astonishing cognitive feats is the ability to
+envision, manipulate, and plan with objects -- all without actually
+perceiving them. This capacity for ``mental simulation'' has been
+widely studied and has sparked intense debate about the underlying
+representation of mental images \TODO{cite}. Despite this vast
+literature, there have been surprisingly few computational models of
+mental imagery.
 
-The text of the paper should be formatted in two columns with an
-overall width of 7 inches (17.8 cm) and length of 9.25 inches (23.5
-cm), with 0.25 inches between the columns. Leave two line spaces
-between the last author listed and the text of the paper. The left
-margin should be 0.75 inches and the top margin should be 1 inch.
-\textbf{The right and bottom margins will depend on whether you use
-  U.S. letter or A4 paper, so you must be sure to measure the width of
-  the printed text.} Use 10~point Times Roman with 12~point vertical
-spacing, unless otherwise specified.
+\section{Background}
 
-The title should be in 14~point, bold, and centered. The title should
-be formatted with initial caps (the first letter of content words
-capitalized and the rest lower case). Each author's name should appear
-on a separate line, 11~point bold, and centered, with the author's
-email address in parentheses. Under each author's name list the
-author's affiliation and postal address in ordinary 10~point type.
+classic result by shepard and metzler
 
-Indent the first line of each paragraph by 1/8~inch (except for the
-first paragraph of a new section). Do not add extra vertical space
-between paragraphs.
+assume people rotate until congruent
 
-\section{First-Level Headings}
+but how do people actually construct the plan for the rotation?
 
-First level headings should be in 12~point, initial caps, bold and
-centered. Leave one line space above the heading and 1/4~ line space
-below the heading.
+evidence that people *don't* actually rotate until congruent
 
-\subsection{Second-Level Headings}
+so what are they actually doing?
 
-Second level headings should be 11~point, initial caps, bold, and
-flush left. Leave one line space above the heading and 1/4~ line
-space below the heading.
+perhaps they are performing mental manipulations of an image, but not
+a single trajectory of rotation
 
-\subsubsection{Third-Level Headings}
+rational analysis \cite{Marr:1983to,anderson90,Shepard:1987tt}
 
-Third-level headings should be 10~point, initial caps, bold, and flush
-left. Leave one line space above the heading, but no space after the
-heading.
+1) the problem: determine something about the spatial transformations
+that an object has undergone
+2) the goal: make this determiniation while using the least amount of
+resources
+3) the solution: perform those rotations and reflections which will
+give the most information (?) about the answer until an answer is
+found
 
-\section{Formalities, Footnotes, and Floats}
+motivate the active searching approach
+\cite{Gureckis:2012gu,Markant:2012uu} \cite{Markant:2012uu,Nelson2007}
 
-Use standard APA citation format. Citations within the text should
-include the author's last name and year. If the authors' names are
-included in the sentence, place only the year in parentheses, as in
-\citeA{NewellSimon1972a}, but otherwise place the entire reference in
-parentheses with the authors and year separated by a comma
-\cite{NewellSimon1972a}. List multiple references alphabetically and
-separate them by semicolons
-\cite{ChalnickBillman1988a,NewellSimon1972a}. Use the
-et~al. construction only after listing all the authors to a
-publication in an earlier reference and for citations with four or
-more authors.
+we focus less on the representation, and more on the process, assuming
+a spatial representation (because you need both to make a claim,
+\cite{Anderson1978})
 
-\subsection{Footnotes}
+problem solving \cite{Hegarty2004, Schwartz1999}
 
-Indicate footnotes with a number\footnote{Sample of the first
-  footnote.} in the text. Place the footnotes in 9~point type at the
-bottom of the page on which they appear. Precede the footnote with a
-horizontal rule.\footnote{Sample of the second footnote.}
+symbolic vs simulation \cite{Schwartz:1996uy}
 
-\subsection{Tables}
+eye tracking \cite{Just1976}
 
-Number tables consecutively; place the table number and title (in
-10~point) above the table with one line space above the caption and
-one line space below it, as in Table~\ref{sample-table}. You may float
-tables to the top or bottom of a column, set wide tables across both
-columns.
+\section{Models of mental rotation}
 
-\begin{table}[!ht]
-\begin{center} 
-\caption{Sample table title.} 
-\label{sample-table} 
-\vskip 0.12in
-\begin{tabular}{ll} 
-\hline
-Error type    &  Example \\
-\hline
-Take smaller        &   63 - 44 = 21 \\
-Always borrow~~~~   &   96 - 42 = 34 \\
-0 - N = N           &   70 - 47 = 37 \\
-0 - N = 0           &   70 - 47 = 30 \\
-\hline
-\end{tabular} 
-\end{center} 
-\end{table}
+In the mental rotation task, people are presented with pairs of
+two-dimensional shapes similar to those used by \cite{Cooper:1975wp}
+(e.g., Figure \ref{fig:stimuli}) and must determine whether the two
+images depict the same shape, or mirror-image shapes.
 
-\subsection{Figures}
+\subsection{Computational-level analysis}
 
-All artwork must be very dark for purposes of reproduction and should
-not be hand drawn. Number figures sequentially, placing the figure
-number and caption, in 10~point, after the figure with one line space
-above the caption and one line space below it, as in
-Figure~\ref{sample-figure}. If necessary, leave extra white space at
-the bottom of the page to avoid splitting the figure and figure
-caption. You may float figures to the top or bottom of a column, or
-set wide figures across both columns.
+Formally, we denote the shapes as $X_a$ and $X_b$ and assume $X_b$ is
+generated by a transformation of $X_a$, i.e. $X_b=f(X_a, \theta, h)$,
+where $\theta$ is a rotation, $h=0$ is the hypothesis that the images
+depict the same object, and $h=1$ is the hypothesis that the images
+depict mirror-image objects. The posterior probability of each
+hypothesis given the observed shapes is then:
+\begin{equation}
+  p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta, h)p(X_a)p(h)p(\theta)\ \mathrm{d}\theta.
+  \label{eq:posterior}
+\end{equation}
+Because we ultimately want to determine which hypothesis is more
+likely, the quantity of interest is a ratio:
+\begin{equation*}
+  \ell := \frac{\int p(X_b\ \vert\ X_a, \theta, \hi)p(X_a)p(\hi)p(\theta)\ \mathrm{d}\theta}{\int p(X_b\ \vert\ X_a, \theta, \hf)p(X_a)p(\hf)p(\theta)\ \mathrm{d}\theta},
+\end{equation*}
+which (assuming the two hypotheses are equally likely \textit{a
+  priori}, and that all rotations are equally likely) simplifies to:
+\begin{equation}
+  \ell = \frac{\int p(X_b\ \vert\ X_a, \theta, \hi)\ \mathrm{d}\theta}{\int p(X_b\ \vert\ X_a, \theta, \hf)\ \mathrm{d}\theta}.
+  \label{eq:lh-ratio}
+\end{equation}
+If $\ell > 1$, then we accept the hypothesis that the images depict
+the same object ($\hi$); if $\ell < 1$, then we accept the hypothesis
+that the images depict flipped objects ($\hf$).
 
-\begin{figure}[ht]
-\begin{center}
-\fbox{CoGNiTiVe ScIeNcE}
-\end{center}
-\caption{This is a figure.} 
-\label{sample-figure}
-\end{figure}
+\subsection{Algorithmic assumptions}
+
+If we represent a shape of $N$ vertices with a $N\times 2$ coordinate
+matrix $\mathbf{X}=[\mathbf{x}_1, \ldots{}, \mathbf{x}_N]$, then the
+transformation $f$ is $f(\mathbf{X}, h,
+\theta):=\mathbf{X}\F_h^T\R^T$, where $\R$ is a rotation matrix,
+$\F_0$ is the identity matrix $\I$, and $\F_1$ is a reflection matrix
+across the $y$-axis. Assuming no computational constraints, the
+simplest solution is to compute the left inverse of $\Xa\F_h^T$ and,
+for each $h$, check whether $(\Xa
+\F_h^T)_\mathrm{left}^{-1}\cdot{}\Xb$ is a valid rotation matrix.
+
+However, we assume that the ability to compute left inverses is not
+available. Instead, the observed shapes may only be transformed by a
+small amount at a time, and each transformation is costly in terms of
+computational resources. The goal, then is to estimate the integrals
+in Eq.~\ref{eq:lh-ratio} by evaluating $p(\Xb\ \vert\ \Xa, \theta, h)$
+as few times as possible.
+
+We define $p(\Xb\ \vert\ \Xa, \theta, h)$ to be the similarity between
+$\Xb$ and a transformation of $\Xa$:
+\begin{equation}
+  p(\Xb\ \vert\ \Xa, \theta, h):= S(\Xb, f(\Xa, h, \theta)).
+  \label{eq:likelihood}
+\end{equation}
+We do not know which vertices of $\Xb$ correspond to which vertices of
+$\Xa$, so $S$ must marginalize over the set of possible mappings,
+$\mathbb{M}$. For brevity, let $\mathbf{X}_m=\M\cdot{}f(\Xa, h,
+\theta)$ where $\M\in\mathbb{M}$ is a permutation matrix. Then:
+\begin{equation}
+  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{|\mathbb{M}|} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma^2)
+  \label{eq:similarity}
+\end{equation}
+
+Additionally, we have the constraint of small
+transformations. Specifically, if the current mental image is $\Xt$,
+then:
+\begin{equation}
+  \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
+      f(\Xt, 0, \epsilon) &\mbox{ rotate right by $\epsilon$,} \\
+      f(\Xt, 0, -\epsilon) &\mbox{ rotate left by $\epsilon$,} \\
+      f(\Xt, 1, 0) &\mbox{ flip,} \\
+      f(\Xa, 0, 0) &\mbox{ reset, or} \\
+      f(\Xa, 1, 0) &\mbox{ reset and flip.} \\
+    \end{array} \right.
+\end{equation}
+
+\subsubsection{Hill climbing model}
+
+An intuitive and simple model is to first rotate in the direction that
+increases similarity, and then stop rotating once a ``match'' is
+found. If all rotations have been exhausted, then flip, and try
+rotating again.
+
+\subsubsection{Bayesian quadrature model}
+
+\section{Methods}
+
+\subsection{Stimuli}
+
+\subsection{Procedure}
+
+\section{Results}
+
+\section{Discussion}
 
@@
 
-Place acknowledgments (including funding information) in a section at
-the end of the paper.
-
-\section{References Instructions}
-
-Follow the APA Publication Manual for citation format, both within the
-text and in the reference list, with the following exceptions: (a) do
-not cite the page numbers of any book, including chapters in edited
-volumes; (b) use the same format for unpublished references as for
-published ones. Alphabetize references by the surnames of the authors,
-with single author entries preceding multiple author entries. Order
-references by the same authors by the year of publication, with the
-earliest first.
-
-Use a first level section heading for the reference list. Use a
-hanging indent style, with the first line of the reference flush
-against the left margin and subsequent lines indented by 1/8~inch.
-Below are example references for a conference paper, book chapter,
-journal article, technical report, dissertation, book, and edited
-volume, respectively.
-
-\nocite{ChalnickBillman1988a}
-\nocite{Feigenbaum1963a}
-\nocite{Hill1983a}
-\nocite{OhlssonLangley1985a}
-\nocite{Lewis1978a}
-\nocite{NewellSimon1972a}
-\nocite{ShragerLangley1990a}
+This research was supported by ONR MURI grant number N00014-13-1-0341,
+and a Berkeley Fellowship awarded to JBH.
 
@@
 
-\bibliography{CogSci_Template}
+\bibliography{references}
 

### 2014-01-19 man/cogsci-2014/cogsci-2014.tex
More work on modeling section

@@
 \newcommand{\hf}[0]{h=1}
+\newcommand{\dif}[0]{\,\mathrm{d}}
 
@@
 \begin{equation}
-  p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta, h)p(X_a)p(h)p(\theta)\ \mathrm{d}\theta.
+  p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta, h)p(X_a)p(h)p(\theta)\dif\theta.
   \label{eq:posterior}
@@
 \begin{equation*}
-  \ell := \frac{\int p(X_b\ \vert\ X_a, \theta, \hi)p(X_a)p(\hi)p(\theta)\ \mathrm{d}\theta}{\int p(X_b\ \vert\ X_a, \theta, \hf)p(X_a)p(\hf)p(\theta)\ \mathrm{d}\theta},
+  \ell := \frac{\int p(X_b\ \vert\ X_a, \theta, \hi)p(X_a)p(\hi)p(\theta)\dif\theta}{\int p(X_b\ \vert\ X_a, \theta, \hf)p(X_a)p(\hf)p(\theta)\dif\theta},
 \end{equation*}
@@
 \begin{equation}
-  \ell = \frac{\int p(X_b\ \vert\ X_a, \theta, \hi)\ \mathrm{d}\theta}{\int p(X_b\ \vert\ X_a, \theta, \hf)\ \mathrm{d}\theta}.
+  \ell = \frac{\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta}{\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta}.
   \label{eq:lh-ratio}
@@
 \begin{equation}
-  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{|\mathbb{M}|} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma^2)
+  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{|\mathbb{M}|} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma^2).
   \label{eq:similarity}
@@
     \end{array} \right.
+  \label{eq:actions}
 \end{equation}
 
-\subsubsection{Hill climbing model}
+\subsubsection{Threshold-based model}
 
@@
 increases similarity, and then stop rotating once a ``match'' is
-found. If all rotations have been exhausted, then flip, and try
-rotating again.
+found, as determined by a threshold on the value of $S$. If all
+rotations have been exhausted, then flip, and try rotating again.
 
-\subsubsection{Bayesian quadrature model}
+The consequence of doing a search for a match is that
+Eq.~\ref{eq:lh-ratio} becomes a likelihood ratio test, where $\theta$
+is set to the MLE value, rather than being marginalized:
+\begin{equation}
+  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)}.
+  \label{eq:mle-lh-ratio}
+\end{equation}
+
+\subsubsection{Hill-climbing model}
+
+Unfortunately, it is not always clear what constitutes a good enough
+``match''. Many shapes may have multimodal similarity functions
+\TODO{refer to figure}, and the maximum possible value of $S$ is
+unlikely to be known \textit{a priori} \TODO{cite something about how
+  shape matching is hard}. In this case, we must weaken our stopping
+criteria: rather than searching for a value above a threshold, we
+search for \textit{any} local maximum.  Similar to the threshold-based
+model, we use the (local) MLE estimate of $\theta$, as in
+Eq.~\ref{eq:mle-lh-ratio}.
+
+\subsubsection{Confidence-based model}
+
+An alternative to the previous two models is to approximate the
+integrals in Eq.~\ref{eq:lh-ratio} by picking locations to sample
+which will most improve accuracy of the estimate. Rather than a point
+estimate of the ratio, or even a simple Monte Carlo estimate, we can
+infer a distribution by placing priors over the similarity functions
+themselves:
+\begin{equation}
+\ell = \frac{\int \left[\int S(\Xb, f(\Xa, \theta, \hi))p(\theta)\dif\theta\right] p(S)\dif S}{\int \left[\int S(\Xb, f(\Xa, \theta, \hf))p(\theta)\dif\theta\right] p(S)\dif S}
+\end{equation}
+This is a technique known in the machine-learning literature as
+\textit{Bayesian Quadrature} (BQ)
+\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}. Specifically, we
+use Gaussian Process (GP) priors \cite{Osborne:2012tm} \TODO{probably
+  need to describe better what a GP is}. Because our domain is on a
+circle, we utilize a periodic kernel for the covariance
+\cite{Rasmussen:2006vz}, i.e. $k(\theta,
+\theta^\prime)=h^2\exp[-\frac{2}{w^2}\sin^2(\frac{1}{2}(\theta-\theta^\prime))]$.
+
+Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$, we first place a GP prior on
+$\log(S_h)$. Because GPs have no guarantee of non-negativity, placing
+the prior over the log enforces positivity after it is exponentiated:
+\begin{equation}
+\mathbb{E}[Z_h] \approx \int \exp(\mu_h(\theta))p(\theta)\dif\theta,
+\end{equation}
+where $\mu_h:=\mu(\log S_h)$ is the mean of the log-GP. However, this
+integral is still intractable. To approximate it, we fit a second GP
+over points sampled from the log-GP.\footnote{Note that the approach
+  presented in \cite{Osborne:2012tm} differs slightly from the one
+  used here. We chose this newer version after a discussion with one
+  of the authors on the original BQ paper \TODO{cite personal
+    communication}.}We will denote these points as
+$\bar{S}_h:=\exp(\mu_h)$. Then, from \TODO{cite personal
+  communication}:
+\begin{align}
+  \mathbb{E}[Z_h] &\approx \int \bar{\mu}_h(\theta)p(\theta)\dif\theta,\\
+  \mathbb{V}(Z_h) &\approx \iint \mathrm{Cov}_h(\theta,
+  \theta^\prime)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\prime)p(\theta)p(\theta^\prime)\dif\theta\dif\theta^\prime,
+\end{align}
+where $\bar{\mu}_h:=\mu(\bar{S}_h)$ is the mean of the second GP, and
+$\mathrm{Cov}_h:=\mathrm{Cov}(\log S_h)$ is the covariance of the
+log-GP.
+
+Now that we have a distribution over $Z_h$, we can additionally
+compute which points which we would expect to strengthen our estimate
+(i.e., decrease the variance). From \cite{Osborne:2012tm}, the
+expected variance of $Z_h$ if we sampled a new observation at
+$\theta_a$ is:
+\begin{align}
+\mathbb{E}&[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) + \mathbb{E}[Z_h] - \\
+&\int \mathbb{E}[Z_h|\theta_{a}]^2 \mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a, \theta_a))\dif\log S_h(\theta_a)\nonumber
+\label{ref:expected-variance}
+\end{align}
+
+Revisiting our likelihood ratio, we now have:
+\begin{equation}
+p(\ell)\approx\frac{\mathcal{N}(\mathbb{E}[Z_0], \mathbb{V}(Z_0))}{\mathcal{N}(\mathbb{E}[Z_1], \mathbb{V}(Z_1))}
+\end{equation}
+\TODO{figure out what to do here.}
 

### 2014-01-20 man/cogsci-2014/cogsci-2014.tex
Start methods section

@@
 
+We randomly generated 20 shapes of 5 or 6 vertices (e.g., $\Xa$ in
+Figure \ref{fig:stimuli}). For each shape, we computed 20 ``same'' and
+20 ``flipped'' stimuli pairs, with $\theta$ spaced at $20^\prime$
+increments between 0 and 360 (with 0 and 180 degree rotations repeated
+twice, in order to gather an equal number of responses for each angle
+between 0 and 180). ``Same'' pairs were created by rotating $\Xa$ by
+$\theta$; ``flipped'' pairs first reflected $\Xa$ across the $y$-axis,
+then rotated by $\theta$.
+
+We generated 5 shapes to be used in the practice block. Across all the
+practice 10 stimuli, each shape and each angle (60, 120, 180, 240, or
+300) was repeated twice (once ``flipped'' and once ``same'') such that
+no shape was presented at the same angle twice. Additionally, we
+generated a sixth shape to be included along with the instructions,
+which had both a ``flipped'' and ``same'' version, each rotated to 320
+degrees.
+
+\subsection{Participants}
+
+We recruited \TODO{N} participants on Amazon's Mechanical Turk. Each
+participant was paid \$1.00 for approximately 15 minutes of work. Out
+of these participantsm \TODO{N\% (N)} were included in our
+analyses. The others were excluded either because of an experimental
+error (\TODO{N\% (N)}), because they had already compelted a previous
+version of the experiment (\TODO{N\% (N)}), or because they failed a
+comprehension check (\TODO{N\% (N)}).
+
+\subsection{Design}
+
+The experiment consisted of one block of 10 practice trials followed
+by two blocks of 100 experiment trials. Within a block, trials were
+presented in a random order.
+
+All participants saw the same 10 practice trials as described
+above. Participants in different conditions saw different experimental
+trials, however. Across participants, there were a total of 800
+experimental stimuli (20 shapes $\times$ 20 angles $\times$ 2
+same/flipped). These stimuli were split across conditions in the
+following manner. First, stimuli were split into four blocks of 200
+trials. Within each block, each shape was repeated 10 times and each
+rotation was repeated 10 times (5 same, 5 flipped), such that across
+all blocks, each stimulus appeared exactly once. Each block was then
+split in half. Each participant then completed two half-blocks (not
+necessarily both from the original full block).
+
 \subsection{Procedure}
+
+Participants were given the following instructions while being shown
+an example ``same'' pair and an example ``flipped'' pair: \textit{``On
+  each trial, you will see two images. Sometimes, they show the
+  \textbf{same} object. Other times, the images show \textbf{flipped}
+  objects. The task is to determine whether the two images show the
+  \textbf{same} object or \textbf{flipped} objects.''}
+
+On each trial, particpants were instructed to press the `b' key to
+begin and to focus on the fixation cross that appeared for \TODO{N}
+milliseconds afterwards. The two images were then presented
+side-by-side, each at 300px $\times$ 300px, and participants could
+press `s' to indicate they thought the images depicted the same
+object, or `d' to indicate they thought the images depicted flipped
+objects.
+
+While there was no time limit on response time, participants were
+urged to repond as quickly as possible while still being
+accurate. Specifically, we asked them to aim for at least 85\% correct
+in the experimental blocks, and provided a score counter to keep them
+informed of their accuracy.
 

### 2014-01-21 lib/mental_rotation/model/__init__.py
Add oracle model

@@ -1,8 +1,10 @@
 from .base_model import BaseModel
 from .gold_standard_model import GoldStandardModel
+from .oracle_model import OracleModel
 from .hill_climbing_model import HillClimbingModel
 from .threshold_model import ThresholdModel
 from .bayesian_quadrature_model import BayesianQuadratureModel
 
-__all__ = ['BaseModel', 'GoldStandardModel', 'HillClimbingModel', 
-           'ThresholdModel', 'BayesianQuadratureModel']
+__all__ = ['BaseModel', 'GoldStandardModel', 'OracleModel',
+           'HillClimbingModel', 'ThresholdModel',
+           'BayesianQuadratureModel']

### 2014-01-21 lib/mental_rotation/model/base_model.py

@@ -147,14 +147,20 @@ class BaseModel(object):
 
     @property
     def log_lh_h0(self):
-        Fi = self.F_i
-        log_S0 = self.log_S_i[Fi == 0].max()
+        Fi = self.F_i == 0
+        if Fi.any():
+            log_S0 = self.log_S_i[Fi].max()
+        else:
+            log_S0 = -np.inf
         return log_S0
 
     @property
     def log_lh_h1(self):
-        Fi = self.F_i
-        log_S1 = self.log_S_i[Fi == 1].max()
+        Fi = self.F_i == 1
+        if Fi.any():
+            log_S1 = self.log_S_i[Fi].max()
+        else:
+            log_S1 = -np.inf
         return log_S1
 
     def hypothesis_test(self):

### 2014-01-21 lib/mental_rotation/model/model.py

@@ -151,7 +151,7 @@ class R(Variable):
     
     @value.setter
     def value(self, val):
-        val = val % (2 * np.pi)
+        val = float(val) % (2 * np.pi)
         if val >= np.pi:
             val -= 2 * np.pi
         self._value = val

### 2014-01-21 lib/mental_rotation/model/oracle_model.py

@@ -1,0 +1,86 @@
+import numpy as np
+import scipy.optimize as optim
+import logging
+
+from . import BaseModel
+
+logger = logging.getLogger("mental_rotation.model.oracle")
+
+
+class OracleModel(BaseModel):
+
+    _iter = 720
+
+    def __init__(self, *args, **kwargs):
+        super(OracleModel, self).__init__(*args, **kwargs)
+        self.direction = None
+        self.target = None
+
+    ##################################################################
+    # Sampling
+
+    def _solve(self):
+        Xa = self.model['Xa'].value
+        Xb = self.model['Xb'].value
+        
+        h = 0
+        R = np.dot(np.linalg.pinv(Xa), Xb).T
+        if np.sign(R[0, 1]) == np.sign(R[1, 0]):
+            h = 1
+            F = np.array([[-1, 0], [0, 1]], dtype=float)
+            R = np.dot(np.linalg.pinv(np.dot(Xa, F.T)), Xb).T
+        
+        costheta = R[0, 0]
+        sintheta = R[1, 0]
+        theta = np.arctan2(sintheta, costheta)
+
+        return theta, h
+
+    def draw(self):
+        if self._current_iter == 0:
+            R, F = self._solve()
+            self.model['R'].value = 0
+            self.model['F'].value = F
+            self.target = R
+            self.direction = np.sign(self._unwrap(R))
+            return
+
+        R = self.model['R'].value
+        if np.abs(self._unwrap(R - self.target)) < self.opts['step']:
+            self.model['R'].value = self.target
+            self.status = 'done'
+
+        else:
+            step = self.direction * np.abs(self._random_step())
+            self.model['R'].value = R + step
+
+    ##################################################################
+    # Plotting methods
+    def plot(self, ax, F=None):
+        Fi = self.F_i
+
+        if F is None or F == 0:
+            Ri0 = self.R_i[Fi == 0]
+            Si0 = self.S_i[Fi == 0]
+            ax.plot(Ri0, Si0, 'r-', lw=2)
+            ax.plot(Ri0, Si0, 'ro', markersize=5)
+
+        if F is None or F == 1:
+            Ri1 = self.R_i[Fi == 1]
+            Si1 = self.S_i[Fi == 1]
+            ax.plot(Ri1, Si1, 'b-', lw=2)
+            ax.plot(Ri1, Si1, 'bo', markersize=5)
+
+    ##################################################################
+    # Copying/Saving
+
+    def __getstate__(self):
[…]

### 2014-01-21 lib/mental_rotation/model/hill_climbing_model.py
Fix bugs in plotting functions of hill climbing and threshold models

@@ -57,17 +57,20 @@ class HillClimbingModel(BaseModel):
     ##################################################################
     # Plotting methods
 
-    def plot(self, ax):
+    def plot(self, ax, F=None):
         Fi = self.F_i
 
-        Ri0 = self.R_i[Fi == 0]
-        Si0 = self.S_i[Fi == 0]
+        if F is None or F == 0:
+            Ri0 = self.R_i[Fi == 0]
+            Si0 = self.S_i[Fi == 0]
+            ax.plot(Ri0, Si0, 'r-', lw=2)
+            ax.plot(Ri0, Si0, 'ro', markersize=5)
 
-        Ri1 = self.R_i[Fi == 1]
-        Si1 = self.S_i[Fi == 1]
-
-        ax.plot(Ri0, Si0, 'ro', markersize=7)
-        ax.plot(Ri1, Si1, 'bo', markersize=7)
+        if F is None or F == 1:
+            Ri1 = self.R_i[Fi == 1]
+            Si1 = self.S_i[Fi == 1]
+            ax.plot(Ri1, Si1, 'b-', lw=2)
+            ax.plot(Ri1, Si1, 'bo', markersize=5)
 
     ##################################################################
     # Copying/Saving

### 2014-01-21 lib/mental_rotation/model/threshold_model.py

@@ -74,17 +74,20 @@ class ThresholdModel(BaseModel):
     ##################################################################
     # Plotting methods
 
-    def plot(self, ax):
+    def plot(self, ax, F=None):
         Fi = self.F_i
 
-        Ri0 = self.R_i[Fi == 0]
-        Si0 = self.S_i[Fi == 0]
+        if F is None or F == 0:
+            Ri0 = self.R_i[Fi == 0]
+            Si0 = self.S_i[Fi == 0]
+            ax.plot(Ri0, Si0, 'r-', lw=2)
+            ax.plot(Ri0, Si0, 'ro', markersize=5)
 
-        Ri1 = self.R_i[Fi == 1]
-        Si1 = self.S_i[Fi == 1]
-
-        ax.plot(Ri0, Si0, 'ro', markersize=7)
-        ax.plot(Ri1, Si1, 'bo', markersize=7)
+        if F is None or F == 1:
+            Ri1 = self.R_i[Fi == 1]
+            Si1 = self.S_i[Fi == 1]
+            ax.plot(Ri1, Si1, 'b-', lw=2)
+            ax.plot(Ri1, Si1, 'bo', markersize=5)
 
     ##################################################################
     # Copying/Saving

### 2014-01-21 man/cogsci-2014/cogsci-2014.tex
Add figures to paper

@@
 \usepackage{amsmath, amsthm, amssymb}
+\usepackage{subfig}
 
@@
 \section{Models of mental rotation}
+
+\begin{figure*}[t]
+  \begin{center}
+    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
+    \caption{\textbf{Model traces.}}
+    \label{fig:model-traces}
+  \end{center}
+\end{figure*}
 
@@
 
+\begin{figure*}[t]
+  \begin{center}
+    \includegraphics[width=\textwidth]{../../figures/D/response-times.pdf}
+    \caption{\textbf{Response times.}}
+    \label{fig:response-times}
+  \end{center}
+\end{figure*}
+
 We recruited \TODO{N} participants on Amazon's Mechanical Turk. Each
@@
 
-The experiment consisted of one block of 10 practice trials followed
-by two blocks of 100 experiment trials. Within a block, trials were
-presented in a random order.
+The experiment consisted 210 total trials: of one block of 10 practice
+trials followed by two blocks of 100 experiment trials. Within a
+block, trials were presented in a random order.
 
-All participants saw the same 10 practice trials as described
-above. Participants in different conditions saw different experimental
-trials, however. Across participants, there were a total of 800
-experimental stimuli (20 shapes $\times$ 20 angles $\times$ 2
-same/flipped). These stimuli were split across conditions in the
-following manner. First, stimuli were split into four blocks of 200
-trials. Within each block, each shape was repeated 10 times and each
-rotation was repeated 10 times (5 same, 5 flipped), such that across
-all blocks, each stimulus appeared exactly once. Each block was then
-split in half. Each participant then completed two half-blocks (not
-necessarily both from the original full block).
+All participants saw the same 10 practice trials (as described
+above). There were a total of 800 experimental stimuli (20 shapes
+$\times$ 20 angles $\times$ 2 same/flipped), which were split into 8
+blocks of 100 across conditions in the following manenr. First,
+stimuli were split into four blocks of 200 trials. Within each block,
+each shape was repeated 10 times and each rotation was repeated 10
+times (5 same, 5 flipped), such that across all blocks, each stimulus
+appeared exactly once. Each block was then split in half. Each
+participant then completed two half-blocks (not necessarily both from
+the original full block).
 
 \subsection{Procedure}
+
+\begin{figure*}[t]
+  \begin{center}
+    \includegraphics[width=0.85\textwidth]{../../figures/D/accuracy.pdf}
+    \caption{\textbf{Accuracy.}}
+    \label{fig:response-times}
+  \end{center}
+\end{figure*}
 
@@
 
-While there was no time limit on response time, participants were
-urged to repond as quickly as possible while still being
-accurate. Specifically, we asked them to aim for at least 85\% correct
-in the experimental blocks, and provided a score counter to keep them
-informed of their accuracy.
+While there was no limit on response time, participants were urged to
+repond as quickly as possible while still being
+accurate. Specifically, we asked them to aim for at least 85\%
+accuracy in the experimental blocks, and provided a counter to keep
+them informed of their score.
 
 \section{Results}
+
+\begin{figure}[t]
+  \centering
+  \subfloat{\includegraphics[width=0.25\textwidth]{../../figures/D/trial-vs-accuracy.pdf}}%
+  \subfloat{\includegraphics[width=0.25\textwidth]{../../figures/D/trial-vs-time.pdf}}
+  \caption{\textbf{Effect of learning.}}
+  \label{fig:learning}
+\end{figure}
+
+response time as a function of rotation
+
+accuracy as a function of rotation
+
+practice effects
 

### 2014-01-26 analysis/analyses/accuracy_corrs.py
Standardize number of bootstrap samples

@@ -27,7 +27,6 @@ def run(data, results_path, seed):
             corr = dict(util.bootcorr(
                 accuracy_means['exp'],
                 accuracy_means[key],
-                nsamples=5000,
                 method='spearman'))
 
             print "exp v. %s: %s" % (key, util.report_spearman.format(**corr))

### 2014-01-26 analysis/analyses/response_time_corrs.py

@@ -28,7 +28,6 @@ def run(data, results_path, seed):
             corr = dict(util.bootcorr(
                 response_means['exp'],
                 response_means[key],
-                nsamples=5000,
                 method='spearman'))
 
             print "exp v. %s: %s" % (key, util.report_spearman.format(**corr))

### 2014-01-26 lib/mental_rotation/analysis.py

@@ -86,7 +86,7 @@ def load_all(version, data_path, human=None):
     return data
 
 
-def bootstrap(x, nsamples=1000):
+def bootstrap(x, nsamples=10000):
     arr = np.asarray(x)
     n, = arr.shape
     boot_idx = np.random.randint(0, n, n * nsamples)
@@ -112,7 +112,7 @@ def beta(x):
     return stats
 
 
-def bootcorr(x, y, nsamples=1000, method='pearson'):
+def bootcorr(x, y, nsamples=10000, method='pearson'):
     arr1 = np.asarray(x)
     arr2 = np.asarray(y)
     n, = arr1.shape

### 2014-01-26 analysis/analyses/response_time.py
Report standard deviation for response time rather than confidence intervals

@@ -14,12 +14,15 @@ def run(data, results_path, seed):
 
         for name in sorted(data.keys()):
             df = data[name]
-            t = dict(util.bootstrap(df['time']))
+            stats = {
+                'mean': df['time'].mean(),
+                'std': df['time'].std()
+            }
 
-            print "%s:\t%s" % (name, util.report_mean.format(**t))
+            print "%s:\t%s" % (name, util.report_mean.format(**stats))
             cmd = util.newcommand(
                 "%sTime" % name.capitalize(),
-                util.latex_mean.format(**t))
+                util.latex_mean.format(**stats))
 
             fh.write(cmd)
 

### 2014-01-26 analysis/analyses/util.py

@@ -24,5 +24,5 @@ latex_pearson = r"$r={median:.2f}$, 95\% CI $[{lower:.2f}, {upper:.2f}]$"
 report_percent = "M={median:.1f}%, 95% CI [{lower:.1f}%, {upper:.1f}%]"
 latex_percent = r"$M={median:.1f}\%$, 95\% CI $[{lower:.1f}\%, {upper:.1f}\%]$"
 
-report_mean = "M={median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
-latex_mean = r"$M={median:.2f}$, 95\% CI $[{lower:.2f}, {upper:.2f}]$"
+report_mean = "M={mean:.2f}, SD={std:.2f}"
+latex_mean = r"$M={mean:.2f}$, $SD={std:.2f}$"

### 2014-01-26 man/cogsci-2014/cogsci-2014.tex
Input analysis tex files
Update figures, more work on methods

@@
 }{\end{itemize}}
+
+\input{../../results/D/response_time.tex}
+\input{../../results/D/response_time_corrs.tex}
+\input{../../results/D/accuracy.tex}
+\input{../../results/D/accuracy_corrs.tex}
+\input{../../results/D/trial_time_corrs.tex}
+\input{../../results/D/trial_accuracy_corrs.tex}
 
@@
 
-\subsection{Participants}
-
 \begin{figure*}[t]
   \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/response-times.pdf}
+    \includegraphics[width=\textwidth]{../../figures/D/response_time.pdf}
     \caption{\textbf{Response times.}}
@@
 
-We recruited \TODO{N} participants on Amazon's Mechanical Turk. Each
-participant was paid \$1.00 for approximately 15 minutes of work. Out
-of these participantsm \TODO{N\% (N)} were included in our
-analyses. The others were excluded either because of an experimental
-error (\TODO{N\% (N)}), because they had already compelted a previous
-version of the experiment (\TODO{N\% (N)}), or because they failed a
-comprehension check (\TODO{N\% (N)}).
-
 \subsection{Design}
 
-The experiment consisted 210 total trials: of one block of 10 practice
-trials followed by two blocks of 100 experiment trials. Within a
-block, trials were presented in a random order.
+We recruited 247 participants on Amazon's Mechanical Turk. Each
+participant was paid \$1.00 for approximately 15 minutes of work,
+consisting of one block of 10 practice trials followed by two blocks
+of 100 experiment trials. Within a block, trials were presented in a
+random order.
 
@@
 On each trial, particpants were instructed to press the `b' key to
-begin and to focus on the fixation cross that appeared for \TODO{N}
+begin and to focus on the fixation cross that appeared for 750
 milliseconds afterwards. The two images were then presented
@@
 
+\section{Analysis}
+
+Out of the 247 participants, 200 (81\%) were included in our
+analyses. Of the rest, we excluded 10 (4\%) because of an experimental
+error, 6 (2.4\%) because they had already compelted a previous version
+of the experiment, and 31 (12.6\%) because they failed a comprehension
+check. A participant was marked as passing the comprehension check if
+they answered at least 85\% of ``easy'' trials correctly. ``Easy''
+trials were defined to be those for which the presented shapes
+differed by a rotation of no more than 20 degrees.
+
+We ran model simulations for the same stimuli that participants
+viewed. We ran 10 samples of each stimulus under the Oracle model, and
+100 samples of each stimulus for the Threshold, Hill Climbing, and
+Bayesian Quadrature models, with the exception of stimuli where
+$\theta=0$ or $\theta=180$, in which case we ran double the number of
+samples.
+
+For all analyses (except those of accuracy), confidence intervals were
+computed using a bootstrap analysis with 10000 bootstrap samples. For
+analyses of accuracy, confidence intervals were computed from a
+binomial proportion with a Jeffrey's beta prior.
+
 \section{Results}
@@
   \centering
-  \subfloat{\includegraphics[width=0.25\textwidth]{../../figures/D/trial-vs-accuracy.pdf}}%
-  \subfloat{\includegraphics[width=0.25\textwidth]{../../figures/D/trial-vs-time.pdf}}
+  \subfloat{\includegraphics[width=0.25\textwidth]{../../figures/D/trial_accuracy.pdf}}%
+  \subfloat{\includegraphics[width=0.25\textwidth]{../../figures/D/trial_time.pdf}}
   \caption{\textbf{Effect of learning.}}

### 2014-01-27 man/cogsci-2014/cogsci-2014.tex
Work on results section

@@
 
+\newcommand{\Oc}[0]{Oracle}
+\newcommand{\Th}[0]{Threshold}
+\newcommand{\Hc}[0]{HC}
+\newcommand{\Bq}[0]{BQ}
+
 \newenvironment{pitemize}{
@@
 \input{../../results/D/trial_accuracy_corrs.tex}
+\input{../../results/D/num_chance.tex}
+\input{../../results/D/theta_time_corrs.tex}
+\input{../../results/D/theta_accuracy_corrs.tex}
 
@@
 \end{abstract}
+
+\TODO{put correlations in figures}
+
+\TODO{need a figure with an example similarity curve, and example
+  stimuli}
 
@@
 \section{Models of mental rotation}
-
-\begin{figure*}[t]
-  \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
-    \caption{\textbf{Model traces.}}
-    \label{fig:model-traces}
-  \end{center}
-\end{figure*}
 
@@
 
-\subsubsection{Threshold-based model}
+\subsubsection{\Oc{} model}
 
-An intuitive and simple model is to first rotate in the direction that
-increases similarity, and then stop rotating once a ``match'' is
-found, as determined by a threshold on the value of $S$. If all
-rotations have been exhausted, then flip, and try rotating again.
+The previous hypothesis in the literature is that people perform a
+mental rotation in the shortest direction until a match is found, and
+that determining this direction is an implementation detail \TODO{cite
+  anderson?}. To reflect this hypothesis, we created an ``oracle''
+model which knows the correct rotation \textit{a priori}, and simply
+performs rotations in the shortest direction until it reaches the
+correct rotation.
+
+\subsubsection{\Th{} model}
+
+An intuitive and simple model which does not assume the direction is
+known might first rotate in the direction that increases similarity,
+and then stop rotating once a ``match'' is found, as determined by a
+threshold on the value of $S$. If all rotations have been exhausted,
+then flip, and try rotating again.
 
@@
 
-\subsubsection{Hill-climbing model}
+\subsubsection{Hill climbing model}
 
@@
 criteria: rather than searching for a value above a threshold, we
-search for \textit{any} local maximum.  Similar to the threshold-based
-model, we use the (local) MLE estimate of $\theta$, as in
-Eq.~\ref{eq:mle-lh-ratio}.
+search for \textit{any} local maximum, as in the Hill Climbing search
+algorithm (\Hc{}). Similar to the \Th{} model, we use the (local) MLE
+estimate of $\theta$, as in Eq.~\ref{eq:mle-lh-ratio}.
 
-\subsubsection{Confidence-based model}
+\subsubsection{Bayesian quadrature model}
 
-An alternative to the previous two models is to approximate the
-integrals in Eq.~\ref{eq:lh-ratio} by picking locations to sample
-which will most improve accuracy of the estimate. Rather than a point
-estimate of the ratio, or even a simple Monte Carlo estimate, we can
-infer a distribution by placing priors over the similarity functions
-themselves:
+An alternative to the previous models is not to perform a search for
+the MLE, but to approximate the integrals in Eq.~\ref{eq:lh-ratio} by
+picking locations to sample which will most improve accuracy of the
+estimate. Rather than a point estimate of the ratio, or even a simple
+Monte Carlo estimate, we can infer a distribution by placing priors
+over the similarity functions themselves:
 \begin{equation}
@@
 over points sampled from the log-GP.\footnote{Note that the approach
-  presented in \cite{Osborne:2012tm} differs slightly from the one
-  used here. We chose this newer version after a discussion with one
-  of the authors on the original BQ paper \TODO{cite personal
-    communication}.}We will denote these points as
-$\bar{S}_h:=\exp(\mu_h)$. Then, from \TODO{cite personal
-  communication}:
+  used here differs slightly from the one presented in
+  \citeA{Osborne:2012tm} and is based on personal communication with
+  one of the authors \cite{Duvenaud:2013td}.}We will denote these
+points as $\bar{S}_h:=\exp(\mu_h)$. Then, from
+\citeA{Duvenaud:2013td}:
 \begin{align}
@@
     \includegraphics[width=\textwidth]{../../figures/D/response_time.pdf}
-    \caption{\textbf{Response times.}}
+    \caption{\textbf{Response times.} \TODO{}}
     \label{fig:response-times}
@@
     \includegraphics[width=0.85\textwidth]{../../figures/D/accuracy.pdf}
-    \caption{\textbf{Accuracy.}}
-    \label{fig:response-times}
+    \caption{\textbf{Accuracy.} \TODO{}}
+    \label{fig:accuracy}
   \end{center}
@@
 We ran model simulations for the same stimuli that participants
-viewed. We ran 10 samples of each stimulus under the Oracle model, and
-100 samples of each stimulus for the Threshold, Hill Climbing, and
-Bayesian Quadrature models, with the exception of stimuli where
+viewed. We ran 10 samples of each stimulus under the \Oc{} model, and
+100 samples of each stimulus for the \Th{}, \Hc{}, and
+\Bq{} models, with the exception of stimuli where
 $\theta=0$ or $\theta=180$, in which case we ran double the number of
@@
 
-For all analyses (except those of accuracy), confidence intervals were
-computed using a bootstrap analysis with 10000 bootstrap samples. For
-analyses of accuracy, confidence intervals were computed from a
-binomial proportion with a Jeffrey's beta prior.
+For analyses of response time, confidence intervals were computed
+using a bootstrap analysis (sampling with replacement) with 1000
+bootstrap samples. Response times were only only considered for
+correct responses.
+
+For analyses of accuracy, confidence intervals were
+computed from a binomial proportion with a Jeffrey's beta prior.  For
+correlations, we performed a bootstrap analysis (sampling with
+replacement) over Spearman correlations using 10000 bootstrap samples.
+
+To test if participants or a model was above chance on a particular
+stimulus, we used the same posterior and checked whether $p>0.05$ for
+the 50\% mark.
 
@@
   \subfloat{\includegraphics[width=0.25\textwidth]{../../figures/D/trial_time.pdf}}
-  \caption{\textbf{Effect of learning.}}
+  \caption{\textbf{Effect of learning.} \TODO{}}
   \label{fig:learning}
@@
 
-response time as a function of rotation
+\subsubsection{Response times}
 
-accuracy as a function of rotation
+The average response time across all stimuli was \ExpTime{} seconds,
+and the average response time for an individual stimulus ranged from
+\ExpTimeMin{} to \ExpTimeMax{} seconds. The average accuracy across
+all stimuli was \ExpAccuracy{}, though there were \ExpNumChance{}
+individual stimuli (out of 720) for which people were not above
+chance.
 
-practice effects
+As expected, the minimum angle of rotation was significanty
+monotonically correlated with average per-stimulus response times,
+both for flipped pairs (\ExpThetaTimeCorrFlipped{}) and same pairs
+(\ExpThetaTimeCorrSame{}). This was also the case for the number of
+actions taken by the \Oc{} model (\OcThetaTimeCorr{}), the \Th{}
+model (\ThThetaTimeCorr{}), and the \Bq{} model
+(\BqThetaTimeCorr{}), but not the \Hc{} model
+(\HcThetaTimeCorr{}). Fig.~\ref{fig:response-times}a
+
+The number of actions taken by the \Oc{} model was the best predictor
+of human response times, with a correlation of \ExpOcTimeCorr{}. The
+\Th{} model was the next best (\ExpThTimeCorr{}), with the \Bq{} model
+not far behind (\ExpBqTimeCorr{}). The \Hc{} model was not
+significantly correlated with people (\ExpHcTimeCorr{}).
+
+\TODO{flipped response times higher than same response times for
+  smaller angles?}
+
+\subsubsection{Accuracy}
+
+Of the models, the \Oc{} and \Th{} models peformed perfectly,
+with 100\% accuracy. The \Bq{} model was very accurate
+overall (\BqAccuracy{}), though there were \BqNumChance{} stimuli for
+which it was not above chance. The \Hc{} model was not
+particularly accurate, and although it did perform above chance
+overall (\HcAccuracy{}), there were \HcNumChance{} individual stimuli
+for which it was not above chance.
+
+The minimum angle was also correlated with the average per-stimulus
+accuracy: for flipped pairs, \ExpThetaAccuracyCorrFlipped{}, and for
+same pairs, \ExpThetaAccuracyCorrSame{}. There was also a significant
+effect for the \Hc{} model (\HcThetaAccuracyCorr{}) and the
+\Bq{} model
+(\BqThetaAccuracyCorr{}). Fig.~\ref{fig:accuracy}
+
+There was a moderate correlation between the accuracy of the \Hc{}
+model and human accuracy (\ExpHcAccuracyCorr{}). The \Bq{} model had a
+similar correlation, with \ExpBqAccuracyCorr{}.
+
+\subsubsection{Practice effects}
+
+There was a significant effect of learning both on response time
+(\ExpTrialTimeCorr{}) and on accuracy (\ExpTrialAccuracyCorr{}),
+though the effect on accuracy was not significant during the second
+half of the experiment (\ExpaTrialAccuracyCorr{} for the first half
+vs. \ExpbTrialAccuracyCorr{} for the second
+half). Fig.~\ref{fig:learning}
 
 \section{Discussion}
+
+\begin{figure*}[t]
+  \begin{center}
+    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
+    \caption{\textbf{Model traces.} \TODO{} \TODO{put pictures of the
+        stimuli here}}
+    \label{fig:model-traces}
+  \end{center}
+\end{figure*}
+
+We replicated the previous monotonic response time result
+
+Also replicated the constant accuracy for flipped stimuli, but not for
+same stimuli, from \TODO{cite cooper}. Why is this? Maybe the strategy
+is ``they are different until proven the same''.
+
+None of the models were particularly good fits to human data. The best
+fit was the \Oc{} model, though this is not satisfying because the
+\Oc{} model knows the answer beforehand, which people obviously do
+not. The next best model, the \Th{} model, is also not a
+particularly satisfying account, because it knows the global maximum
+\textit{a priori}, which is not a reasonable assumption. \TODO{explain
+  why it is faster for 180 than mid angles -- it is because for a lot
+  of stimuli, there is rotational symmetry at 180, so it almost always
+  starts checking the correct version first. For mid angles, it might
+  check the wrong version first, meaning it takes much longer.}
+
+The \Bq{} model makes more reasonable assumptions about the observer's
+knowledge, but has its own problems. Because it is forced to take
+small steps, it often gets stuck sampling the same region over and
+over again, particularly for stimuli with larger rotations
+(Fig.~\ref{fig:model-traces}). As a result, its response time curve is
+convex rather than concave.
+
+\TODO{talk about how these results suggest a model of mental rotation
+  is harder than was previous thought, and that this is a first step
+  towards detangling these issues.}
 

### 2014-01-27 man/cogsci-2014/cogsci-2014.tex
Fill in some citations

@@
 widely studied and has sparked intense debate about the underlying
-representation of mental images \TODO{cite}. Despite this vast
-literature, there have been surprisingly few computational models of
-mental imagery.
+representation of mental images \cite{Kosslyn:2009tj,
+  Pylyshyn:2002vk}. Despite this vast literature, there have been
+surprisingly few computational models of mental imagery.
 
@@
 
+previous models of mental imagery: \cite{Kosslyn:1977tv, Anderson1978,
+  Funt:1983wn, Glasgow:1992tj}
+
 assume people rotate until congruent
@@
 
-evidence that people *don't* actually rotate until congruent
+evidence that people *don't* actually rotate until congruent,
+e.g. rotating away from near-perfect matches \cite{Gardony:2013gn}
 
@@
 In the mental rotation task, people are presented with pairs of
-two-dimensional shapes similar to those used by \cite{Cooper:1975wp}
-(e.g., Figure \ref{fig:stimuli}) and must determine whether the two
-images depict the same shape, or mirror-image shapes.
+two-dimensional shapes similar to those used by \citeA{Cooper:1975wp};
+an example stimulus is shown in Fig.~\ref{fig:stimuli}. The goal is to
+determine whether the two images depict the same shape, or
+mirror-image (flipped) shapes.
 
@@
 \cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}. Specifically, we
-use Gaussian Process (GP) priors \cite{Osborne:2012tm} \TODO{probably
-  need to describe better what a GP is}. Because our domain is on a
-circle, we utilize a periodic kernel for the covariance
+use Gaussian Process (GP) priors, as in \citeA{Osborne:2012tm}
+\TODO{probably need to describe better what a GP is}. Because our
+domain is on a circle, we utilize a periodic kernel for the covariance
 \cite{Rasmussen:2006vz}, i.e. $k(\theta,
@@
 compute which points which we would expect to strengthen our estimate
-(i.e., decrease the variance). From \cite{Osborne:2012tm}, the
+(i.e., decrease the variance). From \citeA{Osborne:2012tm}, the
 expected variance of $Z_h$ if we sampled a new observation at
@@
 \mathbb{E}&[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) + \mathbb{E}[Z_h] - \\
-&\int \mathbb{E}[Z_h|\theta_{a}]^2 \mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a, \theta_a))\dif\log S_h(\theta_a)\nonumber
+&\int \mathbb{E}[Z_h|\theta_{a}]^2 \mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a, \theta_a))\dif\log S_h(\theta_a).\nonumber
 \label{ref:expected-variance}
@@
 \begin{equation}
-p(\ell)\approx\frac{\mathcal{N}(\mathbb{E}[Z_0], \mathbb{V}(Z_0))}{\mathcal{N}(\mathbb{E}[Z_1], \mathbb{V}(Z_1))}
+p(\ell)\approx\frac{\mathcal{N}(\mathbb{E}[Z_0], \mathbb{V}(Z_0))}{\mathcal{N}(\mathbb{E}[Z_1], \mathbb{V}(Z_1))}.
 \end{equation}
-\TODO{figure out what to do here.}
+This distribution does not have a nice form. However, we really are
+just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$: thus, rather than
+computing $p(\ell)$, we use $Z_D=Z_0-Z_1$. Assuming $Z_0$ and $Z_1$
+are independent,\footnote{This assumption is probably not
+  correct. \TODO{explanation?}} we have:
+\begin{equation}
+  p(Z_D)\propto\mathcal{N}(\mathbb{E}[Z_0] - \mathbb{E}[Z_1], \mathbb{V}(Z_0) + \mathbb{V}(Z_1)).
+\end{equation}
+We then sample new observations until we are at least 95\% confident
+that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
+$h=0$, and when $p(Z_D<0)>0.975$, we accept $h=1$.
 
@@
 
-We replicated the previous monotonic response time result
+We replicated the previous monotonic response time result, though not
+the linear resulte. Actually kind of similar to \cite{Gardony:2013gn}
+\TODO{elaborate}
 
 Also replicated the constant accuracy for flipped stimuli, but not for
-same stimuli, from \TODO{cite cooper}. Why is this? Maybe the strategy
-is ``they are different until proven the same''.
+same stimuli, from \cite{Cooper:1975wp}. Why is this? Maybe the
+strategy is ``they are different until proven the same''.
 
@@
 
+Other problems: evidence that rotations aren't actually holistic
+\cite{Yuille:1982tx}, eye movements \cite{Just1976}
+
 \TODO{talk about how these results suggest a model of mental rotation

### 2014-01-28 man/cogsci-2014/cogsci-2014.tex
Some work on background

@@
 
-\title{An active sampling model of mental rotation}
+\title{Modeling mental rotation}
 
@@
 
+\TODO{fix title}
+
 \section{Introduction}
@@
 widely studied and has sparked intense debate about the underlying
-representation of mental images \cite{Kosslyn:2009tj,
+representation of mental images \cite<e.g.,>{Kosslyn:2009tj,
   Pylyshyn:2002vk}. Despite this vast literature, there have been
@@
 
-classic result by shepard and metzler
+In the classic experiment by \citeA{Shepard1971}, participants viewed
+images of three-dimensional objects and had to determine whether the
+images depicted the same object (which differed by a rotation) or two
+separate objects (which differed by a reflection and a rotation). The
+result was that people's response times had a strong linear
+correlation with the minimum angle of rotation. The conclusion was
+that people were solving this task by ``mentally rotating'' the
+objects until they were congruent.
 
-previous models of mental imagery: \cite{Kosslyn:1977tv, Anderson1978,
-  Funt:1983wn, Glasgow:1992tj}
+However, there are some questions regarding this process of problem
+solving that are not explained by the alignment hypothesis: how do
+people know the axis around which to rotate the objects? If the axis
+is known, how do people know which direction to rotate the objects?
+And finally, how do people know when to stop rotating and give a
+response?
 
-assume people rotate until congruent
+Previous models of mental rotation have, for the most part, not
+addressed these questions. \citeA{Kosslyn:1977tv} focus primarily on
+simulating the visual simulation buffer rather than the decision of
+\textit{what} should be simulated. Similarly, \citeA{Glasgow:1992tj}
+are more concerned with the representation of imagery than the
+process. Although \citeA{Anderson1978} emphasizes the importance of
+considering both representation and process, he dismisses the problem
+of determining the direction and axis of rotation as merely a
+``technical difficulty''.
 
-but how do people actually construct the plan for the rotation?
+The only model (of which the authors are aware) that has more
+seriously attempted to address these concerns is the parallel-process
+model of mental rotation by \citeA{Funt:1983wn}. He determines the
+direction and axis of rotation using the method-of-moments method;
+once this target rotation computed, his model rotates one object until
+the target rotation is reached, and then compares the two to see if
+they are congruent.
 
-evidence that people *don't* actually rotate until congruent,
-e.g. rotating away from near-perfect matches \cite{Gardony:2013gn}
+More recently, evidence has come to light that people \textit{do not}
+actually rotate until the objects are congruent. Eye-tracking studies
+suggest people pay attention to certain salient features of the
+object, rather than a holistic representation \cite{Just1976}, and 
+\citeA{Yuille:1982tx} e.g. rotating away from near-perfect matches \cite{Gardony:2013gn}
 
@@
 a single trajectory of rotation
+
+\subsection{A rational analysis of mental rotation}
 
@@
 
+here we use simple 2d models as a starting point, but the general case
+of solving this problem is very difficult, \TODO{cite shape matching
+  literature}
+
 \subsection{Computational-level analysis}
@@
 Because we ultimately want to determine which hypothesis is more
-likely, the quantity of interest is a ratio:
-\begin{equation*}
-  \ell := \frac{\int p(X_b\ \vert\ X_a, \theta, \hi)p(X_a)p(\hi)p(\theta)\dif\theta}{\int p(X_b\ \vert\ X_a, \theta, \hf)p(X_a)p(\hf)p(\theta)\dif\theta},
-\end{equation*}
-which (assuming the two hypotheses are equally likely \textit{a
-  priori}, and that all rotations are equally likely) simplifies to:
+likely, the quantity of interest is a ratio $\ell:=p(h=0\ \vert\ X_a,
+X_b) / p(h=1\ \vert\ X_a, X_b)$ which (assuming the two hypotheses are
+equally likely \textit{a priori}, and that all rotations are equally
+likely) simplifies to:
 \begin{equation}
@@
 performs rotations in the shortest direction until it reaches the
-correct rotation.
+point where the shapes are aligned.
 
@@
 Unfortunately, it is not always clear what constitutes a good enough
-``match''. Many shapes may have multimodal similarity functions
-\TODO{refer to figure}, and the maximum possible value of $S$ is
-unlikely to be known \textit{a priori} \TODO{cite something about how
-  shape matching is hard}. In this case, we must weaken our stopping
-criteria: rather than searching for a value above a threshold, we
-search for \textit{any} local maximum, as in the Hill Climbing search
-algorithm (\Hc{}). Similar to the \Th{} model, we use the (local) MLE
-estimate of $\theta$, as in Eq.~\ref{eq:mle-lh-ratio}.
+``match''. In the current formulation of the problem, we know the
+exact geometry of the shapes, and know that a linear transformation
+exists which will perfectly align the shapes. However, more generally,
+shapes could be more complex and/or three-dimensional, be subject to
+perceptual uncertainty, or even be not exactly identical. In these
+cases, the problem of choosing a threshold becomes much harder: it is
+not clear how to choose a reasonable threshold \textit{a priori}, as
+the global optimum will depend on many factors. Indeed, shape matching
+is a non-trivial problem in computer vision, requiring sophisticated
+algorithms to achieve reasonable performance
+\cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}.
+
+Thus, the \Th{} model will not be robust beyond this domain. One
+obvious alternative would be to perform an exhaustive search or use a
+global optimisation strategy. However, this would not result in the
+linear response time found by \citeA{Shepard1971}. A second
+alternative (which would potentially demonstrate the linear response
+time result) could instead just perform a Hill Climbing (\Hc{}) search
+for a \textit{local} maximum. Similar to the \Th{} model, we use the
+(local) MLE estimate of $\theta$, as in Eq.~\ref{eq:mle-lh-ratio}.
 
@@
 picking locations to sample which will most improve accuracy of the
-estimate. Rather than a point estimate of the ratio, or even a simple
+estimate. \TODO{Really need to motivate this better: why do we need
+  something so complex? Talk about how it's robust to the shape of the
+  function, the scale, noise, etc. It's complicated, but it should be
+  able to handle these things, so let's try it and see if it works at
+  all. If it does, we can work towards somethign more succinct in the
+  future.}
+
+Rather than a point estimate of the ratio, or even a simple
 Monte Carlo estimate, we can infer a distribution by placing priors
@@
 \begin{equation}
-p(\ell)\approx\frac{\mathcal{N}(\mathbb{E}[Z_0], \mathbb{V}(Z_0))}{\mathcal{N}(\mathbb{E}[Z_1], \mathbb{V}(Z_1))}.
+p(\ell)=p\left(\frac{Z_0}{Z_1}\right)\approx\frac{\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0], \mathbb{V}(Z_0))}{\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1], \mathbb{V}(Z_1))}.
 \end{equation}
@@
 
-\subsection{Design}
+\subsection{Participants and Design}
 
@@
 
-\section{Analysis}
+\section{Results}
+
+\subsection{General analysis}
 
@@
 
-\section{Results}
-
 \begin{figure}[t]
@@
 
-\subsubsection{Response times}
+\subsection{Response times}
 
@@
 and the average response time for an individual stimulus ranged from
-\ExpTimeMin{} to \ExpTimeMax{} seconds. The average accuracy across
-all stimuli was \ExpAccuracy{}, though there were \ExpNumChance{}
-individual stimuli (out of 720) for which people were not above
-chance.
-
-As expected, the minimum angle of rotation was significanty
-monotonically correlated with average per-stimulus response times,
-both for flipped pairs (\ExpThetaTimeCorrFlipped{}) and same pairs
+\ExpTimeMin{} to \ExpTimeMax{} seconds.  As expected, the minimum
+angle of rotation was significanty monotonically correlated with
+average per-stimulus response times, both for flipped pairs
+(\ExpThetaTimeCorrFlipped{}) and same pairs
 (\ExpThetaTimeCorrSame{}). This was also the case for the number of
-actions taken by the \Oc{} model (\OcThetaTimeCorr{}), the \Th{}
-model (\ThThetaTimeCorr{}), and the \Bq{} model
-(\BqThetaTimeCorr{}), but not the \Hc{} model
+actions taken by the \Oc{} model (\OcThetaTimeCorr{}), the \Th{} model
+(\ThThetaTimeCorr{}), and the \Bq{} model (\BqThetaTimeCorr{}), but
+not the \Hc{} model
 (\HcThetaTimeCorr{}). Fig.~\ref{fig:response-times}a
@@
 
-\subsubsection{Accuracy}
+\subsection{Accuracy}
 
-Of the models, the \Oc{} and \Th{} models peformed perfectly,
-with 100\% accuracy. The \Bq{} model was very accurate
-overall (\BqAccuracy{}), though there were \BqNumChance{} stimuli for
-which it was not above chance. The \Hc{} model was not
-particularly accurate, and although it did perform above chance
-overall (\HcAccuracy{}), there were \HcNumChance{} individual stimuli
-for which it was not above chance.
+The average accuracy across all stimuli was \ExpAccuracy{}, though
+there were \ExpNumChance{} individual stimuli (out of 720) for which
+people were not above chance. Of the models, the \Oc{} and \Th{}
+models peformed perfectly, with 100\% accuracy. The \Bq{} model was
+very accurate overall (\BqAccuracy{}), though there were
+\BqNumChance{} stimuli for which it was not above chance. The \Hc{}
+model was not particularly accurate, and although it did perform above
+chance overall (\HcAccuracy{}), there were \HcNumChance{} individual
+stimuli for which it was not above chance.
 
@@
 
-\subsubsection{Practice effects}
+\subsection{Practice effects}
 

### 2014-01-28 analysis/analyses/accuracy_corrs.py
Use pearson correlations for human vs model correlations

@@ -27,12 +27,12 @@ def run(data, results_path, seed):
             corr = dict(util.bootcorr(
                 accuracy_means['exp'],
                 accuracy_means[key],
-                method='spearman'))
+                method='pearson'))
 
-            print "exp v. %s: %s" % (key, util.report_spearman.format(**corr))
+            print "exp v. %s: %s" % (key, util.report_pearson.format(**corr))
             cmd = util.newcommand(
                 "Exp%sAccuracyCorr" % key.capitalize(),
-                util.latex_spearman.format(**corr))
+                util.latex_pearson.format(**corr))
             fh.write(cmd)
 
     return pth

### 2014-01-28 analysis/analyses/response_time_corrs.py

@@ -28,12 +28,12 @@ def run(data, results_path, seed):
             corr = dict(util.bootcorr(
                 response_means['exp'],
                 response_means[key],
-                method='spearman'))
+                method='pearson'))
 
-            print "exp v. %s: %s" % (key, util.report_spearman.format(**corr))
+            print "exp v. %s: %s" % (key, util.report_pearson.format(**corr))
             cmd = util.newcommand(
                 "Exp%sTimeCorr" % key.capitalize(),
-                util.latex_spearman.format(**corr))
+                util.latex_pearson.format(**corr))
             fh.write(cmd)
 
     return pth

### 2014-01-29 man/cogsci-2014/cogsci-2014.tex
More work on introduction/background/modeling

@@
 
-\section{Background}
-
 In the classic experiment by \citeA{Shepard1971}, participants viewed
@@
 
-Previous models of mental rotation have, for the most part, not
-addressed these questions. \citeA{Kosslyn:1977tv} focus primarily on
-simulating the visual simulation buffer rather than the decision of
-\textit{what} should be simulated. Similarly, \citeA{Glasgow:1992tj}
-are more concerned with the representation of imagery than the
-process. Although \citeA{Anderson1978} emphasizes the importance of
-considering both representation and process, he dismisses the problem
-of determining the direction and axis of rotation as merely a
-``technical difficulty''.
+In this paper, we attempt to address these questions through a
+rational analysis \cite{Marr:1983to,anderson90,Shepard:1987tt} of four
+models of mental rotation. First, we discuss the previous literature
+on mental imagery. Second, we outline computational- and
+algorithmic-level analyses of the problem of mental rotation. Then, we
+describe a behavioral experiment based on the classic mental rotation
+studies \cite<e.g.,>{Cooper:1975wp}, and compare the results of the
+experiment with each of the models. We conclude with a discussion of
+the strengths and weaknesses of each model, and lay out directions for
+future work.
+
+\section{Background}
+
+Previous models of mental rotation have, for the most part, focused
+mostly on the representation of mental images, rather than how people
+decide \textit{which} mental images to
+simulate. \citeA{Kosslyn:1977tv} propose a model of mental imagery's
+visual buffer, but not the planning process of how it is used;
+similarly, \citeA{Glasgow:1992tj} are mostly concerned with the
+knowledge representation of imagery. Although \citeA{Anderson1978}
+emphasizes the importance of considering both representation and
+process, he dismisses the problem of determining the direction and
+axis of rotation as merely a ``technical difficulty''.
 
 The only model (of which the authors are aware) that has more
-seriously attempted to address these concerns is the parallel-process
-model of mental rotation by \citeA{Funt:1983wn}. He determines the
-direction and axis of rotation using the method-of-moments method;
-once this target rotation computed, his model rotates one object until
-the target rotation is reached, and then compares the two to see if
-they are congruent.
+seriously attempted to address concerns regarding which simulation to
+run is the parallel-process model of mental rotation by
+\citeA{Funt:1983wn}. He determines the direction and axis of rotation
+using the method-of-moments method; once this target rotation
+computed, his model rotates one object until the target rotation is
+reached, and then compares the two to see if they are
+congruent. \TODO{?}
 
-More recently, evidence has come to light that people \textit{do not}
-actually rotate until the objects are congruent. Eye-tracking studies
-suggest people pay attention to certain salient features of the
-object, rather than a holistic representation \cite{Just1976}, and 
-\citeA{Yuille:1982tx} e.g. rotating away from near-perfect matches \cite{Gardony:2013gn}
+However, there is evidence that people do not actually rotate until
+the objects are congruent, but rather pay attention to certain salient
+features of the object, rather than a holistic representation
+\cite{Just1976,Yuille:1982tx}. More recent evidence shows that when
+people perform physical rotations, they do not rotate until congruency
+is reached and may even rotate \textit{away} from near perfect matches
+\cite{Gardony:2013gn}. Moreover, the state-of-the-art in computer
+vision suggests that there is more to shape matching than simply
+checking for congruency, particularly when shapes are complex or when
+the shapes might not be exactly the same
+\cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}.
 
@@
 \begin{equation}
-  p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta, h)p(X_a)p(h)p(\theta)\dif\theta.
+  p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta, h)p(X_a)p(h)p(\theta)\dif\theta,
   \label{eq:posterior}
 \end{equation}
-Because we ultimately want to determine which hypothesis is more
-likely, the quantity of interest is a ratio $\ell:=p(h=0\ \vert\ X_a,
-X_b) / p(h=1\ \vert\ X_a, X_b)$ which (assuming the two hypotheses are
-equally likely \textit{a priori}, and that all rotations are equally
-likely) simplifies to:
+where $p(X_b\ \vert\ X_a, \theta, h)$ is the probability that $X_b$
+was generated from $X_a$, and $p(X_a)$ is the probability of observing
+$X_a$. Because we ultimately want to determine which hypothesis is
+more likely, the quantity of interest is a ratio $\ell:=p(h=0\ \vert\
+X_a, X_b) / p(h=1\ \vert\ X_a, X_b)$ which (assuming that all
+rotations are equally likely) simplifies to:
 \begin{equation}
-  \ell = \frac{\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta}{\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta}.
+  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)p(h=0)}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)p(h=1)}.
   \label{eq:lh-ratio}
@@
 
-If we represent a shape of $N$ vertices with a $N\times 2$ coordinate
-matrix $\mathbf{X}=[\mathbf{x}_1, \ldots{}, \mathbf{x}_N]$, then the
-transformation $f$ is $f(\mathbf{X}, h,
-\theta):=\mathbf{X}\F_h^T\R^T$, where $\R$ is a rotation matrix,
-$\F_0$ is the identity matrix $\I$, and $\F_1$ is a reflection matrix
-across the $y$-axis. Assuming no computational constraints, the
-simplest solution is to compute the left inverse of $\Xa\F_h^T$ and,
-for each $h$, check whether $(\Xa
-\F_h^T)_\mathrm{left}^{-1}\cdot{}\Xb$ is a valid rotation matrix.
-
-However, we assume that the ability to compute left inverses is not
-available. Instead, the observed shapes may only be transformed by a
-small amount at a time, and each transformation is costly in terms of
-computational resources. The goal, then is to estimate the integrals
-in Eq.~\ref{eq:lh-ratio} by evaluating $p(\Xb\ \vert\ \Xa, \theta, h)$
-as few times as possible.
+We represent a shape of $N$ vertices with a $N\times 2$ coordinate
+matrix $\mathbf{X}=[\mathbf{x}_1, \ldots{}, \mathbf{x}_N]$, and denote
+the rotation and/or reflection transformation as $f(\mathbf{X}, h,
+\theta):=\mathbf{X}\F_h^T\R^T$, where $\R$ is a rotation matrix, and
+$\F_h$ is either the identity matrix $\I$ (when $h=0$) or a reflection
+matrix across the $y$-axis (when $h=1$).  We assume that the observed
+shapes may only be transformed by a small amount at a time, and each
+transformation is costly in terms of computational resources. The
+goal, then is to estimate the integrals in Eq.~\ref{eq:lh-ratio} by
+evaluating $p(\Xb\ \vert\ \Xa, \theta, h)$ as few times as possible.
 
@@
 
+To summarize, we approximate the likelihood term of
+Eq.~\ref{eq:lh-ratio} using the similarity function defined in
+Eq.~\ref{eq:similarity}. Because we assume analog rotations, this
+similarity can only be computed at sequential locations, i.e. the
+actions listed in Eq.~\ref{eq:actions}.
+
+\subsection{Specific models}
+
+The last decision we need to make is which algorithm to use to
+approximate Eq.~\ref{eq:lh-ratio} using samples of the similarity
+function. The algorithm must do two things: first, it must choose
+\textit{which} places to sample, and second, it must decide when to
+stop sampling.
+
 \subsubsection{\Oc{} model}
 
-The previous hypothesis in the literature is that people perform a
-mental rotation in the shortest direction until a match is found, and
-that determining this direction is an implementation detail \TODO{cite
-  anderson?}. To reflect this hypothesis, we created an ``oracle''
-model which knows the correct rotation \textit{a priori}, and simply
-performs rotations in the shortest direction until it reaches the
-point where the shapes are aligned.
+One hypothesis in the literature is that people perform a mental
+rotation in the shortest direction until a match is found
+\cite{Shepard1971, Cooper:1975wp}, and that determining this direction
+is an implementation detail \cite{Anderson1978}, or that it is
+computed beforehand \cite{Funt:1983wn}.  To reflect this hypothesis,
+we created an ``oracle'' model which first computes the correct
+rotation and then simply takes actions in the shortest direction until
+it reaches the point where the shapes are aligned.
+
+To determine the correct rotation, we solve for the rotation matrix by
+computing $(\Xa \F_h^T)_\mathrm{left}^{-1}\cdot{}\Xb$, where
+$(\Xa\F_h^T)_\mathrm{left}^{-1}$ is the left inverse of
+$\Xa\F_h^T$. For each $h$, we then check whether the resulting matrix
+is a valid rotation matrix.
 
@@
 
-An intuitive and simple model which does not assume the direction is
-known might first rotate in the direction that increases similarity,
-and then stop rotating once a ``match'' is found, as determined by a
-threshold on the value of $S$. If all rotations have been exhausted,
-then flip, and try rotating again.
+An intuitive and simple model which does not know which way to rotate
+\textit{a priori} might first rotate in the direction that increases
+similarity, and then stop rotating once a ``match'' is found, as
+determined by a threshold on the value of $S$. If all rotations have
+been exhausted, then flip, and try rotating again.
+
+We note, however, that it is not always clear what constitutes a good
+enough ``match''. In the current formulation of the problem, we know
+the exact geometry of the shapes, and know that a linear
+transformation exists which will perfectly align the shapes. However,
+shapes could be more complex and/or three-dimensional, be subject to
+perceptual uncertainty, or even be not exactly identical. In these
+cases, it is not clear how to choose a reasonable threshold \textit{a
+  priori}, as the global optimum will depend on many factors. Thus, it
+is unlikely that the \Th{} model will not be robust beyond this
+domain.
 
@@
 
-Unfortunately, it is not always clear what constitutes a good enough
-``match''. In the current formulation of the problem, we know the
-exact geometry of the shapes, and know that a linear transformation
-exists which will perfectly align the shapes. However, more generally,
-shapes could be more complex and/or three-dimensional, be subject to
-perceptual uncertainty, or even be not exactly identical. In these
-cases, the problem of choosing a threshold becomes much harder: it is
-not clear how to choose a reasonable threshold \textit{a priori}, as
-the global optimum will depend on many factors. Indeed, shape matching
-is a non-trivial problem in computer vision, requiring sophisticated
-algorithms to achieve reasonable performance
-\cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}.
-
-Thus, the \Th{} model will not be robust beyond this domain. One
-obvious alternative would be to perform an exhaustive search or use a
-global optimisation strategy. However, this would not result in the
-linear response time found by \citeA{Shepard1971}. A second
-alternative (which would potentially demonstrate the linear response
-time result) could instead just perform a Hill Climbing (\Hc{}) search
-for a \textit{local} maximum. Similar to the \Th{} model, we use the
-(local) MLE estimate of $\theta$, as in Eq.~\ref{eq:mle-lh-ratio}.
+One way to deal with the problem of choosing a threshold would simply
+be to perform an exhaustive search or use a global optimisation
+strategy. However, these strategies would not result in the linear
+response time found by \citeA{Shepard1971}. A second alternative,
+which could potentially demonstrate the linear response time result,
+might instead just perform a Hill Climbing (\Hc{}) search for a
+\textit{local} maximum. However, it is not necessarily the case that
+all local maxima for the correct hypothesis will be higher than all
+the local maxima for the incorrect hypothesis: thus, we would expect
+the \Hc{} model to have difficulties reliably choosing the correct
+hypothesis.
 
@@
 
-An alternative to the previous models is not to perform a search for
-the MLE, but to approximate the integrals in Eq.~\ref{eq:lh-ratio} by
-picking locations to sample which will most improve accuracy of the
-estimate. \TODO{Really need to motivate this better: why do we need
-  something so complex? Talk about how it's robust to the shape of the
-  function, the scale, noise, etc. It's complicated, but it should be
-  able to handle these things, so let's try it and see if it works at
-  all. If it does, we can work towards somethign more succinct in the
-  future.}
+While the previous few models have all focused on \textit{searching}
+for the global maximum, the goal is really to just approximate
+Eq.~\ref{eq:lh-ratio} while still obeying the constraint of sequential
+rotations. Instead of perfoming a search, we could try maintaining a
+probability distribution over our \textit{estimate}, and then choose
+actions which are expected to improve the estimate of
+Eq.~\ref{eq:lh-ratio}. This strategy has several obvious
+benefits. First, it does not make assumptions about the shape or scale
+of the similarity function. Second, by choosing to sample places which
+are informative, this method implicitly minimizes the amount of
+rotation, but because it is constrained to sequential rotations,
+should still exhibit the linear effect of the angle.
 
@@
 We ran model simulations for the same stimuli that participants
-viewed. We ran 10 samples of each stimulus under the \Oc{} model, and
-100 samples of each stimulus for the \Th{}, \Hc{}, and
-\Bq{} models, with the exception of stimuli where
-$\theta=0$ or $\theta=180$, in which case we ran double the number of
-samples.
+viewed. We ran 10 samples of each stimulus under each of the modesl,
+with the exception of stimuli where $\theta=0$ or $\theta=180$, in
+which case we ran double the number of samples.
 

### 2014-01-29 bin/sims/amazon.py
Checkout git branch when updating repo in amazon.py

@@ -358,7 +358,7 @@ def update_git(args):
     instances = get_instances(conn, active_only=True)
     for inst in instances:
         print "Updating git repo on instance '%s'..." % inst.id
-        cmd = "cd project/optimal-mental-rotation && git pull"
+        cmd = "cd project/optimal-mental-rotation && git checkout cogsci-2014 && git pull"
         ssh_call(zone, inst, cmd, False)
 
 

### 2014-01-29 man/cogsci-2014/cogsci-2014.tex
More work on modeling section

@@
 
-\section{Models of mental rotation}
+\section{Modeling mental rotation}
 
@@
 generated by a transformation of $X_a$, i.e. $X_b=f(X_a, \theta, h)$,
-where $\theta$ is a rotation, $h=0$ is the hypothesis that the images
-depict the same object, and $h=1$ is the hypothesis that the images
+where $\theta$ is a rotation, $\hi$ is the hypothesis that the images
+depict the same object, and $\hf$ is the hypothesis that the images
 depict mirror-image objects. The posterior probability of each
@@
 $X_a$. Because we ultimately want to determine which hypothesis is
-more likely, the quantity of interest is a ratio $\ell:=p(h=0\ \vert\
-X_a, X_b) / p(h=1\ \vert\ X_a, X_b)$ which (assuming that all
+more likely, the quantity of interest is a ratio $\ell:=p(\hi\ \vert\
+X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which (assuming that all
 rotations are equally likely) simplifies to:
 \begin{equation}
-  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)p(h=0)}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)p(h=1)}.
+  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)p(\hi)}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)p(\hf)}.
   \label{eq:lh-ratio}
@@
 \theta):=\mathbf{X}\F_h^T\R^T$, where $\R$ is a rotation matrix, and
-$\F_h$ is either the identity matrix $\I$ (when $h=0$) or a reflection
-matrix across the $y$-axis (when $h=1$).  We assume that the observed
-shapes may only be transformed by a small amount at a time, and each
-transformation is costly in terms of computational resources. The
-goal, then is to estimate the integrals in Eq.~\ref{eq:lh-ratio} by
-evaluating $p(\Xb\ \vert\ \Xa, \theta, h)$ as few times as possible.
+$\F_h$ is either the identity matrix $\I$ (when $\hi$) or a reflection
+matrix across the $y$-axis (when $\hf$).
 
@@
 \begin{equation}
-  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{|\mathbb{M}|} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma^2).
+  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{|\mathbb{M}|} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
   \label{eq:similarity}
 \end{equation}
+where $\sigma_S=0.15$ is the standard deviation of the similarity.
 
-Additionally, we have the constraint of small
-transformations. Specifically, if the current mental image is $\Xt$,
-then:
+Additionally, we assume that the observed shapes may only be
+transformed by a small amount at a time, and each transformation is
+costly in terms of computational resources. Specifically, if the
+current mental image is $\Xt$, then:
 \begin{equation}
   \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
-      f(\Xt, 0, \epsilon) &\mbox{ rotate right by $\epsilon$,} \\
-      f(\Xt, 0, -\epsilon) &\mbox{ rotate left by $\epsilon$,} \\
+      f(\Xt, 0, |\epsilon|) &\mbox{ rotate by $\epsilon$,} \\
       f(\Xt, 1, 0) &\mbox{ flip,} \\
       f(\Xa, 0, 0) &\mbox{ reset, or} \\
-      f(\Xa, 1, 0) &\mbox{ reset and flip.} \\
+      f(\Xa, 1, 0) &\mbox{ reset and flip,} \\
     \end{array} \right.
@@
 \end{equation}
+where $\epsilon\sim \mathcal{N}(0, \sigma_\epsilon)$ and
+$\sigma_\epsilon=0.7$ is the standard deviation of the step size in
+radians.
 
@@
 Eq.~\ref{eq:lh-ratio} using the similarity function defined in
-Eq.~\ref{eq:similarity}. Because we assume analog rotations, this
-similarity can only be computed at sequential locations, i.e. the
+Eq.~\ref{eq:similarity}. Because we assume mental rotations are
+performed sequentially, this similarity can only be computed for the
 actions listed in Eq.~\ref{eq:actions}.
 
-\subsection{Specific models}
+\subsection{Specific models of mental rotation}
 
-The last decision we need to make is which algorithm to use to
-approximate Eq.~\ref{eq:lh-ratio} using samples of the similarity
-function. The algorithm must do two things: first, it must choose
-\textit{which} places to sample, and second, it must decide when to
-stop sampling.
+In order to approximate Eq.~\ref{eq:lh-ratio} using samples of the
+similarity function, we must decide \textit{which} places to sample
+and \textit{when} stop sampling. The models below differ in how they
+make these decisions.
 
-\subsubsection{\Oc{} model}
+\subsubsection{Oracle model}
 
@@
 $(\Xa\F_h^T)_\mathrm{left}^{-1}$ is the left inverse of
-$\Xa\F_h^T$. For each $h$, we then check whether the resulting matrix
-is a valid rotation matrix.
+$\Xa\F_h^T$. We then check each $h$ to see if computation produces a
+valid rotation matrix; the $h$ that does produce a rotation matrix is
+the correct hypothesis. This approach gives us the true value of
+$\theta$, so Eq.~\ref{eq:lh-ratio} becomes a likelihood ratio test,
+where $\theta$ is set to the MLE value, rather than being
+marginalized:
+\begin{equation}
+  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)p(\hi)}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)p(\hf)}.
+  \label{eq:mle-lh-ratio}
+\end{equation}
 
-\subsubsection{\Th{} model}
+\subsubsection{Threshold model}
 
@@
 determined by a threshold on the value of $S$. If all rotations have
-been exhausted, then flip, and try rotating again.
+been exhausted, then flip, and try rotating again. We assume the value
+of $S$ which is greater than the threshold corresponds to the true
+$\theta$, or a point near the true $\theta$. So, as with the \Oc{}
+model, Eq.~\ref{eq:lh-ratio} becomes a likelihood ratio.
 
-We note, however, that it is not always clear what constitutes a good
-enough ``match''. In the current formulation of the problem, we know
-the exact geometry of the shapes, and know that a linear
-transformation exists which will perfectly align the shapes. However,
-shapes could be more complex and/or three-dimensional, be subject to
-perceptual uncertainty, or even be not exactly identical. In these
-cases, it is not clear how to choose a reasonable threshold \textit{a
-  priori}, as the global optimum will depend on many factors. Thus, it
-is unlikely that the \Th{} model will not be robust beyond this
-domain.
-
-The consequence of doing a search for a match is that
-Eq.~\ref{eq:lh-ratio} becomes a likelihood ratio test, where $\theta$
-is set to the MLE value, rather than being marginalized:
-\begin{equation}
-  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)}.
-  \label{eq:mle-lh-ratio}
-\end{equation}
+We note that in the current formulation of the problem, choosing the
+threshold is straightforward because we know both the exact geometry
+of the shapes and that a linear transformation exists which will
+perfectly align them. However, it is not always clear how to choose a
+reasonable threshold \textit{a priori}, as the global optimum will
+depend on many factors (such as shape complexity, dimensionality,
+perceptual uncertainty, and whether the shapes are exactly
+identical). Thus, it is unlikely that the \Th{} model will be robust
+beyond this domain.
 
@@
 might instead just perform a Hill Climbing (\Hc{}) search for a
-\textit{local} maximum. However, it is not necessarily the case that
-all local maxima for the correct hypothesis will be higher than all
-the local maxima for the incorrect hypothesis: thus, we would expect
-the \Hc{} model to have difficulties reliably choosing the correct
-hypothesis.
+\textit{local} maximum. As with the \Oc{} and \Th{} models,
+Eq.~\ref{eq:lh-ratio} becomes a maximum likelihood ratio. Unlike the
+global maximum, however, it is not guaranteed that a local maximum of
+the correct hypothesis will be greater than a local maximum of the
+incorrect hypothesis.
 
@@
 
-Rather than a point estimate of the ratio, or even a simple
-Monte Carlo estimate, we can infer a distribution by placing priors
-over the similarity functions themselves:
+Formally, we denote $Z_h$ as our estimate of the likelihood for
+hypothesis $h$, and write its distribution as:
 \begin{equation}
-\ell = \frac{\int \left[\int S(\Xb, f(\Xa, \theta, \hi))p(\theta)\dif\theta\right] p(S)\dif S}{\int \left[\int S(\Xb, f(\Xa, \theta, \hf))p(\theta)\dif\theta\right] p(S)\dif S}
+  p(Z_h) = \int \left[\int S(\Xb, f(\Xa, \theta, h))p(\theta)\dif\theta\right] p(S)\dif S.
 \end{equation}
@@
 \cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}. Specifically, we
-use Gaussian Process (GP) priors, as in \citeA{Osborne:2012tm}
-\TODO{probably need to describe better what a GP is}. Because our
-domain is on a circle, we utilize a periodic kernel for the covariance
+use a Gaussian Process (GP) prior for $p(S)$, as in
+\citeA{Osborne:2012tm}.\footnote{Because our domain is on a circle, we
+  utilize a periodic kernel for the covariance
   \cite{Rasmussen:2006vz}, i.e. $k(\theta,
-\theta^\prime)=h^2\exp[-\frac{2}{w^2}\sin^2(\frac{1}{2}(\theta-\theta^\prime))]$.
-
+  \theta^\prime)=h^2\exp[-\frac{2}{w^2}\sin^2(\frac{1}{2}(\theta-\theta^\prime))]$.}
 Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$, we first place a GP prior on
-$\log(S_h)$. Because GPs have no guarantee of non-negativity, placing
-the prior over the log enforces positivity after it is exponentiated:
+the log of $S_h$ in order to enforce positivity after it is
+exponentiated:
 \begin{equation}
@@
 integral is still intractable. To approximate it, we fit a second GP
-over points sampled from the log-GP.\footnote{Note that the approach
-  used here differs slightly from the one presented in
-  \citeA{Osborne:2012tm} and is based on personal communication with
-  one of the authors \cite{Duvenaud:2013td}.}We will denote these
-points as $\bar{S}_h:=\exp(\mu_h)$. Then, from
-\citeA{Duvenaud:2013td}:
+over points sampled from the log-GP, which we denote as
+$\bar{S}_h:=\exp(\mu_h)$.\footnote{Note that the approach used here
+  differs slightly from the one presented in \citeA{Osborne:2012tm}
+  and is based on personal communication with one of the authors
+  \cite{Duvenaud:2013td}.} Then, from \citeA{Duvenaud:2013td}:
 \begin{align}
@@
 
-Now that we have a distribution over $Z_h$, we can additionally
-compute which points which we would expect to strengthen our estimate
-(i.e., decrease the variance). From \citeA{Osborne:2012tm}, the
-expected variance of $Z_h$ if we sampled a new observation at
-$\theta_a$ is:
+Assuming independence, we can now write $p(Z_h)\approx\mathcal{N}(Z_h\
+\vert\ \mathbb{E}[Z_h], \mathbb{V}(Z_h))$, which gives us a
+distribution over the likelihood ratio in Eq.~\ref{eq:lh-ratio}:
+\begin{equation}
+p(\ell)\approx\frac{\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0], \mathbb{V}(Z_0))\ p(\hi)}{\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1], \mathbb{V}(Z_1))\ p(\hf)}.
+\end{equation}
+This distribution cannot easily be calculated; however, we really are
+just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. Thus, rather than
+computing $p(\ell)$, we use $Z_D=Z_0-Z_1$:
+\begin{equation}
+  p(Z_D)\propto\mathcal{N}(\mathbb{E}[Z_0] - \mathbb{E}[Z_1], \mathbb{V}(Z_0) + \mathbb{V}(Z_1)).
+\end{equation}
+We then sample new observations until we are at least 95\% confident
+that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
+$\hi$, and when $p(Z_D<0)>0.975$, we accept $\hf$. To determine which
+points to sample, we can compute the expected variance of $Z_h$ if we
+sampled a new observation at $\theta_a$. From \citeA{Osborne:2012tm}:
 \begin{align}
@@
 \end{align}
-
-Revisiting our likelihood ratio, we now have:
-\begin{equation}
-p(\ell)=p\left(\frac{Z_0}{Z_1}\right)\approx\frac{\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0], \mathbb{V}(Z_0))}{\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1], \mathbb{V}(Z_1))}.
-\end{equation}
-This distribution does not have a nice form. However, we really are
-just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$: thus, rather than
-computing $p(\ell)$, we use $Z_D=Z_0-Z_1$. Assuming $Z_0$ and $Z_1$
-are independent,\footnote{This assumption is probably not
-  correct. \TODO{explanation?}} we have:
-\begin{equation}
-  p(Z_D)\propto\mathcal{N}(\mathbb{E}[Z_0] - \mathbb{E}[Z_1], \mathbb{V}(Z_0) + \mathbb{V}(Z_1)).
-\end{equation}
-We then sample new observations until we are at least 95\% confident
-that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
-$h=0$, and when $p(Z_D<0)>0.975$, we accept $h=1$.
+We compute this expected variance for each of the actions in
+Eq.~\ref{eq:actions}, and then choose the action with the lowest
+value.
 

### 2014-01-29 man/cogsci-2014/cogsci-2014.tex
Work on intro/background, formatting references

@@
 
-\title{Modeling mental rotation}
+\title{Modeling mental simulation: a computational analysis of mental rotation}
 
@@
 
-In this paper, we attempt to address these questions through a
-rational analysis \cite{Marr:1983to,anderson90,Shepard:1987tt} of four
-models of mental rotation. First, we discuss the previous literature
-on mental imagery. Second, we outline computational- and
+In this paper, we explore these questions through rational analysis
+\cite{Marr:1983to,anderson90,Shepard:1987tt} and compare four models
+of mental rotation. First, we discuss the previous literature on
+mental imagery. Second, we outline computational- and
 algorithmic-level analyses of the problem of mental rotation. Then, we
@@
 \section{Background}
+
+\TODO{say something here}
+
+\subsection{Mechanisms behind mental rotation}
 
@@
 computed, his model rotates one object until the target rotation is
-reached, and then compares the two to see if they are
-congruent. \TODO{?}
+reached, and then compares the two to see if they are congruent. This
+is a somewhat unsatisfying explanation, however: why would it be
+necessary to perform the rotation if the answer is already known?
 
-However, there is evidence that people do not actually rotate until
-the objects are congruent, but rather pay attention to certain salient
-features of the object, rather than a holistic representation
-\cite{Just1976,Yuille:1982tx}. More recent evidence shows that when
-people perform physical rotations, they do not rotate until congruency
-is reached and may even rotate \textit{away} from near perfect matches
-\cite{Gardony:2013gn}. Moreover, the state-of-the-art in computer
-vision suggests that there is more to shape matching than simply
-checking for congruency, particularly when shapes are complex or when
-the shapes might not be exactly the same
+Perhaps people do not engage in a single trajectory of rotation that
+ends with congruency. Some evidence suggests that people do not
+perform a holistic rotation until the objects are congruent, but
+rather pay attention to certain salient features of the object
+\cite{Just1976,Yuille:1982tx}. A more recent study shows that when
+performing \textit{physical} rotations, people do not rotate until
+congruency is reached, and may even rotate \textit{away} from near
+perfect matches \cite{Gardony:2013gn}. Moreover, the state-of-the-art
+in computer vision suggests that there is more to shape matching than
+simply checking for congruency, particularly when shapes are complex
+or when the shapes might not be exactly the same
 \cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}.
-
-so what are they actually doing?
-
-perhaps they are performing mental manipulations of an image, but not
-a single trajectory of rotation
 
@@
 
-rational analysis \cite{Marr:1983to,anderson90,Shepard:1987tt}
-
-1) the problem: determine something about the spatial transformations
-that an object has undergone
-2) the goal: make this determiniation while using the least amount of
-resources
-3) the solution: perform those rotations and reflections which will
-give the most information (?) about the answer until an answer is
-found
+It may be useful to take a step back and try to analyze mental
+rotation from using rational analysis
+\cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
+level, we could say that the \textit{problem} is to determine
+something about the spatial transformations that an object has
+undergone. At the algorithmic level, we are constrained by the notion
+that mental images must be transformed in an analog manner (or at
+least in a way that is approximately analog), and that mental images
+are time-consuming and effortful to generate. Thus, the \textit{goal}
+is to make this determiniation while using the least amount of
+resources. Due to the sequential constratiant, the minimum amount of
+computation will coincide with the minimum angle of rotation. So, the
+rational \textit{solution} will be something that only rotates through
+the minimum angle.
 
@@
 
-we focus less on the representation, and more on the process, assuming
-a spatial representation (because you need both to make a claim,
-\cite{Anderson1978})
-
 problem solving \cite{Hegarty2004, Schwartz1999}
@@
 symbolic vs simulation \cite{Schwartz:1996uy}
-
-eye tracking \cite{Just1976}
 
@@
 where $\epsilon\sim \mathcal{N}(0, \sigma_\epsilon)$ and
-$\sigma_\epsilon=0.7$ is the standard deviation of the step size in
+$\sigma_\epsilon=0.6$ is the standard deviation of the step size in
 radians.
@@
 rotations. Instead of perfoming a search, we could try maintaining a
-probability distribution over our \textit{estimate}, and then choose
-actions which are expected to improve the estimate of
-Eq.~\ref{eq:lh-ratio}. This strategy has several obvious
-benefits. First, it does not make assumptions about the shape or scale
-of the similarity function. Second, by choosing to sample places which
-are informative, this method implicitly minimizes the amount of
-rotation, but because it is constrained to sequential rotations,
-should still exhibit the linear effect of the angle.
+probability distribution over our \textit{estimate} of
+Eq.~\ref{eq:lh-ratio}, and then choose actions which are expected to
+improve the accuracy of the estimate. This strategy has several
+obvious benefits. First, it does not make assumptions about the shape
+or scale of the similarity function. Second, by choosing to sample
+places which are informative, this method implicitly minimizes the
+amount of rotation, but because it is constrained to sequential
+rotations, should still exhibit the linear effect of the angle.
 
@@
 
-\subsection{General analysis}
-
 Out of the 247 participants, 200 (81\%) were included in our
@@
 \end{figure}
+
+\subsection{Practice effects}
+
+There was a significant effect of learning both on response time
+(\ExpTrialTimeCorr{}) and on accuracy (\ExpTrialAccuracyCorr{}),
+though the effect on accuracy was not significant during the second
+half of the experiment (\ExpaTrialAccuracyCorr{} for the first half
+vs. \ExpbTrialAccuracyCorr{} for the second
+half). Fig.~\ref{fig:learning}
 
@@
 similar correlation, with \ExpBqAccuracyCorr{}.
-
-\subsection{Practice effects}
-
-There was a significant effect of learning both on response time
-(\ExpTrialTimeCorr{}) and on accuracy (\ExpTrialAccuracyCorr{}),
-though the effect on accuracy was not significant during the second
-half of the experiment (\ExpaTrialAccuracyCorr{} for the first half
-vs. \ExpbTrialAccuracyCorr{} for the second
-half). Fig.~\ref{fig:learning}
 

### 2014-01-29 lib/mental_rotation/analysis.py
Load prior of 0.55, not 0.6

@@ -77,7 +77,7 @@ def load_all(version, data_path, human=None):
 
     bq = load_model("BayesianQuadratureModel", version, data_path)
     data['bq'] = bq.groupby('prior').get_group(0.5)
-    data['bqp'] = bq.groupby('prior').get_group(0.6)
+    data['bqp'] = bq.groupby('prior').get_group(0.55)
 
     if human is None:
         data.update(load_human(version, data_path)[1])

### 2014-01-29 man/cogsci-2014/cogsci-2014.tex
Small changes in analysis section

@@
 \end{abstract}
-
-\TODO{put correlations in figures}
 
@@
 trials were defined to be those for which the presented shapes
-differed by a rotation of no more than 20 degrees.
+differed by a rotation of no more than 20 degrees. \TODO{mention
+  excluding trials where people took a really long time to respond}
 
@@
 bootstrap samples. Response times were only only considered for
-correct responses.
+correct responses.  We also used a bootstrap analysis (with 10000
+bootstrap samples) to compute the confidence intervals for both
+Spearman ($r_s$) and Pearson ($r$) correlations.
 
-For analyses of accuracy, confidence intervals were
-computed from a binomial proportion with a Jeffrey's beta prior.  For
-correlations, we performed a bootstrap analysis (sampling with
-replacement) over Spearman correlations using 10000 bootstrap samples.
-
-To test if participants or a model was above chance on a particular
-stimulus, we used the same posterior and checked whether $p>0.05$ for
-the 50\% mark.
+For analyses of accuracy, confidence intervals were computed from a
+binomial proportion with a Jeffrey's beta prior.  To test if
+participants or a model was above chance on a particular stimulus, we
+used the same binomial proportion, and simply tested whether $p>0.05$
+for the 50\% mark.
 

### 2014-01-30 man/cogsci-2014/cogsci-2014.tex
Misc edits

@@
 \documentclass[10pt,letterpaper]{article}
+
+%% For better looking math (but it takes up more space):
+% \AtBeginDocument{\RequirePackage{lmodern, times}}
 
@@
 
+\addtolength{\abovecaptionskip}{-0.25cm}
+\addtolength{\belowcaptionskip}{-0.2cm}
+\addtolength{\textfloatsep}{-2mm}	% space below "Figure 1: ..."
+
 \input{../../results/D/response_time.tex}
@@
 
-\title{Modeling mental simulation: a computational analysis of mental rotation}
+\title{What to simulate? Inferring the direction for mental rotation}
 
@@
 
-\TODO{need a figure with an example similarity curve, and example
-  stimuli}
-
-\TODO{fix title}
-
 \section{Introduction}
@@
 representation of mental images \cite<e.g.,>{Kosslyn:2009tj,
-  Pylyshyn:2002vk}. Despite this vast literature, there have been
-surprisingly few computational models of mental imagery.
+  Pylyshyn:2002vk}. But this debate hasn't addressed one of the most
+fundamental questions about mental simulation: how people decide what
+they should simulate.
 
@@
 
+\begin{figure*}[t]
+  \begin{center}
+    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
+    \caption{\textbf{Model traces.} \TODO{}}
+    \label{fig:model-traces}
+  \end{center}
+\end{figure*}
+
 In order to approximate Eq.~\ref{eq:lh-ratio} using samples of the
@@
 
-We randomly generated 20 shapes of 5 or 6 vertices (e.g., $\Xa$ in
-Figure \ref{fig:stimuli}). For each shape, we computed 20 ``same'' and
-20 ``flipped'' stimuli pairs, with $\theta$ spaced at $20^\prime$
-increments between 0 and 360 (with 0 and 180 degree rotations repeated
-twice, in order to gather an equal number of responses for each angle
-between 0 and 180). ``Same'' pairs were created by rotating $\Xa$ by
-$\theta$; ``flipped'' pairs first reflected $\Xa$ across the $y$-axis,
-then rotated by $\theta$.
+We randomly generated 20 shapes of 5 or 6 vertices (e.g., the stimuli
+in Fig.~\ref{fig:model-traces}). For each shape, we computed 20
+``same'' and 20 ``flipped'' stimuli pairs, with $\theta$ spaced at
+$20^\prime$ increments between 0 and 360 (with 0 and 180 degree
+rotations repeated twice, in order to gather an equal number of
+responses for each angle between 0 and 180). ``Same'' pairs were
+created by rotating $\Xa$ by $\theta$; ``flipped'' pairs first
+reflected $\Xa$ across the $y$-axis, then rotated by $\theta$.
 
@@
 
-\begin{figure*}[t]
-  \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
-    \caption{\textbf{Model traces.} \TODO{} \TODO{put pictures of the
-        stimuli here}}
-    \label{fig:model-traces}
-  \end{center}
-\end{figure*}
-
 We replicated the previous monotonic response time result, though not
@@
 
-\section{Acknowledgments}
+%% TODO: Uncomment this for final version:
 
-This research was supported by ONR MURI grant number N00014-13-1-0341,
-and a Berkeley Fellowship awarded to JBH.
+% \section{Acknowledgments}
+% {\small This research was supported by ONR MURI grant number
+% N00014-13-1-0341, and a Berkeley Fellowship awarded to JBH. }
 
 \bibliographystyle{apacite}
-
+\renewcommand{\bibliographytypesize}{\small}
 \setlength{\bibleftmargin}{.125in}
 \setlength{\bibindent}{-\bibleftmargin}
-
 \bibliography{references}

### 2014-01-30 analysis/plots/util.py
Update model example plot

@@ -49,6 +49,25 @@ def clear_top(ax=None):
     ax.xaxis.set_ticks_position('bottom')
 
 
+def clear_top_bottom(ax=None):
+    """Remove the top and bottom edges of the axis bounding box.
+
+    Parameters
+    ----------
+    ax : axis object (default=pyplot.gca())
+
+    References
+    ----------
+    http://matplotlib.org/examples/pylab_examples/spine_placement_demo.html
+
+    """
+    if ax is None:
+        ax = plt.gca()
+    ax.spines['top'].set_color('none')
+    ax.spines['bottom'].set_color('none')
+    ax.xaxis.set_ticks([])
+
+
 def outward_ticks(ax=None, axis='both'):
     """Make axis ticks face outwards rather than inwards (which is the
     default).
@@ -173,3 +192,72 @@ def sync_ylims(*axes):
     for ax in axes:
         ax.set_ylim(ymin, ymax)
     return ymin, ymax
+
+
+def sync_xlims(*axes):
+    """Synchronize the x-axis data limits for multiple axes. Uses the maximum
+    upper limit and minimum lower limit across all given axes.
+
+    Parameters
+    ----------
+    *axes : axis objects
+        List of matplotlib axis objects to format
+
+    Returns
+    -------
+    out : xmin, xmax
+        The computed bounds
+
+    """
+    xmins, xmaxs = zip(*[ax.get_xlim() for ax in axes])
+    xmin = min(xmins)
+    xmax = max(xmaxs)
+    for ax in axes:
+        ax.set_xlim(xmin, xmax)
+    return xmin, xmax
+
+
+def sync_xlabel_coords(axes, y, x=0.5):
+    """Set the y-coordinate (and optionally the x-coordinate) of the x-axis
+    labels.
+
+    Parameters
+    ----------
+    axes : list
+        list of axis objects
+    y : float
+        y-coordinate for the label
+    x : float (default=0.5)
+        x-coordinate for the label
+    ax : axis object (default=pyplot.gca())
+
+    References
+    ----------
+    http://matplotlib.org/faq/howto_faq.html#align-my-ylabels-across-multiple-subplots
+
+    """
+    for ax in axes:
+        ax.xaxis.set_label_coords(x, y)
+
+
+def sync_ylabel_coords(axes, x, y=0.5):
+    """Set the x-coordinate (and optionally the y-coordinate) of the y-axis
+    labels.
+
+    Parameters
+    ----------
+    axes : list
[…]

### 2014-01-30 lib/mental_rotation/model/base_model.py

@@ -190,15 +190,19 @@ class BaseModel(object):
 
         return lines
 
-    def plot_trace(self, ax, legend=True):
+    def plot_trace(self, ax, legend=True, scale_points=True):
         Fi = self.F_i
-        Si = self.log_S_i - self._log_const
-        Si = (Si - Si.min() + 1) * 10
         Ri = self.R_i
         ti = np.arange(Ri.size)
         ci = np.empty((Ri.size, 3))
         ci[Fi == 0] = np.array([1, 0, 0])
         ci[Fi == 1] = np.array([0, 0, 1])
+
+        if scale_points:
+            Si = self.log_S_i - self._log_const
+            Si = (Si - Si.min() + 1) * 10
+        else:
+            Si = np.zeros(Ri.size) + 30
 
         ax.plot(ti, Ri, 'k-')
         ax.scatter(ti, Ri, c=ci, s=Si, edgecolor=ci)

### 2014-01-30 man/cogsci-2014/cogsci-2014.tex
More work on paper

@@
 
-\textbf{Keywords:} 
-\TODO{}
+\textbf{Keywords:} mental rotation, computational model
 \end{abstract}
@@
 result was that people's response times had a strong linear
-correlation with the minimum angle of rotation. The conclusion was
-that people were solving this task by ``mentally rotating'' the
-objects until they were congruent.
-
-However, there are some questions regarding this process of problem
-solving that are not explained by the alignment hypothesis: how do
-people know the axis around which to rotate the objects? If the axis
-is known, how do people know which direction to rotate the objects?
-And finally, how do people know when to stop rotating and give a
-response?
+correlation with the minimum angle of rotation, leading to the
+conclusion that people solve this task by ``mentally rotating'' the
+objects until they are congruent.  This conclusion leaves several
+questions unanswered, however. How do people know the axis around
+which to rotate the objects? If the axis is known, how do people know
+which direction to rotate the objects?  And finally, how do people
+know how long to rotate?
 
@@
 mental imagery. Second, we outline computational- and
-algorithmic-level analyses of the problem of mental rotation. Then, we
-describe a behavioral experiment based on the classic mental rotation
-studies \cite<e.g.,>{Cooper:1975wp}, and compare the results of the
-experiment with each of the models. We conclude with a discussion of
-the strengths and weaknesses of each model, and lay out directions for
-future work.
+algorithmic-level analyses of the problem of mental rotation. Third,
+we describe a behavioral experiment based on the classic mental
+rotation studies \cite<e.g.,>{Cooper:1975wp}, and compare the results
+of the experiment with each of the models. We conclude with a
+discussion of the strengths and weaknesses of each model, and lay out
+directions for future work.
 
@@
 
-Previous models of mental rotation have, for the most part, focused
-mostly on the representation of mental images, rather than how people
-decide \textit{which} mental images to
-simulate. \citeA{Kosslyn:1977tv} propose a model of mental imagery's
-visual buffer, but not the planning process of how it is used;
-similarly, \citeA{Glasgow:1992tj} are mostly concerned with the
-knowledge representation of imagery. Although \citeA{Anderson1978}
-emphasizes the importance of considering both representation and
-process, he dismisses the problem of determining the direction and
-axis of rotation as merely a ``technical difficulty''.
+Previous models of mental rotation have, for the most part, focused on
+the representation of mental images, rather than how people decide
+\textit{which} mental images to generate. \citeA{Kosslyn:1977tv}
+propose a model of the mental imagery buffer, but not the planning
+process of how it is used. Similarly, \citeA{Julstrom:1985va} and
+\citeA{Glasgow:1992tj} are mostly concerned with modeling the
+representational format underlying imagery. Although
+\citeA{Anderson1978} emphasizes the importance of considering both
+representation and process, he dismisses the problem of determining
+the direction of rotation as a ``technical difficulty''.
 
-The only model (of which the authors are aware) that has more
-seriously attempted to address concerns regarding which simulation to
-run is the parallel-process model of mental rotation by
-\citeA{Funt:1983wn}. He determines the direction and axis of rotation
-using the method-of-moments method; once this target rotation
-computed, his model rotates one object until the target rotation is
-reached, and then compares the two to see if they are congruent. This
-is a somewhat unsatisfying explanation, however: why would it be
-necessary to perform the rotation if the answer is already known?
+The only model (of which the authors are aware) that has seriously
+attempted to address the decision of \textit{what} to simulate is the
+parallel-process model of mental rotation by \citeA{Funt:1983wn}. He
+determines the direction, axis, and extent of rotation by computing
+the objects' principal axes using moments of inertia. One object is
+then rotated through the target rotation, and is checked against the
+other object for congruency. However, this approach assumes that the
+corresponding points on the two objects can be easily identified,
+which is not necessarily the case.
 
@@
 
-It may be useful to take a step back and try to analyze mental
-rotation from using rational analysis
+If people are not performing single-trajectory rotations, what might
+they be doing? To answer this question, we perform a ``rational
+analysis'' of the problem of mental rotation
 \cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
@@
 undergone. At the algorithmic level, we are constrained by the notion
-that mental images must be transformed in an analog manner (or at
-least in a way that is approximately analog), and that mental images
-are time-consuming and effortful to generate. Thus, the \textit{goal}
-is to make this determiniation while using the least amount of
-resources. Due to the sequential constratiant, the minimum amount of
-computation will coincide with the minimum angle of rotation. So, the
-rational \textit{solution} will be something that only rotates through
-the minimum angle.
+that mental images must be transformed in an analog manner (or in a
+way that is at least approximately analog), and that mental images are
+time-consuming and effortful to generate. Thus, the \textit{goal} is
+to make this determination while performing a minimum amount of
+computation (i.e., as few rotations as possible).
 
-motivate the active searching approach
-\cite{Gureckis:2012gu,Markant:2012uu} \cite{Markant:2012uu,Nelson2007}
+The original congruency hypothesis \cite{Shepard1971} actually is a
+rational solution which acheives this goal, in the sense that the
+smallest amount of computation coincides with rotating through the
+minimum angle. However, it relies on assumptions of \textit{a priori}
+knowledge (either of the minimum angle or the correspondence of points
+between shapes). Noting that a rational solution need not necessarily
+maintain a single trajectory of rotation, we explore an alternative
+rational model, which--rather than assuming the angle of rotation
+\textit{a priori}--engages in an ``active sampling'' strategy.
 
-problem solving \cite{Hegarty2004, Schwartz1999}
+Active sampling is the idea that, when given information is
+incomplete, people gather new information in a manner that increases
+certainty about the problem space. This notion has recently gained
+support in several areas of cognitive science
+\cite<e.g.,>{Nelson2007,Gureckis:2012gu}, including other spatial
+domains \cite{Juni:2011vo}. In the case of mental rotation, if people
+do not know the minimum angle beforehand, then actively performing
+rotations may be the best way to gather evidence about the similarity
+between the observed shapes.
 
-symbolic vs simulation \cite{Schwartz:1996uy}
+In the following section, we formalize this rational analysis, and
+propose four models of mental rotation: one based on the original
+hypothesis by \citeA{Shepard1971}; two which are extensions of the
+first, but with relaxes assumptions; and one based on the active
+sampling appraoach.
 
@@
 two-dimensional shapes similar to those used by \citeA{Cooper:1975wp};
-an example stimulus is shown in Fig.~\ref{fig:stimuli}. The goal is to
-determine whether the two images depict the same shape, or
+example stimuli are shown in Fig.~\ref{fig:model-traces}. The goal is
+to determine whether the two images depict the same shape, or
 mirror-image (flipped) shapes.
-
-here we use simple 2d models as a starting point, but the general case
-of solving this problem is very difficult, \TODO{cite shape matching
-  literature}
 
@@
 \end{equation}
+Example similarity curves are shown in Fig.~\ref{fig:model-traces}.
+
 We do not know which vertices of $\Xb$ correspond to which vertices of
@@
 where $\epsilon\sim \mathcal{N}(0, \sigma_\epsilon)$ and
-$\sigma_\epsilon=0.6$ is the standard deviation of the step size in
+$\sigma_\epsilon$ is the standard deviation of the step size in
 radians.
@@
     \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
-    \caption{\textbf{Model traces.} \TODO{}}
+    \caption{\textbf{Example stimuli and model behavior.} On the left
+      are two stimulus pairs, one ``same'' (top) and one ``flipped''
+      (bottom), both with a rotation of $120^\prime$. The grey plots
+      show the actions that each model took for these stimuli. The
+      color of each point (action) indicates whether the model's
+      current mental image was generated under the hypothesis that the
+      objects are the same (red) or flipped (blue). The similarity
+      functions for each of these hypotheses are shown on the right.}
     \label{fig:model-traces}
@@
 and \textit{when} stop sampling. The models below differ in how they
-make these decisions.
+make these decisions (see Fig.~\ref{fig:model-traces} for examples of
+their behavior).
 
@@
 $(\Xa\F_h^T)_\mathrm{left}^{-1}$ is the left inverse of
-$\Xa\F_h^T$. We then check each $h$ to see if computation produces a
-valid rotation matrix; the $h$ that does produce a rotation matrix is
-the correct hypothesis. This approach gives us the true value of
-$\theta$, so Eq.~\ref{eq:lh-ratio} becomes a likelihood ratio test,
-where $\theta$ is set to the MLE value, rather than being
-marginalized:
+$\Xa\F_h^T$. We then check each $h$ to see if the computation produces
+a valid rotation matrix; the $h$ that does is the correct
+hypothesis. This approach gives us the true value of $\theta$, so
+Eq.~\ref{eq:lh-ratio} becomes a maximum likelihood ratio test, where
+$\theta$ is set to the MLE value, rather than being marginalized:
 \begin{equation}
@@
 
-We note that in the current formulation of the problem, choosing the
-threshold is straightforward because we know both the exact geometry
-of the shapes and that a linear transformation exists which will
-perfectly align them. However, it is not always clear how to choose a
-reasonable threshold \textit{a priori}, as the global optimum will
-depend on many factors (such as shape complexity, dimensionality,
-perceptual uncertainty, and whether the shapes are exactly
-identical). Thus, it is unlikely that the \Th{} model will be robust
-beyond this domain.
+In the current formulation of the problem, choosing the threshold is
+straightforward because we know both the exact geometry of the shapes
+and that a linear transformation exists which will perfectly align
+them. However, it is not always clear how to choose a reasonable
+threshold \textit{a priori}, as the global optimum will depend on many
+factors (such as shape complexity, dimensionality, perceptual
+uncertainty, and whether the shapes are exactly identical). Thus, it
+is unlikely that the \Th{} model will be robust beyond this domain.
 
@@
 Eq.~\ref{eq:lh-ratio} while still obeying the constraint of sequential
-rotations. Instead of perfoming a search, we could try maintaining a
+rotations. Based on evidence that people use ``active sampling''
+\cite<e.g.,>{Nelson2007,Gureckis:2012gu}, we hypothesize a model that
+actively monitors the usefulness of the rotations it is
+performing. Instead of searching for a maximum, we maintain a
 probability distribution over our \textit{estimate} of
-Eq.~\ref{eq:lh-ratio}, and then choose actions which are expected to
-improve the accuracy of the estimate. This strategy has several
-obvious benefits. First, it does not make assumptions about the shape
-or scale of the similarity function. Second, by choosing to sample
-places which are informative, this method implicitly minimizes the
-amount of rotation, but because it is constrained to sequential
-rotations, should still exhibit the linear effect of the angle.
+Eq.~\ref{eq:lh-ratio}, and then sample actions which are expected to
+improve that estimate.  This strategy has several benefits. First, it
+does not make assumptions about the shape or scale of the similarity
+function. Second, by choosing to sample places which are informative,
+this method implicitly minimizes the amount of rotation; yet, because
+it is constrained to sequential rotations, it should still exhibit the
+linear effect of the angle.
 
@@
 
+To evaluate the models described previously, we ran a behavioral
+experiment based on classic mental rotation studies
+\cite<e.g.>{Shepard1971, Cooper:1975wp}.
+
 \subsection{Stimuli}
 
-We randomly generated 20 shapes of 5 or 6 vertices (e.g., the stimuli
-in Fig.~\ref{fig:model-traces}). For each shape, we computed 20
-``same'' and 20 ``flipped'' stimuli pairs, with $\theta$ spaced at
+We randomly generated 20 shapes of 5 or 6 vertices (e.g.,
+Fig.~\ref{fig:model-traces}). For each shape, we computed 20 ``same''
+and 20 ``flipped'' stimuli pairs, with rotations ($\theta$) spaced at
 $20^\prime$ increments between 0 and 360 (with 0 and 180 degree
@@
 
-We generated 5 shapes to be used in the practice block. Across all the
-practice 10 stimuli, each shape and each angle (60, 120, 180, 240, or
-300) was repeated twice (once ``flipped'' and once ``same'') such that
-no shape was presented at the same angle twice. Additionally, we
-generated a sixth shape to be included along with the instructions,
-which had both a ``flipped'' and ``same'' version, each rotated to 320
-degrees.
-
-\begin{figure*}[t]
-  \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/response_time.pdf}
-    \caption{\textbf{Response times.} \TODO{}}
-    \label{fig:response-times}
-  \end{center}
-\end{figure*}
+We generated 5 additional shapes to be used in a practice block of 10
+trials. Across these trials, each shape and each angle (60, 120, 180,
+240, or 300) was repeated twice (once ``flipped'' and once ``same'')
+such that no shape was presented at the same angle twice. We also
+generated a sixth shape to be included along with the instructions.
+This shape had both a ``flipped'' and ``same'' version, each rotated
+to 320 degrees.
 
@@
 
-All participants saw the same 10 practice trials (as described
-above). There were a total of 800 experimental stimuli (20 shapes
-$\times$ 20 angles $\times$ 2 same/flipped), which were split into 8
-blocks of 100 across conditions in the following manenr. First,
-stimuli were split into four blocks of 200 trials. Within each block,
-each shape was repeated 10 times and each rotation was repeated 10
-times (5 same, 5 flipped), such that across all blocks, each stimulus
-appeared exactly once. Each block was then split in half. Each
-participant then completed two half-blocks (not necessarily both from
-the original full block).
+All participants saw the same 10 practice trials, as described
+above. There were 720 unique experimental stimuli (20 shapes $\times$
+18 angles $\times$ 2 same/flipped), though because stimuli with
+rotations of $0^\prime$ or $180^\prime$ were repeated twice, there
+were 800 total experimental stimuli. These stimuli were split across 8
+conditions in the following manner: first, stimuli were split into
+four blocks of 200 trials. Within each block, each shape was repeated
+10 times and each rotation was repeated 10 times (5 same, 5 flipped),
+such that across all blocks, each stimulus appeared exactly once. Each
+block was then split in half. Participants completed two half-blocks
+(not necessarily both from the original full block).
 
 \subsection{Procedure}
-
-\begin{figure*}[t]
-  \begin{center}
-    \includegraphics[width=0.85\textwidth]{../../figures/D/accuracy.pdf}
-    \caption{\textbf{Accuracy.} \TODO{}}
-    \label{fig:accuracy}
-  \end{center}
-\end{figure*}
 
@@
 
-\section{Results}
-
-Out of the 247 participants, 200 (81\%) were included in our
-analyses. Of the rest, we excluded 10 (4\%) because of an experimental
-error, 6 (2.4\%) because they had already compelted a previous version
-of the experiment, and 31 (12.6\%) because they failed a comprehension
-check. A participant was marked as passing the comprehension check if
-they answered at least 85\% of ``easy'' trials correctly. ``Easy''
-trials were defined to be those for which the presented shapes
-differed by a rotation of no more than 20 degrees. \TODO{mention
-  excluding trials where people took a really long time to respond}
-
-We ran model simulations for the same stimuli that participants
-viewed. We ran 10 samples of each stimulus under each of the modesl,
-with the exception of stimuli where $\theta=0$ or $\theta=180$, in
-which case we ran double the number of samples.
+\subsection{Analysis}
 
@@
 
+\section{Results}
+
+Out of the 247 participants, 200 (81\%) were included in our
+analyses. Of the rest, we excluded 10 (4\%) because of an experimental
+error, 6 (2.4\%) because they had already compelted a previous version
+of the experiment, and 31 (12.6\%) because they failed a comprehension
+check. A participant was marked as passing the comprehension check if
+they answered at least 85\% of ``easy'' trials correctly. ``Easy''
+trials were defined to be those for which the presented shapes
+differed by a rotation of no more than 20 degrees. Additionally, we
+excluded individual trials for which the response time exceeded 20
+seconds.
+
+For each model, we ran 50 samples for each of the same stimuli that
+participants viewed (with the exception of stimuli with rotations of
+$0^\prime$ or $180^\prime$, in which case we ran double the number of
+samples). The step size parameter ($\sigma_\epsilon$) was
+approximately fit to each of the models, resulting in
+$\sigma_\epsilon=0.6$ for the \Bq{} model, and $\sigma_\epsilon=0.1$
+for all other models. We also ran the models under two different
+priors, $p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the
+``unequal'' prior), though this only had an effect for the \Bq{}
+model. \TODO{actually, did it affect HC?  check this}
+
 \begin{figure}[t]
   \centering
-  \subfloat{\includegraphics[width=0.25\textwidth]{../../figures/D/trial_accuracy.pdf}}%
-  \subfloat{\includegraphics[width=0.25\textwidth]{../../figures/D/trial_time.pdf}}
-  \caption{\textbf{Effect of learning.} \TODO{}}
+  \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_accuracy.pdf}}%
+  \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_time.pdf}}
+  \caption{\textbf{Effect of learning.} There was a clear effect of
+    learning on participants' responses, especially in the first half
+    of the experiment. Left: accuracy as a function of trial. Right:
+    response time as a function of trial.}
   \label{fig:learning}
@@
 vs. \ExpbTrialAccuracyCorr{} for the second
-half). Fig.~\ref{fig:learning}
+half). Fig.~\ref{fig:learning} shows scatterplots illustrating these
+trends.
 
@@
 (\ThThetaTimeCorr{}), and the \Bq{} model (\BqThetaTimeCorr{}), but
-not the \Hc{} model
-(\HcThetaTimeCorr{}). Fig.~\ref{fig:response-times}a
+not the \Hc{} model (\HcThetaTimeCorr{}). The top row of
+Fig.~\ref{fig:response-time-accuracy} shows the overall effect of
+rotation on response times.
 
@@
 of human response times, with a correlation of \ExpOcTimeCorr{}. The
-\Th{} model was the next best (\ExpThTimeCorr{}), with the \Bq{} model
-not far behind (\ExpBqTimeCorr{}). The \Hc{} model was not
-significantly correlated with people (\ExpHcTimeCorr{}).
+\Th{} model \Bq{} model with $p(h=0)=0.55$ were the next best (for
+\Th{}, \ExpThTimeCorr{}; for \Bq{}, \ExpBqpTimeCorr{}). Not far behind
+was the BQ model with equal prior (\ExpBqTimeCorr{}). The \Hc{} model
+was barely correlated with people (\ExpHcTimeCorr{}).
 
-\TODO{flipped response times higher than same response times for
-  smaller angles?}
+Across all stimuli, participants' response times for ``flipped'' pairs
+were significantly higher than those for ``same'' pairs
+\TODO{confidence intervals} \TODO{flipped response times higher than
+  same response times for smaller angles?}
 
@@
 models peformed perfectly, with 100\% accuracy. The \Bq{} model was
-very accurate overall (\BqAccuracy{}), though there were
-\BqNumChance{} stimuli for which it was not above chance. The \Hc{}
-model was not particularly accurate, and although it did perform above
-chance overall (\HcAccuracy{}), there were \HcNumChance{} individual
-stimuli for which it was not above chance.
+very accurate overall (equal prior: \BqAccuracy{}, unequal prior:
+\BqpAccuracy{}). With the equal prior, there was \BqNumChance{}
+stimulus for which it was not above chance; with the unequal prior,
+there were \BqpNumChance{}. The \Hc{} model was not particularly
+accurate, and although it did perform above chance overall
+(\HcAccuracy{}), there were \HcNumChance{} individual stimuli for
+which it was not above chance.
 
-The minimum angle was also correlated with the average per-stimulus
-accuracy: for flipped pairs, \ExpThetaAccuracyCorrFlipped{}, and for
-same pairs, \ExpThetaAccuracyCorrSame{}. There was also a significant
-effect for the \Hc{} model (\HcThetaAccuracyCorr{}) and the
-\Bq{} model
-(\BqThetaAccuracyCorr{}). Fig.~\ref{fig:accuracy}
+The minimum angle was also correlated with participants' average
+per-stimulus accuracy: for flipped pairs,
+\ExpThetaAccuracyCorrFlipped{}, and for same pairs,
+\ExpThetaAccuracyCorrSame{}. There was also a significant effect for
+the \Hc{} model (\HcThetaAccuracyCorr{}) and the \Bq{} model (equal
+prior: \BqThetaAccuracyCorr{}, unequal prior:
+\BqpThetaAccuracyCorr{}). The bottom row of
+Fig.~\ref{fig:response-time-accuracy} shows the effect of angle on
+accuracy for both participants and models.
 
@@
 model and human accuracy (\ExpHcAccuracyCorr{}). The \Bq{} model had a
-similar correlation, with \ExpBqAccuracyCorr{}.
+similar correlation (equal prior: \ExpBqAccuracyCorr{}, unequal prior:
+\ExpBqpAccuracyCorr{}). Correlations for the other models were
+undefined, because they had 100\% accuracy.
+
+\begin{figure*}[t]
+  \begin{center}
+    \includegraphics[width=\textwidth]{../../figures/D/response_time_accuracy.pdf}
+    \caption{\textbf{Response time and accuracy comparison.} Top:
+      response time of correct responses as a function of the minimum
+      angle of rotation. Bottom: accuracy as a function of the minumum
+      angle of rotation. All error bars are 95\% confidence
+      intervals.}
+    \label{fig:response-time-accuracy}
+  \end{center}
+\end{figure*}
 

### 2014-01-30 man/cogsci-2014/cogsci-2014.tex
Add psiTurk reference

@@
 hypothesis by \citeA{Shepard1971}; two which are extensions of the
-first, but with relaxes assumptions; and one based on the active
-sampling appraoach.
+first but with relaxed assumptions; and one based on the active
+sampling approach.
 
@@
 
-We recruited 247 participants on Amazon's Mechanical Turk. Each
-participant was paid \$1.00 for approximately 15 minutes of work,
-consisting of one block of 10 practice trials followed by two blocks
-of 100 experiment trials. Within a block, trials were presented in a
-random order.
+We recruited 247 participants on Amazon's Mechanical Turk using the
+psiTurk experiment framework \cite{Mcdonnell12}. Each participant was
+paid \$1.00 for approximately 15 minutes of work, consisting of one
+block of 10 practice trials followed by two blocks of 100 experiment
+trials. Within a block, trials were presented in a random order.
 

### 2014-01-30 analysis/analyses/theta_time_corrs.py
Use spearman correlations in theta_time_corrs

@@ -22,7 +22,7 @@ def run(data, results_path, seed):
                 thetas = x['modtheta']
                 times = x['time']
 
-                corr = dict(util.bootcorr(thetas, times))
+                corr = dict(util.bootcorr(thetas, times, method='spearman'))
 
                 print "%s (%s): %s" % (
                     key, flipped, util.report_spearman.format(**corr))
@@ -40,7 +40,7 @@ def run(data, results_path, seed):
             thetas = x['modtheta']
             times = x['time']
 
-            corr = dict(util.bootcorr(thetas, times))
+            corr = dict(util.bootcorr(thetas, times, method='spearman'))
 
             print "%s (all): %s" % (
                 key, util.report_spearman.format(**corr))

### 2014-01-30 analysis/plots/__init__.py
Add plots for time and accuracy correlations

@@ -7,6 +7,8 @@ import accuracy_scatters
 import trial_time
 import trial_accuracy
 import response_time_accuracy
+import theta_time_corrs
+import theta_accuracy_corrs
 
 __all__ = [
     'response_time_histograms',
@@ -16,5 +18,7 @@ __all__ = [
     'accuracy_scatters',
     'trial_time',
     'trial_accuracy',
-    'response_time_accuracy'
+    'response_time_accuracy',
+    'theta_time_corrs',
+    'theta_accuracy_corrs',
 ]

### 2014-01-30 analysis/plots/theta_accuracy_corrs.py

@@ -1,0 +1,89 @@
+#!/usr/bin/env python
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import util
+from path import path
+
+
+def plot(data, fig_path, seed):
+    np.random.seed(seed)
+
+    def makecorr(df):
+        x = df.groupby(['stimulus', 'modtheta'])['correct']\
+              .mean()\
+              .reset_index()
+        thetas = x['modtheta']
+        acc = 1 - x['correct']
+        corr = util.bootcorr(thetas, acc, method="spearman")
+        return corr
+
+    keys = ['exp', 'oc', 'th', 'hc', 'bq', 'bqp']
+    corrs = {}
+    for key in keys:
+        for flipped, df in data[key].groupby('flipped'):
+            corrs[(key, flipped)] = makecorr(df)
+
+    df = pd.DataFrame.from_dict(corrs, orient='index').fillna(0)
+    df.index = pd.MultiIndex.from_tuples(df.index, names=['model', 'stimuli'])
+    df = df.unstack('stimuli').reindex(keys).stack('stimuli')
+
+    fig, ax = plt.subplots()
+    width = 1
+    offset = width * 5 / 2.
+    colors = {
+        'same': '#ff5555',
+        'flipped': '#5555ff'
+    }
+    titles = {
+        'exp': "Human",
+        'oc': "Oracle",
+        'th': "Threshold",
+        'hc': "HC",
+        'bq': "BQ\n(equal)",
+        'bqp': "BQ\n(unequal)"
+    }
+
+    order = ['same', 'flipped']
+    for flipped, y in df.groupby(level='stimuli'):
+        i = order.index(flipped)
+        median = y['median']
+        lerr = median - y['lower']
+        uerr = y['upper'] - median
+        ax.bar(
+            i * width + np.arange(len(median)) * offset,
+            median,
+            yerr=[lerr, uerr],
+            color=colors[flipped],
+            ecolor='k',
+            width=width,
+            edgecolor='none')
+
+    ax.set_ylim(0, 1)
+    ax.set_xlim(-width / 2., len(median) * offset)
+    ax.set_xticks(np.arange(len(median)) * offset + width)
+    ax.set_xticklabels([titles[key] for key in keys], fontsize=10)
+
+    util.clear_right(ax)
+    util.clear_top(ax)
+    util.outward_ticks(ax)
+
+    ax.set_title("Correlation between rotation and accuracy", fontsize=14)
[…]

### 2014-01-30 analysis/plots/theta_time_corrs.py

@@ -1,0 +1,90 @@
+#!/usr/bin/env python
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import util
+from path import path
+
+
+def plot(data, fig_path, seed):
+    np.random.seed(seed)
+
+    def makecorr(df):
+        x = df[df['correct']]\
+            .groupby(['stimulus', 'modtheta'])['time']\
+            .mean()\
+            .reset_index()
+        thetas = x['modtheta']
+        times = x['time']
+        corr = util.bootcorr(thetas, times, method="spearman")
+        return corr
+
+    keys = ['exp', 'oc', 'th', 'hc', 'bq', 'bqp']
+    corrs = {}
+    for key in keys:
+        for flipped, df in data[key].groupby('flipped'):
+            corrs[(key, flipped)] = makecorr(df)
+
+    df = pd.DataFrame.from_dict(corrs, orient='index')
+    df.index = pd.MultiIndex.from_tuples(df.index, names=['model', 'stimuli'])
+    df = df.unstack('stimuli').reindex(keys).stack('stimuli')
+
+    fig, ax = plt.subplots()
+    width = 1
+    offset = width * 5 / 2.
+    colors = {
+        'same': '#ff5555',
+        'flipped': '#5555ff'
+    }
+    titles = {
+        'exp': "Human",
+        'oc': "Oracle",
+        'th': "Threshold",
+        'hc': "HC",
+        'bq': "BQ\n(equal)",
+        'bqp': "BQ\n(unequal)"
+    }
+
+    order = ['same', 'flipped']
+    for flipped, y in df.groupby(level='stimuli'):
+        i = order.index(flipped)
+        median = y['median']
+        lerr = median - y['lower']
+        uerr = y['upper'] - median
+        ax.bar(
+            i * width + np.arange(len(median)) * offset,
+            median,
+            yerr=[lerr, uerr],
+            color=colors[flipped],
+            ecolor='k',
+            width=width,
+            edgecolor='none')
+
+    ax.set_ylim(0, 1)
+    ax.set_xlim(-width / 2., len(median) * offset)
+    ax.set_xticks(np.arange(len(median)) * offset + width)
+    ax.set_xticklabels([titles[key] for key in keys], fontsize=10)
+
+    util.clear_right(ax)
+    util.clear_top(ax)
+    util.outward_ticks(ax)
+
+    ax.set_title("Correlation between rotation and response time", fontsize=14)
[…]

### 2014-01-30 analysis/plots/util.py

@@ -4,7 +4,7 @@ import os
 import sys
 
 from mental_rotation.analysis import load_human, load_all
-from mental_rotation.analysis import bootstrap, beta
+from mental_rotation.analysis import bootstrap, beta, bootcorr
 
 
 def load_config(pth):

### 2014-01-31 analysis/analyses/num_chance.py
Create latex files

@@ -5,6 +5,7 @@ import util
 import pandas as pd
 
 filename = "num_chance.csv"
+texname = "num_chance.tex"
 
 
 def run(data, results_path, seed):
@@ -24,6 +25,15 @@ def run(data, results_path, seed):
     results.index.name = 'model'
     pth = results_path.joinpath(filename)
     results.to_csv(pth)
+
+    with open(results_path.joinpath(texname), "w") as fh:
+        fh.write("%% AUTOMATICALLY GENERATED -- DO NOT EDIT!\n")
+        for model, stats in results.iterrows():
+            cmd = util.newcommand(
+                "%sNumChance" % model.capitalize(),
+                stats['num'])
+            fh.write(cmd)
+
     return pth
 
 

### 2014-01-31 analysis/analyses/overall_accuracy.py

@@ -5,6 +5,7 @@ import util
 import pandas as pd
 
 filename = "overall_accuracy.csv"
+texname = "overall_accuracy.tex"
 
 
 def run(data, results_path, seed):
@@ -20,6 +21,15 @@ def run(data, results_path, seed):
     results.index.name = 'model'
     pth = results_path.joinpath(filename)
     results.to_csv(pth)
+
+    with open(results_path.joinpath(texname), "w") as fh:
+        fh.write("%% AUTOMATICALLY GENERATED -- DO NOT EDIT!\n")
+        for model, stats in results.iterrows():
+            cmd = util.newcommand(
+                "%sAccuracy" % model.capitalize(),
+                util.latex_percent.format(**dict(stats)))
+            fh.write(cmd)
+
     return pth
 
 if __name__ == "__main__":

### 2014-01-31 analysis/analyses/overall_response_time.py

@@ -5,6 +5,7 @@ import util
 import pandas as pd
 
 filename = "overall_response_time.csv"
+texname = "overall_response_time.tex"
 
 
 def run(data, results_path, seed):
@@ -18,6 +19,15 @@ def run(data, results_path, seed):
     results.index.name = 'model'
     pth = results_path.joinpath(filename)
     results.to_csv(pth)
+
+    with open(results_path.joinpath(texname), "w") as fh:
+        fh.write("%% AUTOMATICALLY GENERATED -- DO NOT EDIT!\n")
+        for model, stats in results.iterrows():
+            cmd = util.newcommand(
+                "%sTime" % model.capitalize(),
+                util.latex_mean.format(**dict(stats)))
+            fh.write(cmd)
+
     return pth
 
 

### 2014-01-31 analysis/analyses/trial_accuracy_corrs.py

@@ -5,6 +5,7 @@ import util
 import pandas as pd
 
 filename = "trial_accuracy_corrs.csv"
+texname = "trial_accuracy_corrs.tex"
 
 
 def run(data, results_path, seed):
@@ -18,13 +19,22 @@ def run(data, results_path, seed):
         df = means.groupby('model').get_group(key)
         trials = df['trial']
         accuracy = df['median']
-        corr = util.bootcorr(trials, accuracy)
+        corr = util.bootcorr(trials, accuracy, method='spearman')
         results[key] = corr
 
     results = pd.DataFrame.from_dict(results, orient='index')
     results.index.name = 'model'
     pth = results_path.joinpath(filename)
     results.to_csv(pth)
+
+    with open(results_path.joinpath(texname), "w") as fh:
+        fh.write("%% AUTOMATICALLY GENERATED -- DO NOT EDIT!\n")
+        for model, stats in results.iterrows():
+            cmd = util.newcommand(
+                "%sTrialAccuracyCorr" % model.capitalize(),
+                util.latex_spearman.format(**dict(stats)))
+            fh.write(cmd)
+
     return pth
 
 

### 2014-01-31 analysis/analyses/trial_time_corrs.py

@@ -5,6 +5,7 @@ import util
 import pandas as pd
 
 filename = "trial_time_corrs.csv"
+texname = "trial_time_corrs.tex"
 
 
 def run(data, results_path, seed):
@@ -25,6 +26,15 @@ def run(data, results_path, seed):
     results.index.name = 'model'
     pth = results_path.joinpath(filename)
     results.to_csv(pth)
+
+    with open(results_path.joinpath(texname), "w") as fh:
+        fh.write("%% AUTOMATICALLY GENERATED -- DO NOT EDIT!\n")
+        for model, stats in results.iterrows():
+            cmd = util.newcommand(
+                "%sTrialTimeCorr" % model.capitalize(),
+                util.latex_spearman.format(**dict(stats)))
+            fh.write(cmd)
+
     return pth
 
 

### 2014-01-31 analysis/plots/accuracy_corrs.py
Plot formatting

@@ -42,7 +42,8 @@ def plot(results_path, fig_path):
             color=colors[flipped],
             ecolor='k',
             width=width,
-            edgecolor='none')
+            edgecolor='none',
+            capsize=0)
 
     ax.set_ylim(0, 1)
     ax.set_xlim(-width / 2., len(median) * offset)
@@ -55,6 +56,26 @@ def plot(results_path, fig_path):
 
     ax.set_title("Correlation with human accuracy", fontsize=14)
     ax.set_ylabel(r"Pearson correlation ($r$)", fontsize=14)
+
+    p0 = plt.Rectangle(
+        (0, 0), 1, 1,
+        fc=colors['same'],
+        ec=colors['same'])
+    p1 = plt.Rectangle(
+        (0, 0), 1, 1,
+        fc=colors['flipped'],
+        ec=colors['flipped'])
+
+    leg = ax.legend(
+        [p0, p1], ["\"same\" pairs", "\"flipped\" pairs"],
+        numpoints=1, fontsize=12,
+        loc='upper right')
+    frame = leg.get_frame()
+    frame.set_facecolor('0.9')
+    frame.set_edgecolor('#FFFFFF')
+
+    plt.draw()
+    plt.tight_layout()
 
     pths = [fig_path.joinpath("accuracy_corrs.%s" % ext)
             for ext in ('png', 'pdf')]

### 2014-01-31 analysis/plots/response_time_accuracy.py

@@ -33,21 +33,21 @@ def plot(results_path, fig_path):
         ax = axes[0, i]
         df = time_results.groupby('model').get_group(key)
 
-        maxy = -np.inf
-        miny = np.inf
         for flipped, stats in df.groupby('flipped'):
-            lower = stats['median'] - stats['lower']
-            upper = stats['upper'] - stats['median']
+            if key == 'exp':
+                median = stats['median'] / 1000.
+                lower = (stats['median'] - stats['lower']) / 1000.
+                upper = (stats['upper'] - stats['median']) / 1000.
+            else:
+                median = stats['median']
+                lower = stats['median'] - stats['lower']
+                upper = stats['upper'] - stats['median']
+
             ax.errorbar(
-                stats['modtheta'], stats['median'],
+                stats['modtheta'], median,
                 yerr=[lower, upper], lw=3,
                 color=colors[flipped],
                 ecolor=colors[flipped])
-
-            if stats['upper'].max() > maxy:
-                maxy = stats['upper'].max()
-            if stats['lower'].min() < miny:
-                miny = stats['lower'].min()
 
         ax.set_xticks(np.arange(0, 200, 30))
         ax.set_xlim(-10, 190)
@@ -58,8 +58,6 @@ def plot(results_path, fig_path):
             ax.set_ylabel("Response time", fontsize=14)
         else:
             ax.set_ylabel("Number of actions", fontsize=14)
-            buff = np.ceil((np.round(maxy) - np.round(miny)) * 0.05)
-            ax.set_ylim(np.round(miny) - buff, np.round(maxy) + buff)
 
     util.sync_ylims(axes[0, order.index('bq')], axes[0, order.index('bqp')])
 

### 2014-01-31 analysis/plots/response_time_corrs.py

@@ -42,7 +42,8 @@ def plot(results_path, fig_path):
             color=colors[flipped],
             ecolor='k',
             width=width,
-            edgecolor='none')
+            edgecolor='none',
+            capsize=0)
 
     ax.set_ylim(0, 1)
     ax.set_xlim(-width / 2., len(median) * offset)
@@ -55,6 +56,26 @@ def plot(results_path, fig_path):
 
     ax.set_title("Correlation with human response time", fontsize=14)
     ax.set_ylabel(r"Pearson correlation ($r$)", fontsize=14)
+
+    p0 = plt.Rectangle(
+        (0, 0), 1, 1,
+        fc=colors['same'],
+        ec=colors['same'])
+    p1 = plt.Rectangle(
+        (0, 0), 1, 1,
+        fc=colors['flipped'],
+        ec=colors['flipped'])
+
+    leg = ax.legend(
+        [p0, p1], ["\"same\" pairs", "\"flipped\" pairs"],
+        numpoints=1, fontsize=12,
+        loc='upper right')
+    frame = leg.get_frame()
+    frame.set_facecolor('0.9')
+    frame.set_edgecolor('#FFFFFF')
+
+    plt.draw()
+    plt.tight_layout()
 
     pths = [fig_path.joinpath("response_time_corrs.%s" % ext)
             for ext in ('png', 'pdf')]

### 2014-01-31 analysis/plots/theta_accuracy_corrs.py

@@ -43,7 +43,8 @@ def plot(results_path, fig_path):
             color=colors[flipped],
             ecolor='k',
             width=width,
-            edgecolor='none')
+            edgecolor='none',
+            capsize=0)
 
     ax.set_ylim(0, 1)
     ax.set_xlim(-width / 2., len(median) * offset)
@@ -56,6 +57,26 @@ def plot(results_path, fig_path):
 
     ax.set_title("Correlation between rotation and accuracy", fontsize=14)
     ax.set_ylabel(r"Spearman correlation ($r_s$)", fontsize=14)
+
+    p0 = plt.Rectangle(
+        (0, 0), 1, 1,
+        fc=colors['same'],
+        ec=colors['same'])
+    p1 = plt.Rectangle(
+        (0, 0), 1, 1,
+        fc=colors['flipped'],
+        ec=colors['flipped'])
+
+    leg = ax.legend(
+        [p0, p1], ["\"same\" pairs", "\"flipped\" pairs"],
+        numpoints=1, fontsize=12,
+        loc='upper right')
+    frame = leg.get_frame()
+    frame.set_facecolor('0.9')
+    frame.set_edgecolor('#FFFFFF')
+
+    plt.draw()
+    plt.tight_layout()
 
     pths = [fig_path.joinpath("theta_accuracy_corrs.%s" % ext)
             for ext in ('png', 'pdf')]

### 2014-01-31 analysis/plots/theta_time_corrs.py

@@ -43,7 +43,8 @@ def plot(results_path, fig_path):
             color=colors[flipped],
             ecolor='k',
             width=width,
-            edgecolor='none')
+            edgecolor='none',
+            capsize=0)
 
     ax.set_ylim(0, 1)
     ax.set_xlim(-width / 2., len(median) * offset)
@@ -56,6 +57,26 @@ def plot(results_path, fig_path):
 
     ax.set_title("Correlation between rotation and response time", fontsize=14)
     ax.set_ylabel(r"Spearman correlation ($r_s$)", fontsize=14)
+
+    p0 = plt.Rectangle(
+        (0, 0), 1, 1,
+        fc=colors['same'],
+        ec=colors['same'])
+    p1 = plt.Rectangle(
+        (0, 0), 1, 1,
+        fc=colors['flipped'],
+        ec=colors['flipped'])
+
+    leg = ax.legend(
+        [p0, p1], ["\"same\" pairs", "\"flipped\" pairs"],
+        numpoints=1, fontsize=12,
+        loc='upper right')
+    frame = leg.get_frame()
+    frame.set_facecolor('0.9')
+    frame.set_edgecolor('#FFFFFF')
+
+    plt.draw()
+    plt.tight_layout()
 
     pths = [fig_path.joinpath("theta_time_corrs.%s" % ext)
             for ext in ('png', 'pdf')]

### 2014-01-31 man/cogsci-2014/cogsci-2014.tex
Misc changes to paper

@@
 
-\input{../../results/D/response_time.tex}
-\input{../../results/D/response_time_corrs.tex}
-\input{../../results/D/accuracy.tex}
-\input{../../results/D/accuracy_corrs.tex}
+\input{../../results/D/overall_response_time.tex}
+\input{../../results/D/overall_accuracy.tex}
+\input{../../results/D/num_chance.tex}
+\input{../../results/D/trial_accuracy_corrs.tex}
 \input{../../results/D/trial_time_corrs.tex}
-\input{../../results/D/trial_accuracy_corrs.tex}
-\input{../../results/D/num_chance.tex}
-\input{../../results/D/theta_time_corrs.tex}
-\input{../../results/D/theta_accuracy_corrs.tex}
 
@@
 The original congruency hypothesis \cite{Shepard1971} actually is a
-rational solution which acheives this goal, in the sense that the
+rational solution which achieves this goal, in the sense that the
 smallest amount of computation coincides with rotating through the
@@
 One way to deal with the problem of choosing a threshold would simply
-be to perform an exhaustive search or use a global optimisation
+be to perform an exhaustive search or use a global optimization
 strategy. However, these strategies would not result in the linear
@@
 
+\begin{figure}[t]
+  \centering
+  \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_accuracy.pdf}}%
+  \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_time.pdf}}
+  \caption{\textbf{Effect of learning.} There was a clear effect of
+    learning on participants' responses, especially in the first half
+    of the experiment. Left: accuracy as a function of trial. Right:
+    response time in milliseconds as a function of trial.}
+  \label{fig:learning}
+\end{figure}
+
 To evaluate the models described previously, we ran a behavioral
@@
 responses for each angle between 0 and 180). ``Same'' pairs were
-created by rotating $\Xa$ by $\theta$; ``flipped'' pairs first
+created by rotating $\Xa$ by $\theta$; ``flipped'' pairs were first
 reflected $\Xa$ across the $y$-axis, then rotated by $\theta$.
@@
 
-On each trial, particpants were instructed to press the `b' key to
+On each trial, participants were instructed to press the `b' key to
 begin and to focus on the fixation cross that appeared for 750
@@
 While there was no limit on response time, participants were urged to
-repond as quickly as possible while still being
+respond as quickly as possible while still being
 accurate. Specifically, we asked them to aim for at least 85\%
@@
 
-For analyses of response time, confidence intervals were computed
-using a bootstrap analysis (sampling with replacement) with 1000
-bootstrap samples. Response times were only only considered for
-correct responses.  We also used a bootstrap analysis (with 10000
-bootstrap samples) to compute the confidence intervals for both
-Spearman ($r_s$) and Pearson ($r$) correlations.
+\begin{figure*}[t]
+  \begin{center}
+    \includegraphics[width=\textwidth]{../../figures/D/response_time_accuracy.pdf}
+    \caption{\textbf{Response time and accuracy comparison.} Top:
+      response time of correct responses as a function of the minimum
+      angle of rotation. Bottom: accuracy as a function of the minimum
+      angle of rotation. All error bars are 95\% confidence
+      intervals.}
+    \label{fig:response-time-accuracy}
+  \end{center}
+\end{figure*}
+
+For analyses of response time, confidence intervals around harmonic
+means were computed using a bootstrap analysis using 10000 bootstrap
+samples (sampled with replacement). Response times were only
+considered for correct responses.  We also used a bootstrap analysis
+of 10000 bootstrap samples to compute the confidence intervals around
+both Spearman ($r_s$) and Pearson ($r$) correlations.
 
@@
 analyses. Of the rest, we excluded 10 (4\%) because of an experimental
-error, 6 (2.4\%) because they had already compelted a previous version
+error, 6 (2.4\%) because they had already completed a previous version
 of the experiment, and 31 (12.6\%) because they failed a comprehension
@@
 differed by a rotation of no more than 20 degrees. Additionally, we
-excluded individual trials for which the response time exceeded 20
-seconds.
+excluded 82 individual trials (out of 40000) for which the response
+time was either less than 100 milliseconds or greater than 20 seconds.
 
-For each model, we ran 50 samples for each of the same stimuli that
-participants viewed (with the exception of stimuli with rotations of
-$0^\prime$ or $180^\prime$, in which case we ran double the number of
-samples). The step size parameter ($\sigma_\epsilon$) was
-approximately fit to each of the models, resulting in
-$\sigma_\epsilon=0.6$ for the \Bq{} model, and $\sigma_\epsilon=0.1$
-for all other models. We also ran the models under two different
-priors, $p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the
-``unequal'' prior), though this only had an effect for the \Bq{}
-model. \TODO{actually, did it affect HC?  check this}
-
-\begin{figure}[t]
-  \centering
-  \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_accuracy.pdf}}%
-  \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_time.pdf}}
-  \caption{\textbf{Effect of learning.} There was a clear effect of
-    learning on participants' responses, especially in the first half
-    of the experiment. Left: accuracy as a function of trial. Right:
-    response time as a function of trial.}
-  \label{fig:learning}
-\end{figure}
+For each model, we ran 50 samples for each stimulus that participants
+viewed (with the exception of stimuli with rotations of $0^\prime$ or
+$180^\prime$, in which case we ran double the number of samples). The
+step size parameter ($\sigma_\epsilon$) was approximately fit to each
+of the models, resulting in $\sigma_\epsilon=0.6$ for the \Bq{} model,
+and $\sigma_\epsilon=0.1$ for all other models. We also ran the models
+under two different priors, $p(h=0)=0.5$ (the ``equal'' prior) and
+$p(h=0)=0.55$ (the ``unequal'' prior), though this only had an effect
+for the \Bq{} model. \TODO{actually, did it affect HC?  check this}
 
@@
 
-The average response time across all stimuli was \ExpTime{} seconds,
-and the average response time for an individual stimulus ranged from
-\ExpTimeMin{} to \ExpTimeMax{} seconds.  As expected, the minimum
-angle of rotation was significanty monotonically correlated with
-average per-stimulus response times, both for flipped pairs
-(\ExpThetaTimeCorrFlipped{}) and same pairs
-(\ExpThetaTimeCorrSame{}). This was also the case for the number of
-actions taken by the \Oc{} model (\OcThetaTimeCorr{}), the \Th{} model
-(\ThThetaTimeCorr{}), and the \Bq{} model (\BqThetaTimeCorr{}), but
-not the \Hc{} model (\HcThetaTimeCorr{}). The top row of
-Fig.~\ref{fig:response-time-accuracy} shows the overall effect of
-rotation on response times.
+% The average response time across all stimuli was \ExpTime{} seconds.
+% As expected, the minimum angle of rotation was significantly
+% monotonically correlated with average per-stimulus response times,
+% both for flipped pairs (\ExpThetaTimeCorrFlipped{}) and same pairs
+% (\ExpThetaTimeCorrSame{}). This was also the case for the number of
+% actions taken by the \Oc{} model (\OcThetaTimeCorr{}), the \Th{} model
+% (\ThThetaTimeCorr{}), and the \Bq{} model (\BqThetaTimeCorr{}), but
+% not the \Hc{} model (\HcThetaTimeCorr{}). The top row of
+% Fig.~\ref{fig:response-time-accuracy} shows the overall effect of
+% rotation on response times.
 
-The number of actions taken by the \Oc{} model was the best predictor
-of human response times, with a correlation of \ExpOcTimeCorr{}. The
-\Th{} model \Bq{} model with $p(h=0)=0.55$ were the next best (for
-\Th{}, \ExpThTimeCorr{}; for \Bq{}, \ExpBqpTimeCorr{}). Not far behind
-was the BQ model with equal prior (\ExpBqTimeCorr{}). The \Hc{} model
-was barely correlated with people (\ExpHcTimeCorr{}).
+% The number of actions taken by the \Oc{} model was the best predictor
+% of human response times, with a correlation of \ExpOcTimeCorr{}. The
+% \Th{} model \Bq{} model with $p(h=0)=0.55$ were the next best (for
+% \Th{}, \ExpThTimeCorr{}; for \Bq{}, \ExpBqpTimeCorr{}). Not far behind
+% was the BQ model with equal prior (\ExpBqTimeCorr{}). The \Hc{} model
+% was barely correlated with people (\ExpHcTimeCorr{}).
 
-Across all stimuli, participants' response times for ``flipped'' pairs
-were significantly higher than those for ``same'' pairs
-\TODO{confidence intervals} \TODO{flipped response times higher than
-  same response times for smaller angles?}
+% Across all stimuli, participants' response times for ``flipped'' pairs
+% were significantly higher than those for ``same'' pairs
+% \TODO{confidence intervals} \TODO{flipped response times higher than
+%   same response times for smaller angles?}
 
@@
 
-The average accuracy across all stimuli was \ExpAccuracy{}, though
-there were \ExpNumChance{} individual stimuli (out of 720) for which
-people were not above chance. Of the models, the \Oc{} and \Th{}
-models peformed perfectly, with 100\% accuracy. The \Bq{} model was
-very accurate overall (equal prior: \BqAccuracy{}, unequal prior:
-\BqpAccuracy{}). With the equal prior, there was \BqNumChance{}
-stimulus for which it was not above chance; with the unequal prior,
-there were \BqpNumChance{}. The \Hc{} model was not particularly
-accurate, and although it did perform above chance overall
-(\HcAccuracy{}), there were \HcNumChance{} individual stimuli for
-which it was not above chance.
+% The average accuracy across all stimuli was \ExpAccuracy{}, though
+% there were \ExpNumChance{} individual stimuli (out of 720) for which
+% people were not above chance. Of the models, the \Oc{} and \Th{}
+% models performed perfectly, with 100\% accuracy. The \Bq{} model was
+% very accurate overall (equal prior: \BqAccuracy{}, unequal prior:
+% \BqpAccuracy{}). With the equal prior, there was \BqNumChance{}
+% stimulus for which it was not above chance; with the unequal prior,
+% there were \BqpNumChance{}. The \Hc{} model was not particularly
+% accurate, and although it did perform above chance overall
+% (\HcAccuracy{}), there were \HcNumChance{} individual stimuli for
+% which it was not above chance.
 
-The minimum angle was also correlated with participants' average
-per-stimulus accuracy: for flipped pairs,
-\ExpThetaAccuracyCorrFlipped{}, and for same pairs,
-\ExpThetaAccuracyCorrSame{}. There was also a significant effect for
-the \Hc{} model (\HcThetaAccuracyCorr{}) and the \Bq{} model (equal
-prior: \BqThetaAccuracyCorr{}, unequal prior:
-\BqpThetaAccuracyCorr{}). The bottom row of
-Fig.~\ref{fig:response-time-accuracy} shows the effect of angle on
-accuracy for both participants and models.
+% The minimum angle was also correlated with participants' average
+% per-stimulus accuracy: for flipped pairs,
+% \ExpThetaAccuracyCorrFlipped{}, and for same pairs,
+% \ExpThetaAccuracyCorrSame{}. There was also a significant effect for
+% the \Hc{} model (\HcThetaAccuracyCorr{}) and the \Bq{} model (equal
+% prior: \BqThetaAccuracyCorr{}, unequal prior:
+% \BqpThetaAccuracyCorr{}). The bottom row of
+% Fig.~\ref{fig:response-time-accuracy} shows the effect of angle on
+% accuracy for both participants and models.
 
-There was a moderate correlation between the accuracy of the \Hc{}
-model and human accuracy (\ExpHcAccuracyCorr{}). The \Bq{} model had a
-similar correlation (equal prior: \ExpBqAccuracyCorr{}, unequal prior:
-\ExpBqpAccuracyCorr{}). Correlations for the other models were
-undefined, because they had 100\% accuracy.
+% There was a moderate correlation between the accuracy of the \Hc{}
+% model and human accuracy (\ExpHcAccuracyCorr{}). The \Bq{} model had a
+% similar correlation (equal prior: \ExpBqAccuracyCorr{}, unequal prior:
+% \ExpBqpAccuracyCorr{}). Correlations for the other models were
+% undefined, because they had 100\% accuracy.
 
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/response_time_accuracy.pdf}
-    \caption{\textbf{Response time and accuracy comparison.} Top:
-      response time of correct responses as a function of the minimum
-      angle of rotation. Bottom: accuracy as a function of the minumum
-      angle of rotation. All error bars are 95\% confidence
-      intervals.}
-    \label{fig:response-time-accuracy}
+    \subfloat{\includegraphics[width=0.33\textwidth]{../../figures/D/theta_time_corrs.pdf}}
+    \subfloat{\includegraphics[width=0.33\textwidth]{../../figures/D/theta_accuracy_corrs.pdf}}
+    \subfloat{\includegraphics[width=0.33\textwidth]{../../figures/D/response_time_corrs.pdf}}
+    \caption{}
+    \label{fig:response-time-corrs}
   \end{center}
@@
 We replicated the previous monotonic response time result, though not
-the linear resulte. Actually kind of similar to \cite{Gardony:2013gn}
+the linear result. Actually kind of similar to \cite{Gardony:2013gn}
 \TODO{elaborate}

### 2014-01-31 analysis/analyses/accuracy_corrs.py
Add latex to accuracy and response time corrs

@@ -5,6 +5,7 @@ import numpy as np
 import util
 
 filename = "accuracy_corrs.csv"
+texname = "accuracy_corrs.tex"
 
 
 def run(data, results_path, seed):
@@ -38,6 +39,17 @@ def run(data, results_path, seed):
         results.index, names=['model', 'flipped'])
     pth = results_path.joinpath(filename)
     results.to_csv(pth)
+
+    with open(results_path.joinpath(texname), "w") as fh:
+        fh.write("%% AUTOMATICALLY GENERATED -- DO NOT EDIT!\n")
+        for (model, flipped), stats in results.iterrows():
+            if flipped != 'all':
+                continue
+            cmd = util.newcommand(
+                "%sAccuracyCorr" % model.capitalize(),
+                util.latex_pearson.format(**dict(stats)))
+            fh.write(cmd)
+
     return pth
 
 

### 2014-01-31 analysis/analyses/response_time_corrs.py

@@ -5,6 +5,7 @@ import numpy as np
 import util
 
 filename = "response_time_corrs.csv"
+texname = "response_time_corrs.tex"
 
 
 def run(data, results_path, seed):
@@ -38,6 +39,17 @@ def run(data, results_path, seed):
         results.index, names=['model', 'flipped'])
     pth = results_path.joinpath(filename)
     results.to_csv(pth)
+
+    with open(results_path.joinpath(texname), "w") as fh:
+        fh.write("%% AUTOMATICALLY GENERATED -- DO NOT EDIT!\n")
+        for (model, flipped), stats in results.iterrows():
+            if flipped != 'all':
+                continue
+            cmd = util.newcommand(
+                "%sTimeCorr" % model.capitalize(),
+                util.latex_pearson.format(**dict(stats)))
+            fh.write(cmd)
+
     return pth
 
 

### 2014-01-31 analysis/plots/__init__.py
Plot formatting

@@ -12,6 +12,8 @@ import theta_time_corrs
 import theta_accuracy_corrs
 import response_time_corrs
 import accuracy_corrs
+import response_time_stimulus
+import human_model_corrs
 
 __all__ = [
     'response_time_histograms',
@@ -26,4 +28,6 @@ __all__ = [
     'theta_accuracy_corrs',
     'response_time_corrs',
     'accuracy_corrs',
+    'response_time_stimulus',
+    'human_model_corrs'
 ]

### 2014-01-31 analysis/plots/human_model_corrs.py

@@ -8,21 +8,21 @@ from path import path
 
 
 def plot(results_path, fig_path):
-    keys = ['exp', 'oc', 'th', 'hc', 'bq', 'bqp']
-    corrs = pd.read_csv(results_path.joinpath("theta_time_corrs.csv"))\
-              .set_index(['model', 'flipped'])\
-              .unstack('flipped')\
-              .reindex(keys)
+    keys = ['oc', 'th', 'hc', 'bq', 'bqp']
+
+    time = pd.read_csv(results_path.joinpath("response_time_corrs.csv"))\
+             .set_index(['model', 'flipped'])\
+             .unstack('flipped')\
+             .reindex(keys)
+    acc = pd.read_csv(results_path.joinpath("accuracy_corrs.csv"))\
+            .set_index(['model', 'flipped'])\
+            .unstack('flipped')\
+            .reindex(keys)
 
     fig, ax = plt.subplots()
-    width = 1
+    width = 0.8
     offset = width * 5 / 2.
-    colors = {
-        'same': '#ff5555',
-        'flipped': '#5555ff'
-    }
     titles = {
-        'exp': "Human",
         'oc': "Oracle",
         'th': "Threshold",
         'hc': "HC",
@@ -30,21 +30,35 @@ def plot(results_path, fig_path):
         'bqp': "BQ\n(unequal)"
     }
 
-    order = ['same', 'flipped']
-    for i, flipped in enumerate(order):
-        y = corrs.xs(flipped, axis=1, level='flipped')
-        median = y['median']
-        lerr = median - y['lower']
-        uerr = y['upper'] - median
-        ax.bar(
-            i * width + np.arange(len(median)) * offset,
-            median,
-            yerr=[lerr, uerr],
-            color=colors[flipped],
-            ecolor='k',
-            width=width,
-            edgecolor='none',
-            capsize=0)
+    colors = ['#55cc77', '#cc55aa']
+
+    y = time.xs('all', axis=1, level='flipped')
+    median = y['median']
+    lerr = median - y['lower']
+    uerr = y['upper'] - median
+    ax.bar(
+        np.arange(len(median)) * offset,
+        median,
+        yerr=[lerr, uerr],
+        color=colors[0],
+        ecolor='k',
+        width=width,
+        edgecolor='none',
+        capsize=0)
+
[…]

### 2014-01-31 analysis/plots/response_time_corrs.py

@@ -15,12 +15,8 @@ def plot(results_path, fig_path):
               .reindex(keys)
 
     fig, ax = plt.subplots()
-    width = 1
-    offset = width * 5 / 2.
-    colors = {
-        'same': '#ff5555',
-        'flipped': '#5555ff'
-    }
+    width = 0.8
+
     titles = {
         'oc': "Oracle",
         'th': "Threshold",
@@ -29,50 +25,31 @@ def plot(results_path, fig_path):
         'bqp': "BQ\n(unequal)"
     }
 
-    order = ['same', 'flipped']
-    for i, flipped in enumerate(order):
-        y = corrs.xs(flipped, axis=1, level='flipped')
-        median = y['median']
-        lerr = median - y['lower']
-        uerr = y['upper'] - median
-        ax.bar(
-            i * width + np.arange(len(median)) * offset,
-            median,
-            yerr=[lerr, uerr],
-            color=colors[flipped],
-            ecolor='k',
-            width=width,
-            edgecolor='none',
-            capsize=0)
+    y = corrs.xs('all', axis=1, level='flipped')
+    median = y['median']
+    lerr = median - y['lower']
+    uerr = y['upper'] - median
+    ax.bar(
+        np.arange(len(median)),
+        median,
+        yerr=[lerr, uerr],
+        color='0.5',
+        ecolor='k',
+        width=width,
+        edgecolor='none',
+        capsize=0)
 
     ax.set_ylim(0, 1)
-    ax.set_xlim(-width / 2., len(median) * offset)
-    ax.set_xticks(np.arange(len(median)) * offset + width)
+    ax.set_xlim(-width / 2., len(median))
+    ax.set_xticks(np.arange(len(median)) + (width / 2.))
     ax.set_xticklabels([titles[key] for key in keys], fontsize=10)
 
     util.clear_right(ax)
     util.clear_top(ax)
     util.outward_ticks(ax)
 
-    ax.set_title("Correlation with human response time", fontsize=14)
+    ax.set_title("Human vs. model response time", fontsize=14)
     ax.set_ylabel(r"Pearson correlation ($r$)", fontsize=14)
-
-    p0 = plt.Rectangle(
-        (0, 0), 1, 1,
-        fc=colors['same'],
-        ec=colors['same'])
-    p1 = plt.Rectangle(
-        (0, 0), 1, 1,
-        fc=colors['flipped'],
[…]

### 2014-01-31 analysis/plots/response_time_stimulus.py

@@ -8,7 +8,7 @@ from path import path
 
 
 def plot(results_path, fig_path):
-    order = ['exp', 'oc', 'th', 'hc', 'bq', 'bqp']
+    order = ['exp', 'th', 'bqp']
     titles = {
         'exp': "Human",
         'oc': "Oracle",
@@ -23,17 +23,17 @@ def plot(results_path, fig_path):
         'flipped': 'b'
     }
 
-    time_results = pd.read_csv(
-        results_path.joinpath("theta_time.csv"))
-    acc_results = pd.read_csv(
-        results_path.joinpath("theta_accuracy.csv"))
+    means = pd.read_csv(
+        results_path.joinpath("response_time_means.csv"))\
+        .set_index(['stimulus', 'flipped', 'model'])\
+        .groupby(level='stimulus').get_group(2)
 
-    fig, axes = plt.subplots(2, len(order), sharex=True)
+    fig, axes = plt.subplots(1, len(order), sharex=True)
     for i, key in enumerate(order):
-        ax = axes[0, i]
-        df = time_results.groupby('model').get_group(key)
+        ax = axes[i]
+        df = means.groupby(level='model').get_group(key)
 
-        for flipped, stats in df.groupby('flipped'):
+        for flipped, stats in df.groupby(level='flipped'):
             if key == 'exp':
                 median = stats['median'] / 1000.
                 lower = (stats['median'] - stats['lower']) / 1000.
@@ -51,7 +51,7 @@ def plot(results_path, fig_path):
 
         ax.set_xticks(np.arange(0, 200, 30))
         ax.set_xlim(-10, 190)
-
+        ax.set_xlabel("Rotation")
         ax.set_title(titles[key], fontsize=14)
 
         if key == 'exp':
@@ -59,27 +59,7 @@ def plot(results_path, fig_path):
         else:
             ax.set_ylabel("Number of actions", fontsize=14)
 
-    util.sync_ylims(axes[0, order.index('bq')], axes[0, order.index('bqp')])
-
-    for i, key in enumerate(order):
-        ax = axes[1, i]
-        df = acc_results.groupby('model').get_group(key)
-
-        for flipped, stats in df.groupby('flipped'):
-            lower = stats['median'] - stats['lower']
-            upper = stats['upper'] - stats['median']
-            ax.errorbar(
-                stats['modtheta'], stats['median'],
[…]

### 2014-01-31 analysis/plots/theta_accuracy_corrs.py

@@ -55,7 +55,6 @@ def plot(results_path, fig_path):
     util.clear_top(ax)
     util.outward_ticks(ax)
 
-    ax.set_title("Correlation between rotation and accuracy", fontsize=14)
     ax.set_ylabel(r"Spearman correlation ($r_s$)", fontsize=14)
 
     p0 = plt.Rectangle(
@@ -70,10 +69,14 @@ def plot(results_path, fig_path):
     leg = ax.legend(
         [p0, p1], ["\"same\" pairs", "\"flipped\" pairs"],
         numpoints=1, fontsize=12,
-        loc='upper right')
+        loc='upper right',
+        bbox_to_anchor=(1, 1.05))
     frame = leg.get_frame()
     frame.set_facecolor('0.9')
     frame.set_edgecolor('#FFFFFF')
+
+    fig.set_figwidth(5)
+    fig.set_figheight(3)
 
     plt.draw()
     plt.tight_layout()

### 2014-01-31 man/cogsci-2014/cogsci-2014.tex
Finish first whole draft!

@@
 \usepackage{amsmath, amsthm, amssymb}
-\usepackage{subfig}
+\usepackage{subcaption}
+\usepackage{caption}
 
@@
 \input{../../results/D/trial_time_corrs.tex}
+\input{../../results/D/response_time_corrs.tex}
+\input{../../results/D/accuracy_corrs.tex}
 
@@
 \begin{abstract}
-\TODO{}
+  When people use mental imagery to solve tasks, how do they decide
+  \textit{which} images to generate? To answer this question, we
+  explored how mental simulation should used in the classic
+  psychological task of determining if two images depict the same
+  object in different orientations \citeA{Shepard1971}. Through a
+  rational analysis of mental rotation, we formalized four models and
+  compared them to human performance on a 2D mental rotation
+  task. None of the models were a particularly satisfying explanation
+  of human behavior, leading us to suggest that the question of ``what
+  to simulate?'' is more difficult than has previously been
+  assumed. While only exploratory, this work provides the foundation
+  necessary to begin answering this question.
 
@@
 
-\TODO{say something here}
-
-\subsection{Mechanisms behind mental rotation}
-
 Previous models of mental rotation have, for the most part, focused on
@@
 
-The only model (of which the authors are aware) that has seriously
-attempted to address the decision of \textit{what} to simulate is the
-parallel-process model of mental rotation by \citeA{Funt:1983wn}. He
-determines the direction, axis, and extent of rotation by computing
-the objects' principal axes using moments of inertia. One object is
-then rotated through the target rotation, and is checked against the
-other object for congruency. However, this approach assumes that the
-corresponding points on the two objects can be easily identified,
-which is not necessarily the case.
-
-Perhaps people do not engage in a single trajectory of rotation that
-ends with congruency. Some evidence suggests that people do not
-perform a holistic rotation until the objects are congruent, but
-rather pay attention to certain salient features of the object
-\cite{Just1976,Yuille:1982tx}. A more recent study shows that when
-performing \textit{physical} rotations, people do not rotate until
-congruency is reached, and may even rotate \textit{away} from near
-perfect matches \cite{Gardony:2013gn}. Moreover, the state-of-the-art
-in computer vision suggests that there is more to shape matching than
-simply checking for congruency, particularly when shapes are complex
-or when the shapes might not be exactly the same
-\cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}.
+The only models (of which the authors are aware) that have seriously
+attempted to address the decision of \textit{what} to simulate are
+those by \citeA{Funt:1983wn} and \citeA{Just:1985uu}. In both of these
+models, the axis and direction of rotation are computed prior to
+performing the rotation. One object is then rotated through the target
+rotation, and is checked against the other object for
+congruency. However, this approach assumes that the corresponding
+points on the two objects can be easily identified, which is not
+necessarily the case. Indeed, the state-of-the-art in computer vision
+suggests that there is more to shape matching than simply checking for
+congruency, particularly when shapes are complex or when the shapes
+might not be exactly the same
+\cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}. Additionally, recent
+research shows that when performing \textit{physical} rotations,
+people do not rotate until congruency is reached, and may even rotate
+\textit{away} from near perfect matches \cite{Gardony:2013gn}.
 
@@
 
-If people are not performing single-trajectory rotations, what might
-they be doing? To answer this question, we perform a ``rational
-analysis'' of the problem of mental rotation
+If people are not computing the rotation beforehand, what might they
+be doing? To answer this question, we perform a ``rational analysis''
+of the problem of mental rotation
 \cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
-level, we could say that the \textit{problem} is to determine
-something about the spatial transformations that an object has
-undergone. At the algorithmic level, we are constrained by the notion
-that mental images must be transformed in an analog manner (or in a
-way that is at least approximately analog), and that mental images are
-time-consuming and effortful to generate. Thus, the \textit{goal} is
-to make this determination while performing a minimum amount of
-computation (i.e., as few rotations as possible).
+level, we could say that the \textit{problem} is to determine which
+spatial transformations an object has undergone based on two
+depictions of that object which do not include information about point
+correspondences across images. At the algorithmic level, we are
+constrained by the notion that mental images must be transformed in an
+analog manner (or in a way that is at least approximately analog), and
+that mental images are time-consuming and effortful to generate. Thus,
+the \textit{goal} is to make this determination while performing a
+minimum amount of computation (i.e., as few rotations as possible).
 
@@
 smallest amount of computation coincides with rotating through the
-minimum angle. However, it relies on assumptions of \textit{a priori}
-knowledge (either of the minimum angle or the correspondence of points
-between shapes). Noting that a rational solution need not necessarily
-maintain a single trajectory of rotation, we explore an alternative
-rational model, which--rather than assuming the angle of rotation
-\textit{a priori}--engages in an ``active sampling'' strategy.
+minimum angle. However, it violates the constraint that we do not know
+points correspondences, which is exactly what necessitates the use of
+imagery. Noting that a rational solution need not necessarily maintain
+a single trajectory of rotation, we explore an alternative rational
+model, which--rather than computing the angle of rotation ahead of
+time--engages in an ``active sampling'' strategy.
 
@@
 to determine whether the two images depict the same shape, or
-mirror-image (flipped) shapes.
+mirror-image shapes.
 
@@
 
+\begin{figure*}[t]
+  \begin{center}
+    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
+    \caption{\textbf{Example stimuli and model behavior.} On the left
+      are two stimulus pairs, one ``same'' (top) and one ``flipped''
+      (bottom), both with a rotation of $120^\prime$. The grey plots
+      show the actions that each model took for these stimuli. The
+      color of each point (action) indicates whether the model's
+      current mental image was generated under the hypothesis that the
+      objects are the same (red) or flipped (blue). The similarity
+      functions for each of these hypotheses are shown on the right.}
+    \label{fig:model-traces}
+  \end{center}
+\end{figure*}
+
 We represent a shape of $N$ vertices with a $N\times 2$ coordinate
@@
 We do not know which vertices of $\Xb$ correspond to which vertices of
-$\Xa$, so $S$ must marginalize over the set of possible mappings,
-$\mathbb{M}$. For brevity, let $\mathbf{X}_m=\M\cdot{}f(\Xa, h,
-\theta)$ where $\M\in\mathbb{M}$ is a permutation matrix. Then:
+$\Xa$, so $S$ must marginalize over the set of possible mappings. For
+brevity, let $\mathbf{X}_m=\M\cdot{}f(\Xa, h, \theta)$ where $\M$ is a
+permutation matrix. Then:
 \begin{equation}
-  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{|\mathbb{M}|} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
+  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{2N} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
   \label{eq:similarity}
 \end{equation}
-where $\sigma_S=0.15$ is the standard deviation of the similarity.
+where $N$ is the number of vertices, $2N$ is the total number of
+possible mappings,\footnote{It is $2N$ and not $N^2$ because, in polar
+  coordinates, vertices are always connected to their two nearest
+  neighbors in the $\theta$ dimension.}and $\sigma_S=0.15$ is the
+standard deviation of the similarity.
 
@@
 
-\begin{figure*}[t]
-  \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
-    \caption{\textbf{Example stimuli and model behavior.} On the left
-      are two stimulus pairs, one ``same'' (top) and one ``flipped''
-      (bottom), both with a rotation of $120^\prime$. The grey plots
-      show the actions that each model took for these stimuli. The
-      color of each point (action) indicates whether the model's
-      current mental image was generated under the hypothesis that the
-      objects are the same (red) or flipped (blue). The similarity
-      functions for each of these hypotheses are shown on the right.}
-    \label{fig:model-traces}
-  \end{center}
-\end{figure*}
-
 In order to approximate Eq.~\ref{eq:lh-ratio} using samples of the
@@
 computed beforehand \cite{Funt:1983wn}.  To reflect this hypothesis,
-we created an ``oracle'' model which first computes the correct
-rotation and then simply takes actions in the shortest direction until
-it reaches the point where the shapes are aligned.
+we created an ``oracle'' model which is told which points on each
+shape correspond. The \Oc{} model first computes the correct rotation
+and then simply takes actions in the shortest direction until it
+reaches the point where the shapes are aligned.
 
@@
 \subsubsection{Hill climbing model}
+
+\begin{figure*}[t]
+  \begin{center}
+    \includegraphics[width=\textwidth]{../../figures/D/response_time_accuracy.pdf}
+    \caption{\textbf{Response time and accuracy comparison.} Top:
+      response time of correct responses as a function of the minimum
+      angle of rotation. Bottom: accuracy as a function of the minimum
+      angle of rotation. All error bars are 95\% confidence
+      intervals.}
+    \label{fig:response-time-accuracy}
+  \end{center}
+\end{figure*}
 
@@
   utilize a periodic kernel for the covariance
-  \cite{Rasmussen:2006vz}, i.e. $k(\theta,
-  \theta^\prime)=h^2\exp[-\frac{2}{w^2}\sin^2(\frac{1}{2}(\theta-\theta^\prime))]$.}
-Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$, we first place a GP prior on
-the log of $S_h$ in order to enforce positivity after it is
-exponentiated:
+  \cite{Rasmussen:2006vz}.}  Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$,
+we first place a GP prior on the log of $S_h$ in order to enforce
+positivity after it is exponentiated:
 \begin{equation}
@@
 
-\begin{figure}[t]
-  \centering
-  \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_accuracy.pdf}}%
-  \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_time.pdf}}
-  \caption{\textbf{Effect of learning.} There was a clear effect of
-    learning on participants' responses, especially in the first half
-    of the experiment. Left: accuracy as a function of trial. Right:
-    response time in milliseconds as a function of trial.}
-  \label{fig:learning}
-\end{figure}
+% \begin{figure}[t]
+%   \centering
+%   \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_accuracy.pdf}}%
+%   \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_time.pdf}}
+%   \caption{\textbf{Effect of learning.} There was a clear effect of
+%     learning on participants' responses, especially in the first half
+%     of the experiment. Left: accuracy as a function of trial. Right:
+%     response time in milliseconds as a function of trial.}
+%   \label{fig:learning}
+% \end{figure}
 
@@
 
-\subsection{Stimuli}
+\subsubsection{Stimuli}
 
@@
 
-\subsection{Participants and Design}
+\subsubsection{Participants and Design}
 
@@
 above. There were 720 unique experimental stimuli (20 shapes $\times$
-18 angles $\times$ 2 same/flipped), though because stimuli with
+18 angles $\times$ 2 reflections), though because stimuli with
 rotations of $0^\prime$ or $180^\prime$ were repeated twice, there
@@
 four blocks of 200 trials. Within each block, each shape was repeated
-10 times and each rotation was repeated 10 times (5 same, 5 flipped),
-such that across all blocks, each stimulus appeared exactly once. Each
-block was then split in half. Participants completed two half-blocks
-(not necessarily both from the original full block).
+10 times and each rotation was repeated 10 times (5 ``same'', 5
+``flipped''), such that across all blocks, each stimulus appeared
+exactly once. Each block was then split in half. Participants
+completed two half-blocks (not necessarily both from the original full
+block).
 
-\subsection{Procedure}
+\subsubsection{Procedure}
+
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[b]{0.32\textwidth}
+    \includegraphics[width=\textwidth]{../../figures/D/theta_time_corrs.pdf}
+    \caption{\textbf{Angle vs. response time}}
+    \label{fig:theta-time-corrs}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.32\textwidth}
+    \includegraphics[width=\textwidth]{../../figures/D/theta_accuracy_corrs.pdf}
+    \caption{\textbf{Angle vs. accuracy}}
+    \label{fig:theta-accuracy-corrs}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.32\textwidth}
+    \includegraphics[width=\textwidth]{../../figures/D/human_model_corrs.pdf}
+    \caption{\textbf{Human vs. model}}
+    \label{fig:human-model-corrs}
+  \end{subfigure}
+  \vspace{10pt}
+  \caption{\textbf{Correlations.} (a) Spearman correlations between
+    the minimum angle of rotation and response time. (b) Spearman
+    correlations between the minimum angle of rotation and
+    accuracy. The \Oc{} and \Th{} models had 100\% accuracy;
+    therefore, their correlations are undefined. (c) Pearson
+    correlations between human and model, both for response time and
+    accuracy. All error bars are 95\% confidence intervals.}
+\end{figure*}
 
@@
 side-by-side, each at 300px $\times$ 300px, and participants could
-press `s' to indicate they thought the images depicted the same
-object, or `d' to indicate they thought the images depicted flipped
-objects.
+press `s' to indicate they thought the images depicted the ``same''
+object, or `d' to indicate they thought the images depicted
+``flipped'' objects.
 
@@
 
-\subsection{Analysis}
+\section{Results}
 
-\begin{figure*}[t]
-  \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/response_time_accuracy.pdf}
-    \caption{\textbf{Response time and accuracy comparison.} Top:
-      response time of correct responses as a function of the minimum
-      angle of rotation. Bottom: accuracy as a function of the minimum
-      angle of rotation. All error bars are 95\% confidence
-      intervals.}
-    \label{fig:response-time-accuracy}
-  \end{center}
-\end{figure*}
+Out of the 247 participants, 200 (81\%) were included in our
+analyses. Of the rest, we excluded 10 (4\%) because of an experimental
+error, 6 (2.4\%) because they had already completed a previous version
+of the experiment, and 31 (12.6\%) because they failed a comprehension
+check. A participant was marked as passing the comprehension check if
+they answered at least 85\% of ``easy'' trials correctly. ``Easy''
+trials were defined to be those for which the presented shapes
+differed by a rotation of no more than 20 degrees. Additionally, we
+excluded 82 trials for which the response time was either less than
+100 milliseconds or greater than 20 seconds.
+
+For each model, we ran 50 samples for each stimulus that participants
+viewed (with the exception of stimuli with rotations of $0^\prime$ or
+$180^\prime$, in which case we ran double the number of samples). The
+step size parameter ($\sigma_\epsilon$) was approximately fit to each
+of the models, resulting in $\sigma_\epsilon=0.6$ for the \Bq{} model,
+and $\sigma_\epsilon=0.1$ for all other models. We also ran the models
+under two different priors, $p(h=0)=0.5$ (the ``equal'' prior) and
+$p(h=0)=0.55$ (the ``unequal'' prior), though this only had an effect
+for the \Bq{} model. \TODO{actually, did it affect HC?  check this}
+
+\subsubsection{General analysis}
 
@@
 
-\section{Results}
+\subsubsection{Human}
 
-Out of the 247 participants, 200 (81\%) were included in our
-analyses. Of the rest, we excluded 10 (4\%) because of an experimental
-error, 6 (2.4\%) because they had already completed a previous version
-of the experiment, and 31 (12.6\%) because they failed a comprehension
-check. A participant was marked as passing the comprehension check if
-they answered at least 85\% of ``easy'' trials correctly. ``Easy''
-trials were defined to be those for which the presented shapes
-differed by a rotation of no more than 20 degrees. Additionally, we
-excluded 82 individual trials (out of 40000) for which the response
-time was either less than 100 milliseconds or greater than 20 seconds.
+The average response time across all stimuli was \ExpTime{} seconds.
+As illustrated in Fig.~\ref{fig:theta-time-corrs}, the minimum angle
+of rotation was significantly monotonically correlated with average
+per-stimulus response times, both for ``flipped'' and ``same''
+pairs. While this replicates the general result of previous
+experiments \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are
+not as perfectly linear. However, \citeA{Gardony:2013gn} found a
+constant effect for ``flipped'' pairs; an average of their results and
+the classic results would produce something similar to ours.
 
-For each model, we ran 50 samples for each stimulus that participants
-viewed (with the exception of stimuli with rotations of $0^\prime$ or
-$180^\prime$, in which case we ran double the number of samples). The
-step size parameter ($\sigma_\epsilon$) was approximately fit to each
-of the models, resulting in $\sigma_\epsilon=0.6$ for the \Bq{} model,
-and $\sigma_\epsilon=0.1$ for all other models. We also ran the models
-under two different priors, $p(h=0)=0.5$ (the ``equal'' prior) and
-$p(h=0)=0.55$ (the ``unequal'' prior), though this only had an effect
-for the \Bq{} model. \TODO{actually, did it affect HC?  check this}
-
-\subsection{Practice effects}
+The average accuracy across all stimuli was \ExpAccuracy{}, though
+there were \ExpNumChance{} individual stimuli (out of 720) for which
+people were not above chance.  The minimum angle was also correlated
+with participants' average per-stimulus accuracy, as shown in
+Fig.~\ref{fig:theta-accuracy-corrs}, though significantly more so for
+``same'' pairs. This is the same result found both by
+\citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.
 
@@
 half of the experiment (\ExpaTrialAccuracyCorr{} for the first half
-vs. \ExpbTrialAccuracyCorr{} for the second
-half). Fig.~\ref{fig:learning} shows scatterplots illustrating these
-trends.
+vs. \ExpbTrialAccuracyCorr{} for the second half). These practice
+effects may have contributed to the not-quite-linearity of the human
+response times; future work will need to collect more data per
+participant to mitigate this effect.
 
-\subsection{Response times}
+\subsubsection{\Oc{} model}
 
-% The average response time across all stimuli was \ExpTime{} seconds.
-% As expected, the minimum angle of rotation was significantly
-% monotonically correlated with average per-stimulus response times,
-% both for flipped pairs (\ExpThetaTimeCorrFlipped{}) and same pairs
-% (\ExpThetaTimeCorrSame{}). This was also the case for the number of
-% actions taken by the \Oc{} model (\OcThetaTimeCorr{}), the \Th{} model
-% (\ThThetaTimeCorr{}), and the \Bq{} model (\BqThetaTimeCorr{}), but
-% not the \Hc{} model (\HcThetaTimeCorr{}). The top row of
-% Fig.~\ref{fig:response-time-accuracy} shows the overall effect of
-% rotation on response times.
+As we would expect, the number of actions taken by the \Oc{} model was
+perfectly correlated with the minimum angle of rotation
+(Fig.~\ref{fig:theta-time-corrs}), both for ``same'' and ``flipped''
+pairs. Unsurprisingly, then, \Oc{} model was the best fit to human
+response times, with a correlation of \OcTimeCorr{}
+(Fig.~\ref{fig:human-model-corrs}).  However, the \Oc{} model was
+100\% accurate, and therefore the correlation between minimum angle
+and accuracy was undefined (Fig.~\ref{fig:theta-accuracy-corrs}).
 
-% The number of actions taken by the \Oc{} model was the best predictor
-% of human response times, with a correlation of \ExpOcTimeCorr{}. The
-% \Th{} model \Bq{} model with $p(h=0)=0.55$ were the next best (for
-% \Th{}, \ExpThTimeCorr{}; for \Bq{}, \ExpBqpTimeCorr{}). Not far behind
-% was the BQ model with equal prior (\ExpBqTimeCorr{}). The \Hc{} model
-% was barely correlated with people (\ExpHcTimeCorr{}).
+\subsubsection{\Th{} model}
 
-% Across all stimuli, participants' response times for ``flipped'' pairs
-% were significantly higher than those for ``same'' pairs
-% \TODO{confidence intervals} \TODO{flipped response times higher than
-%   same response times for smaller angles?}
+The number of actions taken by the \Th{} model was about as correlated
+with the minimum angle as human response times were
+(Fig.~\ref{fig:theta-time-corrs}), though this effect was less strong
+for individual shapes (e.g.,
+Fig.~\ref{fig:response-time-stimulus}). Like the \Oc{} model, the
+\Th{} model had 100\% accuracy, and thus there was no correlation
+between minimum angle and accuracy. The \Th{} was able to explain a
+moderate amount of the variance in human response times, with a
+correlation of \ThTimeCorr{} (Fig.~\ref{fig:human-model-corrs}).
 
-\subsection{Accuracy}
+\subsubsection{\Hc{} model}
 
-% The average accuracy across all stimuli was \ExpAccuracy{}, though
-% there were \ExpNumChance{} individual stimuli (out of 720) for which
-% people were not above chance. Of the models, the \Oc{} and \Th{}
-% models performed perfectly, with 100\% accuracy. The \Bq{} model was
-% very accurate overall (equal prior: \BqAccuracy{}, unequal prior:
-% \BqpAccuracy{}). With the equal prior, there was \BqNumChance{}
-% stimulus for which it was not above chance; with the unequal prior,
-% there were \BqpNumChance{}. The \Hc{} model was not particularly
-% accurate, and although it did perform above chance overall
-% (\HcAccuracy{}), there were \HcNumChance{} individual stimuli for
-% which it was not above chance.
+\begin{figure}[t]
+  \begin{center}
+    \includegraphics[width=0.48\textwidth]{../../figures/D/response_time_stimulus.pdf}
+    \caption{\textbf{Typical response time curves for a single shape.}
+      These plots correspond to the shape shown in
+      Fig.~\ref{fig:model-traces}. Left: human curves are either
+      linear (as with the ``same'' pairs), or linear and then flat (as
+      with the ``flipped'' pairs). Middle: the \Th{} model tends to
+      have highly inconsitent within-shape response times. Right: the
+      \Bq{} model is roughly linear.}
+    \label{fig:response-time-stimulus}
+  \end{center}
+\end{figure}
 
-% The minimum angle was also correlated with participants' average
-% per-stimulus accuracy: for flipped pairs,
-% \ExpThetaAccuracyCorrFlipped{}, and for same pairs,
-% \ExpThetaAccuracyCorrSame{}. There was also a significant effect for
-% the \Hc{} model (\HcThetaAccuracyCorr{}) and the \Bq{} model (equal
-% prior: \BqThetaAccuracyCorr{}, unequal prior:
-% \BqpThetaAccuracyCorr{}). The bottom row of
-% Fig.~\ref{fig:response-time-accuracy} shows the effect of angle on
-% accuracy for both participants and models.
+The \Hc{} was the only model for which there was no effect of rotation
+on response time (Fig.~\ref{fig:theta-time-corrs}). Although there was
+a significant effect of rotation on accuracy
+(Fig.~\ref{fig:theta-accuracy-corrs}), the \Hc{} model was barely
+above chance (\HcAccuracy{}) and there were \HcNumChance{} individual
+stimuli for which it was not above chance. The \Hc{} model was not a
+particularly good predictor of human response times (\HcTimeCorr{}),
+though it was a moderate predictor of accuracy (\HcAccuracyCorr{}), as
+shown in Fig.~\ref{fig:human-model-corrs}.
 
-% There was a moderate correlation between the accuracy of the \Hc{}
-% model and human accuracy (\ExpHcAccuracyCorr{}). The \Bq{} model had a
-% similar correlation (equal prior: \ExpBqAccuracyCorr{}, unequal prior:
-% \ExpBqpAccuracyCorr{}). Correlations for the other models were
-% undefined, because they had 100\% accuracy.
+\subsubsection{\Bq{} model}
 
-\begin{figure*}[t]
-  \begin{center}
-    \subfloat{\includegraphics[width=0.33\textwidth]{../../figures/D/theta_time_corrs.pdf}}
-    \subfloat{\includegraphics[width=0.33\textwidth]{../../figures/D/theta_accuracy_corrs.pdf}}
-    \subfloat{\includegraphics[width=0.33\textwidth]{../../figures/D/response_time_corrs.pdf}}
-    \caption{}
-    \label{fig:response-time-corrs}
-  \end{center}
-\end{figure*}
+As with most of the other models, there was a significant effect of
+rotation on the number of steps taken by the \Bq{} model
+(Fig.~\ref{fig:theta-time-corrs}). The \Bq{} model explained variance
+in human response times about as well as the \Th{} model
+(Fig.~\ref{fig:human-model-corrs}), with a correlation of
+\BqTimeCorr{} for the equal prior and \BqpTimeCorr{} for the unequal
+prior.
+
+The \Bq{} model was quite accurate overall, both with the equal prior
+(\BqAccuracy{}) and unequal prior (\BqpAccuracy{}). With the equal
+prior, there was \BqNumChance{} stimulus for which it was not above
+chance; with the unequal prior, there were \BqpNumChance{}. There was
+a significant effect of rotation on the \Bq{} model's accuracy
+(Fig.~\ref{fig:theta-accuracy-corrs}), and itt explained about the
+same amount of variance in people's accuracy as the \Hc{} model: for
+the equal prior, \BqAccuracyCorr{}; for the unequal prior,
+\BqpAccuracyCorr{}.
+
+Because the \Bq{} model relied on Eq.~\ref{eq:lh-ratio} for its
+stopping criteria (as opposed to just finding a maximum), the prior
+$p(h)$ had a fairly significant effect, as shown in
+Fig.~\ref{fig:response-time-accuracy}. With just a small bias of
+$p(h=0)=0.55$, there is a clear separation in response times that is
+similar to the separation seen in human response times. The prior also
+had an effect on accuracy, though this effect did not reflect the
+trends in human behavior.
 
@@
 
-We replicated the previous monotonic response time result, though not
-the linear result. Actually kind of similar to \cite{Gardony:2013gn}
-\TODO{elaborate}
+We set out to answer the question: how do people decide \textit{what}
+to simulate when using mental imagery? Focusing on the specific case
+of determining the direction and extent of mental rotation, we
+formalized four models and compared their performance with the results
+of a behavioral experiment. The \Oc{} model was the best fit to human
+response times, but failed to capture anything regarding accuracy. The
+\Th{} model was a slightly worse fit to human response times, and also
+did not capture any trends in accuracy. The \Bq{} model fit human
+reponse time data as well as the \Th{} model, and captured a small
+amount of the variance in people's accuracy. The \Hc{} model barely
+fit human response time data, though it did fit accuracy as well as
+the \Bq{} model with an equal prior.
 
-Also replicated the constant accuracy for flipped stimuli, but not for
-same stimuli, from \cite{Cooper:1975wp}. Why is this? Maybe the
-strategy is ``they are different until proven the same''.
+While it is clear that none of the models fully capture human behavior
+on the mental rotation task, it is worth noting where they shed light
+on aspects that have previously gone unexplained.  For example, the
+\Bq{} model's adaptive stopping rule is sensitive to the prior, and
+thus provides a possible explanation for why people are slower to
+respond on ``flipped'' stimulus pairs.
 
-None of the models were particularly good fits to human data. The best
-fit was the \Oc{} model, though this is not satisfying because the
-\Oc{} model knows the answer beforehand, which people obviously do
-not. The next best model, the \Th{} model, is also not a
-particularly satisfying account, because it knows the global maximum
-\textit{a priori}, which is not a reasonable assumption. \TODO{explain
-  why it is faster for 180 than mid angles -- it is because for a lot
-  of stimuli, there is rotational symmetry at 180, so it almost always
-  starts checking the correct version first. For mid angles, it might
-  check the wrong version first, meaning it takes much longer.}
+Additionally, a closer look at individual shapes reveals several
+interesting points (Fig.~\ref{fig:response-time-stimulus} shows curves
+for one particular shape; these curves are fairly typical relative to
+other shapes). First, the curves for the human response times on
+specific shapes tend to either be linear, or linear first, then
+flat. Second, while the \Th{} model appears to capture the concavity
+of human response times overall, it is actually fairly inconsistent
+within shapes. Third, the linearity of the \Bq{} model's response time
+curves \textit{is} consistent within shapes, though it does not
+exhibit the same ``plateau'' effect that sometimes appears in the
+human curves.
 
-The \Bq{} model makes more reasonable assumptions about the observer's
-knowledge, but has its own problems. Because it is forced to take
-small steps, it often gets stuck sampling the same region over and
-over again, particularly for stimuli with larger rotations
-(Fig.~\ref{fig:model-traces}). As a result, its response time curve is
-convex rather than concave.
+How could the models we evaluated be improved? First, all of the
+models use holistic rotations, however there is evidence that people
+might compare individual features of shapes rather than the entire
+shape at once \cite{Just1976,Yuille:1982tx}. Second, there are other
+possible ``active sampling'' approaches that could combine the best
+features from the \Th{} and \Bq{} models, such as maintaining a
+distribution over the location and value of the global maximum, rather
+than over the integral. Third, studying rotations with 3D objects of
+varying levels of complexity may be more informative than the
+extremely simple 2D objects used in this experiment.
 
-Other problems: evidence that rotations aren't actually holistic
-\cite{Yuille:1982tx}, eye movements \cite{Just1976}
-
-\TODO{talk about how these results suggest a model of mental rotation
-  is harder than was previous thought, and that this is a first step
-  towards detangling these issues.}
+As a parting comment, we note that the question of how people decide
+\textit{which} rotations to perform is more difficult to answer than
+previously assumed. Although we examined four well-motivated models,
+none came close to explaining the full range of human behavior on this
+task. We suggest that such metacognitive questions are fertile areas
+for future research.
 

### 2014-02-01 man/cogsci-2014/cogsci-2014.tex
Tighten wording

@@
-% 
+ 
 % Annual Cognitive Science Conference
@@
 points on the two objects can be easily identified, which is not
-necessarily the case. Indeed, the state-of-the-art in computer vision
-suggests that there is more to shape matching than simply checking for
-congruency, particularly when shapes are complex or when the shapes
-might not be exactly the same
+necessarily the case.
+
+Indeed, the state-of-the-art in computer vision suggests that there is
+more to shape matching than simply checking for congruency,
+particularly when shapes are complex or when the shapes might not be
+exactly the same
 \cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}. Additionally, recent
@@
 
-The original congruency hypothesis \cite{Shepard1971} actually is a
-rational solution which achieves this goal, in the sense that the
+The original ``congruency'' hypothesis \cite{Shepard1971} actually is
+a rational solution which achieves this goal, in the sense that the
 smallest amount of computation coincides with rotating through the
@@
 
-In the following section, we formalize this rational analysis, and
-propose four models of mental rotation: one based on the original
-hypothesis by \citeA{Shepard1971}; two which are extensions of the
-first but with relaxed assumptions; and one based on the active
-sampling approach.
-
 \section{Modeling mental rotation}
 
-In the mental rotation task, people are presented with pairs of
-two-dimensional shapes similar to those used by \citeA{Cooper:1975wp};
-example stimuli are shown in Fig.~\ref{fig:model-traces}. The goal is
-to determine whether the two images depict the same shape, or
-mirror-image shapes.
+In this section, we formalize our rational analysis, and propose four
+models of mental rotation: one based on existing models; two which are
+extensions of the first but with relaxed assumptions; and one based on
+the active sampling approach.
 
@@
 We define $p(\Xb\ \vert\ \Xa, \theta, h)$ to be the similarity between
-$\Xb$ and a transformation of $\Xa$:
-\begin{equation}
-  p(\Xb\ \vert\ \Xa, \theta, h):= S(\Xb, f(\Xa, h, \theta)).
-  \label{eq:likelihood}
-\end{equation}
-Example similarity curves are shown in Fig.~\ref{fig:model-traces}.
-
-We do not know which vertices of $\Xb$ correspond to which vertices of
-$\Xa$, so $S$ must marginalize over the set of possible mappings. For
-brevity, let $\mathbf{X}_m=\M\cdot{}f(\Xa, h, \theta)$ where $\M$ is a
-permutation matrix. Then:
+$\Xb$ and a transformation of $\Xa$: $p(\Xb\ \vert\ \Xa, \theta, h):=
+S(\Xb, f(\Xa, h, \theta))$.  We do not know which vertices of $\Xb$
+correspond to which vertices of $\Xa$, so $S$ must marginalize over
+the set of possible mappings. For brevity, let
+$\mathbf{X}_m=\M\cdot{}f(\Xa, h, \theta)$ where $\M$ is a permutation
+matrix. Then:
 \begin{equation}
@@
 \end{equation}
-where $N$ is the number of vertices, $2N$ is the total number of
-possible mappings,\footnote{It is $2N$ and not $N^2$ because, in polar
-  coordinates, vertices are always connected to their two nearest
-  neighbors in the $\theta$ dimension.}and $\sigma_S=0.15$ is the
-standard deviation of the similarity.
+where $2N$ is the total number of possible mappings,\footnote{It is
+  $2N$ and not $N^2$ because, in polar coordinates, vertices are
+  always connected to their two nearest neighbors in the $\theta$
+  dimension.}and $\sigma_S=0.15$ is the standard deviation of the
+similarity. Example similarity curves are shown in
+Fig.~\ref{fig:model-traces}.
 
@@
   \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
-      f(\Xt, 0, |\epsilon|) &\mbox{ rotate by $\epsilon$,} \\
+      f(\Xt, 0, \epsilon) &\mbox{ continue rotating by $\epsilon$,} \\
       f(\Xt, 1, 0) &\mbox{ flip,} \\
@@
 \end{equation}
-where $\epsilon\sim \mathcal{N}(0, \sigma_\epsilon)$ and
+where $\epsilon\sim \left|\mathcal{N}(0, \sigma_\epsilon)\right|$ and
 $\sigma_\epsilon$ is the standard deviation of the step size in
@@
 
-One hypothesis in the literature is that people perform a mental
-rotation in the shortest direction until a match is found
-\cite{Shepard1971, Cooper:1975wp}, and that determining this direction
-is an implementation detail \cite{Anderson1978}, or that it is
-computed beforehand \cite{Funt:1983wn}.  To reflect this hypothesis,
-we created an ``oracle'' model which is told which points on each
-shape correspond. The \Oc{} model first computes the correct rotation
-and then simply takes actions in the shortest direction until it
-reaches the point where the shapes are aligned.
+One hypothesis in the literature is that people compute the direction
+and extent of rotation beforehand using \textit{a prior} knowledge of
+the correspondence between points in the images
+\cite{Funt:1983wn,Just:1985uu}.  To reflect this hypothesis, we
+created an ``oracle'' model which is told which points on each shape
+correspond. From that correspondence, it computes the correct rotation
+and rotates through it.
 
@@
 
-In the current formulation of the problem, choosing the threshold is
-straightforward because we know both the exact geometry of the shapes
-and that a linear transformation exists which will perfectly align
-them. However, it is not always clear how to choose a reasonable
-threshold \textit{a priori}, as the global optimum will depend on many
-factors (such as shape complexity, dimensionality, perceptual
-uncertainty, and whether the shapes are exactly identical). Thus, it
-is unlikely that the \Th{} model will be robust beyond this domain.
-
 \subsubsection{Hill climbing model}
@@
 
-One way to deal with the problem of choosing a threshold would simply
-be to perform an exhaustive search or use a global optimization
-strategy. However, these strategies would not result in the linear
-response time found by \citeA{Shepard1971}. A second alternative,
-which could potentially demonstrate the linear response time result,
-might instead just perform a Hill Climbing (\Hc{}) search for a
-\textit{local} maximum. As with the \Oc{} and \Th{} models,
-Eq.~\ref{eq:lh-ratio} becomes a maximum likelihood ratio. Unlike the
-global maximum, however, it is not guaranteed that a local maximum of
-the correct hypothesis will be greater than a local maximum of the
-incorrect hypothesis.
+In the current formulation of the problem, choosing the threshold is
+straightforward because we know both the exact geometry of the shapes
+and that a linear transformation exists which will perfectly align
+them. However, this choice is not always clear \textit{a priori}, as
+the global optimum will depend on many factors (e.g., shape
+complexity, dimensionality, perceptual uncertainty, and whether the
+shapes are identical).  One way to deal with the problem of choosing a
+threshold would simply be to perform an exhaustive search or use a
+global optimization strategy. However, these strategies would not
+result in the linear response time found by \citeA{Shepard1971}. A
+second alternative, which could potentially demonstrate the linear
+response time result, might instead just perform a Hill Climbing
+(\Hc{}) search for a \textit{local} maximum. As with the \Oc{} and
+\Th{} models, Eq.~\ref{eq:lh-ratio} becomes a maximum likelihood
+ratio.
 
@@
 \end{equation}
+
 This is a technique known in the machine-learning literature as
@@
 use a Gaussian Process (GP) prior for $p(S)$, as in
-\citeA{Osborne:2012tm}.\footnote{Because our domain is on a circle, we
-  utilize a periodic kernel for the covariance
-  \cite{Rasmussen:2006vz}.}  Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$,
-we first place a GP prior on the log of $S_h$ in order to enforce
-positivity after it is exponentiated:
-\begin{equation}
-\mathbb{E}[Z_h] \approx \int \exp(\mu_h(\theta))p(\theta)\dif\theta,
-\end{equation}
-where $\mu_h:=\mu(\log S_h)$ is the mean of the log-GP. However, this
-integral is still intractable. To approximate it, we fit a second GP
-over points sampled from the log-GP, which we denote as
+\citeA{Osborne:2012tm}.  Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$, we
+first place a GP prior on the log of $S_h$ in order to enforce
+positivity after it is exponentiated, i.e. $\mathbb{E}[Z_h] \approx
+\int \exp(\mu_h(\theta))p(\theta)\dif\theta$, where $\mu_h:=\mu(\log
+S_h)$ is the mean of the log-GP.  However, this integral is still
+intractable. To approximate it, we fit a second GP over points sampled
+from the log-GP, which we denote as
 $\bar{S}_h:=\exp(\mu_h)$.\footnote{Note that the approach used here
@@
 just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. Thus, rather than
-computing $p(\ell)$, we use $Z_D=Z_0-Z_1$:
-\begin{equation}
-  p(Z_D)\propto\mathcal{N}(\mathbb{E}[Z_0] - \mathbb{E}[Z_1], \mathbb{V}(Z_0) + \mathbb{V}(Z_1)).
-\end{equation}
+computing $p(\ell)$, we use $Z_D=Z_0-Z_1$ and compute:
+$p(Z_D)\propto\mathcal{N}(\mathbb{E}[Z_0] - \mathbb{E}[Z_1],
+\mathbb{V}(Z_0) + \mathbb{V}(Z_1))$.
+
 We then sample new observations until we are at least 95\% confident
@@
 of 10000 bootstrap samples to compute the confidence intervals around
-both Spearman ($r_s$) and Pearson ($r$) correlations.
-
-For analyses of accuracy, confidence intervals were computed from a
-binomial proportion with a Jeffrey's beta prior.  To test if
-participants or a model was above chance on a particular stimulus, we
-used the same binomial proportion, and simply tested whether $p>0.05$
-for the 50\% mark.
+both Spearman ($r_s$) and Pearson ($r$) correlations.  For analyses of
+accuracy, confidence intervals were computed from a binomial
+proportion with a Jeffrey's beta prior.  To test if participants or a
+model was above chance on a particular stimulus, we used the same
+binomial proportion, and simply tested whether $p>0.05$ for the 50\%
+mark.
 
@@
 formalized four models and compared their performance with the results
-of a behavioral experiment. The \Oc{} model was the best fit to human
-response times, but failed to capture anything regarding accuracy. The
-\Th{} model was a slightly worse fit to human response times, and also
-did not capture any trends in accuracy. The \Bq{} model fit human
-reponse time data as well as the \Th{} model, and captured a small
-amount of the variance in people's accuracy. The \Hc{} model barely
-fit human response time data, though it did fit accuracy as well as
-the \Bq{} model with an equal prior.
-
-While it is clear that none of the models fully capture human behavior
-on the mental rotation task, it is worth noting where they shed light
-on aspects that have previously gone unexplained.  For example, the
-\Bq{} model's adaptive stopping rule is sensitive to the prior, and
-thus provides a possible explanation for why people are slower to
-respond on ``flipped'' stimulus pairs.
+of a behavioral experiment. While it is clear that none of the models
+fully captured human behavior on the mental rotation task, it is worth
+noting where they shed light on aspects that have previously gone
+unexplained.  For example, the \Bq{} model's adaptive stopping rule is
+sensitive to the prior, and thus provides a possible explanation for
+why people are slower to respond on ``flipped'' stimulus pairs.
 

### 2014-02-01 analysis/analyses/accuracy_corrs.py
Add human vs human to corrs

@@ -34,6 +34,12 @@ def run(data, results_path, seed):
             method='pearson')
         results[(key, 'all')] = corr
 
+    corr = util.bootcorr(
+        means['expA'],
+        means['expB'],
+        method='pearson')
+    results[('exp', 'all')] = corr
+
     results = pd.DataFrame.from_dict(results, orient='index')
     results.index = pd.MultiIndex.from_tuples(
         results.index, names=['model', 'flipped'])

### 2014-02-01 analysis/analyses/response_time_corrs.py

@@ -21,6 +21,7 @@ def run(data, results_path, seed):
     for key in means:
         if key in exclude:
             continue
+
         for flipped in ['same', 'flipped']:
             corr = util.bootcorr(
                 means['exp'].unstack('flipped')[flipped],
@@ -33,6 +34,12 @@ def run(data, results_path, seed):
             means[key],
             method='pearson')
         results[(key, 'all')] = corr
+
+    corr = util.bootcorr(
+        means['expA'],
+        means['expB'],
+        method='pearson')
+    results[('exp', 'all')] = corr
 
     results = pd.DataFrame.from_dict(results, orient='index')
     results.index = pd.MultiIndex.from_tuples(

### 2014-02-01 analysis/plots/human_model_corrs.py

@@ -8,7 +8,7 @@ from path import path
 
 
 def plot(results_path, fig_path):
-    keys = ['oc', 'th', 'hc', 'bq', 'bqp']
+    keys = ['exp', 'oc', 'th', 'hc', 'bq', 'bqp']
 
     time = pd.read_csv(results_path.joinpath("response_time_corrs.csv"))\
              .set_index(['model', 'flipped'])\
@@ -20,9 +20,10 @@ def plot(results_path, fig_path):
             .reindex(keys)
 
     fig, ax = plt.subplots()
-    width = 0.8
+    width = 1
     offset = width * 5 / 2.
     titles = {
+        'exp': "Human",
         'oc': "Oracle",
         'th': "Threshold",
         'hc': "HC",

### 2014-02-01 man/cogsci-2014/cogsci-2014.tex
Add note about human vs human corrs
Update results, add a few TODOs

@@
 \end{abstract}
+
+\TODO{need to compute better human vs human correlations}
 
@@
     correlations between human and model, both for response time and
-    accuracy. All error bars are 95\% confidence intervals.}
+    accuracy. The first set of bars is human vs. human, to give an
+    upper bound. All error bars are 95\% confidence intervals.}
 \end{figure*}
@@
 $180^\prime$, in which case we ran double the number of samples). The
-step size parameter ($\sigma_\epsilon$) was approximately fit to each
-of the models, resulting in $\sigma_\epsilon=0.6$ for the \Bq{} model,
-and $\sigma_\epsilon=0.1$ for all other models. We also ran the models
-under two different priors, $p(h=0)=0.5$ (the ``equal'' prior) and
-$p(h=0)=0.55$ (the ``unequal'' prior), though this only had an effect
-for the \Bq{} model. \TODO{actually, did it affect HC?  check this}
+step size parameter ($\sigma_\epsilon$) was fit to human data for each
+of the models, resulting in $\sigma_\epsilon=0.6$ for the \Th{} and
+\Bq{} models and $\sigma_\epsilon=0.1$ for the \Oc{} and \Hc{}
+models. We also ran the models under two different priors,
+$p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the ``unequal''
+prior). This had an effect on accuracy for the \Hc{} model, and an
+effect on both accuracy and response time for the \Bq{} model. Below,
+we report results for the \Hc{} model with an unequal prior, and the
+\Bq{} model with both priors.
 
@@
 
-The average response time across all stimuli was \ExpTime{} seconds.
-As illustrated in Fig.~\ref{fig:theta-time-corrs}, the minimum angle
-of rotation was significantly monotonically correlated with average
-per-stimulus response times, both for ``flipped'' and ``same''
-pairs. While this replicates the general result of previous
+The average response time across all stimuli was \ExpTime{}
+milliseconds.  As illustrated in Fig.~\ref{fig:theta-time-corrs}, the
+minimum angle of rotation was significantly monotonically correlated
+with average per-stimulus response times, both for ``flipped'' and
+``same'' pairs. While this replicates the general result of previous
 experiments \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are
@@
 (Fig.~\ref{fig:theta-time-corrs}), both for ``same'' and ``flipped''
-pairs. Unsurprisingly, then, \Oc{} model was the best fit to human
+pairs. Unsurprisingly, the \Oc{} model was the best fit to human
 response times, with a correlation of \OcTimeCorr{}
@@
 correlation of \ThTimeCorr{} (Fig.~\ref{fig:human-model-corrs}).
+
+\TODO{say something about the large step size -- actually produces
+  non-intuitive results, even though they are a better fit. with small
+  step size (where it doesn't over-rotate) it does even worse!}
 
@@
 stimuli for which it was not above chance. The \Hc{} model was not a
-particularly good predictor of human response times (\HcTimeCorr{}),
-though it was a moderate predictor of accuracy (\HcAccuracyCorr{}), as
-shown in Fig.~\ref{fig:human-model-corrs}.
+particularly good predictor of human behavior, neither for response
+times (\HcTimeCorr{}) nor accuracy (\HcAccuracyCorr{}), as shown in
+Fig.~\ref{fig:human-model-corrs}.
 
@@
 a significant effect of rotation on the \Bq{} model's accuracy
-(Fig.~\ref{fig:theta-accuracy-corrs}), and itt explained about the
-same amount of variance in people's accuracy as the \Hc{} model: for
-the equal prior, \BqAccuracyCorr{}; for the unequal prior,
+(Fig.~\ref{fig:theta-accuracy-corrs}), and it explained about the same
+amount of variance in people's accuracy as the \Hc{} model: for the
+equal prior, \BqAccuracyCorr{}; for the unequal prior,
 \BqpAccuracyCorr{}.

### 2014-02-03 analysis/analyses/__init__.py
Add better between-subjects correlation analysis

@@ -16,6 +16,7 @@ import response_time_means
 import accuracy_means
 import trial_time_means
 import trial_accuracy_means
+import human_corrs
 
 __all__ = [
     # across all stimuli
@@ -31,6 +32,7 @@ __all__ = [
 
     # correlations between human and model means for each
     # shape/rotation/reflection
+    'human_corrs',
     'response_time_corrs',
     'accuracy_corrs',
 

### 2014-02-03 analysis/analyses/accuracy_corrs.py

@@ -34,12 +34,6 @@ def run(data, results_path, seed):
             method='pearson')
         results[(key, 'all')] = corr
 
-    corr = util.bootcorr(
-        means['expA'],
-        means['expB'],
-        method='pearson')
-    results[('exp', 'all')] = corr
-
     results = pd.DataFrame.from_dict(results, orient='index')
     results.index = pd.MultiIndex.from_tuples(
         results.index, names=['model', 'flipped'])

### 2014-02-03 analysis/analyses/human_corrs.py

@@ -1,0 +1,81 @@
+#!/usr/bin/env python
+
+import pandas as pd
+import numpy as np
+import util
+import scipy.stats
+
+filename = "human_corrs.csv"
+texname = "human_corrs.tex"
+
+
+def run(data, results_path, seed):
+    np.random.seed(seed)
+
+    exp = data['exp'].set_index(['pid', 'trial'])
+
+    pids = pd.Series(exp.index.get_level_values('pid')).drop_duplicates()
+    pids.sort()
+    pids = pids.reset_index(drop=True)
+
+    n = len(pids)
+    m = 100
+    time_corrs = np.empty(m)
+    acc_corrs = np.empty(m)
+
+    for i in xrange(m):
+        idx = np.arange(n)
+        np.random.shuffle(idx)
+
+        p0 = pids[idx[:int(n / 2)]]
+        p1 = pids[idx[int(n / 2):]]
+
+        df0 = exp.drop(p0, axis=0, level='pid').reset_index()
+        df1 = exp.drop(p1, axis=0, level='pid').reset_index()
+
+        tm0 = df0[df0['correct']]\
+            .groupby(['stimulus', 'modtheta', 'flipped'])['time']\
+            .apply(lambda x: 1. / (1. / x).mean())
+        tm1 = df1[df1['correct']]\
+            .groupby(['stimulus', 'modtheta', 'flipped'])['time']\
+            .apply(lambda x: 1. / (1. / x).mean())
+
+        am0 = df0\
+            .groupby(['stimulus', 'modtheta', 'flipped'])['correct']\
+            .apply(util.beta).unstack(-1)['median']
+        am1 = df1\
+            .groupby(['stimulus', 'modtheta', 'flipped'])['correct']\
+            .apply(util.beta).unstack(-1)['median']
+
+        means = pd.DataFrame({'tm0': tm0, 'tm1': tm1, 'am0': am0, 'am1': am1})
+        time_corrs[i] = scipy.stats.pearsonr(means['tm0'], means['tm1'])[0]
+        acc_corrs[i] = scipy.stats.pearsonr(means['am0'], means['am1'])[0]
+
+    time_stats = np.percentile(time_corrs, [2.5, 50, 97.5])
+    acc_stats = np.percentile(acc_corrs, [2.5, 50, 97.5])
+
+    results = pd.DataFrame({
+        'time': time_stats,
+        'accuracy': acc_stats
+    }, index=['lower', 'median', 'upper']).T.reset_index()
+    results['model'] = 'exp'
[…]

### 2014-02-03 analysis/analyses/response_time_corrs.py

@@ -35,12 +35,6 @@ def run(data, results_path, seed):
             method='pearson')
         results[(key, 'all')] = corr
 
-    corr = util.bootcorr(
-        means['expA'],
-        means['expB'],
-        method='pearson')
-    results[('exp', 'all')] = corr
-
     results = pd.DataFrame.from_dict(results, orient='index')
     results.index = pd.MultiIndex.from_tuples(
         results.index, names=['model', 'flipped'])

### 2014-02-03 analysis/plots/human_model_corrs.py

@@ -10,14 +10,21 @@ from path import path
 def plot(results_path, fig_path):
     keys = ['exp', 'oc', 'th', 'hc', 'bq', 'bqp']
 
-    time = pd.read_csv(results_path.joinpath("response_time_corrs.csv"))\
+    human = pd.read_csv(results_path.joinpath("human_corrs.csv"))\
+              .set_index('measure')
+    human['flipped'] = 'all'
+
+    time = pd.read_csv(results_path.joinpath("response_time_corrs.csv"))
+    time = time.append(human.xs('time'))\
+               .set_index(['model', 'flipped'])\
+               .unstack('flipped')\
+               .reindex(keys)
+
+    acc = pd.read_csv(results_path.joinpath("accuracy_corrs.csv"))
+    acc = acc.append(human.xs('accuracy'))\
              .set_index(['model', 'flipped'])\
              .unstack('flipped')\
              .reindex(keys)
-    acc = pd.read_csv(results_path.joinpath("accuracy_corrs.csv"))\
-            .set_index(['model', 'flipped'])\
-            .unstack('flipped')\
-            .reindex(keys)
 
     fig, ax = plt.subplots()
     width = 1

### 2014-02-03 man/cogsci-2014/cogsci-2014.tex
Update manuscript with new analysis

@@
 
+% space between figure and caption
 \addtolength{\abovecaptionskip}{-0.25cm}
-\addtolength{\belowcaptionskip}{-0.2cm}
-\addtolength{\textfloatsep}{-2mm}	% space below "Figure 1: ..."
+
+% space below "Figure 1: ...", but only for inline figures
+\addtolength{\textfloatsep}{-0.5cm}
 
@@
 \input{../../results/D/accuracy_corrs.tex}
+\input{../../results/D/human_corrs.tex}
 
@@
 \end{abstract}
-
-\TODO{need to compute better human vs human correlations}
 
@@
     therefore, their correlations are undefined. (c) Pearson
-    correlations between human and model, both for response time and
-    accuracy. The first set of bars is human vs. human, to give an
-    upper bound. All error bars are 95\% confidence intervals.}
+    correlations between human and model (except the first set of
+    bars, which is human vs. human), both for response time and
+    accuracy. All error bars are 95\% confidence intervals.}
 \end{figure*}
@@
 
+To gauge between-subject reliability, we randomly split subjects into
+two groups, computed the average response time and accuracy per
+stimulus for each group, and then computed correlations for these two
+measures. We repeated this procedure 100 times, and from those samples
+computed the median correlations and 95\% confidence intervals. For
+response time, the between-subjects correlation was \ExpTimeCorr{},
+and for accuracy, it was \ExpAccuracyCorr{} (also shown in
+Fig.~\ref{fig:human-model-corrs}).
+
 There was a significant effect of learning both on response time
@@
 
-\TODO{say something about the large step size -- actually produces
-  non-intuitive results, even though they are a better fit. with small
-  step size (where it doesn't over-rotate) it does even worse!}
+As noted above, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This
+has the interesting effect of actually causing the \Th{} to
+\textit{over}rotate (e.g., the ``flipped'' pair trace in
+Fig.~\ref{fig:model-traces}). This is because the step size is large
+enough that it sometimes misses the global maximum, and has to do
+another full rotation to find it. However, when we use a lower step
+size (so that it does not miss the maximum), the fit to human data is
+slightly worse.
 

### 2014-02-04 analysis/analyses/num_chance.py
Use words for small numbers in num_chance.py

@@ -7,6 +7,11 @@ import pandas as pd
 filename = "num_chance.csv"
 texname = "num_chance.tex"
 
+words = [
+    'zero', 'one', 'two', 'three', 'four',
+    'five', 'six', 'seven', 'eight', 'nine', 'ten'
+]
+
 
 def run(data, results_path, seed):
     np.random.seed(seed)
@@ -29,9 +34,13 @@ def run(data, results_path, seed):
     with open(results_path.joinpath(texname), "w") as fh:
         fh.write("%% AUTOMATICALLY GENERATED -- DO NOT EDIT!\n")
         for model, stats in results.iterrows():
+            if stats['num'] < len(words):
+                num = words[stats['num']]
+            else:
+                num = stats['num']
+
             cmd = util.newcommand(
-                "%sNumChance" % model.capitalize(),
-                stats['num'])
+                "%sNumChance" % model.capitalize(), num)
             fh.write(cmd)
 
     return pth

### 2014-02-04 analysis/plots/accuracy_scatters.py
Fix legends in scatter plots

@@ -28,7 +28,7 @@ def plot(results_path, fig_path):
             ax.plot(
                 results[(key, flipped)],
                 results[('exp', flipped)],
-                '.', alpha=0.8, label=key)
+                '.', alpha=0.8, label=flipped)
 
         ax.set_xlabel("Model accuracy", fontsize=14)
         ax.set_title(titles[key], fontsize=14)

### 2014-02-04 analysis/plots/response_time_scatters.py

@@ -29,7 +29,7 @@ def plot(results_path, fig_path):
             ax.plot(
                 results[(key, flipped)],
                 results[('exp', flipped)],
-                '.', alpha=0.8, label=key)
+                '.', alpha=0.8, label=flipped)
 
         ax.set_xlabel("Model response time", fontsize=14)
         ax.set_title(titles[key], fontsize=14)

### 2014-02-04 man/cogsci-2014/cogsci-2014.tex
Minor changes here and there

@@
 
-% space between figure and caption
-\addtolength{\abovecaptionskip}{-0.25cm}
-
 % space below "Figure 1: ...", but only for inline figures
-\addtolength{\textfloatsep}{-0.5cm}
+%\addtolength{\textfloatsep}{-0.5cm}
 
@@
 
-\title{What to simulate? Inferring the direction for mental rotation}
+\title{What to simulate? Inferring the right direction for mental
+  rotation}
 
@@
 \begin{abstract}
-  When people use mental imagery to solve tasks, how do they decide
-  \textit{which} images to generate? To answer this question, we
-  explored how mental simulation should used in the classic
-  psychological task of determining if two images depict the same
-  object in different orientations \citeA{Shepard1971}. Through a
-  rational analysis of mental rotation, we formalized four models and
-  compared them to human performance on a 2D mental rotation
-  task. None of the models were a particularly satisfying explanation
-  of human behavior, leading us to suggest that the question of ``what
-  to simulate?'' is more difficult than has previously been
-  assumed. While only exploratory, this work provides the foundation
-  necessary to begin answering this question.
+  When people use mental imagery, how do they decide \textit{which}
+  images to generate? To answer this question, we explored how mental
+  simulation should used in the classic psychological task of
+  determining if two images depict the same object in different
+  orientations \cite{Shepard1971}. Through a rational analysis of
+  mental rotation, we formalized four models and compared them to
+  human performance. \TODO{None of the models were a particularly
+    satisfying explanation of human behavior, leading us to suggest
+    that the question of ``what to simulate?'' is more difficult than
+    has previously been assumed.} While only exploratory, this work
+  provides the foundation necessary to begin answering this question.
 
@@
 
-In the classic experiment by \citeA{Shepard1971}, participants viewed
-images of three-dimensional objects and had to determine whether the
-images depicted the same object (which differed by a rotation) or two
-separate objects (which differed by a reflection and a rotation). The
-result was that people's response times had a strong linear
-correlation with the minimum angle of rotation, leading to the
-conclusion that people solve this task by ``mentally rotating'' the
-objects until they are congruent.  This conclusion leaves several
-questions unanswered, however. How do people know the axis around
-which to rotate the objects? If the axis is known, how do people know
-which direction to rotate the objects?  And finally, how do people
-know how long to rotate?
+Mental rotation provides a simple example of the decision problem
+posed by simulation.  In the classic experiment by
+\citeA{Shepard1971}, participants viewed images of three-dimensional
+objects and had to determine whether the images depicted the same
+object (which differed by a rotation) or two separate objects (which
+differed by a reflection and a rotation). The result was that people's
+response times had a strong linear correlation with the minimum angle
+of rotation, leading to the conclusion that people solve this task by
+``mentally rotating'' the objects until they are congruent.  This
+conclusion leaves several questions unanswered, however. How do people
+know the axis around which to rotate the objects? If the axis is
+known, how do people know which direction to rotate the objects?  And
+finally, how do people know how long to rotate?
 
@@
 \cite{Marr:1983to,anderson90,Shepard:1987tt} and compare four models
-of mental rotation. First, we discuss the previous literature on
-mental imagery. Second, we outline computational- and
-algorithmic-level analyses of the problem of mental rotation. Third,
-we describe a behavioral experiment based on the classic mental
-rotation studies \cite<e.g.,>{Cooper:1975wp}, and compare the results
-of the experiment with each of the models. We conclude with a
-discussion of the strengths and weaknesses of each model, and lay out
-directions for future work.
+of mental rotation. We begin the paper by discussing the previous
+literature on mental imagery. Next, we outline computational- and
+algorithmic-level analyses of the problem of mental rotation.  We then
+describe a behavioral experiment based on the classic mental rotation
+studies \cite<e.g.,>{Cooper:1975wp}, and compare the results of the
+experiment with each of the models. We conclude with a discussion of
+the strengths and weaknesses of each model, and lay out directions for
+future work.
 
-\section{Background}
+\section{Modeling mental rotation}
 
@@
 \textit{which} mental images to generate. \citeA{Kosslyn:1977tv}
-propose a model of the mental imagery buffer, but not the planning
+proposed a model of the mental imagery buffer, but not the planning
 process of how it is used. Similarly, \citeA{Julstrom:1985va} and
-\citeA{Glasgow:1992tj} are mostly concerned with modeling the
+\citeA{Glasgow:1992tj} were mostly concerned with modeling the
 representational format underlying imagery. Although
-\citeA{Anderson1978} emphasizes the importance of considering both
-representation and process, he dismisses the problem of determining
+\citeA{Anderson1978} emphasized the importance of considering both
+representation and process, he dismissed the problem of determining
 the direction of rotation as a ``technical difficulty''.
 
-The only models (of which the authors are aware) that have seriously
+The only models (of which the authors are aware) that seriously
 attempted to address the decision of \textit{what} to simulate are
@@
 
-\subsection{A rational analysis of mental rotation}
-
 If people are not computing the rotation beforehand, what might they
@@
 The original ``congruency'' hypothesis \cite{Shepard1971} actually is
-a rational solution which achieves this goal, in the sense that the
-smallest amount of computation coincides with rotating through the
-minimum angle. However, it violates the constraint that we do not know
-points correspondences, which is exactly what necessitates the use of
-imagery. Noting that a rational solution need not necessarily maintain
-a single trajectory of rotation, we explore an alternative rational
-model, which--rather than computing the angle of rotation ahead of
-time--engages in an ``active sampling'' strategy.
+a rational solution to this problem, in the sense that the smallest
+amount of computation coincides with rotating through the minimum
+angle. However, it violates the constraint that we do not know the
+points of correspondence between the images, which is exactly what
+necessitates the use of imagery. Noting that a rational solution need
+not necessarily maintain a single trajectory of rotation, we explore
+an alternative rational model, which--rather than computing the angle
+of rotation ahead of time--engages in an ``active sampling'' strategy.
 
@@
 
-\section{Modeling mental rotation}
+\section{How should we rotate?}
 
@@
 
-\subsection{Algorithmic assumptions}
+\subsection{Algorithmic constraints}
 
@@
 hypothesis. This approach gives us the true value of $\theta$, so
-Eq.~\ref{eq:lh-ratio} becomes a maximum likelihood ratio test, where
-$\theta$ is set to the MLE value, rather than being marginalized:
+Eq.~\ref{eq:lh-ratio} becomes a generalized likelihood ratio test,
+where $\theta$ is set to the MLE value, rather than being
+marginalized:
 \begin{equation}
@@
 
-\subsubsection{Hill climbing model}
+\subsubsection{Hill Climbing model}
 
@@
 (\Hc{}) search for a \textit{local} maximum. As with the \Oc{} and
-\Th{} models, Eq.~\ref{eq:lh-ratio} becomes a maximum likelihood
-ratio.
+\Th{} models, Eq.~\ref{eq:lh-ratio} becomes a generalized likelihood
+ratio test.
 
-\subsubsection{Bayesian quadrature model}
+\subsubsection{Bayesian Quadrature model}
 
@@
 
-We randomly generated 20 shapes of 5 or 6 vertices (e.g.,
+We randomly generated 20 shapes of five or six vertices (e.g.,
 Fig.~\ref{fig:model-traces}). For each shape, we computed 20 ``same''
@@
 
-We generated 5 additional shapes to be used in a practice block of 10
-trials. Across these trials, each shape and each angle (60, 120, 180,
-240, or 300) was repeated twice (once ``flipped'' and once ``same'')
-such that no shape was presented at the same angle twice. We also
-generated a sixth shape to be included along with the instructions.
-This shape had both a ``flipped'' and ``same'' version, each rotated
-to 320 degrees.
+We generated five additional shapes to be used in a practice block of
+10 trials. Across these trials, each shape and each angle (60, 120,
+180, 240, or 300) was repeated twice (once ``flipped'' and once
+``same'') such that no shape was presented at the same angle twice. We
+also generated a sixth shape to be included along with the
+instructions.  This shape had both a ``flipped'' and ``same'' version,
+each rotated to 320 degrees.
 
@@
 rotations of $0^\prime$ or $180^\prime$ were repeated twice, there
-were 800 total experimental stimuli. These stimuli were split across 8
-conditions in the following manner: first, stimuli were split into
-four blocks of 200 trials. Within each block, each shape was repeated
-10 times and each rotation was repeated 10 times (5 ``same'', 5
-``flipped''), such that across all blocks, each stimulus appeared
-exactly once. Each block was then split in half. Participants
-completed two half-blocks (not necessarily both from the original full
-block).
+were 800 total experimental stimuli. These stimuli were split across
+eight conditions in the following manner: first, stimuli were split
+into four blocks of 200 trials. Within each block, each shape was
+repeated ten times and each rotation was repeated ten times (five
+``same'', five ``flipped''), such that across all blocks, each
+stimulus appeared exactly once. Each block was then split in
+half. Participants completed two half-blocks (not necessarily both
+from the original full block).
 
@@
 analyses. Of the rest, we excluded 10 (4\%) because of an experimental
-error, 6 (2.4\%) because they had already completed a previous version
-of the experiment, and 31 (12.6\%) because they failed a comprehension
+error, 6 (2.4\%) because they had already completed a related
+experiment, and 31 (12.6\%) because they failed a comprehension
 check. A participant was marked as passing the comprehension check if
@@
 
-There was a significant effect of learning both on response time
+There was a significant effect of trial number both on response time
 (\ExpTrialTimeCorr{}) and on accuracy (\ExpTrialAccuracyCorr{}),
@@
 
-We set out to answer the question: how do people decide \textit{what}
-to simulate when using mental imagery? Focusing on the specific case
+We set out to answer the question of how people decide \textit{what}
+to simulate when using mental imagery. Focusing on the specific case
 of determining the direction and extent of mental rotation, we
@@
 
-Additionally, a closer look at individual shapes reveals several
-interesting points (Fig.~\ref{fig:response-time-stimulus} shows curves
-for one particular shape; these curves are fairly typical relative to
-other shapes). First, the curves for the human response times on
-specific shapes tend to either be linear, or linear first, then
-flat. Second, while the \Th{} model appears to capture the concavity
-of human response times overall, it is actually fairly inconsistent
-within shapes. Third, the linearity of the \Bq{} model's response time
-curves \textit{is} consistent within shapes, though it does not
-exhibit the same ``plateau'' effect that sometimes appears in the
-human curves.
+Additionally, a closer look at performance on individual shapes
+reveals several interesting points
+(Fig.~\ref{fig:response-time-stimulus} shows curves for one particular
+shape; these curves are fairly typical relative to other
+shapes). First, the curves for the human response times on specific
+shapes tend to either be linear, or linear first, then flat. Second,
+while the \Th{} model appears to capture the concavity of human
+response times overall, it is actually fairly inconsistent within
+shapes. Third, the linearity of the \Bq{} model's response time curves
+\textit{is} consistent within shapes, though it does not exhibit the
+same ``plateau'' effect that sometimes appears in the human curves.
 

### 2014-02-04 lib/mental_rotation/model/model.py
S_sigma should be standard deviation, not variance

@@ -49,8 +49,8 @@ def slow_log_similarity(X0, X1, S_sigma):
     # number of points and number of dimensions
     n, D = X0.shape
     # covariance matrix
-    Sigma = np.eye(D) * S_sigma
-    invSigma = np.eye(D) * (1. / S_sigma)
+    Sigma = np.eye(D) * (S_sigma ** 2)
+    invSigma = np.eye(D) * (1. / (S_sigma ** 2))
     # constants
     Z0 = D * np.log(2 * np.pi)
     Z1 = np.linalg.slogdet(Sigma)[1]

### 2014-02-04 lib/mental_rotation/model/model_c.pyx

@@ -29,7 +29,7 @@ def log_factorial(long n):
 
 
 cpdef float64_t log_const(long n, long d, float64_t S_sigma):
-    cdef float64_t const = -0.5 * (log(S_sigma) + log(2 * M_PI)) * n * d - log(2 * n)
+    cdef float64_t const = -0.5 * (log(S_sigma ** 2) + log(2 * M_PI)) * n * d - log(2 * n)
     return const
 
 
@@ -68,7 +68,7 @@ cpdef float64_t log_similarity(float64_t[:, ::1] X0, float64_t[:, ::1] X1, float
     # covariance matrix
     S[:, :] = 0
     for i in xrange(D):
-        S[i, i] = S_sigma
+        S[i, i] = S_sigma ** 2
     la.cho_factor(S, S)
     logdet = la.logdet(S)
 

### 2014-02-04 man/cogsci-2014/cogsci-2014.tex
Variance, not standard deviation

@@
   always connected to their two nearest neighbors in the $\theta$
-  dimension.}and $\sigma_S=0.15$ is the standard deviation of the
+  dimension.}and $\sigma_S^2=0.15$ is the variance of the
 similarity. Example similarity curves are shown in
@@
 \end{equation}
-where $\epsilon\sim \left|\mathcal{N}(0, \sigma_\epsilon)\right|$ and
-$\sigma_\epsilon$ is the standard deviation of the step size in
-radians.
+where $\epsilon\sim \left|\mathcal{N}(0, \sigma_\epsilon^2)\right|$
+and $\sigma_\epsilon^2$ is the variance of the step size in radians.
 

### 2014-02-04 man/cogsci-2014/cogsci-2014.tex
Misc wording changes

@@
 
-One of our most astonishing cognitive feats is the ability to
+One of the most astonishing cognitive feats is our ability to
 envision, manipulate, and plan with objects -- all without actually
 perceiving them. This capacity for ``mental simulation'' has been
-widely studied and has sparked intense debate about the underlying
+widely studied, including an intense debate about the underlying
 representation of mental images \cite<e.g.,>{Kosslyn:2009tj,
@@
 object (which differed by a rotation) or two separate objects (which
-differed by a reflection and a rotation). The result was that people's
+differed by a reflection and a rotation). They found that people's
 response times had a strong linear correlation with the minimum angle
-of rotation, leading to the conclusion that people solve this task by
-``mentally rotating'' the objects until they are congruent.  This
-conclusion leaves several questions unanswered, however. How do people
-know the axis around which to rotate the objects? If the axis is
-known, how do people know which direction to rotate the objects?  And
-finally, how do people know how long to rotate?
+of rotation, a result which led to the conclusion that people solve
+this task by ``mentally rotating'' the objects until they are
+congruent.  This explanation leaves several questions unanswered,
+however. How do people know the axis around which to rotate the
+objects? If the axis is known, how do people know which direction to
+rotate the objects?  And finally, how do people know how long to
+rotate?
 
@@
 
-Previous models of mental rotation have, for the most part, focused on
-the representation of mental images, rather than how people decide
+Previous models of mental rotation have largely focused on the
+representation of mental images, rather than how people decide
 \textit{which} mental images to generate. \citeA{Kosslyn:1977tv}
@@
 If people are not computing the rotation beforehand, what might they
-be doing? To answer this question, we perform a ``rational analysis''
-of the problem of mental rotation
+be doing? To answer this question, we perform a rational analysis of
+the problem of mental rotation
 \cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
-level, we could say that the \textit{problem} is to determine which
-spatial transformations an object has undergone based on two
-depictions of that object which do not include information about point
-correspondences across images. At the algorithmic level, we are
-constrained by the notion that mental images must be transformed in an
-analog manner (or in a way that is at least approximately analog), and
-that mental images are time-consuming and effortful to generate. Thus,
-the \textit{goal} is to make this determination while performing a
-minimum amount of computation (i.e., as few rotations as possible).
+level, we can say that the \textit{problem} is to determine which
+spatial transformations an object has undergone based on two images of
+that object (which do not include information about point
+correspondences). At the algorithmic level, we are constrained by the
+notion that mental images must be transformed in an analog manner (or
+in a way that is at least approximately analog), and that mental
+images are time-consuming and effortful to generate. Thus, the
+\textit{goal} is to make this determination while performing a minimum
+amount of computation (i.e., as few rotations as possible).
 
@@
 an alternative rational model, which--rather than computing the angle
-of rotation ahead of time--engages in an ``active sampling'' strategy.
+of rotation ahead of time--engages in an \textit{active sampling}
+strategy.
 
-Active sampling is the idea that, when given information is
-incomplete, people gather new information in a manner that increases
-certainty about the problem space. This notion has recently gained
-support in several areas of cognitive science
+Active sampling is the idea that when initial information is
+incomplete, people will gather new information in a manner that
+increases certainty about the problem space. This notion has recently
+gained support in several areas of cognitive science
 \cite<e.g.,>{Nelson2007,Gureckis:2012gu}, including other spatial
@@
 
-In this section, we formalize our rational analysis, and propose four
+In this section, we formalize our rational analysis and propose four
 models of mental rotation: one based on existing models; two which are
@@
 X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which (assuming that all
-rotations are equally likely) simplifies to:
+rotations are equally likely) is equivalent to:
 \begin{equation}
@@
       are two stimulus pairs, one ``same'' (top) and one ``flipped''
-      (bottom), both with a rotation of $120^\prime$. The grey plots
+      (bottom), both with a rotation of $120^\circ$. The grey plots
       show the actions that each model took for these stimuli. The
@@
 
-One hypothesis in the literature is that people compute the direction
-and extent of rotation beforehand using \textit{a prior} knowledge of
-the correspondence between points in the images
+One hypothesis is that people compute the direction and extent of
+rotation beforehand using \textit{a prior} knowledge of the
+correspondence between points in the images
 \cite{Funt:1983wn,Just:1985uu}.  To reflect this hypothesis, we
@@
 
-An intuitive and simple model which does not know which way to rotate
-\textit{a priori} might first rotate in the direction that increases
-similarity, and then stop rotating once a ``match'' is found, as
-determined by a threshold on the value of $S$. If all rotations have
-been exhausted, then flip, and try rotating again. We assume the value
-of $S$ which is greater than the threshold corresponds to the true
-$\theta$, or a point near the true $\theta$. So, as with the \Oc{}
-model, Eq.~\ref{eq:lh-ratio} becomes a likelihood ratio.
+What if the point correspondences are unknown? An intuitive and simple
+model which obeys this constraint first rotate in the direction that
+increases similarity, and then stop rotating once a ``match'' is
+found, as determined by a threshold on the value of $S$. If all
+rotations have been exhausted, then flip, and try rotating again. We
+assume that the locations where $S$ is greater than the threshold
+correspond to the true $\theta$ (or points near the true
+$\theta$). So, as with the \Oc{} model, Eq.~\ref{eq:lh-ratio} becomes
+a likelihood ratio.
 
@@
 them. However, this choice is not always clear \textit{a priori}, as
-the global optimum will depend on many factors (e.g., shape
-complexity, dimensionality, perceptual uncertainty, and whether the
-shapes are identical).  One way to deal with the problem of choosing a
-threshold would simply be to perform an exhaustive search or use a
-global optimization strategy. However, these strategies would not
-result in the linear response time found by \citeA{Shepard1971}. A
-second alternative, which could potentially demonstrate the linear
-response time result, might instead just perform a Hill Climbing
-(\Hc{}) search for a \textit{local} maximum. As with the \Oc{} and
-\Th{} models, Eq.~\ref{eq:lh-ratio} becomes a generalized likelihood
-ratio test.
+the global optimum depends on many factors (e.g., shape complexity,
+dimensionality, perceptual uncertainty, and whether the shapes are
+identical).  One way to deal with the problem of choosing a threshold
+would simply be to perform an exhaustive search or use a global
+optimization strategy. However, these strategies would not result in
+the linear response time found by \citeA{Shepard1971}. A second
+alternative, which could potentially demonstrate the linear response
+time result, might instead just perform a Hill Climbing (\Hc{}) search
+for a \textit{local} maximum. As with the \Oc{} and \Th{} models,
+Eq.~\ref{eq:lh-ratio} becomes a generalized likelihood ratio test.
 
@@
 function. Second, by choosing to sample places which are informative,
-this method implicitly minimizes the amount of rotation; yet, because
-it is constrained to sequential rotations, it should still exhibit the
-linear effect of the angle.
+this method implicitly minimizes the amount of rotation.
 
@@
 \begin{equation}
-  p(Z_h) = \int \left[\int S(\Xb, f(\Xa, \theta, h))p(\theta)\dif\theta\right] p(S)\dif S.
+  p(Z_h) = \int \left[\int S(\Xb, f(\Xa, \theta, h))p(\theta)\dif\theta\right] p(S)\dif S,
 \end{equation}
+where $S$ is the similarity function, and $p(S)$ is a prior over
+similarity functions.  This method of estimating an integral is known
+in the machine-learning literature as \textit{Bayesian Quadrature}
+\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}, or BQ.
 
-This is a technique known in the machine-learning literature as
-\textit{Bayesian Quadrature} (BQ)
-\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}. Specifically, we
-use a Gaussian Process (GP) prior for $p(S)$, as in
-\citeA{Osborne:2012tm}.  Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$, we
-first place a GP prior on the log of $S_h$ in order to enforce
-positivity after it is exponentiated, i.e. $\mathbb{E}[Z_h] \approx
-\int \exp(\mu_h(\theta))p(\theta)\dif\theta$, where $\mu_h:=\mu(\log
-S_h)$ is the mean of the log-GP.  However, this integral is still
-intractable. To approximate it, we fit a second GP over points sampled
-from the log-GP, which we denote as
+Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$, we first place a
+\textit{Gaussian Process} \cite{Rasmussen:2006vz}, or GP, prior on the
+log of $S_h$ in order to enforce positivity after it is exponentiated,
+i.e. $\mathbb{E}[Z_h] \approx \int
+\exp(\mu_h(\theta))p(\theta)\dif\theta$, where $\mu_h:=\mu(\log S_h)$
+is the mean of the log-GP \cite{Osborne:2012tm}.  However, this
+integral is still intractable. To approximate it, we fit a second GP
+over points sampled from the log-GP, which we denote as
 $\bar{S}_h:=\exp(\mu_h)$.\footnote{Note that the approach used here
@@
   \mathbb{V}(Z_h) &\approx \iint \mathrm{Cov}_h(\theta,
-  \theta^\prime)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\prime)p(\theta)p(\theta^\prime)\dif\theta\dif\theta^\prime,
+  \theta^\circ)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\circ)p(\theta)p(\theta^\circ)\dif\theta\dif\theta^\circ,
 \end{align}
@@
 Fig.~\ref{fig:model-traces}). For each shape, we computed 20 ``same''
-and 20 ``flipped'' stimuli pairs, with rotations ($\theta$) spaced at
-$20^\prime$ increments between 0 and 360 (with 0 and 180 degree
-rotations repeated twice, in order to gather an equal number of
-responses for each angle between 0 and 180). ``Same'' pairs were
-created by rotating $\Xa$ by $\theta$; ``flipped'' pairs were first
-reflected $\Xa$ across the $y$-axis, then rotated by $\theta$.
+and 20 ``flipped'' stimuli pairs, with 18 rotations ($\theta$) spaced
+at $20^\circ$ increments between $0^\circ$ and $360^\circ$ (with
+$0^\circ$ and $180^\circ$ repeated twice, in order to gather an equal
+number of responses for each angle between $0^\circ$ and
+$180^\circ$). ``Same'' pairs were created by rotating $\Xa$ by
+$\theta$; ``flipped'' pairs were first reflected $\Xa$ across the
+$y$-axis, then rotated by $\theta$.
 
 We generated five additional shapes to be used in a practice block of
-10 trials. Across these trials, each shape and each angle (60, 120,
-180, 240, or 300) was repeated twice (once ``flipped'' and once
-``same'') such that no shape was presented at the same angle twice. We
-also generated a sixth shape to be included along with the
-instructions.  This shape had both a ``flipped'' and ``same'' version,
-each rotated to 320 degrees.
+10 trials. Across these trials, there was one ``flipped'' and one
+``same'' repetition of each shape and each angle ($60^\circ$,
+$120^\circ$, $180^\circ$, $240^\circ$, or $300^\circ$) such that no
+shape was presented at the same angle twice. We also generated a sixth
+shape to include with the instructions.  This shape had both a
+``flipped'' and ``same'' version, each rotated to $320^\circ$.
 
@@
 
-All participants saw the same 10 practice trials, as described
+All participants saw the same 10 practice trials as described
 above. There were 720 unique experimental stimuli (20 shapes $\times$
 18 angles $\times$ 2 reflections), though because stimuli with
-rotations of $0^\prime$ or $180^\prime$ were repeated twice, there
-were 800 total experimental stimuli. These stimuli were split across
-eight conditions in the following manner: first, stimuli were split
-into four blocks of 200 trials. Within each block, each shape was
-repeated ten times and each rotation was repeated ten times (five
-``same'', five ``flipped''), such that across all blocks, each
-stimulus appeared exactly once. Each block was then split in
-half. Participants completed two half-blocks (not necessarily both
-from the original full block).
+rotations of $0^\circ$ or $180^\circ$ were repeated twice, there were
+800 total experimental stimuli. These stimuli were split across eight
+conditions in the following manner: first, stimuli were split into
+four blocks of 200 trials. Within each block, each shape was repeated
+ten times and each rotation was repeated ten times (five ``same'',
+five ``flipped''), such that across all blocks, each stimulus appeared
+exactly once. Each block was then split in half. Participants
+completed two half-blocks (not necessarily both from the original full
+block).
 
@@
 experiment, and 31 (12.6\%) because they failed a comprehension
-check. A participant was marked as passing the comprehension check if
-they answered at least 85\% of ``easy'' trials correctly. ``Easy''
-trials were defined to be those for which the presented shapes
-differed by a rotation of no more than 20 degrees. Additionally, we
-excluded 82 trials for which the response time was either less than
-100 milliseconds or greater than 20 seconds.
+check. Comprehension was defined as correctly answering at least 85\%
+of stimuli with a minimum rotation of $0^\circ$ or $20^\circ$.
+Additionally, we excluded 82 trials for which the response time was
+either less than 100 milliseconds or greater than 20 seconds.
 
-For each model, we ran 50 samples for each stimulus that participants
-viewed (with the exception of stimuli with rotations of $0^\prime$ or
-$180^\prime$, in which case we ran double the number of samples). The
-step size parameter ($\sigma_\epsilon$) was fit to human data for each
-of the models, resulting in $\sigma_\epsilon=0.6$ for the \Th{} and
-\Bq{} models and $\sigma_\epsilon=0.1$ for the \Oc{} and \Hc{}
-models. We also ran the models under two different priors,
-$p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the ``unequal''
-prior). This had an effect on accuracy for the \Hc{} model, and an
-effect on both accuracy and response time for the \Bq{} model. Below,
-we report results for the \Hc{} model with an unequal prior, and the
-\Bq{} model with both priors.
+For each model, we ran 50 samples for each of the 800 experimental
+stimuli. The step size parameter ($\sigma_\epsilon$) was fit to human
+response times for each of the models, resulting in
+$\sigma_\epsilon=0.6$ for the \Th{} and \Bq{} models and
+$\sigma_\epsilon=0.1$ for the \Oc{} and \Hc{} models. We also ran the
+models under two different priors, $p(h=0)=0.5$ (the ``equal'' prior)
+and $p(h=0)=0.55$ (the ``unequal'' prior). This only had an effect on
+the stopping criteria for the \Bq{} model.\footnote{The prior also had
+  a small effect on the accuracy of the \Hc{} model, because sometimes
+  the local maxima were small enough that Eq.~\ref{eq:mle-lh-ratio}
+  became dominated by the prior.}
 
@@
 of 10000 bootstrap samples to compute the confidence intervals around
-both Spearman ($r_s$) and Pearson ($r$) correlations.  For analyses of
-accuracy, confidence intervals were computed from a binomial
-proportion with a Jeffrey's beta prior.  To test if participants or a
-model was above chance on a particular stimulus, we used the same
-binomial proportion, and simply tested whether $p>0.05$ for the 50\%
-mark.
+both Spearman ($r_s$) and Pearson ($r$) correlations.  Unless
+otherwise specified, all correlations were computed over 720
+datapoints (one for each stimulus). For analyses of accuracy,
+confidence intervals were computed from a binomial proportion with a
+Jeffrey's beta prior.  To test if participants or a model was above
+chance on a particular stimulus, we used the same binomial proportion,
+and simply tested whether $p>0.05$ for the 50\% mark.
+
+\TODO{need to recompute correlations}
 
@@
 The average accuracy across all stimuli was \ExpAccuracy{}, though
-there were \ExpNumChance{} individual stimuli (out of 720) for which
-people were not above chance.  The minimum angle was also correlated
-with participants' average per-stimulus accuracy, as shown in
+there were \ExpNumChance{} stimuli (out of 720) for which people were
+not above chance.  The minimum angle was also correlated with
+participants' average per-stimulus accuracy, as shown in
 Fig.~\ref{fig:theta-accuracy-corrs}, though significantly more so for
@@
 
-The number of actions taken by the \Th{} model was about as correlated
-with the minimum angle as human response times were
-(Fig.~\ref{fig:theta-time-corrs}), though this effect was less strong
-for individual shapes (e.g.,
-Fig.~\ref{fig:response-time-stimulus}). Like the \Oc{} model, the
-\Th{} model had 100\% accuracy, and thus there was no correlation
-between minimum angle and accuracy. The \Th{} was able to explain a
-moderate amount of the variance in human response times, with a
-correlation of \ThTimeCorr{} (Fig.~\ref{fig:human-model-corrs}).
+The number of actions taken by the \Th{} model was significantly
+correlated with the minimum angle (Fig.~\ref{fig:theta-time-corrs}),
+though this effect was less strong for individual shapes (e.g.,
+Fig.~\ref{fig:response-time-stimulus}).  The \Th{} model was able to
+explain a moderate amount of the variance in human response times,
+with a correlation of \ThTimeCorr{}
+(Fig.~\ref{fig:human-model-corrs}). Like the \Oc{} model, the \Th{}
+model had 100\% accuracy, and thus there was no correlation between
+minimum angle and accuracy.
 
@@
 has the interesting effect of actually causing the \Th{} to
-\textit{over}rotate (e.g., the ``flipped'' pair trace in
+\textit{over}rotate (see, for example, the ``flipped'' pair trace in
 Fig.~\ref{fig:model-traces}). This is because the step size is large
-enough that it sometimes misses the global maximum, and has to do
-another full rotation to find it. However, when we use a lower step
-size (so that it does not miss the maximum), the fit to human data is
-slightly worse.
+enough that it sometimes misses the global maximum, and must do
+another full rotation to find it. However, when the step size is small
+enough to always find the maximum, the fit to human data is slightly
+worse.
 
@@
 
-The \Bq{} model was quite accurate overall, both with the equal prior
-(\BqAccuracy{}) and unequal prior (\BqpAccuracy{}). With the equal
-prior, there was \BqNumChance{} stimulus for which it was not above
-chance; with the unequal prior, there were \BqpNumChance{}. There was
-a significant effect of rotation on the \Bq{} model's accuracy
+The \Bq{} model was quite accurate overall (equal prior:
+(\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
+there was \BqNumChance{} stimulus for which it was not above chance;
+with the unequal prior, there were \BqpNumChance{}. There was a
+significant effect of rotation on the \Bq{} model's accuracy
 (Fig.~\ref{fig:theta-accuracy-corrs}), and it explained about the same
-amount of variance in people's accuracy as the \Hc{} model: for the
-equal prior, \BqAccuracyCorr{}; for the unequal prior,
-\BqpAccuracyCorr{}.
+amount of variance in people's accuracy as the \Hc{} model (equal
+prior: \BqAccuracyCorr{}; unequal prior: \BqpAccuracyCorr{}).
 
-Because the \Bq{} model relied on Eq.~\ref{eq:lh-ratio} for its
+Because the \Bq{} model relies on Eq.~\ref{eq:lh-ratio} for its
 stopping criteria (as opposed to just finding a maximum), the prior
-$p(h)$ had a fairly significant effect, as shown in
+$p(h)$ has a fairly significant effect, as shown in
 Fig.~\ref{fig:response-time-accuracy}. With just a small bias of
-$p(h=0)=0.55$, there is a clear separation in response times that is
-similar to the separation seen in human response times. The prior also
-had an effect on accuracy, though this effect did not reflect the
-trends in human behavior.
+$p(h=0)=0.55$, there is a clear separation in response times for
+``same'' versus ``flipped'' stimuli; this separation is similar to the
+trend also observed in human response times. While the prior also had
+an effect on accuracy, this effect did not reflect the trends in human
+behavior.
 

### 2014-02-05 analysis/analyses/theta_accuracy.py
Multiply accuracies by 100

@@ -13,7 +13,7 @@ def run(data, results_path, seed):
     results = {}
     for key, df in data.iteritems():
         y = df.groupby(['modtheta', 'flipped'])['correct']
-        results[key] = y.apply(util.beta)
+        results[key] = y.apply(util.beta) * 100
 
     results = pd.DataFrame.from_dict(results)
     results.index = pd.MultiIndex.from_tuples(

### 2014-02-05 analysis/analyses/theta_accuracy_stimulus.py

@@ -13,7 +13,7 @@ def run(data, results_path, seed):
     results = {}
     for key, df in data.iteritems():
         y = df.groupby(['stimulus', 'modtheta', 'flipped'])['correct']
-        results[key] = y.apply(util.beta)
+        results[key] = y.apply(util.beta) * 100
 
     results = pd.DataFrame.from_dict(results)
     results.index = pd.MultiIndex.from_tuples(

### 2014-02-05 man/cogsci-2014/cogsci-2014.tex
Update paper with new figure

@@
 
-% \begin{figure}[t]
-%   \centering
-%   \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_accuracy.pdf}}%
-%   \subfloat{\includegraphics[width=0.24\textwidth]{../../figures/D/trial_time.pdf}}
-%   \caption{\textbf{Effect of learning.} There was a clear effect of
-%     learning on participants' responses, especially in the first half
-%     of the experiment. Left: accuracy as a function of trial. Right:
-%     response time in milliseconds as a function of trial.}
-%   \label{fig:learning}
-% \end{figure}
-
 To evaluate the models described previously, we ran a behavioral
@@
 \begin{figure*}[t]
-  \centering
-  \begin{subfigure}[b]{0.32\textwidth}
-    \includegraphics[width=\textwidth]{../../figures/D/theta_time_corrs.pdf}
-    \caption{\textbf{Angle vs. response time}}
-    \label{fig:theta-time-corrs}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.32\textwidth}
-    \includegraphics[width=\textwidth]{../../figures/D/theta_accuracy_corrs.pdf}
-    \caption{\textbf{Angle vs. accuracy}}
-    \label{fig:theta-accuracy-corrs}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.32\textwidth}
-    \includegraphics[width=\textwidth]{../../figures/D/human_model_corrs.pdf}
-    \caption{\textbf{Human vs. model}}
-    \label{fig:human-model-corrs}
-  \end{subfigure}
-  \vspace{10pt}
-  \caption{\textbf{Correlations.} (a) Spearman correlations between
-    the minimum angle of rotation and response time. (b) Spearman
-    correlations between the minimum angle of rotation and
-    accuracy. The \Oc{} and \Th{} models had 100\% accuracy;
-    therefore, their correlations are undefined. (c) Pearson
-    correlations between human and model (except the first set of
-    bars, which is human vs. human), both for response time and
-    accuracy. All error bars are 95\% confidence intervals.}
+  \begin{center}
+    \includegraphics[width=\textwidth]{../../figures/D/response_time_scatters.pdf}
+    \caption{\textbf{Model vs. human response times.}}
+    \label{fig:human-model-scatters}
+  \end{center}
 \end{figure*}
@@
 of 10000 bootstrap samples to compute the confidence intervals around
-both Spearman ($r_s$) and Pearson ($r$) correlations.  Unless
+both Spearman ($r_s$) and Pearson ($\rho$) correlations.  Unless
 otherwise specified, all correlations were computed over 720
@@
 
-\TODO{need to recompute correlations}
-
 \subsubsection{Human}
@@
 The average response time across all stimuli was \ExpTime{}
-milliseconds.  As illustrated in Fig.~\ref{fig:theta-time-corrs}, the
-minimum angle of rotation was significantly monotonically correlated
-with average per-stimulus response times, both for ``flipped'' and
-``same'' pairs. While this replicates the general result of previous
-experiments \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are
-not as perfectly linear. However, \citeA{Gardony:2013gn} found a
-constant effect for ``flipped'' pairs; an average of their results and
-the classic results would produce something similar to ours.
+milliseconds.  The minimum angle of rotation was significantly
+monotonically correlated with average per-stimulus response times,
+both for ``flipped'' and ``same'' pairs \TODO{}. While this replicates
+the general result of previous experiments
+\cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are not as
+perfectly linear. However, \citeA{Gardony:2013gn} found a constant
+effect for ``flipped'' pairs; an average of their results and the
+classic results would produce something similar to ours.
 
@@
 not above chance.  The minimum angle was also correlated with
-participants' average per-stimulus accuracy, as shown in
-Fig.~\ref{fig:theta-accuracy-corrs}, though significantly more so for
-``same'' pairs. This is the same result found both by
+participants' average per-stimulus accuracy, though significantly more
+so for ``same'' pairs (\TODO{}). This is the same result found both by
 \citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.
@@
 response time, the between-subjects correlation was \ExpTimeCorr{},
-and for accuracy, it was \ExpAccuracyCorr{} (also shown in
-Fig.~\ref{fig:human-model-corrs}).
+and for accuracy, it was \ExpAccuracyCorr{}.
 
@@
 perfectly correlated with the minimum angle of rotation
-(Fig.~\ref{fig:theta-time-corrs}), both for ``same'' and ``flipped''
-pairs. Unsurprisingly, the \Oc{} model was the best fit to human
-response times, with a correlation of \OcTimeCorr{}
-(Fig.~\ref{fig:human-model-corrs}).  However, the \Oc{} model was
-100\% accurate, and therefore the correlation between minimum angle
-and accuracy was undefined (Fig.~\ref{fig:theta-accuracy-corrs}).
+(Fig.~\ref{fig:response-time-accuracy}). Unsurprisingly, the \Oc{}
+model was the best fit to human response times, with a correlation of
+\OcTimeCorr{} (Fig.~\ref{fig:human-model-scatters}).  However, the
+\Oc{} model was 100\% accurate, and therefore the correlation between
+minimum angle and accuracy was undefined.
 
@@
 
-The number of actions taken by the \Th{} model was significantly
-correlated with the minimum angle (Fig.~\ref{fig:theta-time-corrs}),
-though this effect was less strong for individual shapes (e.g.,
-Fig.~\ref{fig:response-time-stimulus}).  The \Th{} model was able to
-explain a moderate amount of the variance in human response times,
-with a correlation of \ThTimeCorr{}
-(Fig.~\ref{fig:human-model-corrs}). Like the \Oc{} model, the \Th{}
+There was an overall monotonic relationship between the minimum angle
+of rotation and the number of actions taken by the \Th{} model
+(Fig.~\ref{fig:response-time-accuracy}). However, the relationship
+between angle and response time for individual shapes was usually not
+monotonic (e.g., Fig.~\ref{fig:response-time-stimulus}).  The \Th{}
+model was able to explain a moderate amount of the variance in human
+response times, with a correlation of \ThTimeCorr{}
+(Fig.~\ref{fig:human-model-scatters}). Like the \Oc{} model, the \Th{}
 model had 100\% accuracy, and thus there was no correlation between
@@
 
-The \Hc{} was the only model for which there was no effect of rotation
-on response time (Fig.~\ref{fig:theta-time-corrs}). Although there was
-a significant effect of rotation on accuracy
-(Fig.~\ref{fig:theta-accuracy-corrs}), the \Hc{} model was barely
-above chance (\HcAccuracy{}) and there were \HcNumChance{} individual
-stimuli for which it was not above chance. The \Hc{} model was not a
-particularly good predictor of human behavior, neither for response
-times (\HcTimeCorr{}) nor accuracy (\HcAccuracyCorr{}), as shown in
-Fig.~\ref{fig:human-model-corrs}.
+The \Hc{} was the only model for which there was no monotonic
+relationship between rotation and response time
+(Fig.~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model was
+barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
+individual stimuli for which it was not above chance. The \Hc{} model
+was not a particularly good predictor of human response times
+(\HcTimeCorr{}), as shown in Fig.~\ref{fig:human-model-scatters}. It
+was a moderately good predictor of human accuracy (\HcAccuracyCorr{}).
 
@@
 
-As with most of the other models, there was a significant effect of
-rotation on the number of steps taken by the \Bq{} model
-(Fig.~\ref{fig:theta-time-corrs}). The \Bq{} model explained variance
-in human response times about as well as the \Th{} model
-(Fig.~\ref{fig:human-model-corrs}), with a correlation of
+Like the \Oc{} and \Th{} models, there was an overall monotonic
+relationship between rotation and the number of steps taken by the
+\Bq{} model (Fig.~\ref{fig:response-time-accuracy}). Unlike the \Th{}
+model, this relationship existed for individual shapes as well (e.g.,
+Fig.~\ref{fig:response-time-stimulus}).  The \Bq{} model explained
+variance in human response times about as well as the \Th{} model
+(Fig.~\ref{fig:human-model-scatters}), with a correlation of
 \BqTimeCorr{} for the equal prior and \BqpTimeCorr{} for the unequal
@@
 there was \BqNumChance{} stimulus for which it was not above chance;
-with the unequal prior, there were \BqpNumChance{}. There was a
-significant effect of rotation on the \Bq{} model's accuracy
-(Fig.~\ref{fig:theta-accuracy-corrs}), and it explained about the same
-amount of variance in people's accuracy as the \Hc{} model (equal
-prior: \BqAccuracyCorr{}; unequal prior: \BqpAccuracyCorr{}).
+with the unequal prior, there were \BqpNumChance{}. The \Bq{} model
+explained about the same amount of variance in people's accuracy as
+the \Hc{} model (equal prior: \BqAccuracyCorr{}; unequal prior:
+\BqpAccuracyCorr{}).
 

### 2014-02-05 man/cogsci-2014/cogsci-2014.tex
Response time --> RT

@@
     \includegraphics[width=\textwidth]{../../figures/D/response_time_accuracy.pdf}
-    \caption{\textbf{Response time and accuracy comparison.} Top:
-      response time of correct responses as a function of the minimum
-      angle of rotation. Bottom: accuracy as a function of the minimum
-      angle of rotation. All error bars are 95\% confidence
-      intervals.}
+    \caption{\textbf{Response time and accuracy comparison.} Top: RT
+      of correct responses as a function of the minimum angle of
+      rotation. Bottom: accuracy as a function of the minimum angle of
+      rotation. All error bars are 95\% confidence intervals.}
     \label{fig:response-time-accuracy}
@@
     \includegraphics[width=\textwidth]{../../figures/D/response_time_scatters.pdf}
-    \caption{\textbf{Model vs. human response times.}}
+    \caption{\textbf{Model vs. human RTs.}}
     \label{fig:human-model-scatters}
@@
 of stimuli with a minimum rotation of $0^\circ$ or $20^\circ$.
-Additionally, we excluded 82 trials for which the response time was
-either less than 100 milliseconds or greater than 20 seconds.
+Additionally, we excluded 82 trials for which the response time (RT)
+was either less than 100 milliseconds or greater than 20 seconds.
 
@@
 stimuli. The step size parameter ($\sigma_\epsilon$) was fit to human
-response times for each of the models, resulting in
-$\sigma_\epsilon=0.6$ for the \Th{} and \Bq{} models and
-$\sigma_\epsilon=0.1$ for the \Oc{} and \Hc{} models. We also ran the
-models under two different priors, $p(h=0)=0.5$ (the ``equal'' prior)
-and $p(h=0)=0.55$ (the ``unequal'' prior). This only had an effect on
-the stopping criteria for the \Bq{} model.\footnote{The prior also had
-  a small effect on the accuracy of the \Hc{} model, because sometimes
-  the local maxima were small enough that Eq.~\ref{eq:mle-lh-ratio}
-  became dominated by the prior.}
+RTs for each of the models, resulting in $\sigma_\epsilon=0.6$ for the
+\Th{} and \Bq{} models and $\sigma_\epsilon=0.1$ for the \Oc{} and
+\Hc{} models. We also ran the models under two different priors,
+$p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the ``unequal''
+prior). This only had an effect on the stopping criteria for the \Bq{}
+model.\footnote{The prior also had a small effect on the accuracy of
+  the \Hc{} model, because sometimes the local maxima were small
+  enough that Eq.~\ref{eq:mle-lh-ratio} became dominated by the
+  prior.}
 
@@
 
-For analyses of response time, confidence intervals around harmonic
-means were computed using a bootstrap analysis using 10000 bootstrap
-samples (sampled with replacement). Response times were only
-considered for correct responses.  We also used a bootstrap analysis
-of 10000 bootstrap samples to compute the confidence intervals around
-both Spearman ($r_s$) and Pearson ($\rho$) correlations.  Unless
-otherwise specified, all correlations were computed over 720
-datapoints (one for each stimulus). For analyses of accuracy,
-confidence intervals were computed from a binomial proportion with a
-Jeffrey's beta prior.  To test if participants or a model was above
-chance on a particular stimulus, we used the same binomial proportion,
-and simply tested whether $p>0.05$ for the 50\% mark.
+For analyses of RT, confidence intervals around harmonic means were
+computed using a bootstrap analysis using 10000 bootstrap samples
+(sampled with replacement). RTs were only considered for correct
+responses.  We also used a bootstrap analysis of 10000 bootstrap
+samples to compute the confidence intervals around both Spearman
+($r_s$) and Pearson ($\rho$) correlations.  Unless otherwise
+specified, all correlations were computed over 720 datapoints (one for
+each stimulus). For analyses of accuracy, confidence intervals were
+computed from a binomial proportion with a Jeffrey's beta prior.  To
+test if participants or a model was above chance on a particular
+stimulus, we used the same binomial proportion, and simply tested
+whether $p>0.05$ for the 50\% mark.
 
@@
 
-The average response time across all stimuli was \ExpTime{}
-milliseconds.  The minimum angle of rotation was significantly
-monotonically correlated with average per-stimulus response times,
-both for ``flipped'' and ``same'' pairs \TODO{}. While this replicates
-the general result of previous experiments
-\cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are not as
-perfectly linear. However, \citeA{Gardony:2013gn} found a constant
-effect for ``flipped'' pairs; an average of their results and the
-classic results would produce something similar to ours.
+The average RT across all stimuli was \ExpTime{} milliseconds.  The
+minimum angle of rotation was significantly monotonically correlated
+with average per-stimulus RTs, both for ``flipped'' and ``same'' pairs
+\TODO{}. While this replicates the general result of previous
+experiments \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are
+not as perfectly linear. However, \citeA{Gardony:2013gn} found a
+constant effect for ``flipped'' pairs; an average of their results and
+the classic results would produce something similar to ours.
 
@@
 To gauge between-subject reliability, we randomly split subjects into
-two groups, computed the average response time and accuracy per
-stimulus for each group, and then computed correlations for these two
-measures. We repeated this procedure 100 times, and from those samples
-computed the median correlations and 95\% confidence intervals. For
-response time, the between-subjects correlation was \ExpTimeCorr{},
-and for accuracy, it was \ExpAccuracyCorr{}.
+two groups, computed the average RT and accuracy per stimulus for each
+group, and then computed correlations for these two measures. We
+repeated this procedure 100 times, and from those samples computed the
+median correlations and 95\% confidence intervals. For RT, the
+between-subjects correlation was \ExpTimeCorr{}, and for accuracy, it
+was \ExpAccuracyCorr{}.
 
-There was a significant effect of trial number both on response time
+There was a significant effect of trial number both on RT
 (\ExpTrialTimeCorr{}) and on accuracy (\ExpTrialAccuracyCorr{}),
@@
 effects may have contributed to the not-quite-linearity of the human
-response times; future work will need to collect more data per
-participant to mitigate this effect.
+RTs; future work will need to collect more data per participant to
+mitigate this effect.
 
@@
 (Fig.~\ref{fig:response-time-accuracy}). Unsurprisingly, the \Oc{}
-model was the best fit to human response times, with a correlation of
+model was the best fit to human RTs, with a correlation of
 \OcTimeCorr{} (Fig.~\ref{fig:human-model-scatters}).  However, the
@@
 (Fig.~\ref{fig:response-time-accuracy}). However, the relationship
-between angle and response time for individual shapes was usually not
-monotonic (e.g., Fig.~\ref{fig:response-time-stimulus}).  The \Th{}
-model was able to explain a moderate amount of the variance in human
-response times, with a correlation of \ThTimeCorr{}
+between angle and RT for individual shapes was usually not monotonic
+(e.g., Fig.~\ref{fig:response-time-stimulus}).  The \Th{} model was
+able to explain a moderate amount of the variance in human RTs, with a
+correlation of \ThTimeCorr{}
 (Fig.~\ref{fig:human-model-scatters}). Like the \Oc{} model, the \Th{}
@@
     \includegraphics[width=0.48\textwidth]{../../figures/D/response_time_stimulus.pdf}
-    \caption{\textbf{Typical response time curves for a single shape.}
-      These plots correspond to the shape shown in
+    \caption{\textbf{Typical RT curves for a single shape.}  These
+      plots correspond to the shape shown in
       Fig.~\ref{fig:model-traces}. Left: human curves are either
@@
       with the ``flipped'' pairs). Middle: the \Th{} model tends to
-      have highly inconsitent within-shape response times. Right: the
-      \Bq{} model is roughly linear.}
+      have highly inconsitent within-shape RTs. Right: the \Bq{} model
+      is roughly linear.}
     \label{fig:response-time-stimulus}
@@
 The \Hc{} was the only model for which there was no monotonic
-relationship between rotation and response time
+relationship between rotation and RT
 (Fig.~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model was
@@
 individual stimuli for which it was not above chance. The \Hc{} model
-was not a particularly good predictor of human response times
-(\HcTimeCorr{}), as shown in Fig.~\ref{fig:human-model-scatters}. It
-was a moderately good predictor of human accuracy (\HcAccuracyCorr{}).
+was not a particularly good predictor of human RTs (\HcTimeCorr{}), as
+shown in Fig.~\ref{fig:human-model-scatters}. It was a moderately good
+predictor of human accuracy (\HcAccuracyCorr{}).
 
@@
 Fig.~\ref{fig:response-time-stimulus}).  The \Bq{} model explained
-variance in human response times about as well as the \Th{} model
+variance in human RTs about as well as the \Th{} model
 (Fig.~\ref{fig:human-model-scatters}), with a correlation of
@@
 Fig.~\ref{fig:response-time-accuracy}. With just a small bias of
-$p(h=0)=0.55$, there is a clear separation in response times for
-``same'' versus ``flipped'' stimuli; this separation is similar to the
-trend also observed in human response times. While the prior also had
-an effect on accuracy, this effect did not reflect the trends in human
-behavior.
+$p(h=0)=0.55$, there is a clear separation in RTs for ``same'' versus
+``flipped'' stimuli; this separation is similar to the trend also
+observed in human RTs. While the prior also had an effect on accuracy,
+this effect did not reflect the trends in human behavior.
 
@@
 shape; these curves are fairly typical relative to other
-shapes). First, the curves for the human response times on specific
-shapes tend to either be linear, or linear first, then flat. Second,
-while the \Th{} model appears to capture the concavity of human
-response times overall, it is actually fairly inconsistent within
-shapes. Third, the linearity of the \Bq{} model's response time curves
-\textit{is} consistent within shapes, though it does not exhibit the
-same ``plateau'' effect that sometimes appears in the human curves.
+shapes). First, the curves for the human RTs on specific shapes tend
+to either be linear, or linear first, then flat. Second, while the
+\Th{} model appears to capture the concavity of human RTs overall, it
+is actually fairly inconsistent within shapes. Third, the linearity of
+the \Bq{} model's RT curves \textit{is} consistent within shapes,
+though it does not exhibit the same ``plateau'' effect that sometimes
+appears in the human curves.
 

### 2014-02-05 analysis/analyses/theta_accuracy_corrs.py
Save latex for theta vs accuracy and theta vs time correlations

@@ -5,6 +5,7 @@ import util
 import pandas as pd
 
 filename = "theta_accuracy_corrs.csv"
+texname = "theta_accuracy_corrs.tex"
 
 
 def run(data, results_path, seed):
@@ -24,6 +25,17 @@ def run(data, results_path, seed):
         results.index, names=['model', 'flipped'])
     pth = results_path.joinpath(filename)
     results.to_csv(pth)
+
+    with open(results_path.joinpath(texname), "w") as fh:
+        fh.write("%% AUTOMATICALLY GENERATED -- DO NOT EDIT!\n")
+        for (model, flipped), stats in results.iterrows():
+            cmd = util.newcommand(
+                "%sThetaAccuracyCorr%s" % (
+                    model.capitalize(),
+                    flipped.capitalize()),
+                util.latex_spearman.format(**dict(stats)))
+            fh.write(cmd)
+
     return pth
 
 

### 2014-02-05 analysis/analyses/theta_time_corrs.py

@@ -5,6 +5,7 @@ import util
 import pandas as pd
 
 filename = "theta_time_corrs.csv"
+texname = "theta_time_corrs.tex"
 
 
 def run(data, results_path, seed):
@@ -24,6 +25,17 @@ def run(data, results_path, seed):
         results.index, names=['model', 'flipped'])
     pth = results_path.joinpath(filename)
     results.to_csv(pth)
+
+    with open(results_path.joinpath(texname), "w") as fh:
+        fh.write("%% AUTOMATICALLY GENERATED -- DO NOT EDIT!\n")
+        for (model, flipped), stats in results.iterrows():
+            cmd = util.newcommand(
+                "%sThetaTimeCorr%s" % (
+                    model.capitalize(),
+                    flipped.capitalize()),
+                util.latex_spearman.format(**dict(stats)))
+            fh.write(cmd)
+
     return pth
 
 

### 2014-02-05 man/cogsci-2014/cogsci-2014.tex
Update discussion

@@
 \input{../../results/D/human_corrs.tex}
+\input{../../results/D/theta_time_corrs.tex}
+\input{../../results/D/theta_accuracy_corrs.tex}
 
@@
 minimum angle of rotation was significantly monotonically correlated
-with average per-stimulus RTs, both for ``flipped'' and ``same'' pairs
-\TODO{}. While this replicates the general result of previous
-experiments \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are
-not as perfectly linear. However, \citeA{Gardony:2013gn} found a
-constant effect for ``flipped'' pairs; an average of their results and
-the classic results would produce something similar to ours.
+with average per-stimulus RTs, both for ``flipped''
+(\ExpThetaTimeCorrFlipped{}) and ``same'' pairs
+(\ExpThetaTimeCorrSame{}). While this replicates the general result of
+previous experiments \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our
+results are not as perfectly linear. However, \citeA{Gardony:2013gn}
+found a constant effect for ``flipped'' pairs; an average of their
+results and the classic results would produce something similar to
+ours.
 
@@
 participants' average per-stimulus accuracy, though significantly more
-so for ``same'' pairs (\TODO{}). This is the same result found both by
-\citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.
+so for ``same'' pairs (\ExpThetaAccuracyCorrSame{}) than ``flipped''
+pairs (\ExpThetaAccuracyCorrFlipped{}). This is the same result found
+both by \citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.
 
@@
       linear (as with the ``same'' pairs), or linear and then flat (as
-      with the ``flipped'' pairs). Middle: the \Th{} model tends to
-      have highly inconsitent within-shape RTs. Right: the \Bq{} model
-      is roughly linear.}
+      with the ``flipped'' pairs). Middle: the \Th{} model does not
+      have a reliably monotonic relationship with rotation. Right: the
+      \Bq{} model is roughly linear.}
     \label{fig:response-time-stimulus}
@@
 stopping criteria (as opposed to just finding a maximum), the prior
-$p(h)$ has a fairly significant effect, as shown in
+$p(h)$ had a fairly significant effect, as shown in
 Fig.~\ref{fig:response-time-accuracy}. With just a small bias of
-$p(h=0)=0.55$, there is a clear separation in RTs for ``same'' versus
+$p(h=0)=0.55$, there was a clear separation in RTs for ``same'' versus
 ``flipped'' stimuli; this separation is similar to the trend also
@@
 formalized four models and compared their performance with the results
-of a behavioral experiment. While it is clear that none of the models
-fully captured human behavior on the mental rotation task, it is worth
-noting where they shed light on aspects that have previously gone
-unexplained.  For example, the \Bq{} model's adaptive stopping rule is
-sensitive to the prior, and thus provides a possible explanation for
-why people are slower to respond on ``flipped'' stimulus pairs.
+of a behavioral experiment.
+
+The \Oc{} model was the best predictor of human RTs, but is an
+unsatisfying explanation of human behavior because it relies on
+\textit{a prior} knowledge that people are unlikely to have about
+which points on the two shapes correspond. Similarly, while the \Th{}
+was next best in terms of correlation, it makes the unsatisfying
+assumption that the global maximum is known. Finally, neither of these
+models can explain the difference in people's behavior on ``same'' and
+``flipped'' stimuli.  The \Bq{} model was nearly as good as the \Th{}
+model; in contrast to the other models, the \Bq{} model makes no
+assumptions about people's \textit{a prior} knowledge. Furthermore,
+the \Bq{} model's adaptive stopping rule is sensitive to the prior,
+and thus provides a possible explanation for why people are slower to
+respond on ``flipped'' stimulus pairs.
 
@@
 
-How could the models we evaluated be improved? First, all of the
-models use holistic rotations, however there is evidence that people
-might compare individual features of shapes rather than the entire
-shape at once \cite{Just1976,Yuille:1982tx}. Second, there are other
-possible ``active sampling'' approaches that could combine the best
-features from the \Th{} and \Bq{} models, such as maintaining a
-distribution over the location and value of the global maximum, rather
-than over the integral. Third, studying rotations with 3D objects of
-varying levels of complexity may be more informative than the
-extremely simple 2D objects used in this experiment.
+Thus, we suggest that the \Bq{} model offers the most promising
+explanation of people's behavior on the mental rotation task to
+date. It is the only model out of those surveyed which did not make
+assumptions about people's \textit{a priori} knowledge, which was a
+reasonable predictor of people's behavior, and which provided an
+explanation for the longer response times on ``flipped'' stimuli.
+Clearly, though, the \Bq{} model is not a perfect account, and there
+are several immediately obvious ways in which it could be
+improved. For example, while we used holistic rotations in this paper,
+there is evidence that people compare individual features of shapes
+rather than the entire shape at once
+\cite{Just1976,Yuille:1982tx}. Second, there are other possible active
+sampling approaches that could combine the best features from the
+\Th{} and \Bq{} models, such as maintaining a distribution over the
+location and value of the global maximum, rather than over the
+integral.
 
-As a parting comment, we note that the question of how people decide
-\textit{which} rotations to perform is more difficult to answer than
-previously assumed. Although we examined four well-motivated models,
-none came close to explaining the full range of human behavior on this
-task. We suggest that such metacognitive questions are fertile areas
-for future research.
+To conclude, we hypothesize that people decide \textit{what} to
+simulate through an active sampling strategy similar to that used by
+the \Bq{} model. While more research is required to confirm or
+disprove this claim, the work presented in this paper provides the
+foundation necessary to do so.
 

### 2014-02-07 analysis/analyses/num_chance.py
Use bonferroni correction in num_chance

@@ -22,23 +22,22 @@ def run(data, results_path, seed):
         if key in exclude:
             continue
         y = df.groupby(['stimulus', 'theta', 'flipped'])['correct']
-        num = (y.apply(util.beta, [0.05]) <= 0.5).sum()
-        total = len(y.groups)
-        results[key] = pd.Series({'num': num, 'total': total})
+        alpha = 0.05 / len(y.groups)
+        chance = y.apply(util.beta, [alpha]).unstack(-1)[alpha] <= 0.5
+        results[key] = chance
 
-    results = pd.DataFrame.from_dict(results, orient='index')
-    results.index.name = 'model'
+    results = pd.DataFrame.from_dict(results).stack().reset_index()
+    results.columns = ['stimulus', 'theta', 'flipped', 'model', 'chance']
+
     pth = results_path.joinpath(filename)
     results.to_csv(pth)
 
     with open(results_path.joinpath(texname), "w") as fh:
         fh.write("%% AUTOMATICALLY GENERATED -- DO NOT EDIT!\n")
-        for model, stats in results.iterrows():
-            if stats['num'] < len(words):
-                num = words[stats['num']]
-            else:
-                num = stats['num']
-
+        for model, chance in results.groupby('model')['chance']:
+            num = chance.sum()
+            if num < len(words):
+                num = words[num]
             cmd = util.newcommand(
                 "%sNumChance" % model.capitalize(), num)
             fh.write(cmd)

### 2014-02-07 analysis/analyses/overall_response_time.py
Fix misordered confidence intervals in overall_response_time

@@ -17,6 +17,11 @@ def run(data, results_path, seed):
 
     results = pd.DataFrame.from_dict(results, orient='index')
     results.index.name = 'model'
+    # these are out of order, so fix them
+    results = results.rename(columns={
+        'lower': 'upper',
+        'upper': 'lower'})
+
     pth = results_path.joinpath(filename)
     results.to_csv(pth)
 

### 2014-02-07 analysis/analyses/util.py
Fix reporting formatting

@@ -18,17 +18,17 @@ def newcommand(name, val):
     return cmd + "\n"
 
 
-report_spearman = "rs={median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
-latex_spearman = r"$r_s={median:.2f}$, 95\% CI $[{lower:.2f}, {upper:.2f}]$"
+report_spearman = "\rho={median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
+latex_spearman = r"$\rho={median:.2f}$, 95\% CI $[{lower:.2f}, {upper:.2f}]$"
 
-report_pearson = "\rho={median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
-latex_pearson = r"$\rho={median:.2f}$, 95\% CI $[{lower:.2f}, {upper:.2f}]$"
+report_pearson = "r={median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
+latex_pearson = r"$r={median:.2f}$, 95\% CI $[{lower:.2f}, {upper:.2f}]$"
 
 report_percent = "M={median:.1f}%, 95% CI [{lower:.1f}%, {upper:.1f}%]"
 latex_percent = r"$M={median:.1f}\%$, 95\% CI $[{lower:.1f}\%, {upper:.1f}\%]$"
 
-report_mean = "M={median:.1f} [{lower:.1f}, {upper:.1f}]"
-latex_mean = r"$M={median:.1f}$ $[{lower:.1f}, {upper:.1f}]$"
+report_mean = "M={median:.1f}, 95% CI [{lower:.1f}, {upper:.1f}]"
+latex_mean = r"$M={median:.1f}$ 95\% CI $[{lower:.1f}, {upper:.1f}]$"
 
 
 def run_analysis(func):

### 2014-02-07 analysis/plots/__main__.py
Fix plots __main__ behavior

@@ -1,5 +1,6 @@
 from path import path
 from termcolor import colored
+import traceback
 
 from . import __all__
 from . import *
@@ -8,12 +9,13 @@ from . import util
 root = path("../")
 config = util.load_config(root.joinpath("config.ini"))
 version = config.get("global", "version")
-data_path = root.joinpath(config.get("paths", "data"))
-data = util.load_all(version, data_path)
+results_path = root.joinpath(config.get("paths", "results"), version)
 fig_path = root.joinpath(config.get("paths", "figures"), version)
-seed = config.getint("global", "seed")
 
 for name in __all__:
     func = locals()[name]
     print colored("Executing '%s'" % name, 'blue')
-    print func.plot(data, fig_path, seed)
+    try:
+        print func.plot(results_path, fig_path)
+    except:
+        print colored(traceback.format_exc(limit=3), "red")

### 2014-02-07 analysis/plots/accuracy_scatters.py

@@ -3,7 +3,6 @@
 import matplotlib.pyplot as plt
 import pandas as pd
 import util
-from path import path
 
 
 def plot(results_path, fig_path):
@@ -54,10 +53,4 @@ def plot(results_path, fig_path):
 
 
 if __name__ == "__main__":
-    config = util.load_config("config.ini")
-    version = config.get("global", "version")
-    data_path = path(config.get("paths", "data"))
-    data = util.load_all(version, data_path)
-    fig_path = path(config.get("paths", "figures")).joinpath(version)
-    seed = config.getint("global", "seed")
-    print plot(data, fig_path, seed)
+    util.make_plot(plot)

### 2014-02-07 analysis/plots/accuracy_stimuli.py

@@ -3,7 +3,6 @@
 import matplotlib.pyplot as plt
 import util
 import pandas as pd
-from path import path
 
 
 def plot_key(key, results_path, fig_path):
@@ -48,16 +47,10 @@ def plot_key(key, results_path, fig_path):
 
 def plot(results_path, fig_path):
     pths = []
-    for key in ['exp', 'hc', 'bq']:
+    for key in ['exp', 'hc', 'bq', 'bqp']:
         pths.extend(plot_key(key, results_path, fig_path))
     return pths
 
 
 if __name__ == "__main__":
-    config = util.load_config("config.ini")
-    version = config.get("global", "version")
-    data_path = path(config.get("paths", "data"))
-    data = util.load_all(version, data_path)
-    fig_path = path(config.get("paths", "figures")).joinpath(version)
-    seed = config.getint("global", "seed")
-    print plot(data, fig_path, seed)
+    util.make_plot(plot)

### 2014-02-07 analysis/plots/human_model_corrs.py

@@ -4,7 +4,6 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import util
-from path import path
 
 
 def plot(results_path, fig_path):
@@ -111,9 +110,4 @@ def plot(results_path, fig_path):
 
 
 if __name__ == "__main__":
-    config = util.load_config("config.ini")
-    version = config.get("global", "version")
-    data_path = path(config.get("paths", "data"))
-    data = util.load_all(version, data_path)
-    fig_path = path(config.get("paths", "figures")).joinpath(version)
-    print plot(data, fig_path)
+    util.make_plot(plot)

### 2014-02-07 analysis/plots/response_time_accuracy.py

@@ -4,7 +4,6 @@ import matplotlib.pyplot as plt
 import numpy as np
 import util
 import pandas as pd
-from path import path
 
 
 def plot(results_path, fig_path):
@@ -121,9 +120,4 @@ def plot(results_path, fig_path):
 
 
 if __name__ == "__main__":
-    config = util.load_config("config.ini")
-    version = config.get("global", "version")
-    data_path = path(config.get("paths", "data"))
-    data = util.load_all(version, data_path)
-    fig_path = path(config.get("paths", "figures")).joinpath(version)
-    print plot(data, fig_path)
+    util.make_plot(plot)

### 2014-02-07 analysis/plots/response_time_histograms.py

@@ -4,7 +4,6 @@ import matplotlib.pyplot as plt
 import util
 import pickle
 import numpy as np
-from path import path
 
 
 def plot(results_path, fig_path):
@@ -26,11 +25,7 @@ def plot(results_path, fig_path):
     for i, key in enumerate(order):
         ax = axes[i]
 
-        if key == 'exp':
-            bins = 100
-        else:
-            bins = times[key].ptp() + 1
-
+        bins = min(100, times[key].ptp() + 1)
         ax.hist(np.asarray(times[key]), bins=bins, color='k')
         ax.set_title(titles[key], fontsize=14)
 
@@ -59,8 +54,4 @@ def plot(results_path, fig_path):
 
 
 if __name__ == "__main__":
-    config = util.load_config("config.ini")
-    version = config.get("global", "version")
-    results_path = path(config.get("paths", "results")).joinpath(version)
-    fig_path = path(config.get("paths", "figures")).joinpath(version)
-    print plot(results_path, fig_path)
+    util.make_plot(plot)

### 2014-02-07 analysis/plots/response_time_scatters.py

@@ -3,7 +3,6 @@
 import matplotlib.pyplot as plt
 import pandas as pd
 import util
-from path import path
 
 
 def plot(results_path, fig_path):
@@ -73,8 +72,4 @@ def plot(results_path, fig_path):
 
 
 if __name__ == "__main__":
-    config = util.load_config("config.ini")
-    version = config.get("global", "version")
-    results_path = path(config.get("paths", "results")).joinpath(version)
-    fig_path = path(config.get("paths", "figures")).joinpath(version)
-    print plot(results_path, fig_path)
+    util.make_plot(plot)

### 2014-02-07 analysis/plots/response_time_stimuli.py

@@ -3,7 +3,6 @@
 import matplotlib.pyplot as plt
 import util
 import pandas as pd
-from path import path
 
 
 def plot_key(key, results_path, fig_path):
@@ -48,17 +47,10 @@ def plot_key(key, results_path, fig_path):
 
 def plot(results_path, fig_path):
     pths = []
-    for key in ['exp', 'th', 'hc', 'bq']:
+    for key in ['exp', 'th', 'hc', 'bq', 'bqp']:
         pths.extend(plot_key(key, results_path, fig_path))
     return pths
 
 
 if __name__ == "__main__":
-    config = util.load_config("config.ini")
-    version = config.get("global", "version")
-    data_path = path(config.get("paths", "data"))
-    data = util.load_all(version, data_path)
-    fig_path = path(config.get("paths", "figures")).joinpath(version)
-    seed = config.getint("global", "seed")
-
-    print plot(data, fig_path, seed)
+    util.make_plot(plot)

### 2014-02-07 analysis/plots/response_time_stimulus.py

@@ -4,7 +4,6 @@ import matplotlib.pyplot as plt
 import numpy as np
 import util
 import pandas as pd
-from path import path
 
 
 def plot(results_path, fig_path):
@@ -100,9 +99,4 @@ def plot(results_path, fig_path):
 
 
 if __name__ == "__main__":
-    config = util.load_config("config.ini")
-    version = config.get("global", "version")
-    data_path = path(config.get("paths", "data"))
-    data = util.load_all(version, data_path)
-    fig_path = path(config.get("paths", "figures")).joinpath(version)
-    print plot(data, fig_path)
+    util.make_plot(plot)

### 2014-02-07 analysis/plots/theta_accuracy_corrs.py

@@ -4,7 +4,6 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import util
-from path import path
 
 
 def plot(results_path, fig_path):
@@ -89,9 +88,4 @@ def plot(results_path, fig_path):
 
 
 if __name__ == "__main__":
-    config = util.load_config("config.ini")
-    version = config.get("global", "version")
-    data_path = path(config.get("paths", "data"))
-    data = util.load_all(version, data_path)
-    fig_path = path(config.get("paths", "figures")).joinpath(version)
-    print plot(data, fig_path)
+    util.make_plot(plot)

### 2014-02-07 analysis/plots/trial_accuracy.py

@@ -3,7 +3,6 @@
 import matplotlib.pyplot as plt
 import util
 import pandas as pd
-from path import path
 
 
 def plot(results_path, fig_path):
@@ -37,10 +36,4 @@ def plot(results_path, fig_path):
 
 
 if __name__ == "__main__":
-    config = util.load_config("config.ini")
-    version = config.get("global", "version")
-    data_path = path(config.get("paths", "data"))
-    data = util.load_human(version, data_path)[1]
-    fig_path = path(config.get("paths", "figures")).joinpath(version)
-    seed = config.getint("global", "seed")
-    print plot(data, fig_path, seed)
+    util.make_plot(plot)

### 2014-02-07 analysis/plots/trial_time.py

@@ -3,7 +3,6 @@
 import matplotlib.pyplot as plt
 import util
 import pandas as pd
-from path import path
 
 
 def plot(results_path, fig_path):
@@ -36,10 +35,4 @@ def plot(results_path, fig_path):
 
 
 if __name__ == "__main__":
-    config = util.load_config("config.ini")
-    version = config.get("global", "version")
-    data_path = path(config.get("paths", "data"))
-    data = util.load_human(version, data_path)[1]
-    fig_path = path(config.get("paths", "figures")).joinpath(version)
-    seed = config.getint("global", "seed")
-    print plot(data, fig_path, seed)
+    util.make_plot(plot)

### 2014-02-07 analysis/plots/util.py

@@ -2,6 +2,7 @@ from ConfigParser import SafeConfigParser
 import matplotlib.pyplot as plt
 import os
 import sys
+from path import path
 
 from mental_rotation.analysis import load_human, load_all, zscore
 
@@ -262,4 +263,12 @@ def sync_ylabel_coords(axes, x, y=0.5):
         ax.yaxis.set_label_coords(x, y)
 
 
-report_pearson = r"$\rho$ = {median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
+report_pearson = r"$r$ = {median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
+
+
+def make_plot(func):
+    config = load_config("config.ini")
+    version = config.get("global", "version")
+    results_path = path(config.get("paths", "results")).joinpath(version)
+    fig_path = path(config.get("paths", "figures")).joinpath(version)
+    print func(results_path, fig_path)

### 2014-02-07 man/cogsci-2014/cogsci-2014.tex
Misc small changes (including Fig.-->Figure and Eq.-->Equation)

@@
   images to generate? To answer this question, we explored how mental
-  simulation should used in the classic psychological task of
+  simulation should be used in the classic psychological task of
   determining if two images depict the same object in different
@@
     that the question of ``what to simulate?'' is more difficult than
-    has previously been assumed.} While only exploratory, this work
-  provides the foundation necessary to begin answering this question.
+    has previously been assumed. While only exploratory, this work
+    provides the foundation necessary to begin answering this
+    question.}
 
@@
 this task by ``mentally rotating'' the objects until they are
-congruent.  This explanation leaves several questions unanswered,
-however. How do people know the axis around which to rotate the
+congruent.  However, this explanation leaves several questions
+unanswered. How do people know the axis around which to rotate the
 objects? If the axis is known, how do people know which direction to
@@
 similarity. Example similarity curves are shown in
-Fig.~\ref{fig:model-traces}.
+Figure~\ref{fig:model-traces}.
 
@@
 To summarize, we approximate the likelihood term of
-Eq.~\ref{eq:lh-ratio} using the similarity function defined in
-Eq.~\ref{eq:similarity}. Because we assume mental rotations are
+Equation~\ref{eq:lh-ratio} using the similarity function defined in
+Equation~\ref{eq:similarity}. Because we assume mental rotations are
 performed sequentially, this similarity can only be computed for the
-actions listed in Eq.~\ref{eq:actions}.
+actions listed in Equation~\ref{eq:actions}.
 
@@
 
-In order to approximate Eq.~\ref{eq:lh-ratio} using samples of the
-similarity function, we must decide \textit{which} places to sample
-and \textit{when} stop sampling. The models below differ in how they
-make these decisions (see Fig.~\ref{fig:model-traces} for examples of
-their behavior).
+In order to approximate Equation~\ref{eq:lh-ratio} using samples of
+the similarity function, we must decide \textit{which} places to
+sample and \textit{when} stop sampling. The models below differ in how
+they make these decisions (see Figure~\ref{fig:model-traces} for
+examples of their behavior).
 
@@
 hypothesis. This approach gives us the true value of $\theta$, so
-Eq.~\ref{eq:lh-ratio} becomes a generalized likelihood ratio test,
-where $\theta$ is set to the MLE value, rather than being
+Equation~\ref{eq:lh-ratio} becomes a generalized likelihood ratio
+test, where $\theta$ is set to the MLE value, rather than being
 marginalized:
@@
 correspond to the true $\theta$ (or points near the true
-$\theta$). So, as with the \Oc{} model, Eq.~\ref{eq:lh-ratio} becomes
-a likelihood ratio.
+$\theta$). So, as with the \Oc{} model, Equation~\ref{eq:lh-ratio}
+becomes a likelihood ratio.
 
@@
 for a \textit{local} maximum. As with the \Oc{} and \Th{} models,
-Eq.~\ref{eq:lh-ratio} becomes a generalized likelihood ratio test.
+Equation~\ref{eq:lh-ratio} becomes a generalized likelihood ratio
+test.
 
@@
 for the global maximum, the goal is really to just approximate
-Eq.~\ref{eq:lh-ratio} while still obeying the constraint of sequential
-rotations. Based on evidence that people use ``active sampling''
-\cite<e.g.,>{Nelson2007,Gureckis:2012gu}, we hypothesize a model that
-actively monitors the usefulness of the rotations it is
+Equation~\ref{eq:lh-ratio} while still obeying the constraint of
+sequential rotations. Based on evidence that people use ``active
+sampling'' \cite<e.g.,>{Nelson2007,Gureckis:2012gu}, we hypothesize a
+model that actively monitors the usefulness of the rotations it is
 performing. Instead of searching for a maximum, we maintain a
 probability distribution over our \textit{estimate} of
-Eq.~\ref{eq:lh-ratio}, and then sample actions which are expected to
-improve that estimate.  This strategy has several benefits. First, it
-does not make assumptions about the shape or scale of the similarity
-function. Second, by choosing to sample places which are informative,
-this method implicitly minimizes the amount of rotation.
+Equation~\ref{eq:lh-ratio}, and then sample actions which are expected
+to improve that estimate.  This strategy has several benefits. First,
+it does not make assumptions about the shape or scale of the
+similarity function. Second, by choosing to sample places which are
+informative, this method implicitly minimizes the amount of rotation.
 
@@
 in the machine-learning literature as \textit{Bayesian Quadrature}
-\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}, or BQ.
+\cite{Diaconis:1988uo,Osborne:2012tm}, or BQ.
 
@@
 \vert\ \mathbb{E}[Z_h], \mathbb{V}(Z_h))$, which gives us a
-distribution over the likelihood ratio in Eq.~\ref{eq:lh-ratio}:
+distribution over the likelihood ratio in Equation~\ref{eq:lh-ratio}:
 \begin{equation}
@@
 We compute this expected variance for each of the actions in
-Eq.~\ref{eq:actions}, and then choose the action with the lowest
+Equation~\ref{eq:actions}, and then choose the action with the lowest
 value.
@@
 We randomly generated 20 shapes of five or six vertices (e.g.,
-Fig.~\ref{fig:model-traces}). For each shape, we computed 20 ``same''
-and 20 ``flipped'' stimuli pairs, with 18 rotations ($\theta$) spaced
-at $20^\circ$ increments between $0^\circ$ and $360^\circ$ (with
-$0^\circ$ and $180^\circ$ repeated twice, in order to gather an equal
-number of responses for each angle between $0^\circ$ and
-$180^\circ$). ``Same'' pairs were created by rotating $\Xa$ by
-$\theta$; ``flipped'' pairs were first reflected $\Xa$ across the
-$y$-axis, then rotated by $\theta$.
+Figure~\ref{fig:model-traces}). For each shape, we computed 20
+``same'' and 20 ``flipped'' stimuli pairs, with 18 rotations
+($\theta$) spaced at $20^\circ$ increments between $0^\circ$ and
+$360^\circ$ (with $0^\circ$ and $180^\circ$ repeated twice, in order
+to gather an equal number of responses for each angle between
+$0^\circ$ and $180^\circ$). ``Same'' pairs were created by rotating
+$\Xa$ by $\theta$; ``flipped'' pairs were first reflected $\Xa$ across
+the $y$-axis, then rotated by $\theta$.
 
@@
     \includegraphics[width=\textwidth]{../../figures/D/response_time_scatters.pdf}
-    \caption{\textbf{Model vs. human RTs.}}
+    \caption{\textbf{Model vs. human RTs.} Each subplot shows the
+      z-scored model response times ($x$-axis) vs. the z-scored human
+      response times ($y$-axis). Pearson correlations are shown
+      beneath each subplot. The dotted lines are $x=y$.}
     \label{fig:human-model-scatters}
@@
 $p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the ``unequal''
-prior). This only had an effect on the stopping criteria for the \Bq{}
-model.\footnote{The prior also had a small effect on the accuracy of
-  the \Hc{} model, because sometimes the local maxima were small
-  enough that Eq.~\ref{eq:mle-lh-ratio} became dominated by the
-  prior.}
+prior). This only had a major effect on the stopping criteria for the
+\Bq{} model.\footnote{The prior also had a small effect on the
+  accuracy of the \Hc{} model, because sometimes the local maxima were
+  small enough that Equation~\ref{eq:mle-lh-ratio} became dominated by
+  the prior.}
 
@@
 samples to compute the confidence intervals around both Spearman
-($r_s$) and Pearson ($\rho$) correlations.  Unless otherwise
-specified, all correlations were computed over 720 datapoints (one for
-each stimulus). For analyses of accuracy, confidence intervals were
+($\rho$) and Pearson ($r$) correlations.  Unless otherwise specified,
+all correlations were computed over 720 datapoints (one for each
+stimulus). For analyses of accuracy, confidence intervals were
 computed from a binomial proportion with a Jeffrey's beta prior.  To
@@
 stimulus, we used the same binomial proportion, and simply tested
-whether $p>0.05$ for the 50\% mark.
+whether $p>\frac{0.05}{720}$ for the 50\% mark, where the denominator
+is a Bonferroni correction for multiple comparisons.
 
@@
 The average RT across all stimuli was \ExpTime{} milliseconds.  The
-minimum angle of rotation was significantly monotonically correlated
-with average per-stimulus RTs, both for ``flipped''
+minimum angle of rotation was significantly rank-order (Spearman)
+correlated with average per-stimulus RTs, both for ``flipped''
 (\ExpThetaTimeCorrFlipped{}) and ``same'' pairs
@@
 
-To gauge between-subject reliability, we randomly split subjects into
-two groups, computed the average RT and accuracy per stimulus for each
-group, and then computed correlations for these two measures. We
-repeated this procedure 100 times, and from those samples computed the
-median correlations and 95\% confidence intervals. For RT, the
-between-subjects correlation was \ExpTimeCorr{}, and for accuracy, it
-was \ExpAccuracyCorr{}.
+To gauge between-participant reliability, we randomly split
+participants into two groups, computed the average RT and accuracy per
+stimulus for each group, and then computed correlations for these two
+measures. We repeated this procedure 100 times, and from those samples
+computed the median correlations and 95\% confidence intervals. For
+RT, the between-participants correlation was \ExpTimeCorr{}, and for
+accuracy, it was \ExpAccuracyCorr{}.
 
@@
 perfectly correlated with the minimum angle of rotation
-(Fig.~\ref{fig:response-time-accuracy}). Unsurprisingly, the \Oc{}
+(Figure~\ref{fig:response-time-accuracy}). Unsurprisingly, the \Oc{}
 model was the best fit to human RTs, with a correlation of
-\OcTimeCorr{} (Fig.~\ref{fig:human-model-scatters}).  However, the
+\OcTimeCorr{} (Figure~\ref{fig:human-model-scatters}).  However, the
 \Oc{} model was 100\% accurate, and therefore the correlation between
@@
 of rotation and the number of actions taken by the \Th{} model
-(Fig.~\ref{fig:response-time-accuracy}). However, the relationship
+(Figure~\ref{fig:response-time-accuracy}). However, the relationship
 between angle and RT for individual shapes was usually not monotonic
-(e.g., Fig.~\ref{fig:response-time-stimulus}).  The \Th{} model was
+(e.g., Figure~\ref{fig:response-time-stimulus}).  The \Th{} model was
 able to explain a moderate amount of the variance in human RTs, with a
 correlation of \ThTimeCorr{}
-(Fig.~\ref{fig:human-model-scatters}). Like the \Oc{} model, the \Th{}
-model had 100\% accuracy, and thus there was no correlation between
-minimum angle and accuracy.
+(Figure~\ref{fig:human-model-scatters}). Like the \Oc{} model, the
+\Th{} model had 100\% accuracy, and thus there was no correlation
+between minimum angle and accuracy.
 
@@
 \textit{over}rotate (see, for example, the ``flipped'' pair trace in
-Fig.~\ref{fig:model-traces}). This is because the step size is large
+Figure~\ref{fig:model-traces}). This is because the step size is large
 enough that it sometimes misses the global maximum, and must do
@@
       plots correspond to the shape shown in
-      Fig.~\ref{fig:model-traces}. Left: human curves are either
+      Figure~\ref{fig:model-traces}. Left: human curves are either
       linear (as with the ``same'' pairs), or linear and then flat (as
@@
 relationship between rotation and RT
-(Fig.~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model was
-barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
+(Figure~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model
+was barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
 individual stimuli for which it was not above chance. The \Hc{} model
 was not a particularly good predictor of human RTs (\HcTimeCorr{}), as
-shown in Fig.~\ref{fig:human-model-scatters}. It was a moderately good
-predictor of human accuracy (\HcAccuracyCorr{}).
+shown in Figure~\ref{fig:human-model-scatters}. It was a moderately
+good predictor of human accuracy (\HcAccuracyCorr{}).
 
@@
 relationship between rotation and the number of steps taken by the
-\Bq{} model (Fig.~\ref{fig:response-time-accuracy}). Unlike the \Th{}
-model, this relationship existed for individual shapes as well (e.g.,
-Fig.~\ref{fig:response-time-stimulus}).  The \Bq{} model explained
-variance in human RTs about as well as the \Th{} model
-(Fig.~\ref{fig:human-model-scatters}), with a correlation of
+\Bq{} model (Figure~\ref{fig:response-time-accuracy}). Unlike the
+\Th{} model, this relationship existed for individual shapes as well
+(e.g., Figure~\ref{fig:response-time-stimulus}).  The \Bq{} model
+explained variance in human RTs about as well as the \Th{} model
+(Figure~\ref{fig:human-model-scatters}), with a correlation of
 \BqTimeCorr{} for the equal prior and \BqpTimeCorr{} for the unequal
@@
 (\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
-there was \BqNumChance{} stimulus for which it was not above chance;
+there were \BqNumChance{} stimulii for which it was not above chance;
 with the unequal prior, there were \BqpNumChance{}. The \Bq{} model
-explained about the same amount of variance in people's accuracy as
-the \Hc{} model (equal prior: \BqAccuracyCorr{}; unequal prior:
+had about the same correlation with people's accuracy as the \Hc{}
+model (equal prior: \BqAccuracyCorr{}; unequal prior:
 \BqpAccuracyCorr{}).
 
-Because the \Bq{} model relies on Eq.~\ref{eq:lh-ratio} for its
+Because the \Bq{} model relies on Equation~\ref{eq:lh-ratio} for its
 stopping criteria (as opposed to just finding a maximum), the prior
 $p(h)$ had a fairly significant effect, as shown in
-Fig.~\ref{fig:response-time-accuracy}. With just a small bias of
+Figure~\ref{fig:response-time-accuracy}. With just a small bias of
 $p(h=0)=0.55$, there was a clear separation in RTs for ``same'' versus
@@
 unsatisfying explanation of human behavior because it relies on
-\textit{a prior} knowledge that people are unlikely to have about
+\textit{a priori} knowledge that people are unlikely to have about
 which points on the two shapes correspond. Similarly, while the \Th{}
@@
 reveals several interesting points
-(Fig.~\ref{fig:response-time-stimulus} shows curves for one particular
-shape; these curves are fairly typical relative to other
+(Figure~\ref{fig:response-time-stimulus} shows curves for one
+particular shape; these curves are fairly typical relative to other
 shapes). First, the curves for the human RTs on specific shapes tend
@@
 
-To conclude, we hypothesize that people decide \textit{what} to
-simulate through an active sampling strategy similar to that used by
-the \Bq{} model. While more research is required to confirm or
-disprove this claim, the work presented in this paper provides the
-foundation necessary to do so.
-
 %% TODO: Uncomment this for final version:

### 2014-02-07 man/cogsci-2014/cogsci-2014.tex
Remove paragraph on inter-subject reliability

@@
 
-To gauge between-participant reliability, we randomly split
-participants into two groups, computed the average RT and accuracy per
-stimulus for each group, and then computed correlations for these two
-measures. We repeated this procedure 100 times, and from those samples
-computed the median correlations and 95\% confidence intervals. For
-RT, the between-participants correlation was \ExpTimeCorr{}, and for
-accuracy, it was \ExpAccuracyCorr{}.
-
 There was a significant effect of trial number both on RT

### 2014-02-07 analysis/analyses/all_response_times.py
Only use correct response times for histograms

@@ -9,7 +9,7 @@ filename = "all_response_times.pkl"
 def run(data, results_path, seed):
     results = {}
     for key, df in data.iteritems():
-        results[key] = df['time']
+        results[key] = df[df['correct']]['time']
 
     pth = results_path.joinpath(filename)
     with open(pth, "w") as fh:

### 2014-02-07 analysis/plots/response_time_histograms.py
Shorten x-axis of histograms

@@ -29,12 +29,14 @@ def plot(results_path, fig_path):
         if key == 'exp':
             hist, edges = np.histogram(
                 times[key] / 1000., bins=bins,
-                range=(0, times[key].max() / 1000.))
+                range=(0, 20))
         else:
             hist, edges = np.histogram(
                 times[key], bins=bins,
                 range=(0, bins))
 
+        edges = edges[:101]
+        hist = hist[:100]
         hist = hist * 100 / float(len(times[key]))
         width = edges[1] - edges[0]
         ax.bar(edges[:-1], hist, width=width, color='k')
@@ -46,10 +48,9 @@ def plot(results_path, fig_path):
         util.outward_ticks(ax)
 
         if key == 'exp':
-            ax.set_xlabel("Response time", fontsize=14)
+            ax.set_xlabel("RT (seconds)", fontsize=14)
         else:
             ax.set_xlabel("Number of actions", fontsize=14)
-
     axes[0].set_ylabel("Percent", fontsize=14)
 
     fig.set_figheight(2.5)

### 2014-02-07 man/cogsci-2014/cogsci-2014.tex
Add histogram figure to paper

@@
 
+\begin{figure*}[t]
+  \begin{center}
+    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
+    \caption{\textbf{Example stimuli and model behavior.} On the left
+      is a ``flipped'' stimulus pair with a rotation of
+      $120^\circ$. The grey plots show the actions that each model
+      took for this stimulus. The color of each point (action)
+      indicates whether the model's current mental image was generated
+      under the hypothesis that the objects are the same (red) or
+      flipped (blue). The similarity functions for each of these
+      hypotheses are shown on the right.}
+    \label{fig:model-traces}
+  \end{center}
+\end{figure*}
+
 Previous models of mental rotation have largely focused on the
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
-    \caption{\textbf{Example stimuli and model behavior.} On the left
-      are two stimulus pairs, one ``same'' (top) and one ``flipped''
-      (bottom), both with a rotation of $120^\circ$. The grey plots
-      show the actions that each model took for these stimuli. The
-      color of each point (action) indicates whether the model's
-      current mental image was generated under the hypothesis that the
-      objects are the same (red) or flipped (blue). The similarity
-      functions for each of these hypotheses are shown on the right.}
-    \label{fig:model-traces}
+    \includegraphics[width=\textwidth]{../../figures/D/response_time_accuracy.pdf}
+    \caption{\textbf{Response time and accuracy comparison.} Top: RT
+      of correct responses as a function of the minimum angle of
+      rotation. Bottom: accuracy as a function of the minimum angle of
+      rotation. All error bars are 95\% confidence intervals.}
+    \label{fig:response-time-accuracy}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/response_time_accuracy.pdf}
-    \caption{\textbf{Response time and accuracy comparison.} Top: RT
-      of correct responses as a function of the minimum angle of
-      rotation. Bottom: accuracy as a function of the minimum angle of
-      rotation. All error bars are 95\% confidence intervals.}
-    \label{fig:response-time-accuracy}
+    \includegraphics[width=\textwidth]{../../figures/D/response_time_histograms.pdf}
+    \caption{\textbf{Response time histograms.} Each subplot shows a
+      histogram of RTs for people (left) or one of the models (only
+      for trials which were answered correctly).}
+    \label{fig:histograms}
   \end{center}

### 2014-02-07 analysis/analyses/overall_response_time.py
Only use correct responses for overall response time

@@ -13,7 +13,8 @@ def run(data, results_path, seed):
 
     results = {}
     for name in sorted(data.keys()):
-        results[name] = 1. / util.bootstrap_mean(1. / data[name]['time'])
+        correct = data[name][data[name]['correct']]
+        results[name] = 1. / util.bootstrap_mean(1. / correct['time'])
 
     results = pd.DataFrame.from_dict(results, orient='index')
     results.index.name = 'model'

### 2014-02-07 man/cogsci-2014/cogsci-2014.tex
Should be theta prime, not theta circ
Add references to histograms

@@
 % space below "Figure 1: ...", but only for inline figures
-%\addtolength{\textfloatsep}{-0.5cm}
+\addtolength{\textfloatsep}{-0.5cm}
+
+\addtolength{\abovecaptionskip}{-0.25cm}
+\addtolength{\belowcaptionskip}{-0.2cm}
 
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/response_time_histograms.pdf}
-    \caption{\textbf{Response time histograms.} Each subplot shows a
-      histogram of RTs for people (left) or one of the models (only
-      for trials which were answered correctly).}
-    \label{fig:histograms}
+    \includegraphics[width=\textwidth]{../../figures/D/response_time_scatters.pdf}
+    \caption{\textbf{Model vs. human RTs.} Each subplot shows the
+      z-scored model response times ($x$-axis) vs. the z-scored human
+      response times ($y$-axis). Pearson correlations are shown
+      beneath each subplot. The dotted lines are $x=y$.}
+    \label{fig:human-model-scatters}
   \end{center}
@@
   \mathbb{V}(Z_h) &\approx \iint \mathrm{Cov}_h(\theta,
-  \theta^\circ)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\circ)p(\theta)p(\theta^\circ)\dif\theta\dif\theta^\circ,
+  \theta^\prime)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\prime)p(\theta)p(\theta^\prime)\dif\theta\dif\theta^\prime,
 \end{align}
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/response_time_scatters.pdf}
-    \caption{\textbf{Model vs. human RTs.} Each subplot shows the
-      z-scored model response times ($x$-axis) vs. the z-scored human
-      response times ($y$-axis). Pearson correlations are shown
-      beneath each subplot. The dotted lines are $x=y$.}
-    \label{fig:human-model-scatters}
+    \includegraphics[width=\textwidth]{../../figures/D/response_time_histograms.pdf}
+    \caption{\textbf{Response time histograms.} Each subplot shows the
+      distribution of RTs on correct trials for people and the
+      models.}
+    \label{fig:histograms}
   \end{center}
@@
 test if participants or a model was above chance on a particular
-stimulus, we used the same binomial proportion, and simply tested
-whether $p>\frac{0.05}{720}$ for the 50\% mark, where the denominator
-is a Bonferroni correction for multiple comparisons.
+stimulus, we used the same binomial proportion, and tested whether the
+cumulative probability of answering correctly more than 50\% of the
+time was greater than $1-\frac{0.05}{720}$, where the denominator is a
+Bonferroni correction for multiple comparisons.
 
@@
 
-The average RT across all stimuli was \ExpTime{} milliseconds.  The
-minimum angle of rotation was significantly rank-order (Spearman)
-correlated with average per-stimulus RTs, both for ``flipped''
-(\ExpThetaTimeCorrFlipped{}) and ``same'' pairs
-(\ExpThetaTimeCorrSame{}). While this replicates the general result of
-previous experiments \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our
-results are not as perfectly linear. However, \citeA{Gardony:2013gn}
-found a constant effect for ``flipped'' pairs; an average of their
-results and the classic results would produce something similar to
-ours.
+The average RT across all correctly-judged stimuli was \ExpTime{}
+milliseconds; the full histogram of response times can be seen in
+Figure~\ref{fig:histograms}. The minimum angle of rotation was
+significantly rank-order (Spearman) correlated with average
+per-stimulus RTs, both for ``flipped'' (\ExpThetaTimeCorrFlipped{})
+and ``same'' pairs (\ExpThetaTimeCorrSame{}). While this replicates
+the general result of previous experiments
+\cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are not as
+perfectly linear.
 
@@
 model was the best fit to human RTs, with a correlation of
-\OcTimeCorr{} (Figure~\ref{fig:human-model-scatters}).  However, the
-\Oc{} model was 100\% accurate, and therefore the correlation between
-minimum angle and accuracy was undefined.
+\OcTimeCorr{} (Figure~\ref{fig:human-model-scatters}).  However,
+distribution of response times did not match that of people
+(Figure~\ref{fig:histograms}). Moreover, the \Oc{} model was 100\%
+accurate, and therefore could not explain the relationship between
+rotation and accuracy which was demonstrated by people.
 
@@
 correlation of \ThTimeCorr{}
-(Figure~\ref{fig:human-model-scatters}). Like the \Oc{} model, the
-\Th{} model had 100\% accuracy, and thus there was no correlation
+(Figure~\ref{fig:human-model-scatters}). Like the \Oc{} model,
+however, the overall distribution of its response times did not match
+that of people (Figure~\ref{fig:histograms}). Additionally, the \Th{}
+model had 100\% accuracy, and thus did not exhibit a relationship
 between minimum angle and accuracy.
@@
 \BqTimeCorr{} for the equal prior and \BqpTimeCorr{} for the unequal
-prior.
+prior. The RT distribution from the \Bq{} model had the same overall
+shape as that of people (Figure~\ref{fig:histograms}).
 
@@
 (\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
-there were \BqNumChance{} stimulii for which it was not above chance;
+there were \BqNumChance{} stimuli for which it was not above chance;
 with the unequal prior, there were \BqpNumChance{}. The \Bq{} model

### 2014-02-07 man/cogsci-2014/cogsci-2014.tex
Tightening, etc.

@@
 % space below "Figure 1: ...", but only for inline figures
-\addtolength{\textfloatsep}{-0.5cm}
+%\addtolength{\textfloatsep}{-0.5cm}
 
 \addtolength{\abovecaptionskip}{-0.25cm}
-\addtolength{\belowcaptionskip}{-0.2cm}
+\addtolength{\belowcaptionskip}{-0.3cm}
 
@@
   mental rotation, we formalized four models and compared them to
-  human performance. \TODO{None of the models were a particularly
-    satisfying explanation of human behavior, leading us to suggest
-    that the question of ``what to simulate?'' is more difficult than
-    has previously been assumed. While only exploratory, this work
-    provides the foundation necessary to begin answering this
-    question.}
+  human performance. We found that three of the models, which were
+  based on previous hypotheses in the literature, were unable to
+  account for several aspects of human behavior. The fourth, which is
+  a model based on the idea of \textit{active sampling}
+  \cite<e.g.,>{Gureckis:2012gu}, provides plausible explanations where
+  the other models do not. Based on these results, we suggest that the
+  question of ``what to simulate?'' is more difficult than has
+  previously been assumed, and that an active learning approach holds
+  promise for uncovering the answer.
 
@@
 One of the most astonishing cognitive feats is our ability to
-envision, manipulate, and plan with objects -- all without actually
-perceiving them. This capacity for ``mental simulation'' has been
-widely studied, including an intense debate about the underlying
+envision, manipulate, and plan with objects---all without actually
+perceiving them. This \textit{mental simulation} has been widely
+studied, including an intense debate about the underlying
 representation of mental images \cite<e.g.,>{Kosslyn:2009tj,
@@
 differed by a reflection and a rotation). They found that people's
-response times had a strong linear correlation with the minimum angle
-of rotation, a result which led to the conclusion that people solve
-this task by ``mentally rotating'' the objects until they are
+response times (RTs) had a strong linear correlation with the minimum
+angle of rotation, a result which led to the conclusion that people
+solve this task by ``mentally rotating'' the objects until they are
 congruent.  However, this explanation leaves several questions
@@
 not necessarily maintain a single trajectory of rotation, we explore
-an alternative rational model, which--rather than computing the angle
-of rotation ahead of time--engages in an \textit{active sampling}
-strategy.
+an alternative model, which---rather than explicitly computing the
+angle of rotation---engages in an \textit{active sampling} strategy.
 
@@
   \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
-      f(\Xt, 0, \epsilon) &\mbox{ continue rotating by $\epsilon$,} \\
+      f(\Xt, 0, \epsilon) &\mbox{ rotate by $\epsilon$ radians,} \\
       f(\Xt, 1, 0) &\mbox{ flip,} \\
@@
 where $\epsilon\sim \left|\mathcal{N}(0, \sigma_\epsilon^2)\right|$
-and $\sigma_\epsilon^2$ is the variance of the step size in radians.
+and $\sigma_\epsilon^2$ is the variance of the step size.
 
@@
 sample and \textit{when} stop sampling. The models below differ in how
-they make these decisions (see Figure~\ref{fig:model-traces} for
-examples of their behavior).
+they make these decisions; see Figure~\ref{fig:model-traces} for an
+example.
 
@@
 a valid rotation matrix; the $h$ that does is the correct
-hypothesis. This approach gives us the true value of $\theta$, so
+hypothesis. This gives us the true value of $\theta$, so
 Equation~\ref{eq:lh-ratio} becomes a generalized likelihood ratio
@@
     \caption{\textbf{Model vs. human RTs.} Each subplot shows the
-      z-scored model response times ($x$-axis) vs. the z-scored human
-      response times ($y$-axis). Pearson correlations are shown
-      beneath each subplot. The dotted lines are $x=y$.}
+      z-scored model RTs ($x$-axis) vs. the z-scored human RTs
+      ($y$-axis). Pearson correlations are shown beneath each
+      subplot. The dotted lines are $x=y$.}
     \label{fig:human-model-scatters}
@@
 optimization strategy. However, these strategies would not result in
-the linear response time found by \citeA{Shepard1971}. A second
-alternative, which could potentially demonstrate the linear response
-time result, might instead just perform a Hill Climbing (\Hc{}) search
-for a \textit{local} maximum. As with the \Oc{} and \Th{} models,
+the linear RT found by \citeA{Shepard1971}. A second alternative might
+instead just perform a Hill Climbing (\Hc{}) search for a
+\textit{local} maximum. As with the \Oc{} and \Th{} models,
 Equation~\ref{eq:lh-ratio} becomes a generalized likelihood ratio
@@
   and is based on personal communication with one of the authors
-  \cite{Duvenaud:2013td}.} Then, from \citeA{Duvenaud:2013td}:
-\begin{align}
-  \mathbb{E}[Z_h] &\approx \int \bar{\mu}_h(\theta)p(\theta)\dif\theta,\\
-  \mathbb{V}(Z_h) &\approx \iint \mathrm{Cov}_h(\theta,
-  \theta^\prime)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\prime)p(\theta)p(\theta^\prime)\dif\theta\dif\theta^\prime,
-\end{align}
+  \cite{Duvenaud:2013td}.} Then, from \citeA{Duvenaud:2013td}, we have
+$\mathbb{E}[Z_h] \approx \int \bar{\mu}_h(\theta)p(\theta)\dif\theta$
+and $\mathbb{V}(Z_h) \approx \iint \mathrm{Cov}_h(\theta,
+\theta^\prime)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\prime)p(\theta)p(\theta^\prime)\dif\theta\dif\theta^\prime$,
 where $\bar{\mu}_h:=\mu(\bar{S}_h)$ is the mean of the second GP, and
@@
 
-While there was no limit on response time, participants were urged to
-respond as quickly as possible while still being
-accurate. Specifically, we asked them to aim for at least 85\%
-accuracy in the experimental blocks, and provided a counter to keep
-them informed of their score.
+While there was no limit on RT, participants were urged to respond as
+quickly as possible while still being accurate. Specifically, we asked
+them to aim for at least 85\% accuracy in the experimental blocks, and
+provided a counter to keep them informed of their score.
 
@@
 Out of the 247 participants, 200 (81\%) were included in our
-analyses. Of the rest, we excluded 10 (4\%) because of an experimental
-error, 6 (2.4\%) because they had already completed a related
-experiment, and 31 (12.6\%) because they failed a comprehension
-check. Comprehension was defined as correctly answering at least 85\%
-of stimuli with a minimum rotation of $0^\circ$ or $20^\circ$.
-Additionally, we excluded 82 trials for which the response time (RT)
-was either less than 100 milliseconds or greater than 20 seconds.
+analyses. Of the other 47, we excluded 10 (4\%) because of an
+experimental error, 6 (2.4\%) because they had already completed a
+related experiment, and 31 (12.6\%) because they failed a
+comprehension check. Comprehension was defined as correctly answering
+at least 85\% of stimuli with a minimum rotation of $0^\circ$ or
+$20^\circ$.  Additionally, we excluded 82 trials for which the RT was
+either less than 100 milliseconds or greater than 20 seconds.
 
@@
 prior). This only had a major effect on the stopping criteria for the
-\Bq{} model.\footnote{The prior also had a small effect on the
-  accuracy of the \Hc{} model, because sometimes the local maxima were
-  small enough that Equation~\ref{eq:mle-lh-ratio} became dominated by
-  the prior.}
+\Bq{} model.
 
@@
 The average RT across all correctly-judged stimuli was \ExpTime{}
-milliseconds; the full histogram of response times can be seen in
+milliseconds; the full histogram of RTs can be seen in
 Figure~\ref{fig:histograms}. The minimum angle of rotation was
@@
 (Figure~\ref{fig:human-model-scatters}). Like the \Oc{} model,
-however, the overall distribution of its response times did not match
-that of people (Figure~\ref{fig:histograms}). Additionally, the \Th{}
-model had 100\% accuracy, and thus did not exhibit a relationship
-between minimum angle and accuracy.
+however, the overall distribution of its RTs did not match that of
+people (Figure~\ref{fig:histograms}). Additionally, the \Th{} model
+had 100\% accuracy, and thus did not exhibit a relationship between
+minimum angle and accuracy.
 
@@
 enough that it sometimes misses the global maximum, and must do
-another full rotation to find it. However, when the step size is small
-enough to always find the maximum, the fit to human data is slightly
-worse.
+another full rotation to find it. However, when the step size is
+smaller, the fit to human data is slightly worse.
 
@@
 
-The \Oc{} model was the best predictor of human RTs, but is an
-unsatisfying explanation of human behavior because it relies on
-\textit{a priori} knowledge that people are unlikely to have about
-which points on the two shapes correspond. Similarly, while the \Th{}
-was next best in terms of correlation, it makes the unsatisfying
-assumption that the global maximum is known. Finally, neither of these
-models can explain the difference in people's behavior on ``same'' and
-``flipped'' stimuli.  The \Bq{} model was nearly as good as the \Th{}
-model; in contrast to the other models, the \Bq{} model makes no
-assumptions about people's \textit{a prior} knowledge. Furthermore,
-the \Bq{} model's adaptive stopping rule is sensitive to the prior,
-and thus provides a possible explanation for why people are slower to
-respond on ``flipped'' stimulus pairs.
+The \Oc{} and \Th{} models were---in terms of correlations---the best
+predictors of human RTs. However, both are somewhat unsatisfying
+explanations because they rely on \textit{a priori} knowledge that
+people are unlikely to have. Moreover, they offer no explanation of
+several aspects of human behavior. First, their overall RT
+distributions look nothing like people's
+(Figure~\ref{fig:histograms}).  Second, they both are 100\% accurate,
+and so cannot explain the systematic relationship between rotation and
+accuracy found in people
+(Figure~\ref{fig:response-time-accuracy}). Third, neither model can
+explain the difference in people's behavior on ``same'' and
+``flipped'' stimuli.
 
-Additionally, a closer look at performance on individual shapes
-reveals several interesting points
-(Figure~\ref{fig:response-time-stimulus} shows curves for one
-particular shape; these curves are fairly typical relative to other
-shapes). First, the curves for the human RTs on specific shapes tend
-to either be linear, or linear first, then flat. Second, while the
-\Th{} model appears to capture the concavity of human RTs overall, it
-is actually fairly inconsistent within shapes. Third, the linearity of
-the \Bq{} model's RT curves \textit{is} consistent within shapes,
-though it does not exhibit the same ``plateau'' effect that sometimes
-appears in the human curves.
+In contrast, the \Bq{} model was nearly as good as the \Th{} model,
+yet it makes no assumptions about people's \textit{a priori}
+knowledge. Furthermore, the \Bq{} model matches people's behavior
+better than the \Oc{} or \Th{} models in several ways. Its overall RT
+histogram has the same general shape as people's
+(Figure~\ref{fig:histograms}).  Moreover, a closer look shows that the
+\Bq{} model maintains the monotonic relationship between angle and RT
+even on individual stimuli, while the \Th{} model does not
+(Figure~\ref{fig:response-time-stimulus}). Finally, the \Bq{} model's
+adaptive stopping rule is sensitive to the prior, and thus provides a
+possible explanation for why people are slower to respond on
+``flipped'' stimulus pairs.
 
@@
 explanation of people's behavior on the mental rotation task to
-date. It is the only model out of those surveyed which did not make
-assumptions about people's \textit{a priori} knowledge, which was a
-reasonable predictor of people's behavior, and which provided an
-explanation for the longer response times on ``flipped'' stimuli.
-Clearly, though, the \Bq{} model is not a perfect account, and there
-are several immediately obvious ways in which it could be
-improved. For example, while we used holistic rotations in this paper,
-there is evidence that people compare individual features of shapes
-rather than the entire shape at once
-\cite{Just1976,Yuille:1982tx}. Second, there are other possible active
-sampling approaches that could combine the best features from the
-\Th{} and \Bq{} models, such as maintaining a distribution over the
-location and value of the global maximum, rather than over the
-integral.
+date. While it is not a perfect account, there are several ways in
+which it could be improved. For example, while we used holistic
+rotations in this paper, there is evidence that people compare
+individual features of shapes
+\cite{Just1976,Yuille:1982tx}. Additionally, a different active
+sampling approach could maintain a distribution over the location and
+value of the global maximum, rather than over the integral. We intend
+to explore these possibilities in future work, building upon the
+foundation established in this paper and working towards a better
+understanding of \textit{what} people should choose to simulate.
 

### 2014-02-07 analysis/plots/response_time_scatters.py
Make response time scatters a little shorter

@@ -57,12 +57,12 @@ def plot(results_path, fig_path):
     axes[0].set_ylim(-4, 4)
     axes[0].set_ylabel("Human", fontsize=14)
 
-    fig.set_figheight(3)
+    fig.set_figheight(2.5)
     fig.set_figwidth(16)
     plt.draw()
     plt.tight_layout()
 
-    plt.subplots_adjust(wspace=0.1)
+    plt.subplots_adjust(wspace=0.1, bottom=0.2, top=0.9)
 
     pths = [fig_path.joinpath("response_time_scatters.%s" % ext)
             for ext in ('png', 'pdf')]

### 2014-02-07 man/cogsci-2014/cogsci-2014.tex
More tightening, wording

@@
 
-  \textbf{Keywords:} mental rotation, computational model
+  \textbf{Keywords:} mental rotation, computational modeling
 \end{abstract}
@@
   Pylyshyn:2002vk}. But this debate hasn't addressed one of the most
-fundamental questions about mental simulation: how people decide what
-they should simulate.
+fundamental questions about mental simulation: how people decide
+\textit{what} to simulate.
 
@@
 describe a behavioral experiment based on the classic mental rotation
-studies \cite<e.g.,>{Cooper:1975wp}, and compare the results of the
+studies \cite<e.g.,>{Cooper:1975wp}, and compare the results of our
 experiment with each of the models. We conclude with a discussion of
@@
 \textit{which} mental images to generate. \citeA{Kosslyn:1977tv}
-proposed a model of the mental imagery buffer, but not the planning
-process of how it is used. Similarly, \citeA{Julstrom:1985va} and
+proposed a model of the mental imagery buffer, but did not say
+\textit{how} it should be used. Similarly, \citeA{Julstrom:1985va} and
 \citeA{Glasgow:1992tj} were mostly concerned with modeling the
@@
 points on the two objects can be easily identified, which is not
-necessarily the case.
-
-Indeed, the state-of-the-art in computer vision suggests that there is
-more to shape matching than simply checking for congruency,
-particularly when shapes are complex or when the shapes might not be
-exactly the same
+necessarily the case.  Indeed, the state-of-the-art in computer vision
+suggests that there is more to this problem than simply checking for
+congruency, particularly when shapes are complex or when the shapes
+might not be exactly the same
 \cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}. Additionally, recent
 research shows that when performing \textit{physical} rotations,
-people do not rotate until congruency is reached, and may even rotate
-\textit{away} from near perfect matches \cite{Gardony:2013gn}.
+people do not fully rotate until congruency is reached; they may even
+rotate \textit{away} from initially near perfect matches
+\cite{Gardony:2013gn}.
 
@@
 \begin{equation}
-  p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta, h)p(X_a)p(h)p(\theta)\dif\theta,
+  p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta, h)p(h)p(\theta)\dif\theta,
   \label{eq:posterior}
@@
 where $p(X_b\ \vert\ X_a, \theta, h)$ is the probability that $X_b$
-was generated from $X_a$, and $p(X_a)$ is the probability of observing
-$X_a$. Because we ultimately want to determine which hypothesis is
-more likely, the quantity of interest is a ratio $\ell:=p(\hi\ \vert\
-X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which (assuming that all
-rotations are equally likely) is equivalent to:
+was generated from $X_a$. Because we ultimately want to determine
+which hypothesis is more likely, the quantity of interest is a ratio
+$\ell:=p(\hi\ \vert\ X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which
+(assuming that all rotations are equally likely) is equivalent to:
 \begin{equation}
@@
 S(\Xb, f(\Xa, h, \theta))$.  We do not know which vertices of $\Xb$
-correspond to which vertices of $\Xa$, so $S$ must marginalize over
-the set of possible mappings. For brevity, let
+correspond to which vertices of $\Xa$, so the similarity $S$ must
+marginalize over the set of possible mappings. For brevity, let
 $\mathbf{X}_m=\M\cdot{}f(\Xa, h, \theta)$ where $\M$ is a permutation
@@
 
-Additionally, we assume that the observed shapes may only be
-transformed by a small amount at a time, and each transformation is
-costly in terms of computational resources. Specifically, if the
-current mental image is $\Xt$, then:
+Additionally, we assume that the observed shapes can only be
+transformed by a small amount at a time, and each transformation takes
+a non-negligible amount of time. Specifically, if the current mental
+image is $\Xt$, then:
 \begin{equation}
@@
 model which obeys this constraint first rotate in the direction that
-increases similarity, and then stop rotating once a ``match'' is
-found, as determined by a threshold on the value of $S$. If all
+increases local similarity, and then stop rotating once a ``match'' is
+found, as determined by a threshold on the similarity $S$. If all
 rotations have been exhausted, then flip, and try rotating again. We
@@
 to improve that estimate.  This strategy has several benefits. First,
-it does not make assumptions about the shape or scale of the
-similarity function. Second, by choosing to sample places which are
-informative, this method implicitly minimizes the amount of rotation.
+it does not make assumptions about the scale of the similarity
+function. Second, by choosing to sample places which are informative,
+this method implicitly minimizes the amount of rotation.
 
@@
 test if participants or a model was above chance on a particular
-stimulus, we used the same binomial proportion, and tested whether the
-cumulative probability of answering correctly more than 50\% of the
-time was greater than $1-\frac{0.05}{720}$, where the denominator is a
-Bonferroni correction for multiple comparisons.
+stimulus, we used the same binomial proportion, and tested whether
+$p\left(p(\textrm{correct})\leq 0.5\right)\leq \frac{0.05}{720}$,
+where $\frac{1}{720}$ is a Bonferroni correction for multiple
+comparisons.
 
@@
 As noted above, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This
-has the interesting effect of actually causing the \Th{} to
+has the interesting effect of actually causing the \Th{} model to
 \textit{over}rotate (see, for example, the ``flipped'' pair trace in
@@
 \subsubsection{\Hc{} model}
+
+The \Hc{} was the only model for which there was no monotonic
+relationship between rotation and RT
+(Figure~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model
+was barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
+individual stimuli for which it was not above chance. The \Hc{} model
+was not a particularly good predictor of human RTs (\HcTimeCorr{}), as
+shown in Figure~\ref{fig:human-model-scatters}. It was a moderately
+good predictor of human accuracy (\HcAccuracyCorr{}).
+
+\subsubsection{\Bq{} model}
 
@@
 \end{figure}
-
-The \Hc{} was the only model for which there was no monotonic
-relationship between rotation and RT
-(Figure~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model
-was barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
-individual stimuli for which it was not above chance. The \Hc{} model
-was not a particularly good predictor of human RTs (\HcTimeCorr{}), as
-shown in Figure~\ref{fig:human-model-scatters}. It was a moderately
-good predictor of human accuracy (\HcAccuracyCorr{}).
-
-\subsubsection{\Bq{} model}
 

### 2014-02-07 analysis/analyses/util.py
Add comma to mean

@@ -28,7 +28,7 @@ report_percent = "M={median:.1f}%, 95% CI [{lower:.1f}%, {upper:.1f}%]"
 latex_percent = r"$M={median:.1f}\%$, 95\% CI $[{lower:.1f}\%, {upper:.1f}\%]$"
 
 report_mean = "M={median:.1f}, 95% CI [{lower:.1f}, {upper:.1f}]"
-latex_mean = r"$M={median:.1f}$ 95\% CI $[{lower:.1f}, {upper:.1f}]$"
+latex_mean = r"$M={median:.1f}$, 95\% CI $[{lower:.1f}, {upper:.1f}]$"
 
 
 def run_analysis(func):

### 2014-02-07 man/cogsci-2014/cogsci-2014.tex
More tightening

@@
   mental rotation, we formalized four models and compared them to
-  human performance. We found that three of the models, which were
-  based on previous hypotheses in the literature, were unable to
-  account for several aspects of human behavior. The fourth, which is
-  a model based on the idea of \textit{active sampling}
-  \cite<e.g.,>{Gureckis:2012gu}, provides plausible explanations where
-  the other models do not. Based on these results, we suggest that the
-  question of ``what to simulate?'' is more difficult than has
-  previously been assumed, and that an active learning approach holds
-  promise for uncovering the answer.
+  human performance. We found that three models based on previous
+  hypotheses in the literature were unable to account for several
+  aspects of human behavior. The fourth, which is a model based on the
+  idea of \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu},
+  provides plausible explanations where the other models do not. Based
+  on these results, we suggest that the question of ``what to
+  simulate?'' is more difficult than has previously been assumed, and
+  that an active learning approach holds promise for uncovering the
+  answer.
 
@@
 gained support in several areas of cognitive science
-\cite<e.g.,>{Nelson2007,Gureckis:2012gu}, including other spatial
-domains \cite{Juni:2011vo}. In the case of mental rotation, if people
-do not know the minimum angle beforehand, then actively performing
-rotations may be the best way to gather evidence about the similarity
-between the observed shapes.
+\cite<e.g.,>{Gureckis:2012gu}, including other spatial domains
+\cite{Juni:2011vo}. In the case of mental rotation, if people do not
+know the minimum angle beforehand, then actively performing rotations
+may be the best way to gather evidence about the similarity between
+the observed shapes.
 
@@
 depict mirror-image objects. The posterior probability of each
-hypothesis given the observed shapes is then:
-\begin{equation}
-  p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta, h)p(h)p(\theta)\dif\theta,
-  \label{eq:posterior}
-\end{equation}
+hypothesis given the observed shapes is then: $p(h\ \vert\ X_a, X_b)
+\propto \int p(X_b\ \vert\ X_a, \theta, h)p(h)p(\theta)\dif\theta$,
 where $p(X_b\ \vert\ X_a, \theta, h)$ is the probability that $X_b$
@@
 One hypothesis is that people compute the direction and extent of
-rotation beforehand using \textit{a prior} knowledge of the
+rotation beforehand using \textit{a priori} knowledge of the
 correspondence between points in the images
@@
 
-What if the point correspondences are unknown? An intuitive and simple
-model which obeys this constraint first rotate in the direction that
-increases local similarity, and then stop rotating once a ``match'' is
-found, as determined by a threshold on the similarity $S$. If all
-rotations have been exhausted, then flip, and try rotating again. We
-assume that the locations where $S$ is greater than the threshold
-correspond to the true $\theta$ (or points near the true
-$\theta$). So, as with the \Oc{} model, Equation~\ref{eq:lh-ratio}
-becomes a likelihood ratio.
+What if the point correspondences are unknown? A model which obeys
+this constraint could first rotate in the direction that increases
+local similarity, and then stop rotating once a ``match'' is found, as
+determined by a threshold on the similarity $S$. If all rotations have
+been exhausted, then flip, and try rotating again. We assume that the
+locations where $S$ is greater than the threshold correspond to the
+true $\theta$ (or points near the true $\theta$). So, as with the
+\Oc{} model, we use Equation~\ref{eq:mle-lh-ratio}.
 
@@
 instead just perform a Hill Climbing (\Hc{}) search for a
-\textit{local} maximum. As with the \Oc{} and \Th{} models,
-Equation~\ref{eq:lh-ratio} becomes a generalized likelihood ratio
-test.
+\textit{local} maximum. Thus, as with the \Oc{} and \Th{} models, we
+use Equation~\ref{eq:mle-lh-ratio}.
 
@@
 
-While the previous few models have all focused on \textit{searching}
-for the global maximum, the goal is really to just approximate
-Equation~\ref{eq:lh-ratio} while still obeying the constraint of
-sequential rotations. Based on evidence that people use ``active
-sampling'' \cite<e.g.,>{Nelson2007,Gureckis:2012gu}, we hypothesize a
-model that actively monitors the usefulness of the rotations it is
-performing. Instead of searching for a maximum, we maintain a
-probability distribution over our \textit{estimate} of
-Equation~\ref{eq:lh-ratio}, and then sample actions which are expected
-to improve that estimate.  This strategy has several benefits. First,
-it does not make assumptions about the scale of the similarity
-function. Second, by choosing to sample places which are informative,
-this method implicitly minimizes the amount of rotation.
+While the previous few models all focused on \textit{searching} for
+the global maximum, we need only approximate
+Equation~\ref{eq:lh-ratio} with a set of sequential rotations. Based
+on evidence that people use \textit{active sampling}
+\cite<e.g.,>{Gureckis:2012gu}, we hypothesize a model that actively
+monitors the usefulness of the rotations it is performing. Instead of
+searching for a maximum, we maintain a probability distribution over
+our \textit{estimate} of Equation~\ref{eq:lh-ratio}, and then sample
+actions which are expected to improve that estimate.  This strategy
+has several benefits. First, it does not make assumptions about the
+scale of the similarity function. Second, by choosing to sample places
+which are informative, this method implicitly minimizes the amount of
+rotation.
 
 Formally, we denote $Z_h$ as our estimate of the likelihood for
-hypothesis $h$, and write its distribution as:
-\begin{equation}
-  p(Z_h) = \int \left[\int S(\Xb, f(\Xa, \theta, h))p(\theta)\dif\theta\right] p(S)\dif S,
-\end{equation}
-where $S$ is the similarity function, and $p(S)$ is a prior over
-similarity functions.  This method of estimating an integral is known
-in the machine-learning literature as \textit{Bayesian Quadrature}
-\cite{Diaconis:1988uo,Osborne:2012tm}, or BQ.
+hypothesis $h$, and write its distribution as: $p(Z_h) = \int
+\left[\int S(\Xb, f(\Xa, \theta, h))p(\theta)\dif\theta\right]
+p(S)\dif S$, where $S$ is the similarity function, and $p(S)$ is a
+prior over similarity functions.  This method of estimating an
+integral is known in the machine-learning literature as
+\textit{Bayesian Quadrature} \cite{Diaconis:1988uo,Osborne:2012tm}, or
+BQ.
 
@@
 over points sampled from the log-GP, which we denote as
-$\bar{S}_h:=\exp(\mu_h)$.\footnote{Note that the approach used here
-  differs slightly from the one presented in \citeA{Osborne:2012tm}
-  and is based on personal communication with one of the authors
-  \cite{Duvenaud:2013td}.} Then, from \citeA{Duvenaud:2013td}, we have
+$\bar{S}_h:=\exp(\mu_h)$. Then, from \citeA{Duvenaud:2013td}, we have
 $\mathbb{E}[Z_h] \approx \int \bar{\mu}_h(\theta)p(\theta)\dif\theta$
@@
 $p(Z_D)\propto\mathcal{N}(\mathbb{E}[Z_0] - \mathbb{E}[Z_1],
-\mathbb{V}(Z_0) + \mathbb{V}(Z_1))$.
-
-We then sample new observations until we are at least 95\% confident
-that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
-$\hi$, and when $p(Z_D<0)>0.975$, we accept $\hf$. To determine which
-points to sample, we can compute the expected variance of $Z_h$ if we
-sampled a new observation at $\theta_a$. From \citeA{Osborne:2012tm}:
-\begin{align}
-\mathbb{E}&[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) + \mathbb{E}[Z_h] - \\
-&\int \mathbb{E}[Z_h|\theta_{a}]^2 \mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a, \theta_a))\dif\log S_h(\theta_a).\nonumber
-\label{ref:expected-variance}
-\end{align}
-We compute this expected variance for each of the actions in
+\mathbb{V}(Z_0) + \mathbb{V}(Z_1))$.  We then sample new observations
+until we are at least 95\% confident that $Z_D\neq 0$. In other words,
+when $p(Z_D<0)<0.025$, we accept $\hi$, and when $p(Z_D<0)>0.975$, we
+accept $\hf$. To determine which points to sample, we can compute the
+expected variance of $Z_h$ if we sampled a new observation at
+$\theta_a$. From \citeA{Osborne:2012tm}, we compute the expected
+variance as $\mathbb{E}[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) +
+\mathbb{E}[Z_h] - \int \mathbb{E}[Z_h|\theta_{a}]^2
+\mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a,
+qc\theta_a))\dif\log S_h(\theta_a)$ for each of the actions in
 Equation~\ref{eq:actions}, and then choose the action with the lowest
@@
 On each trial, participants were instructed to press the `b' key to
-begin and to focus on the fixation cross that appeared for 750
-milliseconds afterwards. The two images were then presented
-side-by-side, each at 300px $\times$ 300px, and participants could
-press `s' to indicate they thought the images depicted the ``same''
-object, or `d' to indicate they thought the images depicted
-``flipped'' objects.
+begin and to focus on the fixation cross that appeared for 750ms
+afterwards. The two images were then presented side-by-side, each at
+300px $\times$ 300px, and participants could press `s' to indicate
+they thought the images depicted the ``same'' object, or `d' to
+indicate they thought the images depicted ``flipped'' objects.
 
@@
 $20^\circ$.  Additionally, we excluded 82 trials for which the RT was
-either less than 100 milliseconds or greater than 20 seconds.
+either less than 100ms or greater than 20s.
 
@@
 For analyses of RT, confidence intervals around harmonic means were
-computed using a bootstrap analysis using 10000 bootstrap samples
+computed using a bootstrap analysis of 10000 bootstrap samples
 (sampled with replacement). RTs were only considered for correct
@@
 test if participants or a model was above chance on a particular
-stimulus, we used the same binomial proportion, and tested whether
+stimulus, we used the same binomial proportion and tested whether
 $p\left(p(\textrm{correct})\leq 0.5\right)\leq \frac{0.05}{720}$,
@@
 
-The average RT across all correctly-judged stimuli was \ExpTime{}
-milliseconds; the full histogram of RTs can be seen in
+The average RT across all correctly-judged stimuli was \ExpTime{} ms;
+the full histogram of RTs can be seen in
 Figure~\ref{fig:histograms}. The minimum angle of rotation was
@@
 the general result of previous experiments
-\cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are not as
-perfectly linear.
+\cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are not as linear
+(Figure~\ref{fig:response-time-accuracy}).
 
@@
 
-As we would expect, the number of actions taken by the \Oc{} model was
-perfectly correlated with the minimum angle of rotation
-(Figure~\ref{fig:response-time-accuracy}). Unsurprisingly, the \Oc{}
-model was the best fit to human RTs, with a correlation of
-\OcTimeCorr{} (Figure~\ref{fig:human-model-scatters}).  However,
-distribution of response times did not match that of people
+The number of actions taken by the \Oc{} model was perfectly
+correlated with the minimum angle of rotation
+(Figure~\ref{fig:response-time-accuracy}). The \Oc{} model was the
+best fit to human RTs, with a correlation of \OcTimeCorr{}
+(Figure~\ref{fig:human-model-scatters}).  However, the distribution of
+response times did not match that of people
 (Figure~\ref{fig:histograms}). Moreover, the \Oc{} model was 100\%
-accurate, and therefore could not explain the relationship between
-rotation and accuracy which was demonstrated by people.
+accurate, and therefore could not explain the effect of rotation on
+people's accuracy.
 
@@
 
-As noted above, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This
-has the interesting effect of actually causing the \Th{} model to
-\textit{over}rotate (see, for example, the ``flipped'' pair trace in
-Figure~\ref{fig:model-traces}). This is because the step size is large
-enough that it sometimes misses the global maximum, and must do
-another full rotation to find it. However, when the step size is
-smaller, the fit to human data is slightly worse.
+As noted, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This has
+the interesting effect of actually causing the \Th{} model to
+\textit{over}rotate (see Figure~\ref{fig:model-traces}). This is
+because the step size is large enough that it sometimes misses the
+global maximum, and must do another full rotation to find it. However,
+when the step size is smaller, the fit to human data is slightly
+worse.
 
@@
 \subsubsection{\Bq{} model}
+
+Like the \Oc{} and \Th{} models, there was an overall monotonic
+relationship between rotation and the number of steps taken by the
+\Bq{} model (Figure~\ref{fig:response-time-accuracy}). Unlike the
+\Th{} model, this relationship existed for individual shapes as well
+(e.g., Figure~\ref{fig:response-time-stimulus}).  The \Bq{} model
+explained variance in human RTs about as well as the \Th{} model
+(Figure~\ref{fig:human-model-scatters}), with a correlation of
+\BqTimeCorr{} for the equal prior and \BqpTimeCorr{} for the unequal
+prior, and the RT distribution from the \Bq{} model had the same
+overall shape as that of people (Figure~\ref{fig:histograms}).
+
+The \Bq{} model was quite accurate overall (equal prior:
+(\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
+there were \BqNumChance{} stimuli for which it was not above chance;
+with the unequal prior, there were \BqpNumChance{}. The \Bq{} model
+had about the same correlation with people's accuracy as the \Hc{}
+model (equal prior: \BqAccuracyCorr{}; unequal prior:
+\BqpAccuracyCorr{}).
+
+Because the \Bq{} model relies on Equation~\ref{eq:lh-ratio} for its
+stopping criteria (as opposed to just finding a maximum), the prior
+$p(h)$ had a fairly significant effect
+(Figure~\ref{fig:response-time-accuracy}). With just a small bias of
+$p(h=0)=0.55$, there was a clear separation in RTs for ``same'' versus
+``flipped'' stimuli; this separation is similar to the trend also
+observed in human RTs. While the prior also had an effect on accuracy,
+this effect did not reflect the trends in human behavior.
+
+\section{Discussion}
 
@@
 
-Like the \Oc{} and \Th{} models, there was an overall monotonic
-relationship between rotation and the number of steps taken by the
-\Bq{} model (Figure~\ref{fig:response-time-accuracy}). Unlike the
-\Th{} model, this relationship existed for individual shapes as well
-(e.g., Figure~\ref{fig:response-time-stimulus}).  The \Bq{} model
-explained variance in human RTs about as well as the \Th{} model
-(Figure~\ref{fig:human-model-scatters}), with a correlation of
-\BqTimeCorr{} for the equal prior and \BqpTimeCorr{} for the unequal
-prior. The RT distribution from the \Bq{} model had the same overall
-shape as that of people (Figure~\ref{fig:histograms}).
-
-The \Bq{} model was quite accurate overall (equal prior:
-(\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
-there were \BqNumChance{} stimuli for which it was not above chance;
-with the unequal prior, there were \BqpNumChance{}. The \Bq{} model
-had about the same correlation with people's accuracy as the \Hc{}
-model (equal prior: \BqAccuracyCorr{}; unequal prior:
-\BqpAccuracyCorr{}).
-
-Because the \Bq{} model relies on Equation~\ref{eq:lh-ratio} for its
-stopping criteria (as opposed to just finding a maximum), the prior
-$p(h)$ had a fairly significant effect, as shown in
-Figure~\ref{fig:response-time-accuracy}. With just a small bias of
-$p(h=0)=0.55$, there was a clear separation in RTs for ``same'' versus
-``flipped'' stimuli; this separation is similar to the trend also
-observed in human RTs. While the prior also had an effect on accuracy,
-this effect did not reflect the trends in human behavior.
-
-\section{Discussion}
-
 We set out to answer the question of how people decide \textit{what}
@@
 and so cannot explain the systematic relationship between rotation and
-accuracy found in people
-(Figure~\ref{fig:response-time-accuracy}). Third, neither model can
-explain the difference in people's behavior on ``same'' and
-``flipped'' stimuli.
+human accuracy (Figure~\ref{fig:response-time-accuracy}). Third,
+neither model can explain the difference in people's behavior on
+``same'' and ``flipped'' stimuli.
 
@@
 foundation established in this paper and working towards a better
-understanding of \textit{what} people should choose to simulate.
+understanding of \textit{what} people choose to simulate.
 

### 2014-02-08 man/cogsci-2014/cogsci-2014.tex
Make model descriptions a little clearer

@@
   hypotheses in the literature were unable to account for several
-  aspects of human behavior. The fourth, which is a model based on the
-  idea of \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu},
-  provides plausible explanations where the other models do not. Based
-  on these results, we suggest that the question of ``what to
-  simulate?'' is more difficult than has previously been assumed, and
-  that an active learning approach holds promise for uncovering the
-  answer.
+  aspects of human behavior. The fourth, which is based on the idea of
+  \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu}, provides
+  plausible explanations where the other models do not. Based on these
+  results, we suggest that the question of ``what to simulate?'' is
+  more difficult than has previously been assumed, and that an active
+  learning approach holds promise for uncovering the answer.
 
@@
 
-Active sampling is the idea that when initial information is
-incomplete, people will gather new information in a manner that
-increases certainty about the problem space. This notion has recently
-gained support in several areas of cognitive science
+Active sampling is the idea people gather new information in a manner
+that increases certainty about the problem space. This notion has
+recently gained support in several areas of cognitive science
 \cite<e.g.,>{Gureckis:2012gu}, including other spatial domains
@@
 
-What if the point correspondences are unknown? A model which obeys
-this constraint could first rotate in the direction that increases
-local similarity, and then stop rotating once a ``match'' is found, as
-determined by a threshold on the similarity $S$. If all rotations have
-been exhausted, then flip, and try rotating again. We assume that the
-locations where $S$ is greater than the threshold correspond to the
-true $\theta$ (or points near the true $\theta$). So, as with the
-\Oc{} model, we use Equation~\ref{eq:mle-lh-ratio}.
+A model which does not know point correspondences could use the
+following algorithm: (1) pick a random direction; (2) take a single
+step; (3) if that step decreased similarity, then begin rotating in
+the reverse direction, otherwise continue rotating in the original
+direction; (4) continue rotating in the chosen direction until a
+``match'' is found (defined as finding a value of $S$ that exceeds a
+threshold); and (5) if no match was found, flip, and start over from
+step one.  We assume that the locations where $S$ is greater than the
+threshold correspond to the true $\theta$ (or points near the true
+$\theta$). So, as with the \Oc{} model, we use
+Equation~\ref{eq:mle-lh-ratio}.
 
@@
 optimization strategy. However, these strategies would not result in
-the linear RT found by \citeA{Shepard1971}. A second alternative might
-instead just perform a Hill Climbing (\Hc{}) search for a
-\textit{local} maximum. Thus, as with the \Oc{} and \Th{} models, we
-use Equation~\ref{eq:mle-lh-ratio}.
+the linear RT found by \citeA{Shepard1971}. A second alternative is to
+perform a Hill Climbing (\Hc{}) search; i.e., rotate in the direction
+that increases similarity until no further improvement can be
+found. In contrast with the \Th{} model, this results in arriving in a
+\textit{local} maximum (which may or may not be the global
+maximum). Thus, as with the \Oc{} and \Th{} models, we use
+Equation~\ref{eq:mle-lh-ratio}.
 
@@
 
-Out of the 247 participants, 200 (81\%) were included in our
-analyses. Of the other 47, we excluded 10 (4\%) because of an
-experimental error, 6 (2.4\%) because they had already completed a
-related experiment, and 31 (12.6\%) because they failed a
-comprehension check. Comprehension was defined as correctly answering
-at least 85\% of stimuli with a minimum rotation of $0^\circ$ or
-$20^\circ$.  Additionally, we excluded 82 trials for which the RT was
-either less than 100ms or greater than 20s.
+Of the 247 participants, 200 (81\%) were included in our analyses. Of
+the other 47, we excluded 10 (4\%) because of an experimental error, 6
+(2.4\%) because they had already completed a related experiment, and
+31 (12.6\%) because they failed a comprehension check, which was
+defined as correctly answering at least 85\% of stimuli with a minimum
+rotation of $0^\circ$ or $20^\circ$.  Additionally, we excluded 82
+trials for which the RT was either less than 100ms or greater than
+20s.
 
@@
 was barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
-individual stimuli for which it was not above chance. The \Hc{} model
-was not a particularly good predictor of human RTs (\HcTimeCorr{}), as
-shown in Figure~\ref{fig:human-model-scatters}. It was a moderately
-good predictor of human accuracy (\HcAccuracyCorr{}).
+stimuli for which it was not above chance. The \Hc{} model was not a
+good predictor of human RTs (\HcTimeCorr{}), as shown in
+Figure~\ref{fig:human-model-scatters}. It was a moderately good
+predictor of human accuracy (\HcAccuracyCorr{}).
 
@@
 
-The \Oc{} and \Th{} models were---in terms of correlations---the best
-predictors of human RTs. However, both are somewhat unsatisfying
-explanations because they rely on \textit{a priori} knowledge that
-people are unlikely to have. Moreover, they offer no explanation of
-several aspects of human behavior. First, their overall RT
-distributions look nothing like people's
-(Figure~\ref{fig:histograms}).  Second, they both are 100\% accurate,
-and so cannot explain the systematic relationship between rotation and
-human accuracy (Figure~\ref{fig:response-time-accuracy}). Third,
-neither model can explain the difference in people's behavior on
-``same'' and ``flipped'' stimuli.
+The \Oc{} and \Th{} models were the best predictors of human
+RTs. However, both are somewhat unsatisfying explanations because they
+rely on \textit{a priori} knowledge that people are unlikely to
+have. Moreover, they offer no explanation of several aspects of human
+behavior. First, their overall RT distributions look nothing like
+people's (Figure~\ref{fig:histograms}).  Second, they both are 100\%
+accurate, and so cannot explain the systematic relationship between
+rotation and human accuracy
+(Figure~\ref{fig:response-time-accuracy}). Third, neither model can
+explain the difference in people's behavior on ``same'' and
+``flipped'' stimuli.
 

### 2014-03-28 lib/mental_rotation/model/base_model.py
Allow point scaling to be customizable

@@ -192,7 +192,7 @@ class BaseModel(object):
 
         return lines
 
-    def plot_trace(self, ax, legend=True, scale_points=True):
+    def plot_trace(self, ax, legend=True, scale_points=10):
         Fi = self.F_i
         Ri = self.R_i
         ti = np.arange(Ri.size)
@@ -202,7 +202,7 @@ class BaseModel(object):
 
         if scale_points:
             Si = self.log_S_i - self._log_const
-            Si = (Si - Si.min() + 1) * 10
+            Si = (Si - Si.min() + 1) * scale_points
         else:
             Si = np.zeros(Ri.size) + 30
 

### 2014-04-29 man/cogsci-2014/cogsci-2014.tex
Updates in response to reviewer comments

@@
 \addtolength{\abovecaptionskip}{-0.25cm}
-\addtolength{\belowcaptionskip}{-0.3cm}
+\addtolength{\belowcaptionskip}{-0.35cm}
 
@@
   hypotheses in the literature were unable to account for several
-  aspects of human behavior. The fourth, which is based on the idea of
-  \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu}, provides
-  plausible explanations where the other models do not. Based on these
-  results, we suggest that the question of ``what to simulate?'' is
-  more difficult than has previously been assumed, and that an active
-  learning approach holds promise for uncovering the answer.
+  aspects of human behavior. The fourth is based on the idea
+  \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu}, which is a
+  strategy of choosing actions that will provide the most
+  informative. This last model provides a plausible account of how
+  people use mental rotation, where the other models do not. Based on
+  these results, we suggest that the question of ``what to simulate?''
+  is more difficult than has previously been assumed, and that an
+  active learning approach holds promise for uncovering the answer.
 
@@
 
-\begin{figure*}[t]
-  \begin{center}
-    \includegraphics[width=\textwidth]{../../figures/D/model-traces.pdf}
-    \caption{\textbf{Example stimuli and model behavior.} On the left
-      is a ``flipped'' stimulus pair with a rotation of
-      $120^\circ$. The grey plots show the actions that each model
-      took for this stimulus. The color of each point (action)
-      indicates whether the model's current mental image was generated
-      under the hypothesis that the objects are the same (red) or
-      flipped (blue). The similarity functions for each of these
-      hypotheses are shown on the right.}
-    \label{fig:model-traces}
-  \end{center}
-\end{figure*}
-
 Previous models of mental rotation have largely focused on the
@@
 necessarily the case.  Indeed, the state-of-the-art in computer vision
-suggests that there is more to this problem than simply checking for
-congruency, particularly when shapes are complex or when the shapes
-might not be exactly the same
-\cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}. Additionally, recent
-research shows that when performing \textit{physical} rotations,
-people do not fully rotate until congruency is reached; they may even
-rotate \textit{away} from initially near perfect matches
+suggests that there is more to this problem than checking for
+congruency, particularly when the shapes are complex or not exactly
+the same \cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}. Additionally,
+recent research shows that when performing \textit{physical}
+rotations, people do not rotate until congruency is reached; they may
+even rotate \textit{away} from initially near perfect matches
 \cite{Gardony:2013gn}.
@@
 Active sampling is the idea people gather new information in a manner
-that increases certainty about the problem space. This notion has
-recently gained support in several areas of cognitive science
+that increases certainty about the problem space. An everyday example
+of this can be observed in the game of ``20 questions'', in which one
+person thinks of a concept, and another has to guess the concept in 20
+questions or less. The first question is almost always ``person,
+place, or thing?'', because the answer provides the most possible
+information about the concept of interest. Active sampling has
+recently gained support across several areas of cognitive science
 \cite<e.g.,>{Gureckis:2012gu}, including other spatial domains
-\cite{Juni:2011vo}. In the case of mental rotation, if people do not
-know the minimum angle beforehand, then actively performing rotations
-may be the best way to gather evidence about the similarity between
-the observed shapes.
+\cite{Juni:2011vo}. In the case of mental rotation, actively choosing
+rotations may be the best way to gather evidence about the similarity
+between the observed shapes when the angle of rotation is unknown.
 
@@
 \subsection{Computational-level analysis}
+
+\begin{figure}[t]
+  \begin{center}
+    \includegraphics[width=0.48\textwidth]{../../figures/stimuli_and_similarity.pdf}
+    \caption{\textbf{Example stimuli and similarities.} This figure
+      shows a ``flipped'' stimulus pair with a rotation of
+      $120^\circ$, and the corresponding similarity functions for each
+      hypothesis. Arrows indicate where each shape lies on the curve.}
+    \label{fig:shapes}
+  \end{center}
+\end{figure}
+
+The task we are interested in modeling involves observing two images
+and determining whether one image depicts the ``same'' object as the
+other image (differing only by a rotation), or a ``flipped'' version
+of the object in the other image (differing by a reflection and then a
+rotation).
 
@@
 \begin{equation}
-  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)p(\hi)}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)p(\hf)}.
+  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)\cdot{}p_0}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)\cdot{}p_1},
   \label{eq:lh-ratio}
 \end{equation}
-If $\ell > 1$, then we accept the hypothesis that the images depict
-the same object ($\hi$); if $\ell < 1$, then we accept the hypothesis
-that the images depict flipped objects ($\hf$).
+where $p_0=p(h=0)$ and $p_1=p(h=1)$, for brevity. If $\ell > 1$, then
+we accept the hypothesis that the images depict the same object
+($\hi$); if $\ell < 1$, then we accept the hypothesis that the images
+depict flipped objects ($\hf$).
 
@@
 similarity. Example similarity curves are shown in
-Figure~\ref{fig:model-traces}.
+Figure~\ref{fig:shapes}.
 
@@
       f(\Xt, 1, 0) &\mbox{ flip,} \\
-      f(\Xa, 0, 0) &\mbox{ reset, or} \\
+      f(\Xa, 0, 0) &\mbox{ reset to $0^\circ$, or} \\
       f(\Xa, 1, 0) &\mbox{ reset and flip,} \\
@@
 sample and \textit{when} stop sampling. The models below differ in how
-they make these decisions; see Figure~\ref{fig:model-traces} for an
-example.
+they make these decisions.
 
@@
 \begin{equation}
-  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)p(\hi)}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)p(\hf)}.
+  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)\cdot{}p_0}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)\cdot{}p_1}.
   \label{eq:mle-lh-ratio}
 \end{equation}
+
+If we give equal weight to the two hypotheses, then the priors cancel
+out; if we weigh one hypothesis more heavily, then our decision will
+be biased towards that hypothesis. However, unless the likelihood
+ratio is already very close to 1, small biases in the prior will not
+make much of a difference as to whether we accept $\hi$ or $\hf$.
 
@@
 $\theta$). So, as with the \Oc{} model, we use
-Equation~\ref{eq:mle-lh-ratio}.
+Equation~\ref{eq:mle-lh-ratio}. Additionally, we only allow for the
+``flip'' action after no match has been found, because there is no
+particularly principled way for the \Th{} model to choose when to
+flip.
 
@@
 maximum). Thus, as with the \Oc{} and \Th{} models, we use
-Equation~\ref{eq:mle-lh-ratio}.
+Equation~\ref{eq:mle-lh-ratio}. We only allow for the ``flip'' action
+after a local maximum has been reached, because like the \Th{} model,
+there is otherwise no principled way for the \Hc{} model to choose
+when to flip.
 
@@
 \begin{equation}
-p(\ell)\approx\frac{\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0], \mathbb{V}(Z_0))\ p(\hi)}{\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1], \mathbb{V}(Z_1))\ p(\hf)}.
+p(\ell)\approx\frac{\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0], \mathbb{V}(Z_0))\cdot{}p_0}{\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1], \mathbb{V}(Z_1))\cdot{}p_1}.
 \end{equation}
+
 This distribution cannot easily be calculated; however, we really are
-just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. Thus, rather than
-computing $p(\ell)$, we use $Z_D=Z_0-Z_1$ and compute:
-$p(Z_D)\propto\mathcal{N}(\mathbb{E}[Z_0] - \mathbb{E}[Z_1],
-\mathbb{V}(Z_0) + \mathbb{V}(Z_1))$.  We then sample new observations
-until we are at least 95\% confident that $Z_D\neq 0$. In other words,
-when $p(Z_D<0)<0.025$, we accept $\hi$, and when $p(Z_D<0)>0.975$, we
-accept $\hf$. To determine which points to sample, we can compute the
-expected variance of $Z_h$ if we sampled a new observation at
-$\theta_a$. From \citeA{Osborne:2012tm}, we compute the expected
-variance as $\mathbb{E}[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) +
+just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. Instead, we use
+$Z_D=p_0\cdot{}Z_0-p_1\cdot{}Z_1$ and compute
+$p(Z_D)\propto\mathcal{N}(p_0\cdot{}\mathbb{E}[Z_0] -
+p_1\cdot{}\mathbb{E}[Z_1], p_0^2\cdot{}\mathbb{V}(Z_0) +
+p_1^2\cdot{}\mathbb{V}(Z_1))$.
+We then sample new observations until we are at least 95\% confident
+that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
+$\hi$, and when $p(Z_D<0)>0.975$, we accept $\hf$. Because we compare
+the hypotheses in order to determine when to stop sampling, biasing
+the prior towards one hypothesis should result in requiring less
+evidence for that hypothesis before stopping, and more evidence for
+the other hypothesis. This is in contrast to the previous models,
+where we would not expect the prior to have much effect.
+
+To determine which points to sample, we can compute the expected
+variance of $Z_h$ if we sampled a new observation at $\theta_a$. From
+\citeA{Osborne:2012tm}, we compute the expected variance as
+$\mathbb{E}[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) +
 \mathbb{E}[Z_h] - \int \mathbb{E}[Z_h|\theta_{a}]^2
 \mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a,
-qc\theta_a))\dif\log S_h(\theta_a)$ for each of the actions in
-Equation~\ref{eq:actions}, and then choose the action with the lowest
-value.
+qc\theta_a))\dif\log S_h(\theta_a)$
+for each of the actions in Equation~\ref{eq:actions}, and then choose
+the action with the lowest value.
 
@@
 We randomly generated 20 shapes of five or six vertices (e.g.,
-Figure~\ref{fig:model-traces}). For each shape, we computed 20
-``same'' and 20 ``flipped'' stimuli pairs, with 18 rotations
-($\theta$) spaced at $20^\circ$ increments between $0^\circ$ and
-$360^\circ$ (with $0^\circ$ and $180^\circ$ repeated twice, in order
-to gather an equal number of responses for each angle between
-$0^\circ$ and $180^\circ$). ``Same'' pairs were created by rotating
-$\Xa$ by $\theta$; ``flipped'' pairs were first reflected $\Xa$ across
-the $y$-axis, then rotated by $\theta$.
+Figure~\ref{fig:shapes}). For each shape, we computed 20 ``same'' and
+20 ``flipped'' stimuli pairs, with 18 rotations ($\theta$) spaced at
+$20^\circ$ increments between $0^\circ$ and $360^\circ$ (with
+$0^\circ$ and $180^\circ$ repeated twice, in order to gather an equal
+number of responses for each angle between $0^\circ$ and
+$180^\circ$). ``Same'' pairs were created by rotating $\Xa$ by
+$\theta$; ``flipped'' pairs were first reflected $\Xa$ across the
+$y$-axis, then rotated by $\theta$.
 
@@
 31 (12.6\%) because they failed a comprehension check, which was
-defined as correctly answering at least 85\% of stimuli with a minimum
-rotation of $0^\circ$ or $20^\circ$.  Additionally, we excluded 82
-trials for which the RT was either less than 100ms or greater than
-20s.
+defined as correctly answering at least 85\% of stimuli with a
+rotation of either $0^\circ$, $20^\circ$, or $340^\circ$.
+Additionally, we excluded 82 trials for which the RT was either less
+than 100ms or greater than 20s.
 
@@
 $p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the ``unequal''
-prior). This only had a major effect on the stopping criteria for the
-\Bq{} model.
+prior). As expected, this only had a major effect on the stopping
+criteria for the \Bq{} model.
 
@@
 the interesting effect of actually causing the \Th{} model to
-\textit{over}rotate (see Figure~\ref{fig:model-traces}). This is
-because the step size is large enough that it sometimes misses the
-global maximum, and must do another full rotation to find it. However,
-when the step size is smaller, the fit to human data is slightly
-worse.
+\textit{over}rotate. This is because the step size is large enough
+that it sometimes misses the global maximum, and must do another full
+rotation to find it. However, when the step size is smaller, the fit
+to human data is slightly worse.
 
@@
 $p(h)$ had a fairly significant effect
-(Figure~\ref{fig:response-time-accuracy}). With just a small bias of
-$p(h=0)=0.55$, there was a clear separation in RTs for ``same'' versus
-``flipped'' stimuli; this separation is similar to the trend also
-observed in human RTs. While the prior also had an effect on accuracy,
-this effect did not reflect the trends in human behavior.
+(Figure~\ref{fig:response-time-accuracy}). As expected, with just a
+small bias of $p(h=0)=0.55$, there was a clear separation in RTs for
+``same'' versus ``flipped'' stimuli: because of this bias, the model
+needed less evidence before accepting $h=0$ (thus taking less
+time). This separation is similar to the trend also observed in human
+RTs. The prior also had an effect on accuracy: the bias towards $h=0$
+meant that the model was more likely to judge a pair as ``same'',
+thus, accuracy increased for ``same'' pairs, but decreased for
+``flipped'' pairs. However, this effect on accuracy did not reflect
+the trends in human behavior.
 
@@
     \includegraphics[width=0.48\textwidth]{../../figures/D/response_time_stimulus.pdf}
-    \caption{\textbf{Typical RT curves for a single shape.}  These
-      plots correspond to the shape shown in
-      Figure~\ref{fig:model-traces}. Left: human curves are either
-      linear (as with the ``same'' pairs), or linear and then flat (as
-      with the ``flipped'' pairs). Middle: the \Th{} model does not
-      have a reliably monotonic relationship with rotation. Right: the
-      \Bq{} model is roughly linear.}
+    \caption{\textbf{Typical RT curves for a single object.}  These
+      plots correspond to the object shown in
+      Figure~\ref{fig:shapes}. Left: human curves are either linear
+      (as with the ``same'' pairs), or linear and then flat (as with
+      the ``flipped'' pairs). Middle: the \Th{} model does not have a
+      reliably monotonic relationship with rotation. Right: the \Bq{}
+      model is roughly linear.}
     \label{fig:response-time-stimulus}
@@
 
-% \section{Acknowledgments}
-% {\small This research was supported by ONR MURI grant number
-% N00014-13-1-0341, and a Berkeley Fellowship awarded to JBH. }
+\section{Acknowledgments} {\small This research was supported by ONR
+  MURI grant number N00014-13-1-0341, and a Berkeley Fellowship
+  awarded to JBH. }
 

### 2014-04-30 man/cogsci-2014/cogsci-2014.tex
More tweaks, to fit on 6 pages

@@
 \addtolength{\abovecaptionskip}{-0.25cm}
-\addtolength{\belowcaptionskip}{-0.35cm}
+\addtolength{\belowcaptionskip}{-0.5cm}
 
@@
 rotations, people do not rotate until congruency is reached; they may
-even rotate \textit{away} from initially near perfect matches
+even rotate \textit{away} from near perfect matches
 \cite{Gardony:2013gn}.
@@
 notion that mental images must be transformed in an analog manner (or
-in a way that is at least approximately analog), and that mental
-images are time-consuming and effortful to generate. Thus, the
-\textit{goal} is to make this determination while performing a minimum
-amount of computation (i.e., as few rotations as possible).
+in a way that is approximately analog), and that mental images are
+time-consuming and effortful to generate. Thus, the \textit{goal} is
+to make this determination while performing a minimum amount of
+computation (i.e., as few rotations as possible).
 
-The original ``congruency'' hypothesis \cite{Shepard1971} actually is
-a rational solution to this problem, in the sense that the smallest
+The original ``congruency'' hypothesis \cite{Shepard1971} is a
+rational solution to this problem, in the sense that the smallest
 amount of computation coincides with rotating through the minimum
 angle. However, it violates the constraint that we do not know the
-points of correspondence between the images, which is exactly what
+points of correspondence between the images, which is what
 necessitates the use of imagery. Noting that a rational solution need
-not necessarily maintain a single trajectory of rotation, we explore
-an alternative model, which---rather than explicitly computing the
-angle of rotation---engages in an \textit{active sampling} strategy.
+not maintain a single trajectory of rotation, we explore an
+alternative model, which---rather than computing the angle of
+rotation---engages in an \textit{active sampling} strategy.
 
@@
 information about the concept of interest. Active sampling has
-recently gained support across several areas of cognitive science
+gained support across several areas of cognitive science
 \cite<e.g.,>{Gureckis:2012gu}, including other spatial domains
@@
 the active sampling approach.
+
+The task we are interested in modeling involves observing two images
+and determining whether one image depicts the ``same'' object as the
+other image (differing by a rotation), or a ``flipped'' version of the
+object in the other image (differing by a reflection and then a
+rotation).
 
@@
 
-The task we are interested in modeling involves observing two images
-and determining whether one image depicts the ``same'' object as the
-other image (differing only by a rotation), or a ``flipped'' version
-of the object in the other image (differing by a reflection and then a
-rotation).
-
-Formally, we denote the shapes as $X_a$ and $X_b$ and assume $X_b$ is
-generated by a transformation of $X_a$, i.e. $X_b=f(X_a, \theta, h)$,
-where $\theta$ is a rotation, $\hi$ is the hypothesis that the images
-depict the same object, and $\hf$ is the hypothesis that the images
-depict mirror-image objects. The posterior probability of each
-hypothesis given the observed shapes is then: $p(h\ \vert\ X_a, X_b)
-\propto \int p(X_b\ \vert\ X_a, \theta, h)p(h)p(\theta)\dif\theta$,
+We denote the shapes as $X_a$ and $X_b$ and assume $X_b$ is generated
+by a transformation of $X_a$, i.e. $X_b=f(X_a, \theta, h)$, where
+$\theta$ is a rotation, $\hi$ is the hypothesis that the images depict
+the same object, and $\hf$ is the hypothesis that the images depict
+mirror-image objects. The posterior probability of each hypothesis
+given the observed shapes is then:
+$p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta,
+h)p(h)p(\theta)\dif\theta$,
 where $p(X_b\ \vert\ X_a, \theta, h)$ is the probability that $X_b$
-was generated from $X_a$. Because we ultimately want to determine
-which hypothesis is more likely, the quantity of interest is a ratio
+was generated from $X_a$. Because we want to determine which
+hypothesis is more likely, the quantity of interest is a ratio
 $\ell:=p(\hi\ \vert\ X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which
@@
 
-Additionally, we assume that the observed shapes can only be
-transformed by a small amount at a time, and each transformation takes
-a non-negligible amount of time. Specifically, if the current mental
-image is $\Xt$, then:
+We assume that the observed shapes must be transformed by a small
+amount at a time, and each transformation takes a non-negligible
+amount of time. If the current mental image is $\Xt$, then:
 \begin{equation}
@@
 ratio is already very close to 1, small biases in the prior will not
-make much of a difference as to whether we accept $\hi$ or $\hf$.
+make much of a difference.
 
@@
 threshold); and (5) if no match was found, flip, and start over from
-step one.  We assume that the locations where $S$ is greater than the
-threshold correspond to the true $\theta$ (or points near the true
-$\theta$). So, as with the \Oc{} model, we use
-Equation~\ref{eq:mle-lh-ratio}. Additionally, we only allow for the
-``flip'' action after no match has been found, because there is no
-particularly principled way for the \Th{} model to choose when to
-flip.
+step one.  We only allow for the ``flip'' action after no match has
+been found, because there is no particularly principled way for the
+\Th{} model to choose when to flip. We assume that the locations where
+$S$ is greater than the threshold correspond to the true $\theta$ (or
+points near the true $\theta$). So, as with the \Oc{} model, we use
+Equation~\ref{eq:mle-lh-ratio}.
 
@@
 straightforward because we know both the exact geometry of the shapes
-and that a linear transformation exists which will perfectly align
+and that a linear transformation exists which will align
 them. However, this choice is not always clear \textit{a priori}, as
@@
 identical).  One way to deal with the problem of choosing a threshold
-would simply be to perform an exhaustive search or use a global
-optimization strategy. However, these strategies would not result in
-the linear RT found by \citeA{Shepard1971}. A second alternative is to
-perform a Hill Climbing (\Hc{}) search; i.e., rotate in the direction
-that increases similarity until no further improvement can be
-found. In contrast with the \Th{} model, this results in arriving in a
-\textit{local} maximum (which may or may not be the global
-maximum). Thus, as with the \Oc{} and \Th{} models, we use
-Equation~\ref{eq:mle-lh-ratio}. We only allow for the ``flip'' action
-after a local maximum has been reached, because like the \Th{} model,
-there is otherwise no principled way for the \Hc{} model to choose
-when to flip.
+would be use a global optimization strategy; however, this would not
+result in the linear RT found by \citeA{Shepard1971}. A second
+alternative is to perform a Hill Climbing (\Hc{}) search; i.e., rotate
+in the direction that increases similarity until no further
+improvement can be found. In contrast with the \Th{} model, this
+results in arriving in a \textit{local} maximum (which may or may not
+be the global maximum). Thus, as with the \Oc{} and \Th{} models, we
+use Equation~\ref{eq:mle-lh-ratio}. We only allow for the ``flip''
+action after a local maximum has been reached, because like the \Th{}
+model, there is otherwise no principled way for the \Hc{} model to
+choose when to flip.
 
@@
 While the previous few models all focused on \textit{searching} for
-the global maximum, we need only approximate
-Equation~\ref{eq:lh-ratio} with a set of sequential rotations. Based
-on evidence that people use \textit{active sampling}
-\cite<e.g.,>{Gureckis:2012gu}, we hypothesize a model that actively
-monitors the usefulness of the rotations it is performing. Instead of
+the global maximum, we need only \textit{approximate}
+Equation~\ref{eq:lh-ratio}. We hypothesize a model based on the idea
+of \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu}: instead of
 searching for a maximum, we maintain a probability distribution over
@@
 actions which are expected to improve that estimate.  This strategy
-has several benefits. First, it does not make assumptions about the
-scale of the similarity function. Second, by choosing to sample places
-which are informative, this method implicitly minimizes the amount of
-rotation.
+has the benefits that it does not make assumptions about the scale of
+the similarity function; and, by choosing to sample places which are
+informative, this method implicitly minimizes the amount of rotation.
 
-Formally, we denote $Z_h$ as our estimate of the likelihood for
-hypothesis $h$, and write its distribution as: $p(Z_h) = \int
-\left[\int S(\Xb, f(\Xa, \theta, h))p(\theta)\dif\theta\right]
-p(S)\dif S$, where $S$ is the similarity function, and $p(S)$ is a
-prior over similarity functions.  This method of estimating an
-integral is known in the machine-learning literature as
-\textit{Bayesian Quadrature} \cite{Diaconis:1988uo,Osborne:2012tm}, or
-BQ.
-
-Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$, we first place a
-\textit{Gaussian Process} \cite{Rasmussen:2006vz}, or GP, prior on the
-log of $S_h$ in order to enforce positivity after it is exponentiated,
-i.e. $\mathbb{E}[Z_h] \approx \int
-\exp(\mu_h(\theta))p(\theta)\dif\theta$, where $\mu_h:=\mu(\log S_h)$
-is the mean of the log-GP \cite{Osborne:2012tm}.  However, this
-integral is still intractable. To approximate it, we fit a second GP
-over points sampled from the log-GP, which we denote as
+We denote $Z_h$ as our estimate of the likelihood for hypothesis $h$,
+and write its distribution as:
+$p(Z_h) = \int \left[\int S(\Xb, f(\Xa, \theta,
+  h))p(\theta)\dif\theta\right] p(S)\dif S$,
+where $S$ is the similarity function, and $p(S)$ is a prior over
+similarity functions.  This method of estimating an integral is known
+in the machine-learning literature as \textit{Bayesian Quadrature}
+\cite{Diaconis:1988uo,Osborne:2012tm}, or BQ.  Denoting
+$S_h=S(\Xb, f(\Xa, \theta, h))$, we first place a \textit{Gaussian
+  Process} \cite{Rasmussen:2006vz}, or GP, prior on the log of $S_h$
+in order to enforce positivity after it is exponentiated, i.e.
+$\mathbb{E}[Z_h] \approx \int \exp(\mu_h(\theta))p(\theta)\dif\theta$,
+where $\mu_h:=\mu(\log S_h)$ is the mean of the log-GP
+\cite{Osborne:2012tm}.  To approximate this integral, we fit a second
+GP over points sampled from the log-GP, which we denote as
 $\bar{S}_h:=\exp(\mu_h)$. Then, from \citeA{Duvenaud:2013td}, we have
 $\mathbb{E}[Z_h] \approx \int \bar{\mu}_h(\theta)p(\theta)\dif\theta$
-and $\mathbb{V}(Z_h) \approx \iint \mathrm{Cov}_h(\theta,
+and
+$\mathbb{V}(Z_h) \approx \iint \mathrm{Cov}_h(\theta,
 \theta^\prime)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\prime)p(\theta)p(\theta^\prime)\dif\theta\dif\theta^\prime$,
@@
 
-Assuming independence, we can now write $p(Z_h)\approx\mathcal{N}(Z_h\
-\vert\ \mathbb{E}[Z_h], \mathbb{V}(Z_h))$, which gives us a
-distribution over the likelihood ratio in Equation~\ref{eq:lh-ratio}:
-\begin{equation}
-p(\ell)\approx\frac{\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0], \mathbb{V}(Z_0))\cdot{}p_0}{\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1], \mathbb{V}(Z_1))\cdot{}p_1}.
-\end{equation}
-
-This distribution cannot easily be calculated; however, we really are
-just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. Instead, we use
+Assuming independence, we can now write
+$p(Z_h)\approx\mathcal{N}(Z_h\ \vert\ \mathbb{E}[Z_h],
+\mathbb{V}(Z_h))$,
+which gives us a distribution over the likelihood ratio in
+Equation~\ref{eq:lh-ratio}:
+$p(\ell)\approx\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0],
+\mathbb{V}(Z_0))\cdot{}p_0/\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1],
+\mathbb{V}(Z_1))\cdot{}p_1$.
+This distribution cannot easily be calculated, but we are only
+interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. So, we use
 $Z_D=p_0\cdot{}Z_0-p_1\cdot{}Z_1$ and compute
@@
 the hypotheses in order to determine when to stop sampling, biasing
-the prior towards one hypothesis should result in requiring less
-evidence for that hypothesis before stopping, and more evidence for
-the other hypothesis. This is in contrast to the previous models,
-where we would not expect the prior to have much effect.
+the prior should result in requiring less evidence for one hypothesis
+before stopping, and more evidence for the other hypothesis.
 
-To determine which points to sample, we can compute the expected
-variance of $Z_h$ if we sampled a new observation at $\theta_a$. From
-\citeA{Osborne:2012tm}, we compute the expected variance as
+To choose where to sample, we compute the expected variance of $Z_h$
+given a new observation at $\theta_a$. From \citeA{Osborne:2012tm}, we
+compute
 $\mathbb{E}[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) +
@@
 \mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a,
-qc\theta_a))\dif\log S_h(\theta_a)$
-for each of the actions in Equation~\ref{eq:actions}, and then choose
-the action with the lowest value.
+\theta_a))\dif\log S_h(\theta_a)$
+for each of the actions in Eq.~\ref{eq:actions}; we pick the one with
+the lowest value.
 
@@
 psiTurk experiment framework \cite{Mcdonnell12}. Each participant was
-paid \$1.00 for approximately 15 minutes of work, consisting of one
-block of 10 practice trials followed by two blocks of 100 experiment
-trials. Within a block, trials were presented in a random order.
+paid \$1.00 for 15 minutes of work, consisting of one block of 10
+practice trials followed by two blocks of 100 randomly ordered
+experiment trials.
 
@@
 five ``flipped''), such that across all blocks, each stimulus appeared
-exactly once. Each block was then split in half. Participants
-completed two half-blocks (not necessarily both from the original full
-block).
+once. Each block was then split in half, and participants completed
+two half-blocks.
 
@@
 they thought the images depicted the ``same'' object, or `d' to
-indicate they thought the images depicted ``flipped'' objects.
-
-While there was no limit on RT, participants were urged to respond as
-quickly as possible while still being accurate. Specifically, we asked
-them to aim for at least 85\% accuracy in the experimental blocks, and
-provided a counter to keep them informed of their score.
+indicate they thought the images depicted ``flipped'' objects.  While
+there was no limit on RT, we urged participants to answer as quickly
+as possible while maintaining at least 85\% accuracy in the
+experimental blocks.
 
@@
 defined as correctly answering at least 85\% of stimuli with a
-rotation of either $0^\circ$, $20^\circ$, or $340^\circ$.
-Additionally, we excluded 82 trials for which the RT was either less
-than 100ms or greater than 20s.
+rotation of either $0^\circ$, $20^\circ$, or $340^\circ$.  We also
+excluded 82 trials for which the RT was either less than 100ms or
+greater than 20s.
 
@@
 
-For analyses of RT, confidence intervals around harmonic means were
-computed using a bootstrap analysis of 10000 bootstrap samples
-(sampled with replacement). RTs were only considered for correct
-responses.  We also used a bootstrap analysis of 10000 bootstrap
-samples to compute the confidence intervals around both Spearman
-($\rho$) and Pearson ($r$) correlations.  Unless otherwise specified,
-all correlations were computed over 720 datapoints (one for each
-stimulus). For analyses of accuracy, confidence intervals were
-computed from a binomial proportion with a Jeffrey's beta prior.  To
-test if participants or a model was above chance on a particular
+For analyses of RT, confidence intervals around harmonic means of
+correct responses were computed using a bootstrap analysis of 10000
+bootstrap samples (sampled with replacement).  We also used a
+bootstrap analysis of 10000 bootstrap samples to compute the
+confidence intervals around both Spearman ($\rho$) and Pearson ($r$)
+correlations.  Unless otherwise specified, all correlations were
+computed over 720 stimuli. For analyses of accuracy, confidence
+intervals were computed from a binomial proportion with a Jeffrey's
+beta prior.  To test if judgments were above chance on a particular
 stimulus, we used the same binomial proportion and tested whether
@@
 \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are not as linear
-(Figure~\ref{fig:response-time-accuracy}).
-
-The average accuracy across all stimuli was \ExpAccuracy{}, though
-there were \ExpNumChance{} stimuli (out of 720) for which people were
-not above chance.  The minimum angle was also correlated with
-participants' average per-stimulus accuracy, though significantly more
-so for ``same'' pairs (\ExpThetaAccuracyCorrSame{}) than ``flipped''
-pairs (\ExpThetaAccuracyCorrFlipped{}). This is the same result found
-both by \citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.
-
-There was a significant effect of trial number both on RT
+(Figure~\ref{fig:response-time-accuracy}).\\
+\indent The average accuracy across all stimuli was \ExpAccuracy{},
+though there were \ExpNumChance{} stimuli (out of 720) for which
+people were not above chance.  The minimum angle was also correlated
+with participants' average per-stimulus accuracy, though much more so
+for ``same'' pairs (\ExpThetaAccuracyCorrSame{}) than ``flipped''
+pairs (\ExpThetaAccuracyCorrFlipped{}). This is the same
+result found both by \citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.\\
+\indent There was a significant effect of trial number both on RT
 (\ExpTrialTimeCorr{}) and on accuracy (\ExpTrialAccuracyCorr{}),
@@
 half of the experiment (\ExpaTrialAccuracyCorr{} for the first half
-vs. \ExpbTrialAccuracyCorr{} for the second half). These practice
-effects may have contributed to the not-quite-linearity of the human
-RTs; future work will need to collect more data per participant to
-mitigate this effect.
+vs. \ExpbTrialAccuracyCorr{} for the second half). These effects may
+have contributed to the not-quite-linearity of the human RTs; future
+work should collect more data per participant.
 
@@
 best fit to human RTs, with a correlation of \OcTimeCorr{}
-(Figure~\ref{fig:human-model-scatters}).  However, the distribution of
+(Figure~\ref{fig:human-model-scatters}), although the distribution of
 response times did not match that of people
@@
 of rotation and the number of actions taken by the \Th{} model
-(Figure~\ref{fig:response-time-accuracy}). However, the relationship
-between angle and RT for individual shapes was usually not monotonic
-(e.g., Figure~\ref{fig:response-time-stimulus}).  The \Th{} model was
-able to explain a moderate amount of the variance in human RTs, with a
+(Figure~\ref{fig:response-time-accuracy}), though this relationship
+did not hold for \textit{individual} shapes (e.g.,
+Figure~\ref{fig:response-time-stimulus}).  The \Th{} model was able to
+explain a moderate amount of the variance in human RTs, with a
 correlation of \ThTimeCorr{}
-(Figure~\ref{fig:human-model-scatters}). Like the \Oc{} model,
-however, the overall distribution of its RTs did not match that of
-people (Figure~\ref{fig:histograms}). Additionally, the \Th{} model
-had 100\% accuracy, and thus did not exhibit a relationship between
-minimum angle and accuracy.
-
-As noted, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This has
-the interesting effect of actually causing the \Th{} model to
-\textit{over}rotate. This is because the step size is large enough
-that it sometimes misses the global maximum, and must do another full
-rotation to find it. However, when the step size is smaller, the fit
-to human data is slightly worse.
+(Figure~\ref{fig:human-model-scatters}). Like the \Oc{} model, the
+overall distribution of its RTs did not match that of people
+(Figure~\ref{fig:histograms}).  The \Th{} model had 100\% accuracy,
+and thus did not exhibit a relationship between minimum angle and
+accuracy.  As noted, we fit $\sigma_\epsilon=0.6$ for the \Th{}
+model. This had the interesting effect of causing the \Th{} model to
+\textit{over}rotate, because the step size was large enough that it
+sometimes missed the global maximum, and had to do another full
+rotation to find it.
 
@@
 good predictor of human RTs (\HcTimeCorr{}), as shown in
-Figure~\ref{fig:human-model-scatters}. It was a moderately good
-predictor of human accuracy (\HcAccuracyCorr{}).
+Figure~\ref{fig:human-model-scatters}. It was a moderate predictor of
+human accuracy (\HcAccuracyCorr{}).
 
@@
 prior, and the RT distribution from the \Bq{} model had the same
-overall shape as that of people (Figure~\ref{fig:histograms}).
-
-The \Bq{} model was quite accurate overall (equal prior:
+overall shape as that of people (Figure~\ref{fig:histograms}).\\
+\indent The \Bq{} model was quite accurate overall (equal prior:
 (\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
 there were \BqNumChance{} stimuli for which it was not above chance;
-with the unequal prior, there were \BqpNumChance{}. The \Bq{} model
-had about the same correlation with people's accuracy as the \Hc{}
-model (equal prior: \BqAccuracyCorr{}; unequal prior:
-\BqpAccuracyCorr{}).
-
-Because the \Bq{} model relies on Equation~\ref{eq:lh-ratio} for its
-stopping criteria (as opposed to just finding a maximum), the prior
-$p(h)$ had a fairly significant effect
+with the unequal prior, there were \BqpNumChance{}. The correlation
+with people's accuracy was \BqAccuracyCorr{} (equal prior) and
+\BqpAccuracyCorr{} (unequal prior).\\
+\indent Because the \Bq{} model relies on Equation~\ref{eq:lh-ratio}
+for its stopping criteria (as opposed to just finding a maximum), the
+prior $p(h)$ had an observable effect
 (Figure~\ref{fig:response-time-accuracy}). As expected, with just a
@@
 time). This separation is similar to the trend also observed in human
-RTs. The prior also had an effect on accuracy: the bias towards $h=0$
-meant that the model was more likely to judge a pair as ``same'',
-thus, accuracy increased for ``same'' pairs, but decreased for
-``flipped'' pairs. However, this effect on accuracy did not reflect
-the trends in human behavior.
+RTs. The prior also had an effect on accuracy (though this did not
+reflect human behavior): the bias towards $h=0$ meant that the model
+was more likely to judge a pair as ``same'', thus, accuracy increased
+for ``same'' pairs, but decreased for ``flipped'' pairs.
 
@@
       the ``flipped'' pairs). Middle: the \Th{} model does not have a
-      reliably monotonic relationship with rotation. Right: the \Bq{}
-      model is roughly linear.}
+      monotonic relationship with rotation. Right: the \Bq{} model is
+      roughly linear.}
     \label{fig:response-time-stimulus}
@@
 
-\section{Acknowledgments} {\small This research was supported by ONR
+\subsubsection{Acknowledgments} {\small This research was supported by ONR
   MURI grant number N00014-13-1-0341, and a Berkeley Fellowship

### 2014-04-30 man/cogsci-2014/cogsci-2014.tex
Use posterior odds ratio rather than likelihood ratio

@@
 was generated from $X_a$. Because we want to determine which
-hypothesis is more likely, the quantity of interest is a ratio
-$\ell:=p(\hi\ \vert\ X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which
+hypothesis is more likely, the quantity of interest is a posterior
+odds ratio
+$\mathcal{B}:=p(\hi\ \vert\ X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which
 (assuming that all rotations are equally likely) is equivalent to:
 \begin{equation}
-  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)\cdot{}p_0}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)\cdot{}p_1},
-  \label{eq:lh-ratio}
+  \mathcal{B} = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)\cdot{}p_0}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)\cdot{}p_1},
+  \label{eq:odds-ratio}
 \end{equation}
-where $p_0=p(h=0)$ and $p_1=p(h=1)$, for brevity. If $\ell > 1$, then
+where $p_0=p(h=0)$ and $p_1=p(h=1)$, for brevity. If $\mathcal{B} > 1$, then
 we accept the hypothesis that the images depict the same object
-($\hi$); if $\ell < 1$, then we accept the hypothesis that the images
+($\hi$); if $\mathcal{B} < 1$, then we accept the hypothesis that the images
 depict flipped objects ($\hf$).
@@
 To summarize, we approximate the likelihood term of
-Equation~\ref{eq:lh-ratio} using the similarity function defined in
+Equation~\ref{eq:odds-ratio} using the similarity function defined in
 Equation~\ref{eq:similarity}. Because we assume mental rotations are
@@
 
-In order to approximate Equation~\ref{eq:lh-ratio} using samples of
+In order to approximate Equation~\ref{eq:odds-ratio} using samples of
 the similarity function, we must decide \textit{which} places to
@@
 hypothesis. This gives us the true value of $\theta$, so
-Equation~\ref{eq:lh-ratio} becomes a generalized likelihood ratio
+Equation~\ref{eq:odds-ratio} becomes a generalized likelihood ratio
 test, where $\theta$ is set to the MLE value, rather than being
@@
 \begin{equation}
-  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)\cdot{}p_0}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)\cdot{}p_1}.
+  \mathcal{B} = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)\cdot{}p_0}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)\cdot{}p_1}.
   \label{eq:mle-lh-ratio}
@@
 the global maximum, we need only \textit{approximate}
-Equation~\ref{eq:lh-ratio}. We hypothesize a model based on the idea
+Equation~\ref{eq:odds-ratio}. We hypothesize a model based on the idea
 of \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu}: instead of
 searching for a maximum, we maintain a probability distribution over
-our \textit{estimate} of Equation~\ref{eq:lh-ratio}, and then sample
+our \textit{estimate} of Equation~\ref{eq:odds-ratio}, and then sample
 actions which are expected to improve that estimate.  This strategy
@@
 which gives us a distribution over the likelihood ratio in
-Equation~\ref{eq:lh-ratio}:
-$p(\ell)\approx\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0],
+Equation~\ref{eq:odds-ratio}:
+$p(\mathcal{B})\approx\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0],
 \mathbb{V}(Z_0))\cdot{}p_0/\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1],
@@
 \BqpAccuracyCorr{} (unequal prior).\\
-\indent Because the \Bq{} model relies on Equation~\ref{eq:lh-ratio}
+\indent Because the \Bq{} model relies on Equation~\ref{eq:odds-ratio}
 for its stopping criteria (as opposed to just finding a maximum), the

### 2014-05-01 analysis/analyses/overall_response_time.py
Report millseconds in response times

@@ -31,7 +31,7 @@ def run(data, results_path, seed):
         for model, stats in results.iterrows():
             cmd = util.newcommand(
                 "%sTime" % model.capitalize(),
-                util.latex_mean.format(**dict(stats)))
+                util.latex_msec.format(**dict(stats)))
             fh.write(cmd)
 
     return pth

### 2014-05-01 analysis/analyses/util.py

@@ -30,6 +30,9 @@ latex_percent = r"$M={median:.1f}\%$, 95\% CI $[{lower:.1f}\%, {upper:.1f}\%]$"
 report_mean = "M={median:.1f}, 95% CI [{lower:.1f}, {upper:.1f}]"
 latex_mean = r"$M={median:.1f}$, 95\% CI $[{lower:.1f}, {upper:.1f}]$"
 
+report_msec = "M={median:.1f} msec, 95% CI [{lower:.1f} msec, {upper:.1f} msec]"
+latex_msec = r"$M={median:.1f}$ msec, 95\% CI $[{lower:.1f}\textrm{{ msec}}, {upper:.1f}\textrm{{ msec}}]$"
+
 
 def run_analysis(func):
     config = load_config("config.ini")

### 2014-05-01 man/cogsci-2014/cogsci-2014.tex
Misc error fixes

@@
 In this paper, we explore these questions through rational analysis
-\cite{Marr:1983to,anderson90,Shepard:1987tt} and compare four models
+\cite{Marr:1983to,Anderson:1990,Shepard:1987tt} and compare four models
 of mental rotation. We begin the paper by discussing the previous
@@
 
-Active sampling is the idea people gather new information in a manner
-that increases certainty about the problem space. An everyday example
-of this can be observed in the game of ``20 questions'', in which one
-person thinks of a concept, and another has to guess the concept in 20
-questions or less. The first question is almost always ``person,
-place, or thing?'', because the answer provides the most possible
-information about the concept of interest. Active sampling has
-gained support across several areas of cognitive science
+Active sampling is the idea that people gather new information in a
+manner that increases certainty about the problem space. An everyday
+example of this can be observed in the game of ``20 questions'', in
+which one person thinks of a concept, and another has to guess the
+concept in 20 questions or less. The first question is almost always
+``person, place, or thing?'', because the answer provides the most
+possible information about the concept of interest. Active sampling
+has gained support across several areas of cognitive science
 \cite<e.g.,>{Gureckis:2012gu}, including other spatial domains
@@
 We recruited 247 participants on Amazon's Mechanical Turk using the
-psiTurk experiment framework \cite{Mcdonnell12}. Each participant was
+psiTurk experiment framework \cite{McDonnell12}. Each participant was
 paid \$1.00 for 15 minutes of work, consisting of one block of 10

### 2014-05-01 man/cogsci-2014/cogsci-2014.tex
Last few small changes

@@
   strategy of choosing actions that will provide the most
-  informative. This last model provides a plausible account of how
+  information. This last model provides a plausible account of how
   people use mental rotation, where the other models do not. Based on
@@
 In this paper, we explore these questions through rational analysis
-\cite{Marr:1983to,Anderson:1990,Shepard:1987tt} and compare four models
-of mental rotation. We begin the paper by discussing the previous
-literature on mental imagery. Next, we outline computational- and
-algorithmic-level analyses of the problem of mental rotation.  We then
-describe a behavioral experiment based on the classic mental rotation
-studies \cite<e.g.,>{Cooper:1975wp}, and compare the results of our
-experiment with each of the models. We conclude with a discussion of
-the strengths and weaknesses of each model, and lay out directions for
-future work.
+\cite{Marr:1983to,Anderson:1990,Shepard:1987tt} and compare four
+models of mental rotation. We begin the paper by discussing the
+previous literature on mental imagery. Next, we outline computational-
+and algorithmic-level analyses of the problem of mental rotation.  We
+then describe a behavioral experiment based on the classic mental
+rotation studies \cite<e.g.,>{Cooper:1975wp}, and compare the results
+of our experiment with each of the models. We conclude with a
+discussion of the strengths and weaknesses of each model, and lay out
+directions for future work.
 
@@
 the problem of mental rotation
-\cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
+\cite{Marr:1983to,Anderson:1990,Shepard:1987tt}. At the computational
 level, we can say that the \textit{problem} is to determine which
@@
   always connected to their two nearest neighbors in the $\theta$
-  dimension.}and $\sigma_S^2=0.15$ is the variance of the
+  dimension.} and $\sigma_S^2=0.15$ is the variance of the
 similarity. Example similarity curves are shown in
@@
 
-The average RT across all correctly-judged stimuli was \ExpTime{} ms;
-the full histogram of RTs can be seen in
-Figure~\ref{fig:histograms}. The minimum angle of rotation was
-significantly rank-order (Spearman) correlated with average
-per-stimulus RTs, both for ``flipped'' (\ExpThetaTimeCorrFlipped{})
-and ``same'' pairs (\ExpThetaTimeCorrSame{}). While this replicates
-the general result of previous experiments
-\cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are not as linear
+The average RT across all correctly-judged stimuli was \ExpTime{}; the
+full histogram of RTs can be seen in Figure~\ref{fig:histograms}. The
+minimum angle of rotation was significantly rank-order (Spearman)
+correlated with average per-stimulus RTs, both for ``flipped''
+(\ExpThetaTimeCorrFlipped{}) and ``same'' pairs
+(\ExpThetaTimeCorrSame{}). While this replicates the general result of
+previous experiments \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our
+results are not as linear
 (Figure~\ref{fig:response-time-accuracy}).\\
@@
 \setlength{\bibindent}{-\bibleftmargin}
+\vspace{-0.27cm}
 \bibliography{references}

### 2017-04-21 analysis/analyses/theta_time_stimulus.py
Restyle plots

@@ -14,7 +14,7 @@ def run(data, results_path, seed):
     for key, df in data.iteritems():
         y = df[df['correct']].groupby(
             ['stimulus', 'modtheta', 'flipped'])['time']
-        results[key] = y.apply(util.bootstrap_logmean(x))
+        results[key] = y.apply(util.bootstrap_logmean)
 
     results = pd.DataFrame.from_dict(results)
     results.index = pd.MultiIndex.from_tuples(

### 2017-04-21 analysis/analyses/trial_time_means.py

@@ -15,7 +15,7 @@ def run(data, results_path, seed):
     for key in keys:
         df = data[key]
         times = df[df['correct']].groupby('trial')['time']
-        results[key] = times.apply(util.bootstrap_logmean(x))
+        results[key] = times.apply(util.bootstrap_logmean)
 
     results = pd.DataFrame.from_dict(results, orient='index')
     results.index.name = 'model'

### 2017-04-21 analysis/analyses/util.py

@@ -4,7 +4,7 @@ import pandas as pd
 
 from mental_rotation.analysis import load_human, load_model, load_all
 from mental_rotation.analysis import beta, bootcorr, modtheta
-from mental_rotation.analysis import bootstrap_median, bootstrap_mean, bootstrap_logmean
+from mental_rotation.analysis import bootstrap_median, bootstrap_mean, bootstrap_logmean, logmean
 
 
 def load_config(pth):
@@ -19,19 +19,19 @@ def newcommand(name, val):
 
 
 report_spearman = "\rho={median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
-latex_spearman = r"$\rho={median:.2f}$, 95\% CI $[{lower:.2f}, {upper:.2f}]$"
+latex_spearman = r"$\rho={median:.2f}$, 95\%\ \mathrm{{CI}}\ $[{lower:.2f}, {upper:.2f}]$"
 
 report_pearson = "r={median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
-latex_pearson = r"$r={median:.2f}$, 95\% CI $[{lower:.2f}, {upper:.2f}]$"
+latex_pearson = r"$r={median:.2f}$, 95\%\ \mathrm{{CI}}\ $[{lower:.2f}, {upper:.2f}]$"
 
 report_percent = "M={median:.1f}%, 95% CI [{lower:.1f}%, {upper:.1f}%]"
-latex_percent = r"$M={median:.1f}\%$, 95\% CI $[{lower:.1f}\%, {upper:.1f}\%]$"
+latex_percent = r"$M={median:.1f}\%$, 95\%\ \mathrm{{CI}}\ $[{lower:.1f}\%, {upper:.1f}\%]$"
 
 report_mean = "M={median:.1f}, 95% CI [{lower:.1f}, {upper:.1f}]"
-latex_mean = r"$M={median:.1f}$, 95\% CI $[{lower:.1f}, {upper:.1f}]$"
+latex_mean = r"$M={median:.1f}$, 95\%\ \mathrm{{CI}}\ $[{lower:.1f}, {upper:.1f}]$"
 
 report_msec = "M={median:.1f} msec, 95% CI [{lower:.1f} msec, {upper:.1f} msec]"
[…]

### 2017-04-21 analysis/plots/__init__.py

@@ -12,6 +12,8 @@ import theta_time_corrs
 import theta_accuracy_corrs
 import response_time_stimulus
 import human_model_corrs
+import response_times
+import accuracy
 
 __all__ = [
     'response_time_histograms',
@@ -25,5 +27,7 @@ __all__ = [
     'theta_time_corrs',
     'theta_accuracy_corrs',
     'response_time_stimulus',
-    'human_model_corrs'
+    'human_model_corrs',
+    'response_times',
+    'accuracy'
 ]

### 2017-04-21 analysis/plots/accuracy.py

@@ -1,0 +1,80 @@
+#!/usr/bin/env python
+
+import matplotlib.pyplot as plt
+import numpy as np
+import util
+import pandas as pd
+import seaborn as sns
+
+
+def plot(results_path, fig_path):
+    order = ['exp', 'oc', 'th', 'hc', 'bq', 'bqp']
+    titles = {
+        'exp': "Human",
+        'oc': "Oracle",
+        'th': "Threshold",
+        'hc': "HC",
+        'bq': "BQ (equal prior)",
+        'bqp': "BQ (unequal prior)"
+    }
+
+    colors = {
+        'same': [0.16339869, 0.4449827, 0.69750096],
+        'flipped': [0.72848904, 0.1550173, 0.19738562]
+    }
+
+    acc_results = pd.read_csv(
+        results_path.joinpath("theta_accuracy.csv"))
+
+    fig, axes = plt.subplots(2, 3, sharex=False)
+    for i, key in enumerate(order):
+        ax = axes.flat[i]
+        df = acc_results.groupby('model').get_group(key)
+
+        for flipped, stats in df.groupby('flipped'):
+            lower = stats['median'] - stats['lower']
+            upper = stats['upper'] - stats['median']
+            ax.errorbar(
+                stats['modtheta'], stats['median'],
+                yerr=[lower, upper], lw=3,
+                color=colors[flipped],
+                ecolor=colors[flipped])
+
+        ax.set_xlim(-10, 190)
+        ax.set_ylim(-5, 105)
+        ax.set_xticks(np.arange(0, 200, 60))
+        ax.set_xlabel("Rotation")
+        ax.set_yticks([0, 50, 100])
+        ax.set_yticklabels([0.0, 0.5, 1.0])
+        ax.set_title(titles[key])
+        ax.set_ylabel("Accuracy")
+
+    p0 = plt.Rectangle(
+        (0, 0), 1, 1,
+        fc=colors['same'],
+        ec=colors['same'])
+    p1 = plt.Rectangle(
+        (0, 0), 1, 1,
+        fc=colors['flipped'],
+        ec=colors['flipped'])
+
+    leg = axes[0, 0].legend(
+        [p0, p1], ["same", "flipped"],
+        numpoints=1,
+        loc='lower right',
+        bbox_to_anchor=[1.1, 0])
+
+    sns.despine()
+    fig.set_size_inches(6, 4)
+    plt.tight_layout()
[…]

### 2017-04-21 analysis/plots/response_time_histograms.py

@@ -4,6 +4,7 @@ import matplotlib.pyplot as plt
 import util
 import pickle
 import numpy as np
+import seaborn as sns
 
 
 def plot(results_path, fig_path):
@@ -21,9 +22,9 @@ def plot(results_path, fig_path):
     with open(pth, "r") as fh:
         times = pickle.load(fh)
 
-    fig, axes = plt.subplots(1, len(order), sharey=True)
+    fig, axes = plt.subplots(2, 3, sharey=True)
     for i, key in enumerate(order):
-        ax = axes[i]
+        ax = axes.flat[i]
         bins = 200
 
         if key == 'exp':
@@ -38,26 +39,25 @@ def plot(results_path, fig_path):
         edges = edges[:101]
         hist = hist[:100]
         hist = hist * 100 / float(len(times[key]))
-        width = edges[1] - edges[0]
-        ax.bar(edges[:-1], hist, width=width, color='k')
+        ax.fill_between(edges[:-1], hist, np.zeros_like(hist), color='#666666')
 
         ax.set_xlim(0, edges[-1])
-        ax.set_title(titles[key], fontsize=14)
-        util.clear_right(ax)
-        util.clear_top(ax)
-        util.outward_ticks(ax)
+        ax.set_title(titles[key])
 
         if key == 'exp':
-            ax.set_xlabel("RT (seconds)", fontsize=14)
+            ax.set_xlabel("RT (seconds)")
         else:
-            ax.set_xlabel("Number of actions", fontsize=14)
-    axes[0].set_ylabel("Percent", fontsize=14)
+            ax.set_xlabel("Number of actions")
 
-    fig.set_figheight(2.5)
-    fig.set_figwidth(16)
+    axes[0, 0].set_ylabel("Percent")
+    axes[1, 0].set_ylabel("Percent")
 
-    plt.draw()
+    sns.despine()
+
+    fig.set_figheight(4)
+    fig.set_figwidth(6)
     plt.tight_layout()
+    plt.subplots_adjust(left=0.1)
 
     pths = [fig_path.joinpath("response_time_histograms.%s" % ext)
             for ext in ('png', 'pdf')]

### 2017-04-21 analysis/plots/response_time_scatters.py

@@ -3,6 +3,7 @@
 import matplotlib.pyplot as plt
 import pandas as pd
 import util
+import seaborn as sns
 
 
 def plot(results_path, fig_path):
@@ -29,38 +30,36 @@ def plot(results_path, fig_path):
         .unstack(['model', 'flipped'])\
         .xs('all', level='flipped', axis=1)
 
-    fig, axes = plt.subplots(1, len(order), sharey=True, sharex=True)
+    fig, axes = plt.subplots(2, 3, sharey=True, sharex=True)
 
     for i, key in enumerate(order):
-        ax = axes[i]
+        ax = axes.flat[i]
         xmean = results[key]
         ymean = results['exp']
         corr = util.report_pearson.format(**dict(corrs[key]))
 
-        ax.plot([-4, 4], [-4, 4], ls='--', color='#666666')
-        ax.plot(xmean, ymean, '.', alpha=0.9, color='k')
+        ax.plot([-4, 4], [-4, 4], ls='--', color='#666666', lw=1)
+        ax.plot(xmean, ymean, '.', alpha=0.5, color='#333333')
 
         ax.set_xticks([])
         ax.set_yticks([])
+        ax.set_xlim(-4, 4)
+        ax.set_ylim(-4, 4)
+        ax.set_xlabel("Model")
 
-        ax.set_title(titles[key], fontsize=14)
-        ax.set_xlabel(corr, fontsize=14)
-        util.clear_right(ax)
-        util.clear_top(ax)
-        util.outward_ticks(ax)
+        ax.set_title(titles[key])
+        ax.text(4, -4, corr, horizontalalignment="right", verticalalignment="bottom", fontsize=8)
 
-    util.sync_xlims(axes[order.index('bq')], axes[order.index('bqp')])
+    axes[0, 0].set_ylabel("Human")
+    axes[1, 0].set_ylabel("Human")
 
-    axes[0].set_xlim(-4, 4)
-    axes[0].set_ylim(-4, 4)
-    axes[0].set_ylabel("Human", fontsize=14)
+    axes[1, 2].spines['left'].set_color('none')
+    axes[1, 2].spines['bottom'].set_color('none')
 
-    fig.set_figheight(2.5)
-    fig.set_figwidth(16)
-    plt.draw()
+    sns.despine()
+    fig.set_size_inches(6, 4)
     plt.tight_layout()
-
[…]

### 2017-04-21 analysis/plots/response_times.py

@@ -8,7 +8,7 @@ import seaborn as sns
 
 
 def plot(results_path, fig_path):
-    order = ['exp', 'th', 'bqp']
+    order = ['exp', 'oc', 'th', 'hc', 'bq', 'bqp']
     titles = {
         'exp': "Human",
         'oc': "Oracle",
@@ -19,21 +19,19 @@ def plot(results_path, fig_path):
     }
 
     colors = {
-        'same': [ 0.16339869,  0.4449827 ,  0.69750096],
-        'flipped': [ 0.72848904,  0.1550173 ,  0.19738562]
+        'same': [0.16339869, 0.4449827, 0.69750096],
+        'flipped': [0.72848904, 0.1550173, 0.19738562]
     }
 
-    means = pd.read_csv(
-        results_path.joinpath("theta_time_stimulus.csv"))\
-        .set_index(['stimulus', 'flipped', 'model'])\
-        .groupby(level='stimulus').get_group(2)
+    time_results = pd.read_csv(
+        results_path.joinpath("theta_time.csv"))
 
-    fig, axes = plt.subplots(1, len(order), sharex=True)
+    fig, axes = plt.subplots(2, 3, sharex=False)
     for i, key in enumerate(order):
-        ax = axes[i]
-        df = means.groupby(level='model').get_group(key)
+        ax = axes.flat[i]
+        df = time_results.groupby('model').get_group(key)
 
-        for flipped, stats in df.groupby(level='flipped'):
+        for flipped, stats in df.groupby('flipped'):
             if key == 'exp':
                 median = stats['median'] / 1000.
                 lower = (stats['median'] - stats['lower']) / 1000.
@@ -55,13 +53,21 @@ def plot(results_path, fig_path):
         ax.set_title(titles[key])
 
         if key == "exp":
-            ax.set_yticks([0.5, 1, 1.5, 2, 2.5, 3])
-            ax.set_ylabel("Rotation")
+            ax.set_yticks([1, 1.5, 2, 2.5, 3])
+        elif key == "oc":
+            ax.set_yticks([0, 10, 20, 30, 40])
         elif key == "th":
-            ax.set_yticks([0, 10, 20, 30, 40, 50, 60])
-            ax.set_ylabel("# Steps")
+            ax.set_yticks([0, 10, 20, 30, 40])
+        elif key == "hc":
+            ax.set_yticks([10, 15, 20, 25, 30])
+        elif key == "bq":
[…]

### 2017-04-21 analysis/plots/util.py

@@ -263,7 +263,7 @@ def sync_ylabel_coords(axes, x, y=0.5):
         ax.yaxis.set_label_coords(x, y)
 
 
-report_pearson = r"$r$ = {median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
+report_pearson = r"$r = {median:.2f}, 95\%\ \mathrm{{CI}}\ [{lower:.2f}, {upper:.2f}]$"
 
 
 def make_plot(func):

### 2017-04-21 lib/mental_rotation/analysis.py
Rerun analyses

@@ -174,5 +174,5 @@ def bootcorr(x, y, nsamples=10000, method='pearson'):
     return stats
 
 
-def logmean(x, axis=1):
-    return np.exp(np.mean(np.log(x), axis=axis))
+def logmean(x, **kwargs):
+    return np.exp(np.mean(np.log(x), **kwargs))

### 2017-04-21 analysis/analyses/util.py
Fix latex formatting

@@ -19,19 +19,19 @@ def newcommand(name, val):
 
 
 report_spearman = "\rho={median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
-latex_spearman = r"$\rho={median:.2f}$, 95\%\ \mathrm{{CI}}\ $[{lower:.2f}, {upper:.2f}]$"
+latex_spearman = r"$\rho={median:.2f}$, 95\% CI $[{lower:.2f}, {upper:.2f}]$"
 
 report_pearson = "r={median:.2f}, 95% CI [{lower:.2f}, {upper:.2f}]"
-latex_pearson = r"$r={median:.2f}$, 95\%\ \mathrm{{CI}}\ $[{lower:.2f}, {upper:.2f}]$"
+latex_pearson = r"$r={median:.2f}$, 95\% CI $[{lower:.2f}, {upper:.2f}]$"
 
 report_percent = "M={median:.1f}%, 95% CI [{lower:.1f}%, {upper:.1f}%]"
-latex_percent = r"$M={median:.1f}\%$, 95\%\ \mathrm{{CI}}\ $[{lower:.1f}\%, {upper:.1f}\%]$"
+latex_percent = r"$M={median:.1f}\%$, 95\% CI $[{lower:.1f}\%, {upper:.1f}\%]$"
 
 report_mean = "M={median:.1f}, 95% CI [{lower:.1f}, {upper:.1f}]"
-latex_mean = r"$M={median:.1f}$, 95\%\ \mathrm{{CI}}\ $[{lower:.1f}, {upper:.1f}]$"
+latex_mean = r"$M={median:.1f}$, 95\% CI $[{lower:.1f}, {upper:.1f}]$"
 
 report_msec = "M={median:.1f} msec, 95% CI [{lower:.1f} msec, {upper:.1f} msec]"
-latex_msec = r"$M={median:.1f}$ msec, 95\%\ \mathrm{{CI}}\ $[{lower:.1f}\textrm{{ msec}}, {upper:.1f}\textrm{{ msec}}]$"
+latex_msec = r"$M={median:.1f}$ msec, 95\% CI $[{lower:.1f}\textrm{{ msec}}, {upper:.1f}\textrm{{ msec}}]$"
 
 
 def run_analysis(func):
